### 2013-11-19 paper1/intro.tex
paper1 folder added

\section{introduction}
The neighbourhood of the galactic centres are mysterious places. There is very little information available to accurately map the geometric structure close to the horizon of the supermassive black holes at the galactic centre. Observing these places with great spatial resolution is one obvious solution, however, requires high quality data. The future generation telescope, like mm-VLBI, EHT etc, are capable of resolving these structures. Nevertheless, there is an alternative solution for this problem, indirect mapping of the geometry.

Microlensing provides us with an opportunity to detect the signals, comprehensively detectable, and model the signal as the function of the source shape of the source, its internal brightness and the lensing mass distribution. Now, if the source is a quasar, the radiations are mainly dominated by the accretion disc close to the supermassive black hole horizon.

The motivation of this work is to demonstrate if one assumes a parametric form for the shape (surface brightness 2D) of the accretion disk of the supermassive black hole, is it possible to recover the 2D brightness profile by looking at the brightness of the source over a range of time.

### 2013-11-19 paper1/ms.tex
paper1 folder added

\documentclass[usenatbib,twocolumn]{mn2e}

\usepackage{amsmath}
\usepackage{graphicx}

\title{Probing internal structure of distant quasars using Microlensing}

%\author[I. Mohammed et al.] {Irshad Mohammed\thanks{irshad@physik.uzh.ch}$^1$,
%Mihai Tomozeiu$^1$, Manuel Rabold$^1$,  Prasenjit Saha,$^1$ and Joachim Wambsgans$^3$\\
%$^1${Institute for Theoretical Physics, University of Z$\ddot{u}$rich, 8057 Z$\ddot{u}$rich Switzerland}\\
%$^2${University of Heidelberg, Germany}
%}

\begin{document}

\maketitle

\begin{abstract}
We present a toy model to parametrize and constraint the innermost environment of a distant quasar using microlensing. 
\end{abstract}

\begin{keywords}
Supermassive black holes, microlensing, quasars.
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{intro}

\input{theory}

\input{sourcemodels}

\input{discussion}

\citep{2006MNRAS.367.1209L}

\bibliographystyle{mn2e}

\def\apj{ApJ}
\def\apjl{ApJL}
\def\aj{AJ}
\def\mnras{MNRAS}
\def\aap{A\&A}

\bibliography{heap}

\end{document}

### 2013-11-19 paper1/theory.tex
paper1 folder added

\section{Theory}

\subsection{Basic microlensing theory}

We start by introducing in a succint manner the gravitational lensing theory that is relevant for microlensing in general and for the scope of the
present paper in particular. For a detailed presentation of the theory the reader is referrend to the books and articles that represent the references
of the current section (book by Schneider and others).\\

The general microlensing equation can be written as:
\begin{equation}
\vec{y} =   
 \begin{pmatrix}
  1-\gamma - k & 0 \\
  0 & 1 + \gamma -k \\
 \end{pmatrix} 
\vec{x} - \sum_{i=1}^{n} m_i \frac{\vec{x} - \vec{x_i}}{\left( \vec{x} - \vec{x_i} \right)^2}.
\end{equation}

The equation is valid for a set of $\it{n}$ point (Schwarzschild) lenses placed at coordinates $\vec{x_i}$. $\vec{y}$ represents the coordinates in the source plane for the respective ray, while 
$\vec{x}$ represents the coordinates in the lens plane. Furthermore, $\gamma$ and $k$ are the shear and surface mass density.\\

For the lens mapping a Jacobian can be defined:
\begin{equation} 
A(\vec{x}) = 
 \begin{pmatrix}
  \frac{\partial{y_1}}{\partial{x_1}} & \frac{\partial{y_1}}{\partial{x_2}} \\
  \frac{\partial{y_2}}{\partial{x_1}} & \frac{\partial{y_2}}{\partial{x_2}} \\
 \end{pmatrix}. 
\end{equation}
The determinant of $A(\vec{x})$ is related to the local magnification factor through the equation:

\begin{equation}
\mu(\vec{x}) = \left| \frac{1}{det(A(\vec{x}))} \right|
\end{equation}

According to this equation, from an infinitesimally small source positioned at $\vec{x}$ in the lens plane an observed will receive a flux $\mu(\vec{x}) dF$ instead of the flux $dF$ that would be
observed in the absence of the gravitational lens. Therefore the source will dim or brighten depending on whether $\mu(\vec{x})$ is subunitary or supraunitary.\\

In general, for a single infinitesimally small source more than one image can be observed, each with magnification $\mu_j$. For microlensing events the angular sizes of the
images cannot be resolved. The only observable is the flux to which all nearby images contribute. A total magnification
\begin{equation}
 \mu_p = \sum_{j} \mu_j
\end{equation}
 can be defined and account for an increase/decrease of brightness from all images.

\subsection{Magnification of a point source near a fold}

The determinant of the Jacobian $A$ can have both positive and negative values depending in which region it is computed.Such regions are well defined and separated
by critical curves where the determinant vanishes. The lens equation maps the critical curves of the lens plane into caustics in the source plane. An example of the caustic 
shape can be seen in the magnification map of |||Figure 1|||.

||||||||||||||||||||
||||||Figure 1||||||
||||||||||||||||||||

Along these critical curves in the lens plane or caustics in the source plane,  the magnification $\mu$ is infinite, a result that simply follows from equation ( 
(3). This statement holds only fo infinitesimally small sources. For extended sources the maximum amplification is finite since only an infinitesimally small area of the source
overlaps with the caustics (Schneider Weiss 1986 paper others).

The matrix $A$ at the coordinates of the caustic can have rank 1 or rank 0. If the rank is 1 then the coordinates belong to $fold$ or $cusp$ singularities. For the purpose 
of this paper we are interested only in the behaviour around $fold$ singularities. A condition necessary to distinguish between a fold and cusp singularities is that the
eigenvector of $A$ corresponding to the null eigenvalue is not a tangent vector of the critical curve (Dominik 2008, Schneider Weiss 1992).

||||||||||||||||||||
||||||Figure 2||||||
||||||||||||||||||||

The behaviour of the images of a point source and their corresponding magnifications near such a caustic has been thoroughly studied in the past.

### 2013-11-19 paper1/ms.tex
paper1 added
ms.tex changed

@@
 
-%\author[I. Mohammed et al.] {Irshad Mohammed\thanks{irshad@physik.uzh.ch}$^1$,
-%Mihai Tomozeiu$^1$, Manuel Rabold$^1$,  Prasenjit Saha,$^1$ and Joachim Wambsgans$^3$\\
-%$^1${Institute for Theoretical Physics, University of Z$\ddot{u}$rich, 8057 Z$\ddot{u}$rich Switzerland}\\
-%$^2${University of Heidelberg, Germany}
-%}
+\author[I. Mohammed et al.] {Irshad Mohammed\thanks{irshad@physik.uzh.ch}$^1$,
+Mihai Tomozeiu$^1$, Manuel Rabold$^1$,  Prasenjit Saha,$^1$ and Joachim Wambsgans$^3$\\
+$^1${Institute for Theoretical Physics, University of Z$\ddot{u}$rich, 8057 Z$\ddot{u}$rich Switzerland}\\
+$^2${University of Heidelberg, Germany}
+}
 
@@
 
+\input{fitting}
+
 \input{discussion}

### 2013-11-19 paper1/code.tex
code.tex added

\section{Microlensing code and magnification maps}

### 2013-11-22 paper1/code.tex
draft_code_section

@@
 \section{Microlensing code and magnification maps}
+For the numerical computations of lightcurves for the models a and b, the microlensing code by Joachim Wambsgans was used, which is described in Wambsganss 1999.
+This code is based on the hierachical tree technique to calculate the gravitational effect of the lensing masses(--> references).
+The underlying idea is that a microlensing situation can be separated into three planes, the observer, the lens plane and the source plane (assuming the thin lens approximation to be valid).
+The actual effect, namely the distortion of lightrays emitted from a distant sources, on the way to the observer, is numerically computed by treating this situation from the opposite side: lightrays are shot backward from the observer to lens plane, where they become deflected according to the angle, which is given by the 2-dim.
+mass distribution in the lens plane.
+\begin{equation}
+\vec{\alpha}_{i}=\frac{4G}{c^{2}} \frac{D_{LS}}{D_{S}}\sum_{j=1}^{n}M_j \frac{\vec{r}_{ij}}{r^2_{ij}} 
+\end{equation}  
+For the computation of the individual deflection angles the tree idea is used.
+This means that the positions of all lensing masses are sorted into a 2-dim.
+grid (in the lens plane).
+This grid is subdivided into 4 smaller squares until every cell contains only one mass.
+For the actual calculations of the deflection angle not every mass' gravitational influence is treated equally.
+Instead it is made use of the fact that the influence of lenses on the considered lightray falls off with $r^{-1}$ from the point where the ray hits the lens plane.
+Hence masses situated at higher distances, can be grouped together in the calculation of their gravitational potential, by multipole moments.
+This way the amount of computation time is reduced.
+When the deflection angles are determined, the deflected lightrays are traced to the source plane and a magnification map is created by counting the number of rays which hit each of its pixels. 
+Once the map is created, the lightcurve can be obtained by specifying the transition path of the source within the map.
+The code thereby convolves the function describing the shape of the source with the magnification pattern of the map for each time step.
+Note that in the most observational cases in microlensing, it is the lens that is actually moving in front of a fixed background.
+For the model a it was desirable to work with a caustic geometry as simple as possible. Therefore only two point like masses where used as lenses.

### 2013-12-02 paper1/theory.tex
Edits from Mihai

@@
 We start by introducing in a succint manner the gravitational lensing theory that is relevant for microlensing in general and for the scope of the
-present paper in particular. For a detailed presentation of the theory the reader is referrend to the books and articles that represent the references
+present paper in particular. For a detailed presentation of the theory the reader is referred to the books and articles that represent the references
 of the current section (book by Schneider and others).\\
@@
 According to this equation, from an infinitesimally small source positioned at $\vec{x}$ in the lens plane an observed will receive a flux $\mu(\vec{x}) dF$ instead of the flux $dF$ that would be
-observed in the absence of the gravitational lens. Therefore the source will dim or brighten depending on whether $\mu(\vec{x})$ is subunitary or supraunitary.\\
+observed in the absence of the gravitational lens. In other words the source will dim or brighten depending on whether $\mu(\vec{x})$ is subunitary or supraunitary.\\
 
@@
 \begin{equation}
- \mu_p = \sum_{j} \mu_j
+ \mu_t = \sum_{j} \mu_j
 \end{equation}
@@
 
-The determinant of the Jacobian $A$ can have both positive and negative values depending in which region it is computed.Such regions are well defined and separated
+The determinant of the Jacobian $A$ can have both positive and negative values depending in which region it is computed. Such regions are well defined and separated
 by critical curves where the determinant vanishes. The lens equation maps the critical curves of the lens plane into caustics in the source plane. An example of the caustic 
-shape can be seen in the magnification map of |||Figure 1|||.
+shape can be seen in the magnification map of |||Figure 1|||.\\
 
@@
 The matrix $A$ at the coordinates of the caustic can have rank 1 or rank 0. If the rank is 1 then the coordinates belong to $fold$ or $cusp$ singularities. For the purpose 
-of this paper we are interested only in the behaviour around $fold$ singularities. A condition necessary to distinguish between a fold and cusp singularities is that the
+of this paper we are interested only in the behaviour around $fold$ singularities. A condition necessary to distinguish between a fold and cusp singularity is that the
 eigenvector of $A$ corresponding to the null eigenvalue is not a tangent vector of the critical curve (Dominik 2008, Schneider Weiss 1992).
@@
 
-The behaviour of the images of a point source and their corresponding magnifications near such a caustic has been thoroughly studied in the past.
+The behaviour of the images of a point source and their corresponding magnifications near such a caustic has been thoroughly studied in the past. The reader who is interested in 
+understanding the details related to this subject is referred to papers such as Blandford and Narayan 1986, Schneider and Weiss 1986, Gaudi and Petters 2002. For the present paper we only 
+summarize the common result:
+\begin {equation}
+ \mu_{t}(p) = \mu_0 + C_0 \frac{1}{\sqrt{p}} \Theta(p).
+\end {equation}
+As the previous equation describes the magnification of a point source near a caustic is equal to the sum of the magnification due to other reasons $\mu_0$, assumed to be locally constant,  and a decrease with
+the square root of the distance from the fold.
+The later term becomes activated only after the source enters the region interior to the caustic curves when the stepfunction $\Theta(p)$ become unity.
+The proportionality
+constant $C_0$  is sensible only to the local behaviour of Fermat's potential and units used for the lens plane coordinates $\vec{y}(p,q)$.
+
+\subsection{Magnification of an extended source near a fold}
+
+A source of arbitrary shape can be described by a two dimensional brightness function $S_{2D}(p - p_s, q - q_s)$ defined for a coordinate system $p,q$ where $p_s, q_s$ denote the coordinates of the center of source.
+
+For a microlensing event the lightcurve can be written for an undefined source shape as:
+
+\begin{equation}
+ F(t) = \int_{-\infty}^\infty \int_{-\infty}^\infty S_{2D}(p-p_s(t), q-q_s(t)) \mu_t(p) \mathrm{d}q \mathrm{d}p
+\end{equation}
+
+In order to build the previous equation we have considered that the time dependency of the flux $F$ is given only by the motion of the source with respect to a fixed caustic.
+Therefore the only time dependent quatities in the right hand side of the equation
+are the coordinates of the center of the source $p_s,q_s$ and by construct $S_{2D}$.
+
+Due to the choice of the coordinate system the amplification factor has no dependence on the $q$ coordinate. The previous equation can be rewritten as:
+
+\begin{equation}
+ F(t) 
+= \int_{-\infty}^\infty  \mu_t(p) S_{1D}\left(p-p_s(t)\right) \mathrm{d}p,
+\end{equation}
+
+where we have defined the one dimensional flux function as:
+\begin{equation}
+ S_{1D}(p-p_s(t)) = \int_{-\infty}^\infty S_{2D}(p-p_s(t), q-q_s(t)) \mathrm{d}q
+\end{equation}
+
+This representation is a valid approximation only when the apparent size of the source is much smaller than the corresponding Einstein angle of the lens. In this context 
+all the information about the source shape and brightness will that can be contained in the lightcurve is exhaustively given by the 1D flux function.
+In other words, if two sources with $S_{2D}$ have the same $S_{1D}$ they cannot be distinguished by studying their lightcurves.
+
+\section{Models for extended sources}
+
+In the present study we are analysing three types of sources with different surface brighness: 
+\begin{enumerate}
+ \renewcommand{\theenumi}{(\arabic{enumi})}
+  \item a rotationally symmetric source with a bivariate gaussian surface brightness distribution,
+  \item a disk source with constant surface brightness distribution,
+  \item a crescent shaped source with constant surface brightness distribution.
+\end{enumerate}
+The first two sources are the typical choices used in the literature to describe the luminous parts of a quasar (Prasenjit should give some citations here). The third one is a newly proposed
+variant (Kammrudin 2013 Geometric crescent).
+
+\subsection{Rotationally symmetric source with a bivariate gaussian surface brightness distribution}
+
+A symmetric 2D gaussian can be described mathematically as:
+
+\begin{equation}
+ S_{2D}^G(p-p_s, q-q_s) = \frac{S_0^G}{2 \pi \sigma^2} e^{-\frac{(p-p_s)^2}{2 \sigma^2}} e^{-\frac{(q-q_s)^2}{2 \sigma^2}}.
+\end{equation}
+
+The corresponding 1D brightness is:
+
+\begin{equation}
+ S_{1D}^G(p-p_s) = \frac{S_0^G}{\sqrt{2 \pi} \sigma} e^{-\frac{(p-p_s)^2}{2 \sigma^2}}.
+\end{equation}
+
+Other parameters of the model are the total flux $S_0^G$ and $\sigma$.
+
+Although such a definition for a source would have non-zero surface/linear brightness for any coordinate $p,q$, the amount of light received by a detector from outside a $3 \sigma$ disk centered at $p_s, q_s$
+would be insignificant.
+
+\subsection{Disk source with constant surface brightness distribution}
+
+One can construct mathematically a disk source with constant surface brightness and radius $R$ using a stepfunction:
+\begin{equation}
+ S_{2D}^D(p-p_s, q-q_s) = \frac{S_0^D}{\pi R^2} \Theta \left( R^2 - \left( p-p_s \right)^2 - \left( q-q_s \right)^2 \right).
+\end{equation}
+By integrating over $q$ coordinate the linear brightness function is obtained:
+
+\begin{equation}
+ S_{1D}^D(p-p_s) = \frac{2 S_0^D}{\pi R}  \sqrt{1 - \frac{(p-p_s)^2}{R^2} }    \Theta \left( R^2 - \left( p-p_s \right)^2 \right).
+\end{equation}
+
+\subsection{Crescent source with constant surface brightness distribution}
+
+The surface brightness distribution of a geometric crescent can be built by considering two disk sources of constant brightness.
+One larger disk will contribute positively to the total flux, while one smaller disk
+that is interior to the large one will contribute negatively. This superposition can be written for 2D as:\\
+
+\begin{equation}
+ S_{2D}^C =  S_{2D}^{Dp} -  S_{2D}^{Dn}   (n+1)
+\end{equation}
+with\\
+
+\begin{equation}
+ S_{2D}^{Dp}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dp}}{\pi R_p^2} \Theta \left( R_p^2 - \left( p-p_{sp} \right)^2 - \left( q-q_{sp} \right)^2 \right)
+\end{equation}
+
+and\\
+\begin{equation}
+ S_{2D}^{Dn}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dn}}{\pi R_n^2} \Theta \left( R_n^2 - \left( p-p_{sn} \right)^2 - \left( q-q_{sn} \right)^2 \right).
+\end{equation}
+
+The following notations were used: $R_p, (p_{sp}, q_{sp}), R_n, (p_{sn},q_{sn})$ are the radii and coordinate of the center for the larger positive disk and smaller negative disk.  
+$S_0^{Dp},S_0^{Dn}$ represent the total flux of radiation received from the large and small disk. From this point forward we will not use the total flux from each source. Instead we will 
+use the difference which in this case is th total flux from the crescent-shaped source $S_0^C$. \\
+Equation(n+1) can be written as:\\
+
+\begin{align}
+ S_{2D}^C &= \frac{S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \left\{ \Theta \left[ R_p^2 - \left( p-p_{sp} \right)^2 - \left( q-q_{sp} \right)^2 \right] \right.\nonumber\\
+ &\qquad \left. {} -  \Theta \left[ R_n^2 - \left( p-p_{sn} \right)^2 - \left( q-q_{sn} \right)^2 \right] \right\}.
+\end{align}
+
+Analogous for the linear brightness function:
+
+\begin{align}
+ S_{1D}^C &= \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \left\{ \sqrt{R_p^2 - (p-p_{sp})^2}  \Theta \left[ R_p^2 - \left( p-p_{sp} \right)^2 \right] \right.\nonumber\\
+ &\qquad \left. {} - \sqrt{R_n^2 - (p-p_{sn})^2 } \Theta \left[ R_n^2 - \left( p-p_{sn} \right)^2 \right] \right\}.
+\end{align}
+
+--------------------------------------------------------------------------
+Figure with 2d 1d brightness for gaussian, disk, some crescent
+----------------------------------------------------------------------------
+
+There are some constraints on the parameters used to define a crescent in the previously presented manner that need to be stated. First, we must impose the obvious $R_p > R_n$ relation. Secondly, 
+the small disk must always be interior to the large disk:
+\begin{equation}
+ R_p \ge R_n + \sqrt{\left(p_{sp} - p_{sn} \right)^2 + \left(q_{sp} - q_{sn} \right)^2}
+\end{equation}
+
+For the distances between the centers of the two disks we will use the same notations as the one found in the paper [Geometric crescent], $a \equiv p_{sn} - p_{sp}$ and $b \equiv q_{sn} - q_{sp}$.
+
+\section{Lightcurves of the extended sources during fold crossing}
+
+Using equation (7) and the one-dimensional flux function presented in the previous section one can compute numerically the lightcurves of the three extented sources fot the simplified 
+infinite-wall-caustics model.
+
+\subsection{Lightcurve of the gaussian source}
+
+The amount of light received by an observer from a source with a gaussian distributed brightness with $\sigma$ and total flux $S_0^G$ in the absence of any gravitational lensing is:
+\begin{equation}
+ F^G(t) = \int_{-\infty}^\infty  \left( \mu_0 + \frac{C_0}{\sqrt{p}} \Theta \left( p \right) \right) \left( \frac{S_0^G}{\sqrt{2 \pi} \sigma} e^{-\frac{(p-p_s(t))^2}{2 \sigma^2}} \right) \mathrm{d}p.
+\end{equation}
+which can be simplified to:
+\begin{equation}
+ F^G(t) = \mu_0 S_0^G + \frac{C_0 S_0^G}{\sqrt{2\pi} \sigma} \int_{0}^\infty \frac{e^{-\frac{(p-p_s(t))^2}{2 \sigma^2}}}{\sqrt{p}} \mathrm{d}p.
+\end{equation}
+The result is equivalent to the one obtained in papers such as (citations).\\
+
+\begin{figure}
+\includegraphics[bb = 0 0 600 400]{lightcurve_crescent_1.pdf}
+\caption{Lightcurve 1.}
+\end{figure}
+
+\subsection{Lightcurve of the disk shaped source}
+
+Analogous to the gaussian shaped source, the disk source with uniform brightness, radius $R$ and unmagnified flux $S_0^D$ has a lightcurve described by the equation:
+
+\begin{equation}
+ F^D(t) = \int_{-\infty}^\infty  \left( \mu_0 + \frac{C_0}{\sqrt{p}} \Theta \left( p \right) \right) \left( \frac{2 S_0^D}{ \pi R} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}} \Theta \left(R^2 - \left(p-p_s(t) \right)^2 \right) \right) \mathrm{d}p.
+\end{equation}
+
+which is equivalent to:
+\begin{equation}
+ F^D(t) = \mu_0 S_0^D + \frac{2 C_0 S_0^D}{\pi R} \int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \frac{1}{\sqrt{p}} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}} \mathrm{d}p.
+\end{equation}
+As obtained in preciding papers by (citations).\\
+
+\begin{figure}
+\includegraphics[bb = 0 0 600 400]{lightcurve_gaussian_1.pdf}
+\caption{Lightcurve 2.}
+\end{figure}
+
+\subsection{Lightcurve of the crescent shaped source}
+
+The lightcurve of a crescent shaped source with unamplified flux $S_0^C$, radii $R_p$, $R_n$ and center displacement $a(t)$ is:
+\begin{equation}
+ F^c(t) = \mu_0 S_0^C + C_0 \frac{2 S_0^C}{\pi \left( R_p^2 -R_n^2 \right) } 
+\left(\int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) \right)^2 }{p}} \mathrm{d}p 
+  -  \int_{max(0, p_s(t) - a(t) - R)}^{max(0, p_s(t) -a(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) +a(t) \right)^2 }{p}}  \mathrm{d}p \right)
+\end{equation}
+
+In the previous equation we have renamed the coordinate of the center of the large disk $p_{sp}$ to $p_{s}$.  \\
+
+\begin{figure}
+\includegraphics[bb = 0 0 600 400]{lightcurve_disc_1.pdf}
+\caption{Lightcurve 3.}
+\end{figure}
+
+The function $p_s(t)$ can be chosen to be equal to $v_p(t-t_0) + p_{s0}$. Where $p_{s0}$ is the coordinate $p$ of the source at the initial time, and $v_p$ is the component of the velocity
+along the $p$ axis. Such a modeling of the motion of the object in the source plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
+chosing the function $a(t)$ to be constant in time.

### 2014-01-08 paper1/fitting.tex

\section{model fitting}

\begin{figure}
  \includegraphics[width=0.95\hsize]{plots/data.eps}
\caption{data}
\end{figure}

\begin{figure}
  \includegraphics[width=0.5\hsize]{plots/Crescent.eps}
  \includegraphics[width=0.5\hsize]{plots/CrescentReverse.eps}\\
  \includegraphics[width=0.5\hsize]{plots/Disk.eps}
  \includegraphics[width=0.5\hsize]{plots/Gauss.eps}
\caption{Model Fitting}
\end{figure}

### 2014-01-09 paper1/code.tex
08.01.14

@@
 For the numerical computations of lightcurves for the models a and b, the microlensing code by Joachim Wambsgans was used, which is described in Wambsganss 1999.
-This code is based on the hierachical tree technique to calculate the gravitational effect of the lensing masses(--> references).
+This code is based on the hierachical tree technique to calculate the gravitational effect of the lensing masses.
 The underlying idea is that a microlensing situation can be separated into three planes, the observer, the lens plane and the source plane (assuming the thin lens approximation to be valid).
-The actual effect, namely the distortion of lightrays emitted from a distant sources, on the way to the observer, is numerically computed by treating this situation from the opposite side: lightrays are shot backward from the observer to lens plane, where they become deflected according to the angle, which is given by the 2-dim.
+The actual effect, namely the distortion of lightrays emitted from a distant source, on the way to the observer, is numerically computed by treating this situation from the opposite side: lightrays are shot backward from the observer to lens plane, where they become deflected according to the angle, which is given by the 2-dim.
 mass distribution in the lens plane.
@@
 This way the amount of computation time is reduced.
-When the deflection angles are determined, the deflected lightrays are traced to the source plane and a magnification map is created by counting the number of rays which hit each of its pixels. 
+\\
+When the deflection angles are determined, the deflected lightrays are traced to the source plane and a magnification map is created by counting the number of rays which hit each of its pixels. \\\\
 Once the map is created, the lightcurve can be obtained by specifying the transition path of the source within the map.
-The code thereby convolves the function describing the shape of the source with the magnification pattern of the map for each time step.
-Note that in the most observational cases in microlensing, it is the lens that is actually moving in front of a fixed background.
-For the model a it was desirable to work with a caustic geometry as simple as possible. Therefore only two point like masses where used as lenses.
+The code thereby convolves the function describing the shape of the source with the magnification pattern of the map for each time step (see equation (6)).
+Note that in the most observational cases in microlensing, both lens and source are moving and further that the lens configuration, and with it the magnification pattern, is changing with time.
+While the first subtlety i taken care of by a coordinate transformation in this analysis, for the second one the lens configuration is assumed to be constant in time.
+\\\\
+For the model a it was desirable to work with a caustic geometry as simple as possible.
+Therefore only two equal point like masses were used as lenses, to produce the magnification map depicted in figure (...).
+The magnification maps calculated by the microlensing code are in units of Einstein radii, hence they are principally scale free.
+Physical scale is introduced to them by multiplying the pixels by the correct distance factor.
+The Einstein angle is given by
+\begin{equation}
+\theta_{E}=\sqrt{\frac{4GM}{c^{2}} \frac{D_{LS}}{D_{S}D_{L}}} 
+\end{equation}  
+and therefore depends on the four parameters $M,D_{S},D_{L},D_{LS}$. For the cosmological application or quasar lensing $D_{L}$ is of order 100 Mps to Gpc and $D_{S}$ of order Gpc.\\   
+The scale of the magnification map can be converted into a physical length by multiplying it by a factor of $ \sqrt{\frac{D_{LS}}{D_{S}D_{L}}}$.\\\\
+For the next step, the computation of the actual lightcurve, the code was modified to also allow for crescent shaped images specified through the parameter set $R_p,R_n,a,b$.
+In the original version of the code gaussian and disk shaped images where already implemented.
+Those are completely characterized by one parameter $R_p$.
+The values of this parameters are specified in pixel units corresponding to the magnification map.
+Further one needs to specify the start and end point coordinates of the path which the center of the source image follows in the magnification map (see the depiction in figure (...)).
+---Number of timesteps, convolution----

### 2014-01-16 paper1/code.tex
16.1.14

@@
 \\
-When the deflection angles are determined, the deflected lightrays are traced to the source plane and a magnification map is created by counting the number of rays which hit each of its pixels. \\\\
+When the deflection angles are determined, the deflected lightrays are traced to the source plane and a magnification map is created by counting the number of rays which hit each of its pixels.
+This is the inverse of what would happen in a real observation: here an observer on earth would detect light emitted by several sources with her/his telescope.
+Each of these sources emits lightrays which are subject to the deflection by the intervening gravitational potential of the lensing mass.
+However only a fraction of those will end in the observers telescope on earth.
+Interesting for the analysis are hence those lightrays which are originating by the source of interest and further the subset of these which actually converge at the location of the observer.
+So instead of calculating the trajectories of lightrays emitted by the source in all directions and then succeedingly pick out the ones which end in the telescope of the observer, the computational effort can be restricted to these lightrays of interest in the first place.
+This is done by treating the problem from the opposite direction, namely by using the observers location as starting point and then shooting the lightrays backwards till they hit the source plane.
+The result is a pixel map of the number of lightrays which arrive at the source plane.
+This intermediate result is denoted as the magnification map.
+It characterizes the information about the mass distribution in the lens plane and its effect on lightrays, and projects this information onto the source plane.
+\\\\
 Once the map is created, the lightcurve can be obtained by specifying the transition path of the source within the map.
@@
 Note that in the most observational cases in microlensing, both lens and source are moving and further that the lens configuration, and with it the magnification pattern, is changing with time.
-While the first subtlety i taken care of by a coordinate transformation in this analysis, for the second one the lens configuration is assumed to be constant in time.
+While the first subtlety is taken care of by a coordinate transformation in this analysis, for the second one the lens configuration is assumed to be constant in time.
 \\\\
@@
 Therefore only two equal point like masses were used as lenses, to produce the magnification map depicted in figure (...).
+For a description of the parameter set which characterizes the map see (reference).
 The magnification maps calculated by the microlensing code are in units of Einstein radii, hence they are principally scale free.
@@
 For the next step, the computation of the actual lightcurve, the code was modified to also allow for crescent shaped images specified through the parameter set $R_p,R_n,a,b$.
-In the original version of the code gaussian and disk shaped images where already implemented.
-Those are completely characterized by one parameter $R_p$.
-The values of this parameters are specified in pixel units corresponding to the magnification map.
-Further one needs to specify the start and end point coordinates of the path which the center of the source image follows in the magnification map (see the depiction in figure (...)).
----Number of timesteps, convolution----
+Here $R_p$ denotes the outer radius of the crescent and $R_n$ the inner one.
+The orientation of the source image with respect to the magnification map is specified by the parameters $a$ and $b$ as the shift of the center of the inner disk from the center of the outer disk in $x$- and $y$-direction respectively.
+In the original version of the code, gaussian and disk shaped images where already implemented.
+Those are completely characterized by the single parameter $R_p$.
+The values of the parameters are specified in pixel units corresponding to the magnification map.
+Further one needs to specify the start and end point coordinates of the path, which the center of the source image follows through the magnification map (see the depiction in figure (...)).
+Hence the points along this path are specified through the number of timesteps for which the computation of the brightness is to be carried out.
+Those points correspond to the actual measurement of the brightness of an object in the observational case.
+For each timestep the two-dimensional convolution of equation (6) is carried out numerically for the position of the source on the magnification map.
+\\
+For the purpose of this analysis it was desirable to mimic the analytical behavior of a simple fold as much as possible, for comparing the numerical result with the analytical one, therefore the path of the source was chosen, so that it intersects the border of the caustic perpendicularly, and on a point where the border is almost straight.

### 2014-02-11 paper1/fitting.tex
fitting.tex is updated

@@
 \begin{figure}
+\centering
   \includegraphics[width=0.95\hsize]{plots/data.eps}
-\caption{data}
+\caption{\label{fig:datafitting} data}
 \end{figure}
@@
 \begin{figure}
-  \includegraphics[width=0.5\hsize]{plots/Crescent.eps}
-  \includegraphics[width=0.5\hsize]{plots/CrescentReverse.eps}\\
-  \includegraphics[width=0.5\hsize]{plots/Disk.eps}
-  \includegraphics[width=0.5\hsize]{plots/Gauss.eps}
-\caption{Model Fitting}
+\centering
+  \includegraphics[width=0.95\hsize]{plots/image.eps}
+  \includegraphics[width=0.95\hsize]{plots/image2.eps}
+\caption{\label{fig:mcmc} Model Fitting}
 \end{figure}
+
+In this section, we attempt to recover the crescent model parameters ($R_p, R_n, a, b$) from the given dataset generated by different source models - crescent, uniform-disk and Gaussian-disk for a given set of parameters.
+The form of light curves in each source model can be seen in figure \ref{fig:datafitting}.
+There is a fourth dataset included, which is generated through the crescent source model crossing the caustic (see the magnification map) for the second time but in reverse order (indicated as reverse fold in figure \ref{fig:datafitting}).
+This dataset resembles an arbitary caustic crossing of a crescent shape source including noise.
+The data assumed was very  ideal, including 200 different data points at regular interval of time.
+The aim of this exercise was to explore the degeneracies in  various crescent parameters and recover the correct fiducial model assumed.
+Also, it can be inferred if one can simply distinguish between circularly symmetric and asymmetric sources or uniform and Gaussian sources.
+
+We explored the parameter space (for crescent model) using Markov-Chain Monte-Carlo (MCMC), assuming the errors on light curve data point are Gaussian distributed and about 10 percent of the light curve data point.
+Following the same color scheme (as in figure \ref{fig:datafitting}), figure \ref{fig:mcmc} shows the raw MCMC for each case, top row resembles the real parameter space that was explored, however, the bottom row shows the MCMC in the so-called 'derived parameters' which are more intuitive in their physical meaning.
+
+In the first case, we tried to fit the crescent data (blue line in figure \ref{fig:datafitting}) with our crescent model parameters.
+The maximum likelihood (or the best fit) overlaps well with the fiducial values of the parameters assumed to generate the data.
+Also a degeneracy between the parameters is noticeable, particularly between $a$ and $b$.
+The overall shape of the crescent, the outer and inner radius, can be very well constrained.
+
+In second case, we tried to fit the reverse crescent data (red line in figure \ref{fig:datafitting}).
+As already mentioned, this case resemble an arbitary caustic crossing including noise, it is the most interesting case study.
+We see the similar degeneracies in $a$ and $b$ as in the previous case, however, in the parameter space both the parameters changes the sign.
+This is also intuitive as we are assuming the motion of the source in reverse direction as compared to the previous case.
+On the other hand, $R_p$ and $R_n$ are ver well constrained and overlap nicely with the previous case (see red curves in figure \ref{fig:mcmc}).
+Even for $R_p$ and $R_n$ the maximum likelihood does not exactly co-incide with the fiducial values, but it is in the neighbourhod and  within 1-$\sigma$ errors.
+
+In third case, we shifted our analysis towards sources with circularly symmetric brighness profile.
+First, we tested a uniform disk case (green in figure \ref{fig:datafitting} and \ref{fig:mcmc}).
+Also, one can assume the uniform disk as a special case of the crescent model with $R_n = 0$.
+In such scenario, one should expect the maximum likelihood for any arbitary $a$ and $b$, and $R_n = 0$.
+This is exactly what one can see in figure \ref{fig:mcmc}.
+The overall size of the source can be recovered as $R_p$ is well constrained in this case.
+
+In our last case, we tested another circularly symmetric source but with Gaussian brightness profile (Gaussian disk, magenta color in figure \ref{fig:datafitting} and \ref{fig:mcmc}).
+One can easily notice in figure \ref{fig:mcmc}, that the correct crescent parameters cannot be recovered.
+
+In the last two cases (with circularly symmetric sources), $R_n$ can assume a long range of values or it is weakly constrained, however,  $R_p$ is very well constrained.
+This outer radius $R_p$ can be very well recovered in the first case (uniform disk) but assumed a larger value for Gaussian disk case.
+If one can imagine total brighness profile should be conserved, this observation is intuitive.
+
+We finally conclude the following:
+\begin{enumerate}
+ \item If a source have a crescent shape, it can be well recovered. 
+ \item If a source have circularly symmetric brighness profile, it can also be identified by looking at the parameter space.
+ \item It is also possible to distinguish if the circularly symmetric source has a uniform brightness profile or a Gaussian brightness profile.
+
+\end{enumerate}

### 2014-02-11 paper1/plots/plot.py

@@ -15,31 +15,31 @@ markers = 'or'
 fig = figure()
 
 ax0 = fig.add_subplot(1,2,1)
-ax0.plot(x[:,0],x[:,1],'r',label='Reverse Crescent')
-ax0.plot(xx[:,0],xx[:,1],'b',label='Crescent')
-ax0.plot(y[:,0],y[:,1],'g',label='Uniform Disk')
-ax0.plot(z[:,0],z[:,1],'m',label='Gaussian Disk')
+ax0.plot(x[:,4],x[:,5],'r',label='Reverse Crescent')
+ax0.plot(xx[:,4],xx[:,5],'b',label='Crescent')
+ax0.plot(y[:,4],y[:,5],'g',label='Uniform Disk')
+ax0.plot(z[:,4],z[:,5],'m',label='Gaussian Disk')
 legend(loc=2)
-ax0.set_xlim(20.0, 90.0)
-ax0.set_ylim(0.0, 90.0)
+ax0.set_xlim(30.0, 60.0)
+ax0.set_ylim(0.0, 2.0)
 x0, x1 = ax0.get_xlim()
 y0, y1 = ax0.get_ylim()
 ax0.set_aspect((x1-x0)/(y1-y0))
-xlabel('Rp')
-ylabel('Rn')
+xlabel('R_half')
+ylabel('Rn/Rp')
 
 ax1 = fig.add_subplot(1,2,2)
-ax1.plot(x[:,4],x[:,5],'r',label='Reverse Crescent')
-ax1.plot(xx[:,4],xx[:,5],'b',label='Crescent')
-ax1.plot(y[:,4],y[:,5],'g',label='Disk')
-ax1.plot(z[:,4],z[:,5],'m',label='Gaussian Disk')
+ax1.plot(x[:,2]/x[:,0],x[:,3]/x[:,0],'r',label='Reverse Crescent')
+ax1.plot(xx[:,2]/xx[:,0],xx[:,3]/xx[:,0],'b',label='Crescent')
+ax1.plot(y[:,2]/y[:,0],y[:,3]/y[:,0],'g',label='Disk')
+ax1.plot(z[:,2]/z[:,0],z[:,3]/z[:,0],'m',label='Gaussian Disk')
 legend(loc=2)
-ax1.set_xlim(30.0, 60.0)
-ax1.set_ylim(0.0, 2.0)
+ax1.set_xlim(-1.0, 1.0)
[…]

### 2014-02-17 paper1/intro.tex
t

@@
 \section{introduction}
-The neighbourhood of the galactic centres are mysterious places.
-There is very little information available to accurately map the geometric structure close to the horizon of the supermassive black holes at the galactic centre.
-Observing these places with great spatial resolution is one obvious solution, however, requires high quality data.
-The future generation telescope, like mm-VLBI, EHT etc, are capable of resolving these structures.
-Nevertheless, there is an alternative solution for this problem, indirect mapping of the geometry.
+AMore than half a century after their discovery, quasars still maintain the interest of a productive part of the astrophysical community. Although several mechanisms have been proposed 
+to explain how such systems function a single mechanism is widely accepted at the present time.
+The model was first proposed by Salpeter in 1964 (citation), folowed by Zel'dovich and Novikov in the same year (citation) and Lynden-Bell five years later.
+The present name of the model is the Super-Massive Black Hole (SMBH) paradigm.\\
 
-Microlensing provides us with an opportunity to detect the signals, comprehensively detectable, and model the signal as the function of the source shape of the source, its internal brightness and the lensing mass distribution.
-Now, if the source is a quasar, the radiations are mainly dominated by the accretion disc close to the supermassive black hole horizon.
+According to the paradigm, the object is powered through the accretion of matter from the proximal environment into the black hole.
+The radiation emitted excites the surrounding medium which becomes detectable as the narrow line regions,
+broad band regions and torus.
+Moreover, in the direction perpendicular to the accretion disk, where the medium is more transparent, jets will appear.
+When the medium between and observer and such an object is transparent due to the observation direction
+a quasar can be observed (citation). In other words an active galactic nucleus (AGN) with one of the poles orientated towards Earth would be detected as a quasar by an observer. \\
 
-The motivation of this work is to demonstrate if one assumes a parametric form for the shape (surface brightness 2D) of the accretion disk of the supermassive black hole, is it possible to recover the 2D brightness profile by looking at the brightness of the source over a range of time.
+Through extensive observations done in the past 50 years, an important part of the assumptions and details  regarding the mechanism have been coomfirmed and uncovered.
+Still, the black hole shadow and the surrounding luminous acrretting material (the black hole silouette) remains to be probed.
+Advancements in the very long baseline interferometer observations have made possible since 2007 the detection of structures of a few Schwarzschild radii at the SgrA* object at the center of our galaxy (Doeleman 2008).
+Furthermore, jet launching structure near the supermasve black hole in M87 were resolved with the same observation tool (Doeleman 2012 Jet launching).
+With all the advancement, the current number of baselines does not allow the direct creation of images from the observations still it does allow the fitting of the date to predefined models(Kamruddin 2013).
+A whole range of models have been applied starting from simple geometric models to more complex, physically sensible models (Doeleman et al 2008, Fish et al 2011, Broderick et al 2011, Moscibrodzka et al 2009, Dexter 2010 and others).
+A significant fraction of the later models predict a crescent shaped silouette of the black hole.
+This motivated Kamruddin and Dexter (2013) to use a simple geometric crescent model to fit the data.
+The succesfull results encouraged the authors to speculate that the shadow of the black hole will be observed in the near fututre with the previously mentioned observation method.
+\\
+
+Hundreds of thousands of quasars have been detected up to date, mainly from the Sloan Digital Sky Survey.
+The great majority of them present a redshift larger than 2.15 of their spectra (Paris et al 20Nov 2013 SDSS catalogue).
+Therefore the typical distance to the objects is in the order of gigaparsecs or above.
+A value that is orders of magnitude
+larger than the distance to M87, estimated to be 50 Mpc? (citation). Direct observation of the black hole silouette of quasars require significant technological advances. We argue in the present paper
+ that one can probe and study the black hole shadow and its proximal environment belonging to a quasar without having to resolve sub-event horizon-scales.
+We consider that this is possible by analyzing the data from microlensing-events.
+Moreover, unlike with the few nearby AGNs, the study of the very large quasar population allows for the statistical refinement of the global properties of AGN by highlighting any particular
+characteristic that is present only in SGA* and M87 central region.
+Furthermore, by probing the event-horizon-scale inner regions of of a large number of objects will facilitate the study of the evolution of the previously menttioned region.
+
+Gravitational microlensing of QSOs have been predicted as early as 1981 when Gott suggested (Gott 1981) that if haloes of galaxies were composed of stars with masses less than `` 0.1 $M_{sol}$, then these stars acting as individual lenses whould produce fluctuations of the order of
+unity in the intensities of the QSO images on time scales of 1 -14 years '' (Gott 1981).
+What creates the magnification event is the relative and independent motion of the lens and source with respect to the observer.
+Since the respective lens is surrounded by similar objects one would expect to observe during a long enough time interval
+multiple microlensing events of the same source, the QSO.
+The sequential alignment of the source and observer with different lensing objects will result in an aparently random increase of the QSO light flux analogous to the scintilation events in a meter of radioactivity.
+
+Estimates of the Einstein radius for the particular case of a QSO with redshift 2 lensed by a compact stellar mass object with redshift 0.5 was found to be $10^-6 \sqrt{M/M_{sol}} arcsec$ (citation Wambsganss Quasar microlensing 527).
+The Schwarzschil radius of the Sgr A* black hole, situated at 8 kpc from Earth [Reid M.J.
+1993 The distance to the center of the galaxy] spans an angle of 10 microarcsec on the sky [Doeleman 2008].
+If the quasars observed at z =2 would have similar linear sizes, the aparent size of the  luminous part of the QSO would be many orders of magnitude below the 1 microarcsec level.
+Therefore the aparent size would be significantly smaller than the corresponding Einstein radius of a solar mass lens.
+The magnitude differences between the Einstein angle and the aparrent size of the bright part of the QSO motivates the introduction in the first part of the paper of simplified models for     
+calculating the lightcurves of QSO magnification events.
+The simplified models are introduced in the theory section of the present paper.
+In the theory (II) section of the present article with an introduction to the basic microlensing theory in which we present the general microlensing equation followed by the concept of magnification.
+We underline that for the particular case of micro gravitational lensing
+the images are not resolved and thus the magnifications of each images contribute to a total magnification.
+Furthermore, the concept of critical curves and caustics are introduced.
+Afterwards we focus on a particular singularity, namely the fold singularity.
+We choose to study events related to the previously mentioned singularity since the magnification events caused by the crossing of such boundaries are most probable.
+A quick inspection of figure 1 by the reader should motivatethe previous statement.
+In continuation we introduce the commonly used result involving the magnification of a point-like source near a fold singularity.
+Namely that the magnification is aproximately linear outside the caustic and decreases from infinity with the square root of the distance to the boundary when inside the caustic.
+In the last section of the Introduction (II) we present the general equations for estimating the flux received from an extended source near a fold.
+The only elements required to compute the flux as a function of time are the two dimensional brightness function of the object and the magnification of a point-like source.
+One of the core
+assumptions made when developing the simple models is that the einstein angle is orders of magnitude greater than the apparent size of the source.
+Due to this assumption one can imagine the caustic as an infinite wall to be crossed by the source as presented in figure 2.
+The consequence of the limit is that the information regarding the profiles of the source along any direction parallel to the boundary value is reduced to the corresponding integral value.
+The effect allows the introduction of a one dimensional flux function, which is the two dimensional flux integrated along the axis paralel to the boundary.
+The one dimensional profile contains exhaustively all the information that can be extracted  from the lens generated light-curve \\
+
+In the third section we introduce the three source profiles used in our models and simulations.
+We start with two commonly used and simple sources.
+Namely the gaussian brightness distribution and constant intensity disk.
+In addition we use a geometric crescent-shaped source effectively the same as the one introduced by (geometric crescent paper) with constant surface brightness.
+\\
+The fourth section contains the results of our simplified models calculations, including the numerical equation for lightcurve calculations and the corresponding plots.
+In order to facilitate comparison bewteen the lightcurves of the three distinct sources, we constrain the luminosity and half-light radius of all sources to have the same value.
+
+We continue to the second part of the paper, where we have advanced the level of complexity of our models.
+Using the Wambsganss gravitational lensing code (citation) a realistic magnification map is generated.
+The magnification map replaces the simplified magnification function used previously.
+\\
+The numerical method, hierachical tree structure and backwards reatracing, underlying the microlensing code is treated in section five.
+Aside from a basic review of the physical and numerical principles of the code, the description of its generalization from disk shaped source images to crescent ones can be found here as well.
+Further, the mass distribution in the lens plane, chosen to produce the magnification map used in this analysis, is introduced and motivated.

### 2014-03-20 paper1/theory.tex
whatever
Partially included the half-light radii equations
minor change
t

@@
 
+For a gaussian distributed surface brightness source the half-light radius is directly proportional to the parameter $\sigma$ according to the equation:
+\begin{equation}
+r_{1/2} = \sqrt{ln(4)} \sigma.
+\end{equation}
+
 \subsection{Disk source with constant surface brightness distribution}
@@
 \end{equation}
+
+The half-light radius of a uniform disk source is $R/\sqrt{2}$.
 
@@
 
+The half-light radius of any source is invariant to any rotational transformation.
+In the present case of a crescent source the effective radius is dependent on the parameters $R_p$, $R_n$ and $\sqrt{a^2+b^2}$ exclusively.
+From symmetry considerations the centroid of the source is colinear with the centers of the two disks and it is situated at a distance $d_c$ from the center of the brigth disk.
+$d_c$ can be computed numerically by solving the equation:
+
+\begin{equation}
+\frac{S_0^C}{2} =  
+\end{equation}
+
 \section{Lightcurves of the extended sources during fold crossing}
@@
 \begin{figure}
-\includegraphics[bb = 0 0 600 400]{lightcurve_crescent_1.pdf}
+\includegraphics[width = .8\textwidth]{lightcurve_gaussian_1.png}
 \caption{Lightcurve 1.}
@@
 \begin{figure}
-\includegraphics[bb = 0 0 600 400]{lightcurve_gaussian_1.pdf}
+\includegraphics[width = .8\textwidth]{lightcurve_disc_1.png}
 \caption{Lightcurve 2.}
@@
 \begin{figure}
-\includegraphics[bb = 0 0 600 400]{lightcurve_disc_1.pdf}
+\includegraphics[width = .8\textwidth]{lightcurve_crescent_1.png}
 \caption{Lightcurve 3.}

### 2014-03-21 paper1/chains/circles.py
circles

@@ -1,0 +1,37 @@
+from numpy import linspace, pi, cos, sin, arccos
+from scipy.integrate import odeint
+from pylab import figure, show
+
+def theta_lohi(Rn,c,d,r):
+    ctheta = (r*r + d*d - 1)/(2*d*r)
+    if ctheta > 1:
+	hi = 0
+    elif ctheta < -1:
+        hi = pi
+    else:
+	hi = arccos(ctheta)
+    d = (c*c + d*d)**.5
+    ctheta = (r*r + d*d - Rn*Rn)/(2*d*r)
+    if ctheta > 1:
+	lo = 0
+    elif ctheta < -1:
+        lo = pi
+    else:
+	lo = arccos(ctheta)
+    return (lo,hi)
+
+panel = figure().add_subplot(1,1,1)
+panel.set_aspect('equal')
+
+Rn = 0.4
+c = 0.4
+d = 0.2
+for r in linspace(.02,1+d,50):
+    lo,hi = theta_lohi(Rn,c,d,r)
+    u = linspace(lo,hi,20)
+    x,y = r*cos(u)-d,r*sin(u)
+    panel.plot(x,y)
+
+show()
+
+

### 2014-03-21 paper1/intro.tex

@@
 \section{introduction}
-AMore than half a century after their discovery, quasars still maintain the interest of a productive part of the astrophysical community. Although several mechanisms have been proposed 
+More than half a century after their discovery, quasars still maintain the interest of a productive part of the astrophysical community. Although several mechanisms have been proposed 
 to explain how such systems function a single mechanism is widely accepted at the present time.

### 2014-03-22 paper1/chains/circles.py
circles.py

@@ -3,14 +3,17 @@ from scipy.integrate import odeint
 from pylab import figure, show
 
 def theta_lohi(Rn,c,d,r):
-    ctheta = (r*r + d*d - 1)/(2*d*r)
-    if ctheta > 1:
-	hi = 0
-    elif ctheta < -1:
+    if abs(d) < 1e-10:
         hi = pi
     else:
-	hi = arccos(ctheta)
-    d = (c*c + d*d)**.5
+        ctheta = (r*r + d*d - 1)/(2*d*r)
+        if ctheta > 1:
+            hi = 0
+        elif ctheta < -1:
+            hi = pi
+        else:
+            hi = arccos(ctheta)
+    d += c
     ctheta = (r*r + d*d - Rn*Rn)/(2*d*r)
     if ctheta > 1:
 	lo = 0
@@ -20,18 +23,68 @@ def theta_lohi(Rn,c,d,r):
 	lo = arccos(ctheta)
     return (lo,hi)
 
+def Rhalf(Rp,Rn,a,b):
+    Rn /= Rp
+    c = (a*a + b*b)**0.5 / Rp
+    r = linspace(0.001,1,100)
+
+    def numer(f,r):
+        lo,hi = theta_lohi(Rn,c,0,r)
+        return r*r * (sin(lo)-sin(hi))
+
+    def denom(f,r):
+        lo,hi = theta_lohi(Rn,c,0,r)
+        return r * (hi - lo)
+
+    def arc(f,r):
+        lo,hi = theta_lohi(Rn,c,d,r)
+        return r * (hi - lo)
+
+    up = odeint(numer,[0],r)
+    dn = odeint(denom,[0],r)
+
+    d = up[-1,0]/dn[-1,0]
+    area = odeint(arc,[0],r)
+
+    for i in range(1,len(r)):
+        if area[i-1] < 0.5*area[-1] and area[i] > 0.5*area[-1]:
+            return Rp * r[i]
+
+
 panel = figure().add_subplot(1,1,1)
 panel.set_aspect('equal')
 
-Rn = 0.4
-c = 0.4
-d = 0.2
+Rn = 0.8
+c = 0.2
+d = 0.038
 for r in linspace(.02,1+d,50):
     lo,hi = theta_lohi(Rn,c,d,r)
     u = linspace(lo,hi,20)
     x,y = r*cos(u)-d,r*sin(u)
     panel.plot(x,y)
 
-show()
+# show()
 
[…]

### 2014-03-22 paper1/chains/circles.py
Rhalf

@@ -72,7 +72,7 @@ def cfile(fname):
     fname = fname[0]+'2.'+fname[1]
     print fname
     ouf = open(fname,'w')
-    for l in range(10):
+    while True:
         lyne = inf.readline()
         if not lyne:
             break
@@ -82,9 +82,21 @@ def cfile(fname):
                         float(lyne[4]),float(lyne[5])
             Rh = Rhalf(Rp,Rn,a,b)
         if flag == 'G':
+            Rh = 0.39*float(lyne[2])
+        if flag == 'D':
+            Rh = 2**-.5 * float(lyne[2])
         lyne[2] = '   %7.4f' % Rh
         lyne = '  '.join(lyne)
         ouf.write(lyne+'\n')
     ouf.close()
 
 cfile('CC.chain')
+cfile('CD.chain')
+cfile('DC.chain')
+cfile('DG.chain')
+cfile('GD.chain')
+cfile('CG.chain')
+cfile('DD.chain')
+cfile('GC.chain')
+cfile('GG.chain')
+

### 2014-03-25 paper1/chains/analysismcmc.py
Updated plots and makefile

@@ -10,27 +10,27 @@ def plottinghist(chain,specs):
 	print
 	print mean(data),std(data),chi2,'for',chain
 	density = gaussian_kde(data)
-	xs = linspace(25,75,1000)
+	xs = linspace(10,65,1000)
 	density.covariance_factor = lambda : .25
 	density._compute_covariance()
 	plot(xs,density(xs),specs,label=chain+'   %1.2f'%chi2,linewidth=2.0)
 
 
 print '# mean, st. dev., best $\chi^2$'
-plottinghist('cc.chain','r')
-plottinghist('dc.chain','--r')
-plottinghist('gc.chain','.-.r')
-plottinghist('cd.chain','b')
-plottinghist('dd.chain','--b')
-plottinghist('gd.chain','.-.b')
-plottinghist('cg.chain','g')
-plottinghist('dg.chain','--g')
-plottinghist('gg.chain','.-.g')
+plottinghist('cc2.chain','r')
+plottinghist('dc2.chain','--r')
+plottinghist('gc2.chain','.-.r')
+plottinghist('cd2.chain','b')
+plottinghist('dd2.chain','--b')
+plottinghist('gd2.chain','.-.b')
+plottinghist('cg2.chain','g')
+plottinghist('dg2.chain','--g')
+plottinghist('gg2.chain','.-.g')
 
 pylab.legend(loc='upper center')
-pylab.xlim(25,75)
-pylab.xlabel('$\mathtt{R_p}$',fontsize=24)
+pylab.xlim(10,65)
+pylab.xlabel('$\mathtt{R_{1/2}}$',fontsize=24)
 
 
-pylab.savefig('Rp4all.eps')
-#pylab.show()
+pylab.savefig('Rhalf4all.eps')
+pylab.show()

### 2014-03-25 paper1/chains/plotcontours.py

@@ -1,0 +1,115 @@
+import numpy
+import pylab
+
+
+def plotcontours(array1,array2,CL,spec,opaque,description):
+	import numpy as np
+
+	import matplotlib.pyplot as plt
+	from matplotlib.patches import Ellipse
+
+	def plot_point_cov(points, nstd=2, ax=None, **kwargs):
+	    """
+	    Plots an `nstd` sigma ellipse based on the mean and covariance of a point
+	    "cloud" (points, an Nx2 array).
+
+	    Parameters
+	    ----------
+	        points : An Nx2 array of the data points.
+	        nstd : The radius of the ellipse in numbers of standard deviations.
+	            Defaults to 2 standard deviations.
+	        ax : The axis that the ellipse will be plotted on. Defaults to the 
+	            current axis.
+	        Additional keyword arguments are pass on to the ellipse patch.
+
+	    Returns
+	    -------
+	        A matplotlib ellipse artist
+	    """
+	    pos = points.mean(axis=0)
+	    cov = np.cov(points, rowvar=False)
+	    return plot_cov_ellipse(cov, pos, nstd, ax, **kwargs)
+
+	def plot_cov_ellipse(cov, pos, nstd=2, ax=None, **kwargs):
+	    """
+	    Plots an `nstd` sigma error ellipse based on the specified covariance
+	    matrix (`cov`). Additional keyword arguments are passed on to the 
+	    ellipse patch artist.
+
+	    Parameters
+	    ----------
+	        cov : The 2x2 covariance matrix to base the ellipse on
+	        pos : The location of the center of the ellipse. Expects a 2-element
+	            sequence of [x0, y0].
+	        nstd : The radius of the ellipse in numbers of standard deviations.
+	            Defaults to 2 standard deviations.
+	        ax : The axis that the ellipse will be plotted on. Defaults to the 
+	            current axis.
+	        Additional keyword arguments are pass on to the ellipse patch.
+
+	    Returns
+	    -------
+	        A matplotlib ellipse artist
+	    """
+	    def eigsorted(cov):
+	        vals, vecs = np.linalg.eigh(cov)
[…]

### 2014-03-27 paper1/ms.tex
code.tex added
Edits from Mihai
whatever
re-arranged by Irshad
minor

@@
-\documentclass[usenatbib,twocolumn]{mn2e}
+\documentclass[usenatbib,onecolumn]{mn2e}
 
@@
 \author[I. Mohammed et al.] {Irshad Mohammed\thanks{irshad@physik.uzh.ch}$^1$,
-Mihai Tomozeiu$^1$, Manuel Rabold$^1$,  Prasenjit Saha,$^1$ and Joachim Wambsgans$^3$\\
+Mihai Tomozeiu$^1$, Manuel Rabold$^1$,  Prasenjit Saha,$^1$ and Joachim Wambsgans$^2$\\
 $^1${Institute for Theoretical Physics, University of Z$\ddot{u}$rich, 8057 Z$\ddot{u}$rich Switzerland}\\
@@
 
-\input{sourcemodels}
+\input{code}
 

### 2014-03-28 paper1/discussion.tex
first edits

\section{Discussion}

An interesting case for future study would be crescent model with Gaussian brightness profile.

### 2014-03-28 paper1/fitting.tex
Updated plots and makefile
minor
first edits

--- 2014-01-08 paper1/fitting.tex
@@
-\section{model fitting}
+\section{Model fitting and likelihood analysis}
 
 \begin{figure}
-  \includegraphics[width=0.95\hsize]{plots/data.eps}
-\caption{data}
+\centering
+  \includegraphics[width=0.95\hsize]{plots/data_with_error.eps}
+\caption{\label{fig:datafitting} data}
 \end{figure}
@@
 \begin{figure}
-  \includegraphics[width=0.5\hsize]{plots/Crescent.eps}
-  \includegraphics[width=0.5\hsize]{plots/CrescentReverse.eps}\\
-  \includegraphics[width=0.5\hsize]{plots/Disk.eps}
-  \includegraphics[width=0.5\hsize]{plots/Gauss.eps}
-\caption{Model Fitting}
+\centering
+  \includegraphics[width=0.45\hsize]{plots/Rp4all.eps}
+  \includegraphics[width=0.45\hsize]{plots/Rhalf4all.eps}\\
+  \includegraphics[width=0.45\hsize]{plots/Rhalf_RnRp.eps}
+  \includegraphics[width=0.45\hsize]{plots/aRp_bRp.eps}
+
+\caption{\label{fig:mcmc} Model Fitting}
 \end{figure}
+
+In this section, we attempt to fit and recover the model parameters from the given dataset generated by different source models - crescent, uniform-disk and Gaussian-disk.
+So broadly we have 9 cases to discuss.
+We carried out a likelihood analysis on all three parameter set using Markov-chain Monte-Carlo (MCMC).
+
+\subsection{Mock datasets}
+The datasets for each source model is in the form of light curves, the brightness of the source observed at different time scale.
+Figure \ref{fig:datafitting} shows the three datasets.
+These datasets are generated using the magnification map shown in figure \ref{fig:magnification} between point B and C (direction from C to B).
+The data assumed was very  ideal, including 200 different data points at regular interval of time.
+It is worth noticing that the datasets generated using crescent model and uniform disc model resembles much more than the one generated by Gaussian disc model source.
+This is also very intuitive as uniform disc is a special case in crescent model where $R_n=0$, The fiducial model assumed to generate these mock datasets are:
+\begin{equation}
+	[R_p, R_n, a, b] = [50.0, 30.0, 15.0, 10.0]  \ \ \ \rm{for\ Crescent\ model},
+	\label{eqn:cp}
+\end{equation}
+\begin{equation}
+	R = 50.0 \ \ \ \rm{for\ Uniform\ disc\ model}
+\end{equation}
+
+\begin{equation}
+	R = 50.0 \ \ \ \rm{for\ Gaussian\ disk\ model}
+\end{equation}
+\\
+where, $R$ being the size of the uniform disc and Gaussian disc source model.
+
+The aim of this exercise was to explore the degeneracies in  various model parameters and recover the correct fiducial model assumed.
+Also, it can be inferred if one can simply distinguish between circularly symmetric and asymmetric sources or uniform and Gaussian sources.
+The errorbars on each data point, in all three datasets, is about 10 percent its value.
+Also, because we are fitting these datasets with different caustic crossing (point A to B in figure \ref{fig:magnification}), it adds a natural systematic errors on each data point.
+We also assume the errors are Gaussian distributed.
+
+\subsection{Fitting with crescent model}
+First we explored the parameter space for crescent source model and run its likelihood exploration for three different datasets.
+Here we have four parameters as given by equation \ref{eqn:cp}.
+Three red curves in top row of figure \ref{fig:mcmc} represents the resulting likelihood of $R_p$, the overall size of the crescent source for three different datasets.
+Solid red curve in top row and red contours in bottom row of figure \ref{fig:mcmc} represents the crescent fitting with crescent data (cc.chain in figure \ref{fig:mcmc}).
+All four parameters can be well recovered in this case and the overall likelihood coincides well with the fiducial values assumed and stated in equation \ref{eqn:cp}.
+It can also be well noticed that the half light radius is degenerate with $R_n/R_p$ or in other words, $R_p$ and $R_n$ are correlated.
+
+In second case, we tried to fit crescent model to disk data (dc.chain in figure \ref{fig:mcmc}).
+The dashed red line in top row and blue contours in bottom row represents this case.
+Here one can recover the overall shape of the source, slightly lower than what is recovered in cc.chain case.
+Infact, the likelihood is higher than the previous case as the data is generated by one parameter only and fitting with four parameter that can also model the noise.
+Also, the maximum likelihood is achieved close to $R_n = 0$, as already mentioned, this is intuitive as well.
+The other paramters, $a$ and $b$ are quite arbitary.
+
+In the third case, Gaussian disc data is fit with crescent model (gc.chain).
+The dot-dashed red line in top row and green contours in bottom row represents this case.
+In this case, the extent of $R_p$ is broad and the likelihood is smaller than other cases.
+
+Please notice that in top row of figure \ref{fig:mcmc}, the height of the peaks does not represents in likelihood, it is just ensuring that the area under the curve is unity.
+The likelihood can be estimated from the number in the legend which is the best $\chi^2$.
+
+\subsection{Fitting with uniform disc model}
+In this case, we have only one parameter to fit, the outer radius of the uniform disc with uniform intensity.
+When fitting with crescent data (solid blue line), the maximum likelihood shifts towards higher values of $R$ as compared to the fiducial value (which was 50).
+Also the maximum likelihood and is very low and the contraints are also very weak.
+However, when fitting with Gaussian disc (dot-dashed blue line), the likelihood is still very low, but the parameter is tightly constrained.
+The third case is ideal when fitting uniform disc data (blue dashed line), where the constraints are ideal and also co-incide well with the fiducial value assumed and also likelihood is comparatively better.
+
+\subsection{Fitting with Gaussian disc model}
+In the last case, we fit Gaussian disc model to the three datasets (green curves in top row of figure \ref{fig:mcmc}).
+In this case, when fitted with Gaussian disc data, the likelihood coincide with the fiducial value assumed, in other cases the likelihood is away but weak and very well constrained.
+
+\section{Conclusion and interpretation}
+The main results of this work is given in figure \ref{fig:mcmc} and described in the previous section. Here we try to interpret those results and conclude the take away message.
+
+Lets assume that an ideal telescope observes a quasar for a long time and we get the lightcurve with good precision and one can clearly see the event of caustic crossing in the light curve.
+Now one wants extract the true shape of the source.
+To comment more specifically, lets assume three cases when the true shape of the source is: crescent, uniform disc and Gaussian disc:
+
+\begin{enumerate}
+
+\item Crescent shape source: if one tries to fit this curve with crescent model, the maximum likelihood will be good and one can recover the true parameters with good precision.
+In the other two cases, the maximum likelihood will be very bad.
+Hence this shape can be recovered very well by trusting the maximum likelihood analysis.
+\item Uniform disc shape source: if one tries to fit this curve with Gaussian disk model, the likelihood will be very little to disprove it, however, the other two models are inditinguishable with nearly the same likelihood.
+At this point it does not matter as one can obtain the true result from both disk and crescent model.
+\item Gaussian disc shape source: if one tries to fit this curve with uniform disk model, again likelihood will discard this choice.
+Again the maximum likelihood analysis will point towards the best fitted true model.
+
+\end{enumerate}
+
+At this point, we conclude that maximum likelihood analysis can be well trusted in order to recover true shape and recovering the corresponding parameters.

### 2014-03-28 paper1/theory.tex
Updated plots and makefile
first edits

--- 2013-12-02 paper1/theory.tex
@@
 present paper in particular. For a detailed presentation of the theory the reader is referred to the books and articles that represent the references
-of the current section (book by Schneider and others).\\
-
-The general microlensing equation can be written as:
+of the current section (book by Schneider and others). The general microlensing equation can be written as:
 \begin{equation}
@@
 \vec{x} - \sum_{i=1}^{n} m_i \frac{\vec{x} - \vec{x_i}}{\left( \vec{x} - \vec{x_i} \right)^2}.
+\label{eqn:microlensingequation}
 \end{equation}
 
-The equation is valid for a set of $\it{n}$ point (Schwarzschild) lenses placed at coordinates $\vec{x_i}$. $\vec{y}$ represents the coordinates in the source plane for the respective ray, while 
-$\vec{x}$ represents the coordinates in the lens plane. Furthermore, $\gamma$ and $k$ are the shear and surface mass density.\\
-
-For the lens mapping a Jacobian can be defined:
+The equation is valid for a set of $\it{n}$ point (Schwarzschild) lenses placed at coordinates $\vec{x_i}$.
+$\vec{y}$ represents the coordinates in the source plane for the respective ray, while $\vec{x}$ represents the coordinates in the lens plane.
+Furthermore, $\gamma$ and $k$ are the shear and surface mass density.
+For the lens mapping, a Jacobian can be defined:
 \begin{equation} 
@@
 \mu(\vec{x}) = \left| \frac{1}{det(A(\vec{x}))} \right|
+\label{eqn:mu_x}
 \end{equation}
@@
 by critical curves where the determinant vanishes. The lens equation maps the critical curves of the lens plane into caustics in the source plane. An example of the caustic 
-shape can be seen in the magnification map of |||Figure 1|||.\\
+shape can be seen in figure \ref{fig:magnification}.\\
 
-||||||||||||||||||||
-||||||Figure 1||||||
-||||||||||||||||||||
+\begin{figure}
+\centering
+  \includegraphics[width=0.95\hsize]{plots/magnification.eps}
+\caption{\label{fig:magnification} Magnification map}
+\end{figure}
 
-Along these critical curves in the lens plane or caustics in the source plane,  the magnification $\mu$ is infinite, a result that simply follows from equation ( 
-(3). This statement holds only fo infinitesimally small sources. For extended sources the maximum amplification is finite since only an infinitesimally small area of the source
-overlaps with the caustics (Schneider Weiss 1986 paper others).
+Along these critical curves in the lens plane or caustics in the source plane,  the magnification $\mu$ is infinite, a result that simply follows from equation \ref{eqn:mu_x}.
+This statement holds only for infinitesimally small sources.
+For extended sources the maximum amplification is finite since only an infinitesimally small area of the source
+overlaps with the caustics \citep{1986BAAS...18T.907S,}.
 
-The matrix $A$ at the coordinates of the caustic can have rank 1 or rank 0. If the rank is 1 then the coordinates belong to $fold$ or $cusp$ singularities. For the purpose 
+The matrix $A$ at the coordinates of the caustic can have rank 1 or 0. If the rank is 1 then the coordinates belong to $fold$ or $cusp$ singularities. For the purpose 
 of this paper we are interested only in the behaviour around $fold$ singularities. A condition necessary to distinguish between a fold and cusp singularity is that the
-eigenvector of $A$ corresponding to the null eigenvalue is not a tangent vector of the critical curve (Dominik 2008, Schneider Weiss 1992).
+eigenvector of $A$ corresponding to the null eigenvalue is not a tangent vector of the critical curve \citep{1992A&A...260....1S,}(Dominik 2008).
 
@@
 The behaviour of the images of a point source and their corresponding magnifications near such a caustic has been thoroughly studied in the past. The reader who is interested in 
-understanding the details related to this subject is referred to papers such as Blandford and Narayan 1986, Schneider and Weiss 1986, Gaudi and Petters 2002. For the present paper we only 
+understanding the details related to this subject is referred to papers such as \cite{1986ApJ...310..568B}, \cite{1992A&A...260....1S,}, \cite{2002ApJ...574..970G}, \cite{2002ApJ...580..468G}.
+For the present paper we only
 summarize the common result:
@@
  F(t) = \int_{-\infty}^\infty \int_{-\infty}^\infty S_{2D}(p-p_s(t), q-q_s(t)) \mu_t(p) \mathrm{d}q \mathrm{d}p
+ \label{eqn:ft2d}
 \end{equation}
@@
 = \int_{-\infty}^\infty  \mu_t(p) S_{1D}\left(p-p_s(t)\right) \mathrm{d}p,
+\label{eqn:ft}
 \end{equation}
-
+\\
 where we have defined the one dimensional flux function as:
@@
 
-In the present study we are analysing three types of sources with different surface brighness: 
+In the present study we are analysing three types of sources with different surface brightness: 
 \begin{enumerate}
@@
 The first two sources are the typical choices used in the literature to describe the luminous parts of a quasar (Prasenjit should give some citations here). The third one is a newly proposed
-variant (Kammrudin 2013 Geometric crescent).
+variant \citep{2013MNRAS.434..765K}.
 
@@
 \end{equation}
-
+\\
 The corresponding 1D brightness is:
@@
 \end{equation}
-
+\\
 Other parameters of the model are the total flux $S_0^G$ and $\sigma$.
@@
 Although such a definition for a source would have non-zero surface/linear brightness for any coordinate $p,q$, the amount of light received by a detector from outside a $3 \sigma$ disk centered at $p_s, q_s$
-would be insignificant.
+would be insignificant. For a gaussian distributed surface brightness source the half-light radius is directly proportional to the parameter $\sigma$ according to the equation:
+\begin{equation}
+r_{1/2} = \sqrt{ln(4)} \sigma.
+\end{equation}
 
@@
 \end{equation}
+\\
+The half-light radius of a uniform disk source is $R/\sqrt{2}$.
 
@@
  S_{2D}^C =  S_{2D}^{Dp} -  S_{2D}^{Dn}   (n+1)
+ \label{eqn:s2d}
 \end{equation}
@@
 \end{equation}
-
-and\\
+\\
+and
 \begin{equation}
@@
 \end{equation}
-
-The following notations were used: $R_p, (p_{sp}, q_{sp}), R_n, (p_{sn},q_{sn})$ are the radii and coordinate of the center for the larger positive disk and smaller negative disk.  
-$S_0^{Dp},S_0^{Dn}$ represent the total flux of radiation received from the large and small disk. From this point forward we will not use the total flux from each source. Instead we will 
+\\
+The following notations were used: $R_p, (p_{sp}, q_{sp}), R_n, (p_{sn},q_{sn})$ are the radii and coordinate of the center for the larger positive disk and smaller negative disk respectively.
+$S_0^{Dp},S_0^{Dn}$ represent the total flux of radiation received from the large and small disk.
+From this point forward we will not use the total flux from each source.
+Instead we will
 use the difference which in this case is th total flux from the crescent-shaped source $S_0^C$. \\
-Equation(n+1) can be written as:\\
+Equation \ref{eqn:s2d} can be written as:\\
 
@@
 \end{align}
-
+\\
 Analogous for the linear brightness function:
@@
 
---------------------------------------------------------------------------
-Figure with 2d 1d brightness for gaussian, disk, some crescent
-----------------------------------------------------------------------------
-
 There are some constraints on the parameters used to define a crescent in the previously presented manner that need to be stated. First, we must impose the obvious $R_p > R_n$ relation. Secondly, 
@@
 
-For the distances between the centers of the two disks we will use the same notations as the one found in the paper [Geometric crescent], $a \equiv p_{sn} - p_{sp}$ and $b \equiv q_{sn} - q_{sp}$.
+For the distances between the centers of the two disks we will use the same notations as the one found in the paper \citep{2013MNRAS.434..765K}, $a \equiv p_{sn} - p_{sp}$ and $b \equiv q_{sn} - q_{sp}$.
 
@@
 
-Using equation (7) and the one-dimensional flux function presented in the previous section one can compute numerically the lightcurves of the three extented sources fot the simplified 
+Using equation \ref{eqn:ft} and the one-dimensional flux function presented in the previous section one can compute numerically the lightcurves of the three extented sources for the simplified 
 infinite-wall-caustics model.
@@
 \end{equation}
-The result is equivalent to the one obtained in papers such as (citations).\\
+Figure \ref{fig:lightcurve_gauss} shows the corresponding lightcurve. The result is equivalent to the one obtained in papers such as (citations).\\
 
 \begin{figure}
-\includegraphics[bb = 0 0 600 400]{lightcurve_crescent_1.pdf}
-\caption{Lightcurve 1.}
+\includegraphics[width = .8\textwidth]{plots/lightcurve_gaussian_1.eps}
+\caption{\label{fig:lightcurve_gauss} Lightcurve 1.}
 \end{figure}
@@
 \end{equation}
-As obtained in preciding papers by (citations).\\
+Figure \ref{fig:lightcurve_disk} shows the corresponding lightcurve. As obtained in preciding papers by (citations).\\
 
 \begin{figure}
-\includegraphics[bb = 0 0 600 400]{lightcurve_gaussian_1.pdf}
-\caption{Lightcurve 2.}
+\includegraphics[width = .8\textwidth]{plots/lightcurve_disc_1.eps}
+\caption{\label{fig:lightcurve_disk} Lightcurve 2.}
 \end{figure}
@@
 \end{equation}
-
+Figure \ref{fig:lightcurve_crescent} shows the corresponding lightcurve.
 In the previous equation we have renamed the coordinate of the center of the large disk $p_{sp}$ to $p_{s}$.  \\
@@
 \begin{figure}
-\includegraphics[bb = 0 0 600 400]{lightcurve_disc_1.pdf}
-\caption{Lightcurve 3.}
+\includegraphics[width = .8\textwidth]{plots/lightcurve_crescent_1.eps}
+\caption{\label{fig:lightcurve_crescent} Lightcurve 3.}
 \end{figure}

### 2014-04-06 paper1/code.tex

@@
 \section{Microlensing code and magnification maps}
-For the numerical computations of lightcurves for the models a and b, the microlensing code by Joachim Wambsgans was used, which is described in Wambsganss 1999.
+For the numerical computations of lightcurves for the models a and b, the microlensing code by Joachim Wambsgans was used, which is described in \cite{1999A&A...346L...5W}.
 This code is based on the hierachical tree technique to calculate the gravitational effect of the lensing masses.
@@
 For the computation of the individual deflection angles the tree idea is used.
-This means that the positions of all lensing masses are sorted into a 2-dim.
-grid (in the lens plane).
+This means that the positions of all lensing masses are sorted into a 2-dimensional grid (in the lens plane).
 This grid is subdivided into 4 smaller squares until every cell contains only one mass.
-For the actual calculations of the deflection angle not every mass' gravitational influence is treated equally.
+For the actual calculations of the deflection angle not every mass gravitational influence is treated equally.
 Instead it is made use of the fact that the influence of lenses on the considered lightray falls off with $r^{-1}$ from the point where the ray hits the lens plane.
@@
 Once the map is created, the lightcurve can be obtained by specifying the transition path of the source within the map.
-The code thereby convolves the function describing the shape of the source with the magnification pattern of the map for each time step (see equation (6)).
+The code thereby convolves the function describing the shape of the source with the magnification pattern of the map for each time step (see equation \ref{eqn:ft2d}).
 Note that in the most observational cases in microlensing, both lens and source are moving and further that the lens configuration, and with it the magnification pattern, is changing with time.
@@
 For the model a it was desirable to work with a caustic geometry as simple as possible.
-Therefore only two equal point like masses were used as lenses, to produce the magnification map depicted in figure (...).
+Therefore only two equal point like masses were used as lenses, to produce the magnification map depicted in figure \ref{fig:magnification}.
 For a description of the parameter set which characterizes the map see (reference).
@@
 The values of the parameters are specified in pixel units corresponding to the magnification map.
-Further one needs to specify the start and end point coordinates of the path, which the center of the source image follows through the magnification map (see the depiction in figure (...)).
+Further one needs to specify the start and end point coordinates of the path, which the center of the source image follows through the magnification map (see the depiction in figure \ref{fig:magnification}).
 Hence the points along this path are specified through the number of timesteps for which the computation of the brightness is to be carried out.
 Those points correspond to the actual measurement of the brightness of an object in the observational case.
-For each timestep the two-dimensional convolution of equation (6) is carried out numerically for the position of the source on the magnification map.
+For each timestep the two-dimensional convolution of equation \ref{eqn:ft2d} is carried out numerically for the position of the source on the magnification map.
 \\

### 2014-04-06 paper1/intro.tex

@@
 More than half a century after their discovery, quasars still maintain the interest of a productive part of the astrophysical community. Although several mechanisms have been proposed 
-to explain how such systems function a single mechanism is widely accepted at the present time.
-The model was first proposed by Salpeter in 1964 (citation), folowed by Zel'dovich and Novikov in the same year (citation) and Lynden-Bell five years later.
+to explain how such systems function, a single mechanism is widely accepted at the present time.
+The model was first proposed by \cite{1964ApJ...140..796S}, followed by \cite{1964SPhD....9..246Z} in the same year and \cite{1969Natur.223..690L} five years later.
 The present name of the model is the Super-Massive Black Hole (SMBH) paradigm.\\
@@
 
-Through extensive observations done in the past 50 years, an important part of the assumptions and details  regarding the mechanism have been coomfirmed and uncovered.
-Still, the black hole shadow and the surrounding luminous acrretting material (the black hole silouette) remains to be probed.
-Advancements in the very long baseline interferometer observations have made possible since 2007 the detection of structures of a few Schwarzschild radii at the SgrA* object at the center of our galaxy (Doeleman 2008).
-Furthermore, jet launching structure near the supermasve black hole in M87 were resolved with the same observation tool (Doeleman 2012 Jet launching).
-With all the advancement, the current number of baselines does not allow the direct creation of images from the observations still it does allow the fitting of the date to predefined models(Kamruddin 2013).
+Through extensive observations done in the past 50 years, an important part of the assumptions and details  regarding the mechanism have been confirmed and uncovered.
+Still, the black hole shadow and the surrounding luminous accreting material (the black hole silouette) remains to be probed.
+Advancements in the very long baseline interferometer observations have made possible since 2007, the detection of structures of a few Schwarzschild radii at the Sagitarius A object at the center of our galaxy \citep{2008JPhCS.131a2055D}.
+Furthermore, jet launching structure near the supermasve black hole in M87 were resolved with the same observation tool \citep{2012Sci...338..355D}.
+With all the advancement, the current number of baselines does not allow the direct creation of images from the observations still it does allow the fitting of the data to predefined models \citep{2013MNRAS.434..765K}.
+
 A whole range of models have been applied starting from simple geometric models to more complex, physically sensible models (Doeleman et al 2008, Fish et al 2011, Broderick et al 2011, Moscibrodzka et al 2009, Dexter 2010 and others).
 A significant fraction of the later models predict a crescent shaped silouette of the black hole.
-This motivated Kamruddin and Dexter (2013) to use a simple geometric crescent model to fit the data.
+This motivated \cite{2013MNRAS.434..765K} to use a simple geometric crescent model to fit the data.
 The succesfull results encouraged the authors to speculate that the shadow of the black hole will be observed in the near fututre with the previously mentioned observation method.
@@
 The great majority of them present a redshift larger than 2.15 of their spectra (Paris et al 20Nov 2013 SDSS catalogue).
-Therefore the typical distance to the objects is in the order of gigaparsecs or above.
+Therefore the typical distance to the objects is in the order of Gigaparsecs (Gpc) or above.
 A value that is orders of magnitude
 larger than the distance to M87, estimated to be 50 Mpc? (citation). Direct observation of the black hole silouette of quasars require significant technological advances. We argue in the present paper
- that one can probe and study the black hole shadow and its proximal environment belonging to a quasar without having to resolve sub-event horizon-scales.
+ that one can probe and study the black hole shadow and its proximal environment belonging to a quasar without having to resolve sub-event horizon scales.
 We consider that this is possible by analyzing the data from microlensing-events.
@@
 characteristic that is present only in SGA* and M87 central region.
-Furthermore, by probing the event-horizon-scale inner regions of of a large number of objects will facilitate the study of the evolution of the previously menttioned region.
+Furthermore, by probing the event-horizon-scale inner regions of of a large number of objects will facilitate the study of the evolution of the previously mentioned region.
 
-Gravitational microlensing of QSOs have been predicted as early as 1981 when Gott suggested (Gott 1981) that if haloes of galaxies were composed of stars with masses less than `` 0.1 $M_{sol}$, then these stars acting as individual lenses whould produce fluctuations of the order of
+Gravitational microlensing of QSOs have been predicted as early as 1981 when \cite{1981ApJ...243..140G} suggested that ''if haloes of galaxies were composed of stars with masses less than  0.1 $M_{sol}$, then these stars acting as individual lenses whould produce fluctuations of the order of
 unity in the intensities of the QSO images on time scales of 1 -14 years '' (Gott 1981).
@@
 
-Estimates of the Einstein radius for the particular case of a QSO with redshift 2 lensed by a compact stellar mass object with redshift 0.5 was found to be $10^-6 \sqrt{M/M_{sol}} arcsec$ (citation Wambsganss Quasar microlensing 527).
-The Schwarzschil radius of the Sgr A* black hole, situated at 8 kpc from Earth [Reid M.J.
-1993 The distance to the center of the galaxy] spans an angle of 10 microarcsec on the sky [Doeleman 2008].
+Estimates of the Einstein radius for the particular case of a QSO at redshift 2 lensed by a compact stellar mass object at redshift 0.5 was found to be $10^-6 \sqrt{M/M_{sol}} arcsec$ \citep{2001PASA...18..207W}.
+The Schwarzschild radius of the Sgr A* black hole, situated at 8 kpc from Earth \citep{1993ARA&A..31..345R} spans an angle of 10 microarcsec on the sky \citep{2008JPhCS.131a2055D}.
 If the quasars observed at z =2 would have similar linear sizes, the aparent size of the  luminous part of the QSO would be many orders of magnitude below the 1 microarcsec level.
@@
 In the theory (II) section of the present article with an introduction to the basic microlensing theory in which we present the general microlensing equation followed by the concept of magnification.
-We underline that for the particular case of micro gravitational lensing
+We underline that for the particular case of micro gravitational lensing,
 the images are not resolved and thus the magnifications of each images contribute to a total magnification.
@@
 We choose to study events related to the previously mentioned singularity since the magnification events caused by the crossing of such boundaries are most probable.
-A quick inspection of figure 1 by the reader should motivatethe previous statement.
+A quick inspection of figure 1 by the reader should motivate the previous statement.
 In continuation we introduce the commonly used result involving the magnification of a point-like source near a fold singularity.
@@
 Namely the gaussian brightness distribution and constant intensity disk.
-In addition we use a geometric crescent-shaped source effectively the same as the one introduced by (geometric crescent paper) with constant surface brightness.
+In addition we use a geometric crescent-shaped source effectively the same as the one introduced by \citep{2013MNRAS.434..765K} with constant surface brightness.
 \\

### 2014-04-06 paper1/theory.tex
mag map with track
added points in magnification map

@@
 \centering
-  \includegraphics[width=0.95\hsize]{plots/magnification.eps}
+  \includegraphics[width=0.95\hsize]{plots/IRIS567_path_2.eps}
 \caption{\label{fig:magnification} Magnification map}

### 2014-04-08 paper1/theory.tex
, corrections

@@
 For extended sources the maximum amplification is finite since only an infinitesimally small area of the source
-overlaps with the caustics \citep{1986BAAS...18T.907S,}.
+overlaps with the caustics \citep{1986BAAS...18T.907S}.
 
@@
 of this paper we are interested only in the behaviour around $fold$ singularities. A condition necessary to distinguish between a fold and cusp singularity is that the
-eigenvector of $A$ corresponding to the null eigenvalue is not a tangent vector of the critical curve \citep{1992A&A...260....1S,}(Dominik 2008).
+eigenvector of $A$ corresponding to the null eigenvalue is not a tangent vector of the critical curve \citep{1992A&A...260....1S}(Dominik 2008).
 
@@
 The behaviour of the images of a point source and their corresponding magnifications near such a caustic has been thoroughly studied in the past. The reader who is interested in 
-understanding the details related to this subject is referred to papers such as \cite{1986ApJ...310..568B}, \cite{1992A&A...260....1S,}, \cite{2002ApJ...574..970G}, \cite{2002ApJ...580..468G}.
+understanding the details related to this subject is referred to papers such as \cite{1986ApJ...310..568B}, \cite{1992A&A...260....1S}, \cite{2002ApJ...574..970G}, \cite{2002ApJ...580..468G}.
 For the present paper we only

### 2014-04-15 paper1/intro.tex
first correction intro.tex

@@
 Moreover, in the direction perpendicular to the accretion disk, where the medium is more transparent, jets will appear.
-When the medium between and observer and such an object is transparent due to the observation direction
-a quasar can be observed (citation). In other words an active galactic nucleus (AGN) with one of the poles orientated towards Earth would be detected as a quasar by an observer. \\
+When the medium between and observer and such an object is transparent a quasar can be observed (citation).
+In other words an active galactic nucleus (AGN) with one of the poles orientated towards Earth would be detected as a quasar by an observer.
+\\
 
@@
 Furthermore, jet launching structure near the supermasve black hole in M87 were resolved with the same observation tool \citep{2012Sci...338..355D}.
-With all the advancement, the current number of baselines does not allow the direct creation of images from the observations still it does allow the fitting of the data to predefined models \citep{2013MNRAS.434..765K}.
+With all the advancement, the current number of baselines does not allow the direct generation of images from the observations still it does allow the fitting of the data to predefined models \citep{2013MNRAS.434..765K}.
 
@@
 The great majority of them present a redshift larger than 2.15 of their spectra (Paris et al 20Nov 2013 SDSS catalogue).
-Therefore the typical distance to the objects is in the order of Gigaparsecs (Gpc) or above.
-A value that is orders of magnitude
-larger than the distance to M87, estimated to be 50 Mpc? (citation). Direct observation of the black hole silouette of quasars require significant technological advances. We argue in the present paper
+Therefore the typical distance to the objects is in the order of gigaparsecs  or above.
+A value that is orders of magnitude larger than the distance to M87.
+Direct observation of the black hole silouette of quasars require significant technological advances.
+We argue in the present paper
  that one can probe and study the black hole shadow and its proximal environment belonging to a quasar without having to resolve sub-event horizon scales.
@@
 characteristic that is present only in SGA* and M87 central region.
-Furthermore, by probing the event-horizon-scale inner regions of of a large number of objects will facilitate the study of the evolution of the previously mentioned region.
+Furthermore, by probing the event-horizon-scale inner regions of a large number of objects will facilitate the study of the evolution of the previously mentioned region.
 
@@
 We start with two commonly used and simple sources.
-Namely the gaussian brightness distribution and constant intensity disk.
+Namely the gaussian brightness distribution and the constant intensity disk.
 In addition we use a geometric crescent-shaped source effectively the same as the one introduced by \citep{2013MNRAS.434..765K} with constant surface brightness.

### 2014-04-15 paper1/theory.tex
first correction theory.tex

@@
 
-||||||||||||||||||||
-||||||Figure 2||||||
-||||||||||||||||||||
-
 The behaviour of the images of a point source and their corresponding magnifications near such a caustic has been thoroughly studied in the past. The reader who is interested in 
@@
 the square root of the distance from the fold.
-The later term becomes activated only after the source enters the region interior to the caustic curves when the stepfunction $\Theta(p)$ become unity.
+The later term becomes activated only after the source enters the region interior to the caustic curves when the values of the stepfunction $\Theta(p)$ become unity.
 The proportionality
@@
 In order to build the previous equation we have considered that the time dependency of the flux $F$ is given only by the motion of the source with respect to a fixed caustic.
-Therefore the only time dependent quatities in the right hand side of the equation
+Therefore the only time dependent quantities in the right hand side of the equation
 are the coordinates of the center of the source $p_s,q_s$ and by construct $S_{2D}$.
@@
 This representation is a valid approximation only when the apparent size of the source is much smaller than the corresponding Einstein angle of the lens. In this context 
-all the information about the source shape and brightness will that can be contained in the lightcurve is exhaustively given by the 1D flux function.
-In other words, if two sources with $S_{2D}$ have the same $S_{1D}$ they cannot be distinguished by studying their lightcurves.
+all the information about the source shape and brightness that can be contained in the lightcurve is exhaustively given by the 1D flux function.
+In other words, if two sources with different $S_{2D}$ have the same $S_{1D}$ they cannot be distinguished by studying their lightcurves.
 
@@
 \end{enumerate}
-The first two sources are the typical choices used in the literature to describe the luminous parts of a quasar (Prasenjit should give some citations here). The third one is a newly proposed
+The first two sources are the typical choices used in the literature to describe the luminous parts of a quasar (Prasenjit should give some citations here). The third one is a recently proposed
 variant \citep{2013MNRAS.434..765K}.
@@
 \end{equation}
-Figure \ref{fig:lightcurve_gauss} shows the corresponding lightcurve. The result is equivalent to the one obtained in papers such as (citations).\\
 
@@
 \end{equation}
-Figure \ref{fig:lightcurve_disk} shows the corresponding lightcurve. As obtained in preciding papers by (citations).\\
 
@@
 \end{equation}
-Figure \ref{fig:lightcurve_crescent} shows the corresponding lightcurve.
-In the previous equation we have renamed the coordinate of the center of the large disk $p_{sp}$ to $p_{s}$.  \\
 

### 2014-04-15 paper1/ms.tex
first edits
all figures added

@@
 
-\author[I. Mohammed et al.] {Irshad Mohammed\thanks{irshad@physik.uzh.ch}$^1$,
+\author[TBD] {TBD\thanks{}$^1$,
 Mihai Tomozeiu$^1$, Manuel Rabold$^1$,  Prasenjit Saha,$^1$ and Joachim Wambsgans$^2$\\
@@
 \begin{abstract}
+
 We present a toy model to parametrize and constraint the innermost environment of a distant quasar using microlensing. 
@@
 
-\citep{2006MNRAS.367.1209L}
-
 \bibliographystyle{mn2e}
@@
 \def\aap{A\&A}
-
+\def\nat{nature}
+\def\araa{ARAA}
+\def\pasa{PASA}
 \bibliography{heap}

### 2014-04-15 paper1/theory.tex
simple captions added

@@
 
+\begin{figure}
+\includegraphics[width = .8\textwidth]{plots/figure_rhalf.eps}
+\caption{\label{fig:lightcurve_crescent} Geometry of crescent sources.}
+\end{figure}
+
 \section{Lightcurves of the extended sources during fold crossing}
@@
 \begin{figure}
-\includegraphics[width = .8\textwidth]{plots/lightcurve_gaussian_1.eps}
-\caption{\label{fig:lightcurve_gauss} Lightcurve 1.}
+\includegraphics[width = .8\textwidth]{plots/4source_magnification.eps}
+\caption{\label{fig:lightcurve_gauss} Lightcurves of crescents, gaussian and disk for sources with identical $S_0$ and $r_{1/2}$.
+Crescent source has $r_{1/2}$ = 0.72754 $R_p$, $R_n$ = 0.4 $R_p$, a = 0.3/-0.3 $R_p$ }
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width = .8\textwidth]{plots/lightcurve_disc_1.eps}
-\caption{\label{fig:lightcurve_disk} Lightcurve 2.}
+\includegraphics[width = .8\textwidth]{plots/4avar_magnification.eps}
+\caption{\label{fig:lightcurve_disk} Lightcurves of crescent sources with $R_n$ = 0.4 $R_p$.}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width = .8\textwidth]{plots/lightcurve_crescent_1.eps}
-\caption{\label{fig:lightcurve_crescent} Lightcurve 3.}
+\includegraphics[width = .8\textwidth]{plots/4rn_magnification.eps}
+\caption{\label{fig:lightcurve_crescent} Lightcurves of backwards crescents with  a = 0.3 $R_p$.}
 \end{figure}
@@
 chosing the function $a(t)$ to be constant in time.
+
+\begin{figure}
+\includegraphics[width = .8\textwidth]{plots/4rn2_magnification.eps}
+\caption{\label{fig:lightcurve_crescent} Lightcurve of forwards crescents with a = -0.3 $R_p$.}
+\end{figure}

### 2014-04-15 paper1/ms.tex

@@
 
-\title{Probing internal structure of distant quasars using Microlensing}
+\title{Microlensing as a possible probe of the internal structure of quasars}
 
-\author[TBD] {TBD\thanks{}$^1$,
-Mihai Tomozeiu$^1$, Manuel Rabold$^1$,  Prasenjit Saha,$^1$ and Joachim Wambsgans$^2$\\
+\author[I. Mohammed et al.] {Irshad Mohammed\thanks{irshad@physik.uzh.ch}$^1$,
+Mihai Tomozeiu$^1$, Manuel Rabold$^1$,
+\newauthor
+Prasenjit Saha,$^1$ and Joachim Wambsgans$^2$\\
 $^1${Institute for Theoretical Physics, University of Z$\ddot{u}$rich, 8057 Z$\ddot{u}$rich Switzerland}\\
@@
 
-We present a toy model to parametrize and constraint the innermost environment of a distant quasar using microlensing. 
+In quasars that have been lensed by galaxies, the point-like images
+sometimes show sharp brightness variations (microlensing).  These
+brightness changes are associated with the innermost region of the
+quasar passing behind a complicated pattern of caustics due to the
+stars in the lensing galaxy.  In this paper, we study whether the
+universal properties of optical caustics could enable extraction of
+shape information about the central engine of quasars.  We present a
+toy model with a crescent-shaped source crossing a fold caustic.  The
+silhouette of a black hole over an accretion disk tends to produce
+roughly crescent sources.  When a crescent source crosses a fold
+caustic, the resulting light curve is noticeably different from the
+case of a disc or Gaussian source.  The crescent parameters, apart
+from one degeneracy, can be recovered.
 \end{abstract}

### 2014-04-17 paper1/discussion.tex
Edited Irshad's sections

@@
 
-An interesting case for future study would be crescent model with Gaussian brightness profile.
+This work presents a toy model to put useful constraints on the interior structure of quasars.
+To carry out this analysis on real data, one needs two situations: First, a quasar which can be observed with a reasonably good telescope with high signal to noise and second, a caustic crossing event of the quasar.
+These events are not rare, however, to get an isolated caustic crossing might be.
+However, as the size of the source is negligible as compared to big caustic structures in the sky, even the single fold caustic crossing is not that rare.
+
+In this toy analysis, we used very ideal data, which might not be the case with real source.
+So, we propose a more sophisticated analysis of this kind with different combinations of random and systematic noise and selection of data points for future.
+
+The main emphasis of this work was the crescent model parameters, which we show can be well recovered by the maximum likelihood analysis, apart from highly degenerate $a$ and $b$ parameters.
+The reason of this degeneracy is also very intuitive, if one integrates parallel to the caustic to get the integrated 1D brightness profile, so if the motion of the source is either parallel or perpendicular to the caustic, one of the two parameters ($a$ and $b$) is indistinguishable and in case of angled motion of the source, these two are highly degenerate.
+This degeneracy can be broken if multiple crossing events of the same quasar is available.
+So for example, one can use the first crossing to estimate the likelihood of the four parameters and use them as priors in the second crossing and so on.
+In this way, one can accurately and precisely recover all four parameters.
+
+Further, one can also study another interesting case, where the shape of the source is crescent like but the brightness of the crescent is not uniform but Gaussian.

### 2014-04-17 paper1/fitting.tex
Edited Irshad's sections

@@
 We carried out a likelihood analysis on all three parameter set using Markov-chain Monte-Carlo (MCMC).
+The main results of the MCMC are shown in figure \ref{fig:mcmc}.
+In the top row, we show the nine curves (likelihood) of recovered size of the source and its half light radius (left and right respectively).
+The legend of top row plots has information about the datasets, fitting model and best fit.
+The legend are named as, for example cd.chain, where first alphabet is the dataset and second alphabet is the fitting model.
+So, cd.chain represents the case where we fit disk model to crescent dataset.
+The number accompanying the name of the legend is the best fit $-2log(likelihood)$.
+We follow the convention of red, blue and green color for crescent, uniform disc and Gaussian disk datasets respectively whereas the fitting model is represented by different kind of curves: solid, dashed and dot-dashed respectively.
+
+ In the bottom rows we show three contours, for three different datasets (same color convention) fitted with crescent parameters only because this is the only case with more than one relevant parameter.
+Note that we plotted the 2D ellipses with principal component analysis of the MCMC and this is the reason they are perfect ellipses and not arbitary shape.
 
@@
 It is worth noticing that the datasets generated using crescent model and uniform disc model resembles much more than the one generated by Gaussian disc model source.
-This is also very intuitive as uniform disc is a special case in crescent model where $R_n=0$, The fiducial model assumed to generate these mock datasets are:
+This is also very intuitive as uniform disc is a special case in crescent model where $R_n=0$.
+The fiducial model assumed to generate these mock datasets are:
 \begin{equation}
@@
 where, $R$ being the size of the uniform disc and Gaussian disc source model.
+We also have some nuisance parameters in all these cases: the beginning of the event, the length of the event, normalisation of the lightcurve, however in this analysis, we marginalised over only one nuisance parameter, the beginning of the event, and kept the other two fixed.
 
@@
 Also, it can be inferred if one can simply distinguish between circularly symmetric and asymmetric sources or uniform and Gaussian sources.
-The errorbars on each data point, in all three datasets, is about 10 percent its value.
-Also, because we are fitting these datasets with different caustic crossing (point A to B in figure \ref{fig:magnification}), it adds a natural systematic errors on each data point.
+The errorbars on each data point (shaded region in figure \ref{fig:datafitting}), in all three datasets, is about 10 percent its value.
+Also, because we are fitting these datasets with different (arbitrary) caustic crossing (point A to B in figure \ref{fig:magnification}), it adds a natural systematic errors on each data point.
+We did not add any random noise to the data points, but because the noise is dominated by the systematic part, in this case, the overall contribution of random noise is negligible.
 We also assume the errors are Gaussian distributed.
@@
 Here we have four parameters as given by equation \ref{eqn:cp}.
-Three red curves in top row of figure \ref{fig:mcmc} represents the resulting likelihood of $R_p$, the overall size of the crescent source for three different datasets.
+Three solid curves in top row of figure \ref{fig:mcmc} represents the resulting likelihood of $R_p$, the overall size of the crescent source for three different datasets.
 Solid red curve in top row and red contours in bottom row of figure \ref{fig:mcmc} represents the crescent fitting with crescent data (cc.chain in figure \ref{fig:mcmc}).
@@
 It can also be well noticed that the half light radius is degenerate with $R_n/R_p$ or in other words, $R_p$ and $R_n$ are correlated.
+This degeneracy is not very strong, partially broken, and can be further broken by precise measurements of the light curve.
+Also, parameters $a$ and $b$ have changed their sign (but almost at the correct numerical value), this is due to fitting a reverse light curve in an opposite direction.
+These two parameters shows much stronger degeneracy.
 
 In second case, we tried to fit crescent model to disk data (dc.chain in figure \ref{fig:mcmc}).
-The dashed red line in top row and blue contours in bottom row represents this case.
+The solid blue in top row and blue contours in bottom row represents this case.
 Here one can recover the overall shape of the source, slightly lower than what is recovered in cc.chain case.
-Infact, the likelihood is higher than the previous case as the data is generated by one parameter only and fitting with four parameter that can also model the noise.
+Infact, the likelihood is higher than the previous case as the data is generated by one parameter only and fitting with four parameter that can also model the noise as well.
 Also, the maximum likelihood is achieved close to $R_n = 0$, as already mentioned, this is intuitive as well.
@@
 In the third case, Gaussian disc data is fit with crescent model (gc.chain).
-The dot-dashed red line in top row and green contours in bottom row represents this case.
+The solid green line in top row and green contours in bottom row represents this case.
 In this case, the extent of $R_p$ is broad and the likelihood is smaller than other cases.
 
-Please notice that in top row of figure \ref{fig:mcmc}, the height of the peaks does not represents in likelihood, it is just ensuring that the area under the curve is unity.
-The likelihood can be estimated from the number in the legend which is the best $\chi^2$.
+Please notice that in top row of figure \ref{fig:mcmc}, the height of the peaks does not represents the likelihood, it is just ensuring that the area under the curve is unity.
+The likelihood can be estimated from the number in the legend which is the best $-2log(likelihood)$.
 
@@
 In this case, we have only one parameter to fit, the outer radius of the uniform disc with uniform intensity.
-When fitting with crescent data (solid blue line), the maximum likelihood shifts towards higher values of $R$ as compared to the fiducial value (which was 50).
-Also the maximum likelihood and is very low and the contraints are also very weak.
-However, when fitting with Gaussian disc (dot-dashed blue line), the likelihood is still very low, but the parameter is tightly constrained.
-The third case is ideal when fitting uniform disc data (blue dashed line), where the constraints are ideal and also co-incide well with the fiducial value assumed and also likelihood is comparatively better.
+When fitting with crescent data (dashed red line), the maximum likelihood shifts towards higher values of $R$ as compared to the fiducial value (which was 50).
+Also the maximum likelihood is very low and the contraints are also very weak.
+However, when fitting with Gaussian disc (dashed green line), the likelihood is still very low, but the parameter is tightly constrained.
+The third case is ideal when fitting uniform disc data (dashed blue line), where the constraints are ideal and also co-incide well with the fiducial value assumed and also likelihood is comparatively better.
 
 \subsection{Fitting with Gaussian disc model}
-In the last case, we fit Gaussian disc model to the three datasets (green curves in top row of figure \ref{fig:mcmc}).
+In the last case, we fit Gaussian disc model to the three datasets (dat-dashed curves in top row of figure \ref{fig:mcmc}).
 In this case, when fitted with Gaussian disc data, the likelihood coincide with the fiducial value assumed, in other cases the likelihood is away but weak and very well constrained.
 
-\section{Conclusion and interpretation}
+\subsection{Conclusions and interpretations}
 The main results of this work is given in figure \ref{fig:mcmc} and described in the previous section. Here we try to interpret those results and conclude the take away message.
@@
 Lets assume that an ideal telescope observes a quasar for a long time and we get the lightcurve with good precision and one can clearly see the event of caustic crossing in the light curve.
-Now one wants extract the true shape of the source.
+Now one wants to extract the true shape of the source.
 To comment more specifically, lets assume three cases when the true shape of the source is: crescent, uniform disc and Gaussian disc:
@@
 At this point it does not matter as one can obtain the true result from both disk and crescent model.
-\item Gaussian disc shape source: if one tries to fit this curve with uniform disk model, again likelihood will discard this choice.
+\item Gaussian disc shape source: if one tries to fit this curve with uniform disk model or crescent model, again likelihood will discard this choice.
 Again the maximum likelihood analysis will point towards the best fitted true model.
-
 \end{enumerate}
@@
 At this point, we conclude that maximum likelihood analysis can be well trusted in order to recover true shape and recovering the corresponding parameters.
+This will provide the accurate recovery of the model parameters, however, for more precise recovery, one rely on precise measurements of the lightcurve.

### 2014-04-17 paper1/chains/analysismcmc.py

@@ -10,27 +10,27 @@ def plottinghist(chain,specs):
 	print
 	print mean(data),std(data),chi2,'for',chain
 	density = gaussian_kde(data)
-	xs = linspace(10,65,1000)
+	xs = linspace(10,75,1000)
 	density.covariance_factor = lambda : .25
 	density._compute_covariance()
 	plot(xs,density(xs),specs,label=chain+'   %1.2f'%chi2,linewidth=2.0)
 
 
 print '# mean, st. dev., best $\chi^2$'
-plottinghist('cc2.chain','r')
-plottinghist('dc2.chain','--r')
-plottinghist('gc2.chain','.-.r')
-plottinghist('cd2.chain','b')
-plottinghist('dd2.chain','--b')
-plottinghist('gd2.chain','.-.b')
-plottinghist('cg2.chain','g')
-plottinghist('dg2.chain','--g')
-plottinghist('gg2.chain','.-.g')
+plottinghist('cc.chain','r')
+plottinghist('dc.chain','b')
+plottinghist('gc.chain','g')
+plottinghist('cd.chain','--r')
+plottinghist('dd.chain','--b')
+plottinghist('gd.chain','--g')
+plottinghist('cg.chain','.-.r')
+plottinghist('dg.chain','.-.b')
+plottinghist('gg.chain','.-.g')
 
 pylab.legend(loc='upper center')
-pylab.xlim(10,65)
-pylab.xlabel('$\mathtt{R_{1/2}}$',fontsize=24)
+pylab.xlim(25,75)
+pylab.xlabel('$\mathtt{R_{p}}$',fontsize=24)
 
 
-pylab.savefig('Rhalf4all.eps')
+pylab.savefig('Rp4all.eps')
 pylab.show()

### 2014-04-17 paper1/ms.tex
.
small edits

@@
 
-\author[I. Mohammed et al.] {Irshad Mohammed\thanks{irshad@physik.uzh.ch}$^1$,
-Mihai Tomozeiu$^1$, Manuel Rabold$^1$,
+\author[Various authors]{[Order TBD]Irshad Mohammed,$^1$ Mihai Tomozeiu$^2$, Manuel Rabold$^2$,
 \newauthor
-Prasenjit Saha,$^1$ and Joachim Wambsgans$^2$\\
-$^1${Institute for Theoretical Physics, University of Z$\ddot{u}$rich, 8057 Z$\ddot{u}$rich Switzerland}\\
-$^2${University of Heidelberg, Germany}
+Prasenjit Saha,$^1$ and Joachim Wambsgans$^3$\\
+$^1${Physik-Institut, University of Zurich, Winterthurerstrasse 190,
+  8057 Zurich, Switzerland} \\
+$^2${Institute for Computational Science, University of Zurich,
+  Winterthurerstrasse 190, 8057 Zurich, Switzerland} \\
+$^3${Zentrum f\"ur Astronomie der Universi\"at Heidelberg,
+  M\"onchhofstrasse 12--14, 69120 Heidelberg, Germany}
 }

### 2014-04-17 paper1/theory.tex
small edits

@@
 \includegraphics[width = .8\textwidth]{plots/figure_rhalf.eps}
-\caption{\label{fig:lightcurve_crescent} Geometry of crescent sources.}
+\caption{\label{fig:geom_crescent} Geometry of crescent sources.}
 \end{figure}
@@
 \includegraphics[width = .8\textwidth]{plots/4rn_magnification.eps}
-\caption{\label{fig:lightcurve_crescent} Lightcurves of backwards crescents with  a = 0.3 $R_p$.}
+\caption{\label{fig:lightcurve_crescent_back} Lightcurves of backwards crescents with  a = 0.3 $R_p$.}
 \end{figure}

### 2014-09-15 paper1/chains/analysismcmc.py
some more edits

@@ -1,36 +1,49 @@
 import numpy
 import pylab
-def plottinghist(chain,specs):
+def plottinghist(chain,specs,burn):
 	from numpy import genfromtxt,mean,std,log,max,linspace
 	from scipy.stats import gaussian_kde
 	from pylab import plot,legend
-	datafull = genfromtxt(chain,skip_header = 100)
+	datafull = genfromtxt(chain,skip_header = burn)
 	data = datafull[:,2]
 	chi2 = -2.0*(max(datafull[:,0]))
-	print
-	print mean(data),std(data),chi2,'for',chain
+	index = list(datafull[:,0]).index(-chi2/2.0)
+	bestfit = datafull[index,:]
+#	print mean(data),std(data),chi2,'for',chain
+#	print bestfit,chain
+
 	density = gaussian_kde(data)
-	xs = linspace(10,75,1000)
+	xs = linspace(10,110,1000)
+#	xs = linspace(10,52,250)
 	density.covariance_factor = lambda : .25
 	density._compute_covariance()
-	plot(xs,density(xs),specs,label=chain+'   %1.2f'%chi2,linewidth=2.0)
 
+	bname = '../data/'+chain.lower().replace('chain','bestfit')
+	b = numpy.genfromtxt(bname)
+	nop = len(b)
+	print bname,nop
 
+	plot(xs,density(xs),specs,\
+		label=chain+'   %1.2f'%(chi2/nop),linewidth=2.0)
+
+burn = 3500
 print '# mean, st. dev., best $\chi^2$'
-plottinghist('CC.chain','r')
-plottinghist('DC.chain','b')
-plottinghist('GC.chain','g')
-plottinghist('CD.chain','--r')
-plottinghist('DD.chain','--b')
-plottinghist('GD.chain','--g')
-plottinghist('CG.chain','.-.r')
-plottinghist('DG.chain','.-.b')
-plottinghist('GG.chain','.-.g')
+plottinghist('CC.chain','r',900)
+plottinghist('DC.chain','b',burn)
+plottinghist('GC.chain','g',burn)
+plottinghist('CD.chain','--r',700)
+plottinghist('DD.chain','--b',700)
+plottinghist('GD.chain','--g',900)
+plottinghist('CG.chain','.-.r',8500)
+plottinghist('DG.chain','.-.b',burn)
[…]

### 2014-09-15 paper1/chains/burn_in.py

@@ -1,0 +1,11 @@
+import numpy
+import pylab
+
+chain = ['CC','CD','CG','DC','DD','DG','GC','GD','GG']
+spec = ['r','b','g','k','m','g','k','y','c']
+for i in range(9):
+	data = numpy.genfromtxt('%s.chain'%chain[i])
+	x = numpy.linspace(1,len(data)+1,len(data))
+	pylab.plot(x,-2.0*data[:,0],spec[i],label=chain[i])
+pylab.legend(loc=1,fontsize=12)
+pylab.show()

### 2014-09-15 paper1/chains/plotcontours.py

@@ -88,13 +88,17 @@ def plotcontours(array1,array2,CL,spec,opaque,description):
 	    plt.xlim(x-(CL+0.2)*xx,x+(CL+0.2)*xx)
 	    plt.ylim(y-(CL+0.2)*yy,y+(CL+0.2)*yy)
 
-burn=100
+burn=[8500,700,900]
+burn = burn[::-1]
 files = ['CG','CD','CC']
-spec = ['g','b','r']
+files = files[::-1]
+spec = ['#999933','#009999','#cc6633']
+spec = spec[::-1]
 labels = ['Gaussian Disk','Uniform Disk','Crescent']
+labels = labels[::-1]
 for i in range(len(files)):
-	f1 = numpy.genfromtxt(files[i]+'.chain',skip_header=burn)
-	f2 = numpy.genfromtxt(files[i]+'2.chain',skip_header=burn)
+	f1 = numpy.genfromtxt(files[i]+'.chain',skip_header=burn[i])
+	f2 = numpy.genfromtxt(files[i]+'2.chain',skip_header=burn[i])
 
 	Rp = f1[:,2]
 	Rn = f1[:,3]
@@ -104,10 +108,10 @@ for i in range(len(files)):
 
 	points = [Rhalf,Rn/Rp]
 	points = numpy.transpose(points)
-	plotcontours(Rhalf,Rn/Rp,2,spec[i],0.6,labels[i])
-pylab.legend(loc=1)
-pylab.ylim(0.4,1.0)
-pylab.xlim(15,40)
+	plotcontours(Rhalf,Rn/Rp,2,spec[i],1.0,labels[i])
+pylab.legend(loc=3)
+pylab.ylim(0.0,1.0)
+pylab.xlim(25,42)
 pylab.xlabel('$\mathtt{R_{1/2}}$',fontsize=18)
 pylab.ylabel('$\mathtt{R_n/R_p}$',fontsize=18)
 pylab.savefig('Rhalf_RnRp.eps')
@@ -115,8 +119,8 @@ pylab.show()
 
 
 for i in range(len(files)):
-	f1 = numpy.genfromtxt(files[i]+'.chain',skip_header=burn)
-	f2 = numpy.genfromtxt(files[i]+'2.chain',skip_header=burn)
+	f1 = numpy.genfromtxt(files[i]+'.chain',skip_header=burn[i])
[…]

### 2014-09-15 paper1/data/plotdata.py

@@ -1,30 +1,25 @@
 import numpy
 import pylab
-from matplotlib import pyplot as pl
 
-cr = numpy.genfromtxt('data_crescent_reverse_corr.txt')
-dr = numpy.genfromtxt('data_disk_reverse_corr.txt')
-gr = numpy.genfromtxt('data_gauss_reverse_corr.txt')
 
-err = 0.05
-low = 1-err
-high= 1+err
+def plot(filename,filename_bestfit1,filename_bestfit2,filename_bestfit3,name):
+	x = numpy.genfromtxt(filename)
+	y1 = numpy.genfromtxt(filename_bestfit1)
+	y2 = numpy.genfromtxt(filename_bestfit2)
+	y3 = numpy.genfromtxt(filename_bestfit3)
+	pylab.errorbar(x[:,0],x[:,1],x[:,1]*0.1,color='k',\
+		label='$\mathtt{%s}$'%name)
+	pylab.plot(y1[:,0]-y1[1,0]+y1[1,2],y1[:,1],'r',lw=2,label='$\mathtt{Crescent\ fit}$')
+	pylab.plot(y2[:,0]-y2[1,0]+y2[1,2],y2[:,1],'b',lw=2,label='$\mathtt{Uniform\ disk\ fit}$')
+	pylab.plot(y3[:,0]-y3[1,0]+y3[1,2],y3[:,1],'g',lw=2,label='$\mathtt{Gaussian\ disk\ fit}$')
 
-pylab.plot(cr[:,0]+800,cr[:,1],'r',label='Crescent')
-pylab.fill_between(cr[:,0]+800,cr[:,1]*low,cr[:,1]*high,\
-		alpha=0.5,facecolor='r',edgecolor=None,interpolate=True)
 
-pylab.plot(dr[:,0]+800,dr[:,1],'b',label='Uniform Disk')
-pylab.fill_between(dr[:,0]+800,dr[:,1]*low,dr[:,1]*high,\
-		alpha=0.5,facecolor='b',edgecolor=None,interpolate=True)
-
-pylab.plot(gr[:,0]+800,gr[:,1],'g',label='Gaussian Disk')
-pylab.fill_between(gr[:,0]+800,gr[:,1]*low,gr[:,1]*high,\
-		alpha=0.5,facecolor='g',edgecolor=None,interpolate=True)
-
-pylab.legend(loc=2)
-pylab.xlim(0,200)
-pylab.xlabel('Time')
[…]

### 2014-09-16 paper1/fitting.tex
small edits
some more edits
final fitting text edited

@@
 \centering
-  \includegraphics[width=0.95\hsize]{plots/data_with_error.eps}
-\caption{\label{fig:datafitting} data}
+\includegraphics[width=0.48\hsize]{plots/data_cc.eps}
+\includegraphics[width=0.48\hsize]{plots/data_dd.eps}
+\includegraphics[width=0.48\hsize]{plots/data_gg.eps}
+\includegraphics[width=0.48\hsize]{plots/burnin_cg.eps}
+\caption{\label{fig:datafitting} Three different mock datasets (in black) -- crescent, uniform-disk and Gaussian disk (in reading order).
+In each panel showing the best fit for each of the three models - crescent (red), uniform-disk (blue) and Gaussian disk (green).
+Bottom-right panel shows the $\chi^2$ of each step of the full MCMC for CG case (Gaussian disk data fitted with crescent source model)}.
 \end{figure}
@@
 \centering
-  \includegraphics[width=0.45\hsize]{plots/Rp4all.eps}
-  \includegraphics[width=0.45\hsize]{plots/Rhalf4all.eps}\\
-  \includegraphics[width=0.45\hsize]{plots/Rhalf_RnRp.eps}
-  \includegraphics[width=0.45\hsize]{plots/aRp_bRp.eps}
-
-\caption{\label{fig:mcmc} Model Fitting}
+  \includegraphics[width=0.48\hsize]{plots/Rp4all.eps}
+  \includegraphics[width=0.48\hsize]{plots/Rhalf4all.eps}\\
+  \includegraphics[width=0.48\hsize]{plots/Rhalf_RnRp.eps}
+  \includegraphics[width=0.48\hsize]{plots/aRp_bRp.eps}
+\caption{\label{fig:mcmc} Top panels showing the 1D probability density function for size of the source, $R_p$ (left), and half light radius of the source, $R_{1/2}$ (right).
+Same color represeents the same dataset whereas same line style corresspond to same model fit.
+In the legend showing the reduced $\chi^2$ of the best fit in each case.
+Bottom two panels showing the 2$\sigma$ contours (or error ellipses) of the crescent model when fitting with the three different datasets.}
 \end{figure}
 
-In this section, we attempt to fit and recover the model parameters from the given dataset generated by different source models - crescent, uniform-disk and Gaussian-disk.
-So broadly we have 9 cases to discuss.
-We carried out a likelihood analysis on all three parameter set using Markov-chain Monte-Carlo (MCMC).
-The main results of the MCMC are shown in figure \ref{fig:mcmc}.
-In the top row, we show the nine curves (likelihood) of recovered size of the source and its half light radius (left and right respectively).
-The legend of top row plots has information about the datasets, fitting model and best fit.
-The legend are named as, for example cd.chain, where first alphabet is the dataset and second alphabet is the fitting model.
-So, cd.chain represents the case where we fit disk model to crescent dataset.
-The number accompanying the name of the legend is the best fit $-2log(likelihood)$.
-We follow the convention of red, blue and green color for crescent, uniform disc and Gaussian disk datasets respectively whereas the fitting model is represented by different kind of curves: solid, dashed and dot-dashed respectively.
+In this section, we attempt to mock three datasets, each for crescent, uniform-disk and Gaussian disk case for a given fiducial parameters and contaminate it with systematic and random noise, and fit \& recover the model parameters.
+So broadly we have 9 cases to discuss, we name these nine cases as different permutations (couplet) of three letters C, D and G for crescent, uniform-disk and Gaussian disk case respectively.
+The first letter of the couplet represents the fit model and the second represents the dataset.
+For example, CG represents the case where we fit Gaussian disk data to a crescent model.
+Similarly, DC respresents the case when crescent data is being fit with uniform-disk model.
+We explored the parameter space and carried out a likelihood analysis on all three parameter sets using Markov-chain Monte-Carlo (MCMC).
 
- In the bottom rows we show three contours, for three different datasets (same color convention) fitted with crescent parameters only because this is the only case with more than one relevant parameter.
-Note that we plotted the 2D ellipses with principal component analysis of the MCMC and this is the reason they are perfect ellipses and not arbitary shape.
+\subsection{Mock datasets and model parameters}
+The datasets for each source model is in the form of light curves, the brightness of the source observed at different time.
+Figure \ref{fig:datafitting} shows the three datasets.
+These datasets are generated using the magnification map shown in figure \ref{fig:magnification} if the source of the respective model moves from point C to B.
+The data assumed was very  ideal, including 250 different data points at regular interval of time.
+We added Gaussian random noise to this dataset of width 10\% of the brightness at the respective time, which makes the signal to noise of about $\sim$150.
+After adding random noise, the three datasets are looking highly contaminated and no clear intuitive behaviour of fit can be seen.
 
-\subsection{Mock datasets}
-The datasets for each source model is in the form of light curves, the brightness of the source observed at different time scale.
-Figure \ref{fig:datafitting} shows the three datasets.
-These datasets are generated using the magnification map shown in figure \ref{fig:magnification} between point B and C (direction from C to B).
-The data assumed was very  ideal, including 200 different data points at regular interval of time.
-It is worth noticing that the datasets generated using crescent model and uniform disc model resembles much more than the one generated by Gaussian disc model source.
-This is also very intuitive as uniform disc is a special case in crescent model where $R_n=0$.
 The fiducial model assumed to generate these mock datasets are:
@@
 \end{equation}
+\\
+where, $R_p$ and $R_n$ are the radius of outer and inner circles of the crescent respectively and $a$ and $b$ are the coordinates of the inner circle center with center of outer circle as the origin.
+
 \begin{equation}
-	R = 50.0 \ \ \ \rm{for\ Uniform\ disc\ model}
+	R_p = 50.0 \ \ \ \rm{for\ Uniform\ disc\ model}
 \end{equation}
@@
 \begin{equation}
-	R = 50.0 \ \ \ \rm{for\ Gaussian\ disk\ model}
+	R_p = 50.0 \ \ \ \rm{for\ Gaussian\ disk\ model}
 \end{equation}
 \\
-where, $R$ being the size of the uniform disc and Gaussian disc source model.
-We also have some nuisance parameters in all these cases: the beginning of the event, the length of the event, normalisation of the lightcurve, however in this analysis, we marginalised over only one nuisance parameter, the beginning of the event, and kept the other two fixed.
+where, $R_p$ being the size of the uniform disc and Gaussian disc source model.
+We also have some nuisance parameters in all these cases: the beginning of the event (I), the end of the event (E), normalisation of the lightcurve (N).
+We marginalised over all these nuisance parameters.
+When we kept these three nuisance parameters free to vary in MCMC, it adapts to the amount of data for which the maximum likelihood is most optimal to find out.
+We talk more about this adaptaion in the next section.
+Secondly, We try to fit the these datasets for a crossing of the source from a different caustic, in this case, from point A to B in figure \ref{fig:magnification}.
+This adds a natural systematic in the dataset and the data is no more ideal.
+With these three characterstics -- random noise, systematic noise and nuisance parameters -- we make the dataset very general and highly contaminated to mock the real observations.
 
-The aim of this exercise was to explore the degeneracies in  various model parameters and recover the correct fiducial model assumed.
-Also, it can be inferred if one can simply distinguish between circularly symmetric and asymmetric sources or uniform and Gaussian sources.
-The errorbars on each data point (shaded region in figure \ref{fig:datafitting}), in all three datasets, is about 10 percent its value.
-Also, because we are fitting these datasets with different (arbitrary) caustic crossing (point A to B in figure \ref{fig:magnification}), it adds a natural systematic errors on each data point.
-We did not add any random noise to the data points, but because the noise is dominated by the systematic part, in this case, the overall contribution of random noise is negligible.
-We also assume the errors are Gaussian distributed.
+The aim of this exercise was to explore the degeneracies in  various model parameters and recover the correct fiducial model assumed from a highly contaminated dataset.
+Also, if it is possible that one can simply distinguish between circularly symmetric and asymmetric sources or uniform and Gaussian sources on the basis of maximum likelihoods.
 
 \subsection{Fitting with crescent model}
+
+The main results of the MCMC are shown in figure \ref{fig:mcmc}.
+In the top row, we show the nine probability density function of recovered size of the source ($R_p$) and its half light radius ($R_{1/2}$) (left and right respectively).
+The legend of top-left panel has information about the datasets, fitting model and best fit also applicable to top-right panel for $R_{1/2}$.
+The name of the legend has the same convention as explained in the previous section.
+The number accompanying the name of the legend is the best fit reduced $\chi^2$.
+We follow the convention of red, blue and green color for crescent, uniform disc and Gaussian disk datasets respectively whereas the fitting model is represented by different kind of curves: solid, dashed and dot-dashed respectively.
+
 First we explored the parameter space for crescent source model and run its likelihood exploration for three different datasets.
-Here we have four parameters as given by equation \ref{eqn:cp}.
-Three solid curves in top row of figure \ref{fig:mcmc} represents the resulting likelihood of $R_p$, the overall size of the crescent source for three different datasets.
-Solid red curve in top row and red contours in bottom row of figure \ref{fig:mcmc} represents the crescent fitting with crescent data (cc.chain in figure \ref{fig:mcmc}).
-All four parameters can be well recovered in this case and the overall likelihood coincides well with the fiducial values assumed and stated in equation \ref{eqn:cp}.
-It can also be well noticed that the half light radius is degenerate with $R_n/R_p$ or in other words, $R_p$ and $R_n$ are correlated.
-This degeneracy is not very strong, partially broken, and can be further broken by precise measurements of the light curve.
-Also, parameters $a$ and $b$ have changed their sign (but almost at the correct numerical value), this is due to fitting a reverse light curve in an opposite direction.
-These two parameters shows much stronger degeneracy.
+Here we have four interesting parameters as given by equation \ref{eqn:cp} and we marginalise over all other nuisance parameters.
+Three solid curves in top row of figure \ref{fig:mcmc} represents the resulting likelihood of $R_p$, the overall size of the crescent source fitted with three different datasets.
+We also show results for half light radius $R_{1/2}$ which is calculated using equation \ref{eqn:}.
 
-In second case, we tried to fit crescent model to disk data (dc.chain in figure \ref{fig:mcmc}).
-The solid blue in top row and blue contours in bottom row represents this case.
-Here one can recover the overall shape of the source, slightly lower than what is recovered in cc.chain case.
-Infact, the likelihood is higher than the previous case as the data is generated by one parameter only and fitting with four parameter that can also model the noise as well.
-Also, the maximum likelihood is achieved close to $R_n = 0$, as already mentioned, this is intuitive as well.
-The other paramters, $a$ and $b$ are quite arbitary.
+\begin{enumerate}
 
-In the third case, Gaussian disc data is fit with crescent model (gc.chain).
-The solid green line in top row and green contours in bottom row represents this case.
-In this case, the extent of $R_p$ is broad and the likelihood is smaller than other cases.
+\item Crescent data: This is the true case for crescent model fitting.
+The red solid line represents this case in upper panels of \ref{fig:mcmc} as 1D probability density function marginalised over all other parameters.
+The true value of $R_p$ can be recovered within 2$\sigma$ level with an accuracy up to 10\%.
+Also the best fit to the data is shown in red line in upper-left panel of figure \ref{fig:datafitting}.
+It is a very good fit which also represents a reduced $\chi^2$ close to unity (1.04).
+The orange contour in the lower panels of figure \ref{fig:mcmc} showing the 2D marginalised error ellipses.
+There is a certain degeneracy in $R_{1/2}$ and $R_n/R_p$ parameters but all the values are well recovered.
+
+\item Uniform disk data: In this case, we tried to fit crescent model to disk data.
+The solid blue line in top row panels and cyan contours in bottom row panel represent this case.
+The best fit reduced $\chi^2$ (0.90) is better than the previous case as the data is generated by one parameter only and fitting with four parameter that can also model the noise.
+This can also be seen in the best fit to the data, figure \ref{fig:datafitting} red line in top-right panel, where the curve seems to fit the noise as well.
+The maximum likelihood (and also average likelihood) is achieved close to $R_n = 0$, as already mentioned, this is intuitive as well.
+The other paramters, $a$ and $b$ are quite arbitary and highly uncertain.
+
+\item Gaussian disk data: In this case, Gaussian disc data is fited with crescent model.
+The solid green line in top row panels and green contours in bottom row panels of figure \ref{fig:mcmc} represent this case.
+By looking at figure \ref{fig:mcmc}, it can be seen that all the parameters can be constrained really well, even the best fit $\chi^2$ is much better than other two cases (0.85) which also shows a bit of over-fitting problem.
+However, this is misleading.
+The reason for these strong constraints is that the full data is not at all fitted, in this case the nuisance parameters are rejecting most of the data during the progress of MCMC.
+This can be seen in figure \ref{fig:datafitting} bottom-left panel in red line, where the best fit model is fitted only in very small range of data, even before the actual incidence happened.
+In other words, MCMC is unable to converge in this case and the only way to achieve better likelihood is to reject data points.
+This can be seen in bottom-left panel of figure \ref{fig:datafitting}.
+The $\chi^2$ keep on reducing through the span of MCMC chain and not saturating anywhere.
+
+\end{enumerate}
+
+Therefore we can conclude that by fitting with crescent source model, one can recover the parameters very well if the dataset is generated by crescent source model.
+The uniform disk source model can also be recovered very well for the overall shape of the disk, as the latter is the special case of the former, but other parameters like $a$ and $b$ are arbitary and $R_n \sim 0$.
+No convergence can be achieved for data generated by Gaussian disk model.
 
 Please notice that in top row of figure \ref{fig:mcmc}, the height of the peaks does not represents the likelihood, it is just ensuring that the area under the curve is unity.
-The likelihood can be estimated from the number in the legend which is the best $-2log(likelihood)$.
+The likelihood can be estimated from the number in the legend which is the best fit $\chi^2$.
+
+ In the bottom rows we show three contours, for three different datasets (same color convention) fitted with crescent parameters only because this is the only case with more than one relevant parameter.
+Note that we plotted the 2D ellipses with principal component analysis of the MCMC and this is the reason they are perfect ellipses and not arbitary shape.
 
 \subsection{Fitting with uniform disc model}
-In this case, we have only one parameter to fit, the outer radius of the uniform disc with uniform intensity.
-When fitting with crescent data (dashed red line), the maximum likelihood shifts towards higher values of $R$ as compared to the fiducial value (which was 50).
-Also the maximum likelihood is very low and the contraints are also very weak.
-However, when fitting with Gaussian disc (dashed green line), the likelihood is still very low, but the parameter is tightly constrained.
-The third case is ideal when fitting uniform disc data (dashed blue line), where the constraints are ideal and also co-incide well with the fiducial value assumed and also likelihood is comparatively better.
+In this case, we have only one interesting parameter to fit, the outer radius of the uniform disc with uniform intensity.
+When fitting with the true case, the uniform disk data itself (dashed blue line in upper panels of figure \ref{fig:mcmc}), the best fit $\chi^2$ is very good (0.91) and also it recovers the true value of $R_p=50$.
+When fitting with crescent data (dashed red line), the maximum likelihood shifts towards higher values of $R_p$ as compared to the fiducial value.
+This is justified and intuitive as $R_n \sim > 0$.
+Also the best fit reduced $\chi^2$ is 1.29 which is much worse than the true case.
+However, when fitting with Gaussian disc data (dashed green line), the best fit reduced $\chi^2$ is still very low but still worse than the true case.
+However, it can be seen in figure \ref{fig:datafitting} (blue line bottom-left panel) that in this case also many points are rejected in order to gain better likelihood.
+
+Therefore we conclude here that looking at the maximum likelihood one can identity the true dataset and also recover the true values of the parameter.
 
 \subsection{Fitting with Gaussian disc model}
-In the last case, we fit Gaussian disc model to the three datasets (dat-dashed curves in top row of figure \ref{fig:mcmc}).
-In this case, when fitted with Gaussian disc data, the likelihood coincide with the fiducial value assumed, in other cases the likelihood is away but weak and very well constrained.
+In the last case, we fit Gaussian disc model to the three datasets (dot-dashed curves in top panels of figure \ref{fig:mcmc}).
+In this case, when fitted with the true case, Gaussian disc dataset, the true parameters can be recovered at the correct value and also the best fit reduced $\chi^2$ is close to unity (0.98).
+In all other cases the maximum likelihood is much worse and also the recovered values of $R_p$ are many $\sigma$ away.
+
+Hence, in this case also by looking at the maximum likelihood true case can be identified and parameters can be restored correctly.
 
@@
 
-Lets assume that an ideal telescope observes a quasar for a long time and we get the lightcurve with good precision and one can clearly see the event of caustic crossing in the light curve.
+Lets assume that an ideal telescope observes a quasar for a long time and we get the lightcurve  and one can clearly see the event of caustic crossing in the light curve.
 Now one wants to extract the true shape of the source.
@@
 \item Crescent shape source: if one tries to fit this curve with crescent model, the maximum likelihood will be good and one can recover the true parameters with good precision.
-In the other two cases, the maximum likelihood will be very bad.
+In the other two cases, the maximum likelihood will be bad.
 Hence this shape can be recovered very well by trusting the maximum likelihood analysis.
 \item Uniform disc shape source: if one tries to fit this curve with Gaussian disk model, the likelihood will be very little to disprove it, however, the other two models are inditinguishable with nearly the same likelihood.
-At this point it does not matter as one can obtain the true result from both disk and crescent model.
+At this point it does not matter as one can obtain the true result from both uniform disk and crescent model.
 \item Gaussian disc shape source: if one tries to fit this curve with uniform disk model or crescent model, again likelihood will discard this choice.

### 2014-09-16 paper1/chains/analysismcmc.py

@@ -13,37 +13,40 @@ def plottinghist(chain,specs,burn):
 #	print bestfit,chain
 
 	density = gaussian_kde(data)
-	xs = linspace(10,110,1000)
-#	xs = linspace(10,52,250)
+#	xs = linspace(10,110,1000)
+	xs = linspace(10,52,250)
 	density.covariance_factor = lambda : .25
 	density._compute_covariance()
 
-	bname = '../data/'+chain.lower().replace('chain','bestfit')
-	b = numpy.genfromtxt(bname)
-	nop = len(b)
-	print bname,nop
+#	bname = '../data/'+chain.lower().replace('chain','bestfit')
+#	b = numpy.genfromtxt(bname)
+#	nop = len(b)
+#	print bname,nop
 
-	plot(xs,density(xs),specs,\
-		label=chain+'   %1.2f'%(chi2/nop),linewidth=2.0)
+	plot(xs,density(xs),specs,linewidth=2.0)
+#	plot(xs,density(xs),specs,\
+#		label=chain+'   %1.2f'%(chi2/nop),linewidth=2.0)
+
 
 burn = 3500
 print '# mean, st. dev., best $\chi^2$'
-plottinghist('CC.chain','r',900)
-plottinghist('DC.chain','b',burn)
-plottinghist('GC.chain','g',burn)
-plottinghist('CD.chain','--r',700)
-plottinghist('DD.chain','--b',700)
-plottinghist('GD.chain','--g',900)
-plottinghist('CG.chain','.-.r',8500)
-plottinghist('DG.chain','.-.b',burn)
-plottinghist('GG.chain','.-.g',500)
+plottinghist('CC2.chain','r',900)
+plottinghist('DC2.chain','--r',burn)
+plottinghist('GC2.chain','.-.r',burn)
+plottinghist('CD2.chain','b',700)
+plottinghist('DD2.chain','--b',700)
+plottinghist('GD2.chain','.-.b',900)
+plottinghist('CG2.chain','g',8500)
+plottinghist('DG2.chain','--g',burn)
+plottinghist('GG2.chain','.-.g',500)
 
 pylab.legend(loc=9,fontsize=12)
-pylab.xlim(25,110)
-#pylab.xlim(10,52)
+#pylab.xlim(25,110)
+pylab.xlim(10,52)
 pylab.xlabel('$\mathtt{R_{p}}$',fontsize=20)
[…]

### 2014-09-16 paper1/chains/burn_in.py

@@ -2,10 +2,15 @@ import numpy
 import pylab
 
 chain = ['CC','CD','CG','DC','DD','DG','GC','GD','GG']
-spec = ['r','b','g','k','m','g','k','y','c']
-for i in range(9):
+chain = ['CG']
+spec = ['k','b','g','k','m','g','k','y','c']
+for i in range(len(chain)):
 	data = numpy.genfromtxt('%s.chain'%chain[i])
 	x = numpy.linspace(1,len(data)+1,len(data))
-	pylab.plot(x,-2.0*data[:,0],spec[i],label=chain[i])
-pylab.legend(loc=1,fontsize=12)
+	pylab.semilogx(x,-2.0*data[:,0],spec[i],label='$\mathtt{CrescentFit\ to\ GaussianDiskData}$')
+pylab.xlabel('$\mathtt{Number\ of\ MCMC\ steps}$',fontsize=22)
+pylab.ylabel('$\mathtt{\chi^2}$',fontsize=22)
+pylab.xlim(min(x),max(x))
+pylab.legend(loc=3,fontsize=20)
+pylab.savefig('burnin_cg.eps')
 pylab.show()

### 2014-09-16 paper1/data/plotdata.py

@@ -7,16 +7,17 @@ def plot(filename,filename_bestfit1,filename_bestfit2,filename_bestfit3,name):
 	y1 = numpy.genfromtxt(filename_bestfit1)
 	y2 = numpy.genfromtxt(filename_bestfit2)
 	y3 = numpy.genfromtxt(filename_bestfit3)
-	pylab.errorbar(x[:,0],x[:,1],x[:,1]*0.1,color='k',\
+	pylab.errorbar(x[:,0]+850,x[:,1],x[:,1]*0.1,color='k',\
 		label='$\mathtt{%s}$'%name)
-	pylab.plot(y1[:,0]-y1[1,0]+y1[1,2],y1[:,1],'r',lw=2,label='$\mathtt{Crescent\ fit}$')
-	pylab.plot(y2[:,0]-y2[1,0]+y2[1,2],y2[:,1],'b',lw=2,label='$\mathtt{Uniform\ disk\ fit}$')
-	pylab.plot(y3[:,0]-y3[1,0]+y3[1,2],y3[:,1],'g',lw=2,label='$\mathtt{Gaussian\ disk\ fit}$')
+	pylab.plot(y1[:,0]-y1[1,0]+y1[1,2]+850,y1[:,1],'r',lw=2)
+	pylab.plot(y2[:,0]-y2[1,0]+y2[1,2]+850,y2[:,1],'b',lw=2)
+	pylab.plot(y3[:,0]-y3[1,0]+y3[1,2]+850,y3[:,1],'g',lw=2)
 
 
-plot('data_gauss_reverse_noise.txt','cg.bestfit','dg.bestfit','gg.bestfit','Gaussian\ disk\ data')
-pylab.legend(loc=2,fontsize=16)
-pylab.xlim(-850,-600)
+plot('data_gauss_reverse_noise.txt','cg.bestfit','dg.bestfit','gg.bestfit','Gaussian\ Disk\ data')
+pylab.legend(loc=2,fontsize=22)
+#pylab.axes().set_aspect(20.0)
+pylab.xlim(0,250)
 pylab.ylim(ymin=2)
 pylab.xlabel('$\mathtt{Time}$',fontsize=22)
 pylab.ylabel('$\mathtt{Brightness}$',fontsize=22)

### 2014-10-05 paper1/discussion.tex
Theory edits

@@
 \section{Discussion}
+[Mihai's contribution]\\
+In the current paper we simulate and study the resulting microlensing lightcurves of geometric crescent-shaped sources and compare them with the microlensing lightcurves of other simple mathematically describable source profiles.
+In order to mimic the behaviour of flux of light from the source in the proximity of a fold caustic we make use of the simple approximation described in equation (5).
+The equation would exhaustively describe the magnification map and offer a good aproximation for the particular microlensing regime that we consider.
+Namely, the shape of the caustic boundary in the proximity of the source in the respective plane can be approximated with a line due to reason that the local radius of curvature of caustic is orders of magnitude greater than the half-light radius of the studied source.
+In particular cases in which the previously mentioned approximation loses its validity the impact on the quality of the lightcurves is not evenly distributed.
+The shape of the lightcurve will be maintained though the datapoints corresponding to the source position before and during overlapping the caustic will be affected by smaller errors than the datapoints corresponding to later times.
+\\
+
+The first two source profiles that we consider are the uniform disk and symmetric gaussian source.
+Both of them can be described by a half-light radius $r_{1/2}$ and a total unlensed light flux $S_0$.
+With the two parameters constrained the one dimensional profiles as well as the lightcurves of the two source are completely determined, since no free parameter remains.
+The previous statement does not hold for a crescent source.
+In the case of the crescent source there are in total five parameters: the integrated flux of the source $S_0$, the radii of the bright/dark disk $R_p$/$R_n$ and the displacement of the centers of the two disks on the axes perpendicular and parallel to the caustic $a$ and $b$.
+Two of the parameters can be reduced by expressing the results in terms of $S_0$ and $r_{1/2}$.
+The later being unique for any set of parameters $R_p, R_n$ and $a^2+b^2$.
+Moreover, one of the displacemement parameters $b$ has no impact on the one dimensional profile of the source that results from the projection of the source image on an axis perpendicular to the caustic.
+Since the one dimensional source profile that correpsonds to an axis perpendicular to the caustic contains exhaustively all the information regarding the source that can be revealed by the lightcurve, the value of the parameter $b$ does not effect the shape of the lighcurve.
+Nevertheless the $b$ parameter is relevant to the calculation of $r_{1/2}$.
+It's qualitative effect is to decrease the value of the half-light radius when the absolute value of the parameter is increased.
+With two parameters constrained and another irrelevant to the shape of the lightcurve, two free parameters remain $R_n$ and $a$.
+Figure 4 reveals that the lighcurve of a crescent source has more visible features than the other two light-curves corresponding to the disc and guasssian shape.
+There are three regions where $S_1D$ and by inheritance the lighcurve have distinguishable behaviour.
+The first region that would be recorded on a lightcurve plot represents the preiod of time when the bright disk begins to be overlapped by the caustic and stops when just before the dark disk reaches the caustic.
+During this period of time the flux of light from the source is increasingly magnified.
+The second period starts and ends with the overlapping of the dark disc.
+As a boundary of the two regions there is a distinguishable point where slope of the magnification is drastically changed.
+This apparent discontinuity in the first derivative of the magnification function is caused by the caustic amplification of the sudden drop in the $S_1d$ function.
+During the respective period the magnification growth slows or even reverses during the first part of the period and starts to grow faster again as the dark ends its overlap with the caustic.
+At the point where the dark disc clears the caustic the growth of the magnification is infinite, which appears as a saddle point on the lightcurve.
+Next, the final period correspond to the case when the dark disc has cleared the caustic and the bright disc continues to overlap with the caustic.
+During this period the lighcurve reaches a peak that for most of the parameter space is global and for the rest of the parameter space local.
+At the end of the period the growth of the magnification is negative infinite.
+the respective point appears as a second saddle point on the lightcurve.
+Past this point the magnification of any finite source will decay in roughly the same manner.
+\\
+
+The impact of parameter $a$ on the shape of the lightcurve can be observed in figure 5 (a-var).
+Sources where the center of the dark disc reaches the caustic before the center of the bright disc are characterized by a smoother broad peak in contrast to the cases where the center of the dark disc reaches the caustic after the brigh disk.
+In the case of the latter the instance when the dark disc reaches the caustic corresponds to larger and larger magnifications untill it becomes a local and even global peak.
+The effect of the $R_n$  parameter on the lighcurve is presented in figures 6 and 7.
+For the particular set of parameters where $R_p = R_n +a$ the third period of time discussed previously does not exist.
+Particularizing further, if the value of the radius of the dark disc is comparable to the value of the radius of the bright one then the position and shape of the maximum magnification are strikingly different.
+In case the crescent reaches the caustic with the bright region first, the peak magnification happens when the dark region reaches the caustic and it is characterized by a sharp variation in magnification growth, In the opposite case the peak appears before the end of the bright disc reaches the caustic and shape is smoother.
+\\
+
+In addition to the previously studied case of a geometric crescent source created from the overlap of a bright and dark disc we have generated the lightcurves of a constant luminosity elliptical crescent and a gaussian elliptical crescent.
+The first shows striking similarities to the disk crescent source, being characterized by the similar characteristic periods and instances of time.
+The second, gaussian elliptical crescent, similarly to the symetric gaussian source has a continuous one dimensional flux profile except in the region of the elliptic shadow.
+The respective region is detected as a period during which the magnification behaves similarly to the corresponding shadow overlap periods of constant brightness sources previously studied.
+The similarity between the shapes of the lightcurves of the geometric crescent, elliptical crescent and guassian profile suggests that the geometric crescent is a usefull first approximation of the elliptical sources.
+\\
+
+With the different source profiles and their corresponding lightcurves studied we can change our point of view of the system to that of an observer.
+The observer would basically detect only the lightcurve of such as source.
+As described in section ??
+the timing of the onset and offset of the previously described periods can be used to estimate the values of the radii and one of the displacement parameters when assuming a geometric crescent, the halfwidths and one displacement parameters when assuming a constant brightness elliptical crescent and the halfwidth of the dark ellipse when assuming an elliptical gaussian crescent.
+All quantities can be estimated in terms of the relative velocity of the source in a direction perpendicular to the caustic.\\
+
+The previously mentioned abstract parameters can be related to physical quantities specific to the central region of a quasar.
+Thus the luminous region would correspond to the bright accretion disc that surrounds the black hole.
+The later's gravity would cause a shadow in the bright region limited by the extent of the event horizon of the black hole.
+Therefore the radius of the bright disk or halfwidth of the gaussian crescent would provide an estimate of the size of the accretion disk and the radius of the dark disk or halfwidth of the dark ellipse projection would provide and estimate of the Schwarzschild radius of the black hole.
+Moreover a value of the Schwarzschild radius can be used to determine the mass of the black hole $M \approx \frac{\left(t_3 - t_2 \right) v_p c^2}{2G}$.
+
+The parameters whose values cannot be determined due to the loss of information from the directions parallel to the caustic could be obtained in the eventuality in which the same source crosses multiple caustics that are not parallel.
+Multiple crossing of caustics can reveal details of the one dimensional flux profile corresponding to multiple distinct directions which would allow the reconstruction of the two dimesnional profile analogous to the process through which a picture of a CT scan is obtained.
+
+In case of a high quality lightcurve with insignificant noise and measurement errors, the parameters can be obtained by simply identifying the characterstic instances of time withought making use of the actual values of the magnification map.
+If the effect of the errors and the noise distorts the magnification time function enough so that the characteristic points are not identifiable with the characteristic periods still visible, the boundaries of the periods can be roughly estimated.
+To study if good estimates of the parameters can still be obtained when datapoints are strongly affected by the noise we propose ....
+
+[Irshad's contribution]\\
 
@@
 Further, one can also study another interesting case, where the shape of the source is crescent like but the brightness of the crescent is not uniform but Gaussian.
+
+[Contributions from authors]\\
+Prasenjit Saha provided the original idea and plan for the research project as well as multiple contributions to the analysis.
+Mihai Tomozeiu simulated and studied the ideal behaviour of the microlensing lightcurves for the different source profiles discussed.
+Joachim Wambsganss provided the numerical code used by Manuel Rabold to create the magnification map and corresponding lighcurves used in the MCMC analysis, analisis that was performed by Irshad Mohammed in the last part of the presented work.

### 2014-10-05 paper1/theory.tex
Theory edits

@@
  &\qquad \left. {} - \sqrt{R_n^2 - (p-p_{sn})^2 } \Theta \left[ R_n^2 - \left( p-p_{sn} \right)^2 \right] \right\}.
+\label{eqn:s1_d}
 \end{align}
@@
 The half-light radius of any source is invariant to any rotational transformation.
-In the present case of a crescent source the effective radius is dependent on the parameters $R_p$, $R_n$ and $\sqrt{a^2+b^2}$ exclusively.
+In the present case of a crescent source the effective radius is dependent on the parameters $R_p$, $R_n$ and $\sqrt{a^2+b^2}\equiv c $ exclusively.
 From symmetry considerations the centroid of the source is colinear with the centers of the two disks and it is situated at a distance $d_c$ from the center of the brigth disk.
-$d_c$ can be computed numerically by solving the equation:
+$d_c$ can be computed numerically with the use of a variation of equation \ref{eqn:s1_d}:
 
 \begin{equation}
-\frac{S_0^C}{2} =  
+\frac{S_0^C}{2} =  \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{d_c}^{R_p} \left[ \sqrt{R_p^2 - p^2} - \sqrt{R_p^2 - \left(p-c\right)^2} \Theta \left(R_p^2 - \left(p-c\right)^2 \right) \right] dp.
+\end{equation}
+
+With the position of the centroid determined, the half-light radius can be also be computed numerically:
+
+\begin{equation}
+\frac{S_0^C}{2} =  \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{0}^{r_{1/2}} \left[ \Theta_2  - \Theta_1  \right] p dp;
+\end{equation}
+
+\begin{equation}
+\Theta_1 = \left[ \pi - \arccos \frac{R_p^2 + p^2 -d_c^2}{2 R_p p} - \arccos \frac{R_p^2 - p^2 + d_c^2}{2 R_p d_c} \right] \Theta \left( p - R_p + d_c \right); 
+\end{equation}
+
+\begin{equation}
+\Theta_2 =  \pi - \arccos \frac{ p^2  + (c+d_c)^2 - R_n^2}{2 p \left(c + d_c \right)} \Theta \left( p + R_n - d_c -c \right). 
 \end{equation}
@@
 \begin{figure}
-\includegraphics[width = .8\textwidth]{plots/4source_magnification.eps}
+\centering
+	\includegraphics[width = 0.48\textwidth]{plots/4source_profile.eps}
+	\includegraphics[width = 0.48\textwidth]{plots/4source_magnification.eps}
 \caption{\label{fig:lightcurve_gauss} Lightcurves of crescents, gaussian and disk for sources with identical $S_0$ and $r_{1/2}$.
@@
 \begin{figure}
-\includegraphics[width = .8\textwidth]{plots/4avar_magnification.eps}
+\centering
+	\includegraphics[width = 0.48\textwidth]{plots/4avar_profile.eps}
+	\includegraphics[width = 0.48\textwidth]{plots/4avar_magnification.eps}
 \caption{\label{fig:lightcurve_disk} Lightcurves of crescent sources with $R_n$ = 0.4 $R_p$.}
@@
 \begin{figure}
-\includegraphics[width = .8\textwidth]{plots/4rn_magnification.eps}
+\centering
+	\includegraphics[width = 0.48\textwidth]{plots/4rn_profile.eps}
+	\includegraphics[width = 0.48\textwidth]{plots/4rn_magnification.eps}
 \caption{\label{fig:lightcurve_crescent_back} Lightcurves of backwards crescents with  a = 0.3 $R_p$.}
@@
 \begin{figure}
-\includegraphics[width = .8\textwidth]{plots/4rn2_magnification.eps}
+\centering
+	\includegraphics[width = 0.48\textwidth]{plots/4rn2_profile.eps}
+	\includegraphics[width = 0.48\textwidth]{plots/4rn2_magnification.eps}
 \caption{\label{fig:lightcurve_crescent} Lightcurve of forwards crescents with a = -0.3 $R_p$.}
 \end{figure}
+
+There are four characteristic points visible on the resulting one dimensional light profile of the crescent shaped source.
+Two of the points, p1 and p4, mark the outer boundaries of the luminous disc component.
+The other two, p2 and p3, mark the boundaries of the dark disc component.
+Since $S_1D$ is a projection of $S_2D$ on a line perpendicular to the caustic, the following relation holds:
+\begin{equation}
+	p_2-p_1 = R_p -R_n - a.
+\end{equation}
+In addition there are two other obvious relations the characteristic points which are independent of the projection:
+\begin{equation}
+	p_4 -p_1 = 2 R_p,
+\end{equation}
+
+\begin{equation}
+        p_3 -p_2 = 2 R_n.
+\end{equation}
+All four points mark the positions where derivative $\frac{dS_{1D}}{dp}$ is discontinuous.
+The points can be used to define three regions: $p_1 - p_2$ where $S_1D$ is convex, $p_2 - p_3$ where $S_1D$ is concave, and $p_3 - p_4$ where $S_1D$ is convex again.
+\\
+
+Due to the nature of the caustic and the monotonic behaviour of the magnification map on both sides of the caustic.
+, the previously mentioned characteristic points are inherited by the microlensing lightcurve.
+The points on the temporal dimension $t_1, t_2, t_3$ and $t_4$ corresnpond to instances in time when the fold is aligned with $p_1, p_2, p_3$ and $p_4$, respectively.
+For a constant relative velocity $v_p$ between the source and the caustic there is a simple relation between the points $p_i$ and instances $t_i$:
+
+\begin{equation}
+	t_j - t_i = \frac{p_j - p_i}{v_p}.
+\end{equation}
+
+With the use of the previous four equations the following identities can be written:
+
+\begin{equation}
+	R_p = \frac{v_p \left( t_4 -t_1 \right)}{2}, 
+\end{equation}
+
+\begin{equation}
+        R_n = \frac{v_p \left( t_3 -t_2 \right)}{2}, 
+\end{equation}
+
+\begin{equation}
+        a = \frac{v_p \left( t_4 +t_1 - t_3 - t_2 \right)}{2}. 
+\end{equation}
+
+\begin{figure}
+\centering
+	\includegraphics[width = 0.48\textwidth]{plots/ch_points.eps}
+        \includegraphics[width = 0.48\textwidth]{plots/ch_instances.eps}
+\caption{\label{fig:char_points} The characteristic points and instances of a crescent source's one dimensional profile and microlensing lightcurve.}
+\end{figure}
+
+\begin{figure}
+\centering
+	\includegraphics[width = 0.48\textwidth]{plots/infinite_fold_ellipsess.eps}
+\caption{\label{fig:ellipse_profile} Elliptical sources with constant and gaussian distributed brightness.}
+\end{figure}
+
+\subsection{Extension to ellipticaly shaped crescent sources}
+
+We further consider a source built from the superposition of two unalligned ellipses of constant surface brightness.
+Analogous to the previously presented crescent source the dark ellipse is interior to the bright ellipse and the distance between their centers projected on the lines perpendicular and parallel to the fold are $a$ and $b$.
+The semidimensions are ($a_p, b_p$) for thebright ellipse and ($a_n, b_n$) for the dark ellipse.
+The semimajor axis of the ellipses for the angles $\delta_p$ and $\delta_n$ with the caustic resulting in projected widths $2 h_p$ and $2h_n$ of the two ellipses on the axis perpendicular to the caustic.
+
+The two dimensional description of the source is:
+
+\begin{equation}
+S_{2D}^{CEC} = S_{2D}^{CEP} - S_{2D}^{CEN}
+\end{equation}
+
+Where
+
+\begin {equation}
+S_{2D}^{CEP} = \frac{S_0^{CEC}}{2 \pi \left(a_p b_p - a_n b_n \right)} \Theta \left( 1 - \frac{\left( \left( p -p_{sp} \right) \sin{\delta_p} + \left( q - q_{sp} \right) \cos{\delta_p} \right)^2}{a_p^2}  - \frac{\left( \left( p -p_{sp} \right) \cos{\delta_p} + \left( q - q_{sp} \right) \sin{\delta_p} \right)^2}{ b_p^2} \right),
+\end {equation}
+and
+\begin {equation}
+S_{2D}^{CEN} = \frac{S_0^{CEC}}{2 \pi \left(a_p b_p - a_n b_n \right)} \Theta \left( 1 - \frac{\left( \left( p -p_{sn} \right) \sin{\delta_n} + \left( q - q_{sn} \right) \cos{\delta_n} \right)^2}{a_n^2}  - \frac{\left( \left( p -p_{sn} \right) \cos{\delta_n} + \left( q - q_{sn} \right) \sin{\delta_n} \right)^2}{ b_n^2} \right),
+\end {equation}.
+
+Analogous to the previous three equations the following relations can be obtained:
+
+\begin{equation}
+        h_p = \frac{v_p \left( t_4 -t_1 \right)}{2}, 
+\end{equation}
+
+\begin{equation}
+        h_n = \frac{v_p \left( t_3 -t_2 \right)}{2}, 
+\end{equation}
+
+\begin{equation}
+        a = \frac{v_p \left( t_4 +t_1 - t_3 - t_2 \right)}{2}. 
+\end{equation}
+
+The time instances used in the previous equations are analogous to the ones considered in the crescent source formed from the superposition of constant surface brightness disks.
+
+Moreover we consider a source described by an elliptical gaussian brightness function shadowed by a dark ellipse. In this case the two dimensional description of the source is given by:
+
+\begin {equation}
+S_{2D}^{GEC} = \frac{S_0^{GEC}}{2 \pi a_p b_p} \exp \left( - \frac{\left( \left( p -p_{sp} \right) \sin{\delta_p} + \left( q - q_{sp} \right) \cos{\delta_p} \right)^2}{2 a_p^2}  - \frac{\left( \left( p -p_{sp} \right) \cos{\delta_p} + \left( q - q_{sp} \right) \sin{\delta_p} \right)^2}{2 b_p^2} \right) \Pi \left(p,q\right)
+\end {equation}
+
+where the function
+\begin{equation}
+\Pi \left(p,q\right) = 1 - \Theta \left( 1 - \frac{\left( \left( p -p_{sn} \right) \sin{\delta_n} + \left( q - q_{sn} \right) \cos{\delta_n} \right)^2}{2 a_n^2}  - \frac{\left( \left( p -p_{sn} \right) \cos{\delta_n} + \left( q - q_{sn} \right) \sin{\delta_n} \right)^2}{2 b_n^2} \right).
+\end{equation}
+
+\begin{figure}
+\centering
+	\includegraphics[width = 0.48\textwidth]{plots/ellipse_profile.eps}
+        \includegraphics[width = 0.48\textwidth]{plots/ellipse_lightcurve.eps}
+\caption{\label{fig:ellipse_profile} The one dimensional profile and resulting lightcurves of the elliptical sources.}
+\end{figure}

### 2014-10-05 paper1/intro.tex
figure add
small edits

@@
 Moreover, in the direction perpendicular to the accretion disk, where the medium is more transparent, jets will appear.
-When the medium between and observer and such an object is transparent a quasar can be observed (citation).
+When the medium between and observer and such an object is transparent a quasar can be observed \citep[e.g.,][]{1984RvMP...56..255B}.
 In other words an active galactic nucleus (AGN) with one of the poles orientated towards Earth would be detected as a quasar by an observer.
@@
 Through extensive observations done in the past 50 years, an important part of the assumptions and details  regarding the mechanism have been confirmed and uncovered.
-Still, the black hole shadow and the surrounding luminous accreting material (the black hole silouette) remains to be probed.
-Advancements in the very long baseline interferometer observations have made possible since 2007, the detection of structures of a few Schwarzschild radii at the Sagitarius A object at the center of our galaxy \citep{2008JPhCS.131a2055D}.
+Still, the black hole shadow and the surrounding luminous accreting material (the black hole silhouette) remains to be probed.
+Advancements in the very long baseline interferometer observations have made possible since 2007, the detection of structures of a few Schwarzschild radii of Sgr~A* at the center of our Galaxy \citep{2008JPhCS.131a2055D}.
 Furthermore, jet launching structure near the supermasve black hole in M87 were resolved with the same observation tool \citep{2012Sci...338..355D}.
@@
 A whole range of models have been applied starting from simple geometric models to more complex, physically sensible models (Doeleman et al 2008, Fish et al 2011, Broderick et al 2011, Moscibrodzka et al 2009, Dexter 2010 and others).
-A significant fraction of the later models predict a crescent shaped silouette of the black hole.
+A significant fraction of the later models predict a crescent shaped silhouette of the black hole.
 This motivated \cite{2013MNRAS.434..765K} to use a simple geometric crescent model to fit the data.
-The succesfull results encouraged the authors to speculate that the shadow of the black hole will be observed in the near fututre with the previously mentioned observation method.
+The succesful results encouraged the authors to speculate that the shadow of the black hole will be observed in the near fututre with the previously mentioned observation method.
 \\
@@
 A value that is orders of magnitude larger than the distance to M87.
-Direct observation of the black hole silouette of quasars require significant technological advances.
+Direct observation of the black hole silhouette of quasars require significant technological advances.
 We argue in the present paper
@@
 Moreover, unlike with the few nearby AGNs, the study of the very large quasar population allows for the statistical refinement of the global properties of AGN by highlighting any particular
-characteristic that is present only in SGA* and M87 central region.
+characteristic that is present only in Sgr~A* and M87 central region.
 Furthermore, by probing the event-horizon-scale inner regions of a large number of objects will facilitate the study of the evolution of the previously mentioned region.
 
-Gravitational microlensing of QSOs have been predicted as early as 1981 when \cite{1981ApJ...243..140G} suggested that ''if haloes of galaxies were composed of stars with masses less than  0.1 $M_{sol}$, then these stars acting as individual lenses whould produce fluctuations of the order of
+Gravitational microlensing of QSOs have been predicted as early as 1981 when \cite{1981ApJ...243..140G} suggested that ``if haloes of galaxies were composed of stars with masses less than  0.1 $M_\odot$, then these stars acting as individual lenses whould produce fluctuations of the order of
 unity in the intensities of the QSO images on time scales of 1 -14 years '' (Gott 1981).
@@
 
-Estimates of the Einstein radius for the particular case of a QSO at redshift 2 lensed by a compact stellar mass object at redshift 0.5 was found to be $10^-6 \sqrt{M/M_{sol}} arcsec$ \citep{2001PASA...18..207W}.
-The Schwarzschild radius of the Sgr A* black hole, situated at 8 kpc from Earth \citep{1993ARA&A..31..345R} spans an angle of 10 microarcsec on the sky \citep{2008JPhCS.131a2055D}.
-If the quasars observed at z =2 would have similar linear sizes, the aparent size of the  luminous part of the QSO would be many orders of magnitude below the 1 microarcsec level.
+Estimates of the Einstein radius for the particular case of a QSO at redshift 2 lensed by a compact stellar mass object at redshift 0.5 was found to be $10^{-6} \sqrt{M/M_\odot}\rm\,arcsec$ \citep{2001PASA...18..207W}.
+The Schwarzschild radius of the Sgr A* black hole, situated at 8~kpc from Earth \citep{1993ARA&A..31..345R} spans an angle of 10 microarcsec on the sky \citep{2008JPhCS.131a2055D}.
+If the quasars observed at $z=2$ would have similar linear sizes, the aparent size of the  luminous part of the QSO would be many orders of magnitude below the 1 microarcsec level.
 Therefore the aparent size would be significantly smaller than the corresponding Einstein radius of a solar mass lens.
@@
 Furthermore, the concept of critical curves and caustics are introduced.
-Afterwards we focus on a particular singularity, namely the fold singularity.
+Afterwards we focus on a particular catastrophe, namely the fold.
 We choose to study events related to the previously mentioned singularity since the magnification events caused by the crossing of such boundaries are most probable.
@@
 The one dimensional profile contains exhaustively all the information that can be extracted  from the lens generated light-curve \\
+
+\begin{figure}
+\includegraphics[width = .8\textwidth]{plots/infinite_fold.eps}
+\caption{\label{fig:infinite_fold}  Source profiles and magnification map for an infinite fold. Objects have the same $S_{0}$ and $r_{1/2}$ }
+\end{figure}
 

### 2015-02-03 paper1/theory.tex

@@
 
-The matrix $A$ at the coordinates of the caustic can have rank 1 or 0. If the rank is 1 then the coordinates belong to $fold$ or $cusp$ singularities. For the purpose 
-of this paper we are interested only in the behaviour around $fold$ singularities. A condition necessary to distinguish between a fold and cusp singularity is that the
+The matrix $A$ at the coordinates of the caustic can have rank 1 or 0. If the rank is 1 then the coordinates belong to {\em fold\/} or {\em cusp\/} catastrophes. For the purpose 
+of this paper we are interested only in the behaviour around {\em fold\/} catastrophe. A condition necessary to distinguish between a fold and cusp singularity is that the
 eigenvector of $A$ corresponding to the null eigenvalue is not a tangent vector of the critical curve \citep{1992A&A...260....1S}(Dominik 2008).

### 2015-02-04 paper1/intro.tex
intro polishing
Intro polishing

@@
 \section{introduction}
-More than half a century after their discovery, quasars still maintain the interest of a productive part of the astrophysical community. Although several mechanisms have been proposed 
-to explain how such systems function, a single mechanism is widely accepted at the present time.
+More than half a century after their discovery, quasars still maintain the interest of a productive part of the astrophysical community. 
+Although several mechanisms have been proposed to explain how such systems function, a single mechanism is widely accepted at the present time. 
 The model was first proposed by \cite{1964ApJ...140..796S}, followed by \cite{1964SPhD....9..246Z} in the same year and \cite{1969Natur.223..690L} five years later.
@@
 According to the paradigm, the object is powered through the accretion of matter from the proximal environment into the black hole. 
-The radiation emitted excites the surrounding medium which becomes detectable as the narrow line regions,
-broad band regions and torus.
+The radiation emitted excites the surrounding medium which becomes detectable as the narrow line regions, broad band regions and torus. 
 Moreover, in the direction perpendicular to the accretion disk, where the medium is more transparent, jets will appear.
 When the medium between and observer and such an object is transparent a quasar can be observed \citep[e.g.,][]{1984RvMP...56..255B}. 
-In other words an active galactic nucleus (AGN) with one of the poles orientated towards Earth would be detected as a quasar by an observer.
-\\
+In other words an active galactic nucleus (AGN) with one of the poles orientated towards Earth would be detected as a quasar by an observer. \\
 
@@
 
-A whole range of models have been applied starting from simple geometric models to more complex, physically sensible models (Doeleman et al 2008, Fish et al 2011, Broderick et al 2011, Moscibrodzka et al 2009, Dexter 2010 and others).
+A whole range of models have been applied starting from simple geometric models to more complex, physically sensible models (\citep{2008Natur.455...78D}, \citep{2011ApJ...738...38B}, \citep{2009ApJ...706..497M}, \citep{2010ApJ...717.1092D}).
 A significant fraction of the later models predict a crescent shaped silhouette of the black hole. 
 This motivated \cite{2013MNRAS.434..765K} to use a simple geometric crescent model to fit the data. 
-The succesful results encouraged the authors to speculate that the shadow of the black hole will be observed in the near fututre with the previously mentioned observation method.
-\\
+The succesful results encouraged the authors to speculate that the shadow of the black hole will be observed in the near fututre with the previously mentioned observation method. \\
 
 Hundreds of thousands of quasars have been detected up to date, mainly from the Sloan Digital Sky Survey.
-The great majority of them present a redshift larger than 2.15 of their spectra (Paris et al 20Nov 2013 SDSS catalogue).
+The great majority of them present a redshift larger than 2.15 of their spectra (\citep{2014A&A...563A..54P}).
 Therefore the typical distance to the objects is in the order of gigaparsecs or above. 
-A value that is orders of magnitude larger than the distance to M87.
-Direct observation of the black hole silhouette of quasars require significant technological advances.
-We argue in the present paper
- that one can probe and study the black hole shadow and its proximal environment belonging to a quasar without having to resolve sub-event horizon scales.
+A value that is orders of magnitude larger than the distance to M87. Direct observation of the black hole silhouette of quasars require significant technological advances.
+We argue in the present paper that one can probe and study the black hole shadow and its proximal environment belonging to a quasar without having to resolve sub-event horizon scales. 
 We consider that this is possible by analyzing the data from microlensing-events. 
@@
 characteristic that is present only in Sgr~A* and M87 central region. 
-Furthermore, by probing the event-horizon-scale inner regions of a large number of objects will facilitate the study of the evolution of the previously mentioned region.
+Furthermore, the probing of the event-horizon-scale inner regions of a large number of objects will facilitate the study of the evolution of the previously mentioned region.
 
-Gravitational microlensing of QSOs have been predicted as early as 1981 when \cite{1981ApJ...243..140G} suggested that ``if haloes of galaxies were composed of stars with masses less than  0.1 $M_\odot$, then these stars acting as individual lenses whould produce fluctuations of the order of
-unity in the intensities of the QSO images on time scales of 1 -14 years '' (Gott 1981).
+Gravitational microlensing of QSOs have been predicted as early as 1981 when \cite{1981ApJ...243..140G} suggested that 
+``if haloes of galaxies were composed of stars with masses less than  0.1 $M_\odot$, then these stars acting as individual lenses whould produce fluctuations of the order of 
+unity in the intensities of the QSO images on time scales of 1 -14 years '' . 
 What creates the magnification event is the relative and independent motion of the lens and source with respect to the observer.
@@
 The simplified models are introduced in the theory section of the present paper.
-In the theory (II) section of the present article with an introduction to the basic microlensing theory in which we present the general microlensing equation followed by the concept of magnification.
+The theory (II) section of the present article starts with an introduction to the basic microlensing theory in which we present the general microlensing equation followed by the concept of magnification.
 We underline that for the particular case of micro gravitational lensing,
 the images are not resolved and thus the magnifications of each images contribute to a total magnification. 
-Furthermore, the concept of critical curves and caustics are introduced.
-Afterwards we focus on a particular catastrophe, namely the fold.
+Furthermore, the concept of critical curves and caustics are introduced. Afterwards we focus on a particular catastrophe, namely the fold. 
 We choose to study events related to the previously mentioned singularity since the magnification events caused by the crossing of such boundaries are most probable.
-A quick inspection of figure 1 by the reader should motivate the previous statement.
+A quick inspection of figure \ref{fig:magnification_map} by the reader should motivate the previous statement.
 In continuation we introduce the commonly used result involving the magnification of a point-like source near a fold singularity. 
 Namely that the magnification is aproximately linear outside the caustic and decreases from infinity with the square root of the distance to the boundary when inside the caustic.  
-In the last section of the Introduction (II) we present the general equations for estimating the flux received from an extended source near a fold.
+In the last section of the theory (II) we present the general equations for estimating the flux received from an extended source near a fold. 
 The only elements required to compute the flux as a function of time are the two dimensional brightness function of the object and the magnification of a point-like source. 
-One of the core
-assumptions made when developing the simple models is that the einstein angle is orders of magnitude greater than the apparent size of the source.
-Due to this assumption one can imagine the caustic as an infinite wall to be crossed by the source as presented in figure 2.
+One of the core assumptions made when developing the simple models is that the Einstein angle is orders of magnitude greater than the apparent size of the source. 
+Due to this assumption one can imagine the caustic as an infinite wall to be crossed by the source as presented in \ref{infinite_fild}. 
 The consequence of the limit is that the information regarding the profiles of the source along any direction parallel to the boundary value is reduced to the corresponding integral value. 
@@
 The one dimensional profile contains exhaustively all the information that can be extracted  from the lens generated light-curve \\
-
-\begin{figure}
-\includegraphics[width = .8\textwidth]{plots/infinite_fold.eps}
-\caption{\label{fig:infinite_fold}  Source profiles and magnification map for an infinite fold. Objects have the same $S_{0}$ and $r_{1/2}$ }
-\end{figure}
 
@@
 In addition we use a geometric crescent-shaped source effectively the same as the one introduced by \citep{2013MNRAS.434..765K} with constant surface brightness.
-\\
-The fourth section contains the results of our simplified models calculations, including the numerical equation for lightcurve calculations and the corresponding plots.
+\\ The fourth section contains the results of our simplified models calculations, including the numerical equation for lightcurve calculations and the corresponding plots.
 In order to facilitate comparison bewteen the lightcurves of the three distinct sources, we constrain the luminosity and half-light radius of all sources to have the same value. 
+Furthermore, towards the end of the section, the microlensing light curves were generated for more complex images of black hole silhouettes. 
+Namely the images of Sgr A* and M87 compiled in the paper \citep{2015MNRAS.446.1973R} and based on simulations presented in \citep{2009MNRAS.394L.126M} article were used as sources.
 
-We continue to the second part of the paper, where we have advanced the level of complexity of our models.
-Using the Wambsganss gravitational lensing code (citation) a realistic magnification map is generated.
-The magnification map replaces the simplified magnification function used previously.
-\\
-The numerical method, hierachical tree structure and backwards reatracing, underlying the microlensing code is treated in section five.
-Aside from a basic review of the physical and numerical principles of the code, the description of its generalization from disk shaped source images to crescent ones can be found here as well.
-Further, the mass distribution in the lens plane, chosen to produce the magnification map used in this analysis, is introduced and motivated.
+We continue to the second part of the paper, where we have advanced the level of complexity of the amplification map.
+Using the Wambsganss gravitational lensing code \citep{1990LNP...360..186W} a realistic magnification map is generated.
+The numerical method, hierachical tree structure and backwards raytracing, underlying the microlensing code is treated in section five.
+Aside from a basic review of the physical and numerical principles of the code, the description of its generalization from disk shaped source images to crescent ones can be found in the Microlensing code and magnification map as well.
+Furthermore, the mass distribution in the lens plane, chosen to produce the magnification map used in this analysis, is introduced and motivated.
+
+%Irshad - you can add here your paragraphs
+
+Finally, in the Discussion (VII) section, the main results of the current work and their implications are discussed.

### 2015-02-08 paper1/discussion.tex
released draft

--- 2014-04-17 paper1/discussion.tex
@@
 \section{Discussion}
+[Mihai's contribution]\\
+In the current paper we simulate and study the resulting microlensing lightcurves of geometric crescent-shaped sources 
+and compare them with the microlensing lightcurves of other simple mathematically describable source profiles. 
+In order to mimic the behaviour of flux of light from the source in the proximity of a fold caustic we make use of the simple approximation described in equation (5). 
+The equation would exhaustively describe the magnification map and offer a good universal aproximation for the 
+particular microlensing regime that we consider. 
+Namely, the shape of the caustic boundary in the proximity of the source in the respective plane can be approximated with
+ a line due to reason that the local radius of curvature of caustic is orders of magnitude greater than the half-light
+ radius of the studied source. 
+In particular cases in which the previously mentioned approximation loses its validity the impact on the quality of the 
+lightcurves is not evenly distributed. The shape of the lightcurve will be maintained. 
+The datapoints corresponding to the source position before and during the overlapping of the caustic will be affected by
+ smaller errors than the datapoints corresponding to later times. \\
+
+The first two source profiles that we consider are the uniform disk and symmetric gaussian source. 
+Both of them can be described by a half-light radius $r_{1/2}$ and a total unlensed light flux $S_0$. 
+With the two parameters constrained the one dimensional profiles as well as the lightcurves of the two source are completely determined, since no free parameter remains.
+
+The previous statement does not hold for a crescent source.
+In the case of the crescent source there are in total five parameters: the integrated flux of the source $S_0$,
+ the radii of the bright/dark disk $R_p$/$R_n$ and the displacement of the centers of the two disks on the axes perpendicular and parallel to the caustic $a$ and $b$.
+
+Two of the parameters can be reduced by expressing the results in terms of $S_0$ and $r_{1/2}$. 
+The later being determinable for any set of parameters $R_p, R_n$ and $a^2+b^2$. Moreover, one of the displacemement 
+parameters $b$ has no impact on the one dimensional profile of the source that results from the projection of the source image on an axis perpendicular to the caustic.
+
+Since the one dimensional source profile that corresponds to an axis perpendicular to the caustic contains exhaustively all the information regarding the source that can be revealed by the lightcurve,
+the value of the parameter $b$ does not effect the shape of the lighcurve.
+
+Nevertheless the $b$ parameter is relevant to the calculation of $r_{1/2}$. 
+It's qualitative effect is to decrease the value of the half-light radius when the absolute value of the parameter is increased.  
+With two parameters constrained and another irrelevant to the shape of the lightcurve, two free parameters remain $R_n$ and $a$. 
+Figure 4 reveals that the lighcurve of a crescent source has more visible features than the other two light-curves 
+corresponding to the disc and guasssian shape. The parameters $R_n$ and $a$ have strong influences on the shape of 
+the microlensed lightcurve, as can be seen in figures 5, 6 and 7.
+Moreover, the one dimensional source profile corresponding to the direction perpendicular to the caustic reveal four characteristic points.
+The overlap of each of these points with the caustic leaves visible features on the lightcurve at the corresponding instances of time.
+In timely order the instances correspond to the start of the overlap between the caustic and the bright disk, the start of the overlap between the caustic and the dark disk, the end of the overlap between the caustic and dark disk and finally the end of the overlap between the caustic and the bright disk.
+
+With the different source profiles and their corresponding lightcurves studied we can change our point of view of the system to that of an observer.
+The observer would basically detect only the lightcurve of such as source.
+As described in section (Lightcurve of a crescent source) the timing of the onset and offset of the previously described periods can be used to estimate the values of the radii and one of the displacement parameters when assuming a geometric crescent, the halfwidths and one displacement parameters when assuming a constant brightness elliptical crescent and the halfwidth of the dark ellipse when assuming an elliptical gaussian crescent.
+All quantities can be estimated in terms of the relative velocity of the source in a direction perpendicular to the caustic.\\
+
+The previously mentioned abstract parameters can be related to physical quantities specific to the central region of a quasar.
+Thus the luminous region would correspond to the bright accretion disc that surrounds the black hole.
+The later's gravity would cause a shadow in the bright region limited by the extent of the event horizon of the black hole.
+Therefore the radius of the bright disk would provide an estimate of the size of the accretion disk and the radius of the dark disc would provide and estimate of the Schwarzschild radius of the black hole.
+Moreover a value of the Schwarzschild radius
+can be used to estimate the mass of the black hole $M \approx \frac{\Delta t_{dark}  v_p c^2}{2G}$. In the previous 
+expression the $\Delta t_{dark}$ denotes the period of overlap between the black hole shadow and fold caustic. 
+$v_p$ denotes the component of the relative velocity of the source and fold which is perpendicular to the caustic.
+Furthermore, a simulated image of M87 presented in \citep{2012MNRAS.421.1517D} (figure 9) has been microlensed.
+On the resulting lightcurve the instances corresponding to the start and end of the black hole shadow and caustic overlap were distinguishable as can be seen in figure 10.
+
+The parameters whose values cannot be determined due to the loss of information from the directions parallel to the 
+caustic could be obtained in the eventuality in which the same source crosses multiple caustics that are not parallel. 
+Multiple crossing of caustics can reveal details of the one dimensional flux profile corresponding to multiple distinct 
+directions which would allow the reconstruction of the two dimensional profile analogous to the process through which an image of a CT scan is obtained.
+
+In case of a high quality lightcurve with insignificant noise and measurement errors, the parameters can be obtained 
+by simply identifying the characterstic instances of time withought making use of the actual values of the magnification
+ map. If the effect of the errors and the noise distorts the magnification time function enough so that the 
+characteristic points are not identifiable with the characteristic periods still visible, the boundaries of the periods 
+can be roughly estimated. To study if good estimates of the parameters can still be obtained when datapoints are
+ strongly affected by the noise we propose ....
+
+[Irshad's contribution]\\
 

### 2015-02-08 paper1/theory.tex
Intro polishing
released draft

@@
   \includegraphics[width=0.95\hsize]{plots/IRIS567_path_2.eps}
-\caption{\label{fig:magnification} Magnification map}
+\caption{\label{fig:magnification_map} Magnification map}
 \end{figure}
@@
 eigenvector of $A$ corresponding to the null eigenvalue is not a tangent vector of the critical curve \citep{1992A&A...260....1S}(Dominik 2008).
+
+\begin{figure}
+\includegraphics[width = .8\textwidth]{plots/infinite_fold.eps}
+\caption{\label{fig:infinite_fold}  Source profiles for gaussian, disk, crescents (right) and magnification map for an infinite fold (left). Objects have the same $S_{0}$ and $r_{1/2}$. }
+\end{figure}
 
@@
 Due to the nature of the caustic and the monotonic behaviour of the magnification map on both sides of the caustic.
-, the previously mentioned characteristic points are inherited by the microlensing lightcurve.
-The points on the temporal dimension $t_1, t_2, t_3$ and $t_4$ corresnpond to instances in time when the fold is aligned with $p_1, p_2, p_3$ and $p_4$, respectively.
+The previously mentioned characteristic points are inherited by the microlensing lightcurve.
+The points on the temporal dimension $t_1, t_2, t_3$ and $t_4$ correspond to instances in time when the fold is aligned with $p_1, p_2, p_3$ and $p_4$, respectively.
 For a constant relative velocity $v_p$ between the source and the caustic there is a simple relation between the points $p_i$ and instances $t_i$:
@@
         \includegraphics[width = 0.48\textwidth]{plots/ch_instances.eps}
-\caption{\label{fig:char_points} The characteristic points and instances of a crescent source's one dimensional profile and microlensing lightcurve.}
+\caption{\label{fig:char_points} The characteristic points and instances of a crescent source's one dimensional profile and microlensing lightcurve. }
 \end{figure}
+
+Figure 4 reveals that the lighcurve of a crescent source has more visible features than the other two light-curves
+corresponding to the disc and guasssian shape. There are three regions where $S_{1D}$ and by inheritance the lighcurve
+ have distinguishable behaviour. The first region that would be recorded on a lightcurve plot represents the preiod
+of time when the bright disk begins to be overlapped by the caustic and stops when just before the dark disk reaches
+the caustic. During this period of time the flux of light from the source is increasingly magnified. The second
+period starts and ends with the overlapping of the dark disc. As a boundary of the two regions there is a
+distinguishable point where slope of the magnification is drastically changed. This apparent discontinuity in the
+first derivative of the magnification function is caused by the caustic amplification of the sudden drop in the $S_{1d}$
+function. During the respective period the magnification growth slows or even reverses during the first part of the
+period and starts to grow faster again as the dark disk ends its overlap with the caustic. At the point where
+the dark disc clears the caustic the growth of the magnification is infinite, which appears as a saddle point
+on the lightcurve. Next, the final period correspond to the case when the dark disc has cleared the caustic
+and the bright disc continues to overlap with the caustic. During this period the lighcurve reaches a peak
+that for most of the parameter space is global and for the rest of the parameter space local.
+At the end of the period the growth of the magnification is negative infinite. the respective point appears 
+as a second saddle point on the lightcurve. Past this point the magnification of any finite source will 
+decay in roughly the same manner. \\
+
+The impact of parameter $a$ on the shape of the lightcurve can be observed in figure 5 (a-var). Sources where 
+the center of the dark disc reaches the caustic before the center of the bright disc are characterized by a smoother
+broad peak in contrast to the cases where the center of the dark disc reaches the caustic after the brigh disk.
+In the case of the latter the instance when the dark disc reaches the caustic corresponds to larger and
+larger magnifications untill it becomes a local and even global peak. The effect of the $R_n$  parameter on the
+lighcurve is presented in figures 6 and 7. For the particular set of parameters where $R_p = R_n +a$ the third
+period of time discussed previously does not exist. Particularizing further, if the value of the radius of the dark
+disc is comparable to the value of the radius of the bright one then the position and shape of the maximum magnification
+ are strikingly different. In case the crescent reaches the caustic with the bright region first, the peak magnification
+ happens when the dark region reaches the caustic and it is characterized by a sharp variation in magnification growth,
+In the opposite case the peak appears before the end of the bright disc reaches the caustic and shape is smoother. \\
+
+\subsection{Microlensing a simulated image of M87}
 
@@
 \centering
-	\includegraphics[width = 0.48\textwidth]{plots/infinite_fold_ellipsess.eps}
-\caption{\label{fig:ellipse_profile} Elliptical sources with constant and gaussian distributed brightness.}
+        \includegraphics[width = 0.48\textwidth]{plots/M87_image.eps}
+\caption{\label{fig:M87_image} The simulated image of M87's black hole sillouette obtained by \citep{2012MNRAS.421.1517D} from a GRMHD simulation presented in \citep{2009MNRAS.394L.126M}.
+The color bar is linear and scaled to reach value 1 for the maximum brightness.
+Points p2 and p3, mark the start and end of the equivalent dark disk and are associated to the moment of their overlap with the caustic t2 and t3.}
 \end{figure}
 
-\subsection{Extension to ellipticaly shaped crescent sources}
-
-We further consider a source built from the superposition of two unalligned ellipses of constant surface brightness.
-Analogous to the previously presented crescent source the dark ellipse is interior to the bright ellipse and the distance between their centers projected on the lines perpendicular and parallel to the fold are $a$ and $b$.
-The semidimensions are ($a_p, b_p$) for thebright ellipse and ($a_n, b_n$) for the dark ellipse.
-The semimajor axis of the ellipses for the angles $\delta_p$ and $\delta_n$ with the caustic resulting in projected widths $2 h_p$ and $2h_n$ of the two ellipses on the axis perpendicular to the caustic.
-
-The two dimensional description of the source is:
-
-\begin{equation}
-S_{2D}^{CEC} = S_{2D}^{CEP} - S_{2D}^{CEN}
-\end{equation}
-
-Where
-
-\begin {equation}
-S_{2D}^{CEP} = \frac{S_0^{CEC}}{2 \pi \left(a_p b_p - a_n b_n \right)} \Theta \left( 1 - \frac{\left( \left( p -p_{sp} \right) \sin{\delta_p} + \left( q - q_{sp} \right) \cos{\delta_p} \right)^2}{a_p^2}  - \frac{\left( \left( p -p_{sp} \right) \cos{\delta_p} + \left( q - q_{sp} \right) \sin{\delta_p} \right)^2}{ b_p^2} \right),
-\end {equation}
-and
-\begin {equation}
-S_{2D}^{CEN} = \frac{S_0^{CEC}}{2 \pi \left(a_p b_p - a_n b_n \right)} \Theta \left( 1 - \frac{\left( \left( p -p_{sn} \right) \sin{\delta_n} + \left( q - q_{sn} \right) \cos{\delta_n} \right)^2}{a_n^2}  - \frac{\left( \left( p -p_{sn} \right) \cos{\delta_n} + \left( q - q_{sn} \right) \sin{\delta_n} \right)^2}{ b_n^2} \right),
-\end {equation}.
-
-Analogous to the previous three equations the following relations can be obtained:
-
-\begin{equation}
-        h_p = \frac{v_p \left( t_4 -t_1 \right)}{2}, 
-\end{equation}
-
-\begin{equation}
-        h_n = \frac{v_p \left( t_3 -t_2 \right)}{2}, 
-\end{equation}
-
-\begin{equation}
-        a = \frac{v_p \left( t_4 +t_1 - t_3 - t_2 \right)}{2}. 
-\end{equation}
-
-The time instances used in the previous equations are analogous to the ones considered in the crescent source formed from the superposition of constant surface brightness disks.
-
-Moreover we consider a source described by an elliptical gaussian brightness function shadowed by a dark ellipse. In this case the two dimensional description of the source is given by:
-
-\begin {equation}
-S_{2D}^{GEC} = \frac{S_0^{GEC}}{2 \pi a_p b_p} \exp \left( - \frac{\left( \left( p -p_{sp} \right) \sin{\delta_p} + \left( q - q_{sp} \right) \cos{\delta_p} \right)^2}{2 a_p^2}  - \frac{\left( \left( p -p_{sp} \right) \cos{\delta_p} + \left( q - q_{sp} \right) \sin{\delta_p} \right)^2}{2 b_p^2} \right) \Pi \left(p,q\right)
-\end {equation}
-
-where the function
-\begin{equation}
-\Pi \left(p,q\right) = 1 - \Theta \left( 1 - \frac{\left( \left( p -p_{sn} \right) \sin{\delta_n} + \left( q - q_{sn} \right) \cos{\delta_n} \right)^2}{2 a_n^2}  - \frac{\left( \left( p -p_{sn} \right) \cos{\delta_n} + \left( q - q_{sn} \right) \sin{\delta_n} \right)^2}{2 b_n^2} \right).
-\end{equation}
+\citep{2012MNRAS.421.1517D} have created a radiative image of M87 based on the GRMHD simulations presented in \citep{2009MNRAS.394L.126M}.
+The respective image, reproduced in figure \ref{fig:M87_image}, has been projected to a 1D profile associated to the perpendicular direction to a fold caustic approaching the image from the right.
+The projection is presented in the left panel of figure 10.
+The amplification values of the flux of light corresponding to a microlensing event are presented in the right panel of figure 10.
+In general the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions surrounding the luminous parts of the image have nonzero flux and thus are more extended than the simplified source model.
 
@@
 \centering
-	\includegraphics[width = 0.48\textwidth]{plots/ellipse_profile.eps}
-        \includegraphics[width = 0.48\textwidth]{plots/ellipse_lightcurve.eps}
-\caption{\label{fig:ellipse_profile} The one dimensional profile and resulting lightcurves of the elliptical sources.}
+	\includegraphics[width = 0.48\textwidth]{plots/M87_shape.eps}
+        \includegraphics[width = 0.48\textwidth]{plots/M87_lc.eps}
+\caption{\label{fig:M87_plots} The one dimensional profile obtained from the numerical integration along the ordonata axis of the source image (figure 9) is presented in the left subplot.
+The corresponding lightcurve associated to the 1D profile is presented in the right subplot.
+The points p2 and p3 mark the start and end of the equivalent dark disk and are associated to the moments of their overlap with the caustic: t2 and t3.}
 \end{figure}

### 2015-02-15 paper1/code.tex
Irshad edits

@@
 For the model a it was desirable to work with a caustic geometry as simple as possible.
-Therefore only two equal point like masses were used as lenses, to produce the magnification map depicted in figure \ref{fig:magnification}.
+Therefore only two equal point like masses were used as lenses, to produce the magnification map depicted in figure \ref{fig:magnification_map}.
 For a description of the parameter set which characterizes the map see (reference).
@@
 The values of the parameters are specified in pixel units corresponding to the magnification map.
-Further one needs to specify the start and end point coordinates of the path, which the center of the source image follows through the magnification map (see the depiction in figure \ref{fig:magnification}).
+Further one needs to specify the start and end point coordinates of the path, which the center of the source image follows through the magnification map (see the depiction in figure \ref{fig:magnification_map}).
 Hence the points along this path are specified through the number of timesteps for which the computation of the brightness is to be carried out.

### 2015-02-15 paper1/discussion.tex
Irshad edits

@@
 
-This work presents a toy model to put useful constraints on the interior structure of quasars.
-To carry out this analysis on real data, one needs two situations: First, a quasar which can be observed with a reasonably good telescope with high signal to noise and second, a caustic crossing event of the quasar.
+This work presents a toy model to put useful constraints on the interior structure of quasars, particularly the environment geometry of the event horizon.
+To carry out this analysis on real data, one needs two situations: First, a quasar which can be observed with a reasonably good telescope with high signal to noise ($\sim 150$) and second, a caustic crossing event of the quasar.
 These events are not rare, however, to get an isolated caustic crossing might be.
-However, as the size of the source is negligible as compared to big caustic structures in the sky, even the single fold caustic crossing is not that rare.
+As the size of the source is negligible as compared to big caustic structures in the sky, even the single fold caustic crossing is not that rare.
 
-In this toy analysis, we used very ideal data, which might not be the case with real source.
-So, we propose a more sophisticated analysis of this kind with different combinations of random and systematic noise and selection of data points for future.
+In this toy analysis, we used very ideal data, which might not be the case with real source but we contaminated it to match a reasonable signal to noise.
 
@@
 In this way, one can accurately and precisely recover all four parameters.
-
-Further, one can also study another interesting case, where the shape of the source is crescent like but the brightness of the crescent is not uniform but Gaussian.
+Using our toy analysis we conclude that it is feasible, in practice, to recover 2D shape of the source using a microlensing event and distinguish between symmetric and asymmetric sources.
 

### 2015-02-15 paper1/fitting.tex
Irshad edits

@@
-\section{Model fitting and likelihood analysis}
+\section{Model fitting and likelihood analysis}\label{sec:fitting}
 
@@
 In this section, we attempt to mock three datasets, each for crescent, uniform-disk and Gaussian disk case for a given fiducial parameters and contaminate it with systematic and random noise, and fit \& recover the model parameters.
-So broadly we have 9 cases to discuss, we name these nine cases as different permutations (couplet) of three letters C, D and G for crescent, uniform-disk and Gaussian disk case respectively.
+Broadly we have 9 cases to discuss, we name these nine cases as different permutations (couplet) of three letters C, D and G for Crescent, Uniform-Disk and Gaussian Disk case respectively.
 The first letter of the couplet represents the fit model and the second represents the dataset.
@@
 Figure \ref{fig:datafitting} shows the three datasets.
-These datasets are generated using the magnification map shown in figure \ref{fig:magnification} if the source of the respective model moves from point C to B.
+These datasets are generated using the magnification map shown in figure \ref{fig:magnification_map} if the source of the respective model moves from point C to B.
 The data assumed was very  ideal, including 250 different data points at regular interval of time.
@@
 We talk more about this adaptaion in the next section.
-Secondly, We try to fit the these datasets for a crossing of the source from a different caustic, in this case, from point A to B in figure \ref{fig:magnification}.
+Secondly, We try to fit the these datasets for a crossing of the source from a different caustic, in this case, from point A to B in figure \ref{fig:magnification_map}.
 This adds a natural systematic in the dataset and the data is no more ideal.
@@
 Three solid curves in top row of figure \ref{fig:mcmc} represents the resulting likelihood of $R_p$, the overall size of the crescent source fitted with three different datasets.
-We also show results for half light radius $R_{1/2}$ which is calculated using equation \ref{eqn:}.
+We also show results for half light radius $R_{1/2}$.
 
@@
 \item Crescent data: This is the true case for crescent model fitting.
-The red solid line represents this case in upper panels of \ref{fig:mcmc} as 1D probability density function marginalised over all other parameters.
+The red solid line represents this case in upper panels of figure \ref{fig:mcmc} as 1D probability density function marginalised over all other parameters.
 The true value of $R_p$ can be recovered within 2$\sigma$ level with an accuracy up to 10\%.
@@
 
-Therefore we conclude here that looking at the maximum likelihood one can identity the true dataset and also recover the true values of the parameter.
+Therefore we conclude here that looking at the maximum likelihood one can identify the true dataset and also recover the true values of the parameters.
 
@@
 Now one wants to extract the true shape of the source.
-To comment more specifically, lets assume three cases when the true shape of the source is: crescent, uniform disc and Gaussian disc:
+To comment more specifically, lets assume three cases when the true shape of the source is: crescent, uniform disc or Gaussian disc:
 
@@
 
-At this point, we conclude that maximum likelihood analysis can be well trusted in order to recover true shape and recovering the corresponding parameters.
+Therefor, we conclude that maximum likelihood analysis can be well trusted in order to recover true shape and recovering the corresponding parameters.
 This will provide the accurate recovery of the model parameters, however, for more precise recovery, one rely on precise measurements of the lightcurve.

### 2015-02-15 paper1/intro.tex
Irshad edits

@@
 
+In section \ref{sec:fitting}, we carried out a likelihood analysis for the parameterisation of the different source shapes and study the degeneracies involved amongst them.
+To do so, we mocked the datasets corressponding to each source shape and contaminate them with different noise components.
+Finally we run MCMC to explore the parameter space of the source shape and some nuisance parameters.
+The motivation of this part is to explore the possibility if using a simple toy model is it possible to differentiate between symmetric and asymmetric sources based on maximum likelihood analysis.
+
 Finally, in the Discussion (VII) section, the main results of the current work and their implications are discussed.

### 2015-05-12 paper1/fitting.tex
two columns?
editing

@@
 \centering
-\includegraphics[width=0.48\hsize]{plots/data_cc.eps}
-\includegraphics[width=0.48\hsize]{plots/data_dd.eps}
-\includegraphics[width=0.48\hsize]{plots/data_gg.eps}
-\includegraphics[width=0.48\hsize]{plots/burnin_cg.eps}
-\caption{\label{fig:datafitting} Three different mock datasets (in black) -- crescent, uniform-disk and Gaussian disk (in reading order).
-In each panel showing the best fit for each of the three models - crescent (red), uniform-disk (blue) and Gaussian disk (green).
-Bottom-right panel shows the $\chi^2$ of each step of the full MCMC for CG case (Gaussian disk data fitted with crescent source model)}.
+\includegraphics[width=0.9\hsize]{plots/data_cc.eps}
+\includegraphics[width=0.9\hsize]{plots/data_dd.eps}
+\includegraphics[width=0.9\hsize]{plots/data_gg.eps}
+\caption{\label{fig:datafitting} Three different mock datasets with
+  three different parameterisations fitted to each --- crescent (red),
+  uniform-disk (blue) and Gaussian disk (green).}
 \end{figure}
@@
 \centering
-  \includegraphics[width=0.48\hsize]{plots/Rp4all.eps}
-  \includegraphics[width=0.48\hsize]{plots/Rhalf4all.eps}\\
-  \includegraphics[width=0.48\hsize]{plots/Rhalf_RnRp.eps}
-  \includegraphics[width=0.48\hsize]{plots/aRp_bRp.eps}
+\includegraphics[width=0.9\hsize]{plots/burnin_cg.eps}
+\caption{\label{fig:datafitting} The $\chi^2$ of each step of the full
+  MCMC for one case.}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\hsize]{plots/Rhalf_RnRp.eps}
+  \includegraphics[width=0.9\hsize]{plots/aRp_bRp.eps}
 \caption{\label{fig:mcmc} Top panels showing the 1D probability density function for size of the source, $R_p$ (left), and half light radius of the source, $R_{1/2}$ (right).
@@
 Bottom two panels showing the 2$\sigma$ contours (or error ellipses) of the crescent model when fitting with the three different datasets.}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\hsize]{plots/Rp4all.eps}
+\caption{\label{fig:mcmc} Posterior probability distribution for size
+  of the source, $R_p$. Same color represeents the same dataset
+  whereas same line style corresponds to same model fit. In the legend
+  showing the reduced $\chi^2$ of the best fit in each case.}
 \end{figure}

### 2015-07-08 paper1/discussion.tex
+text discussion.tex

@@
 $v_p$ denotes the component of the relative velocity of the source and fold which is perpendicular to the caustic.  
+The respective velocity is an unknown, though it can be constrained on a case by case basis to an order of magnitude or even better. This would require the 
+study of the dynamics of the stellar structure which contain the gravitational lens.
+A better estimate of the relative velocity would facillitate a better estimate of the effective non-rotatating black hole mass associated to the black hole shadow.
+
 Furthermore, a simulated image of M87 presented in \citep{2012MNRAS.421.1517D} (figure 9) has been microlensed.

### 2015-07-08 paper1/discussion.tex
ch theory.tex, +text discussion.tex

--- 2015-02-15 paper1/discussion.tex
@@
 $v_p$ denotes the component of the relative velocity of the source and fold which is perpendicular to the caustic.  
+The respective velocity is an unknown, though it can be constrained on a case by case basis to an order of magnitude or even better. This would require the 
+study of the dynamics of the stellar structure which contains the gravitational lens.
+A better estimate of the relative velocity would facillitate a better estimate of the effective non-rotating black hole mass associated to the black hole shadow.
+
 Furthermore, a simulated image of M87 presented in \citep{2012MNRAS.421.1517D} (figure 9) has been microlensed.

### 2015-07-09 paper1/fitting.tex
some edits

@@
   \includegraphics[width=0.9\hsize]{plots/aRp_bRp.eps}
-\caption{\label{fig:mcmc} Top panels showing the 1D probability density function for size of the source, $R_p$ (left), and half light radius of the source, $R_{1/2}$ (right).
-Same color represeents the same dataset whereas same line style corresspond to same model fit.
-In the legend showing the reduced $\chi^2$ of the best fit in each case.
-Bottom two panels showing the 2$\sigma$ contours (or error ellipses) of the crescent model when fitting with the three different datasets.}
+\caption{\label{fig:mcmc} The 2$\sigma$ contours (or error ellipses)
+  of the crescent model when fitting with the three different
+  datasets.}
 \end{figure}

### 2015-07-09 paper1/intro.tex

@@
 We argue in the present paper that one can probe and study the black hole shadow and its proximal environment belonging to a quasar without having to resolve sub-event horizon scales. 
-We consider that this is possible by analyzing the data from microlensing-events. 
+We consider that this is possible by analyzing the data from microlensing-events as suggested by Agol and Krolik 1999 (citation).
+They argue that if the magnification function $\mu_t(p)$ from equation (6) corresponds to a perfect fold then from the function $F(t)$ present in the same equation one can extract the source profile.
+We argue in the current article that one can partially
+reconstruct the black hole silloute up to a simplified crescent geometry. The crescent fit facillitates an estimation of the black hole mass.       
 Moreover, unlike with the few nearby AGNs, the study of the very large quasar population allows for the statistical refinement of the global properties of AGN by highlighting any particular

### 2015-07-09 paper1/ms.tex
Theory edits
two columns?
some edits
affiliation changes

--- 2014-04-15 paper1/ms.tex
@@
-\documentclass[usenatbib,onecolumn]{mn2e}
+\documentclass[usenatbib]{mn2e}
 
@@
 \usepackage{graphicx}
+\usepackage{epstopdf}
+\usepackage{multicol}
 
@@
 
-\author[I. Mohammed et al.] {Irshad Mohammed\thanks{irshad@physik.uzh.ch}$^1$,
-Mihai Tomozeiu$^1$, Manuel Rabold$^1$,
+\author[Tomozeiu et al]{Mihai Tomozeiu$^{1,2}$,Irshad Mohammed$^1$, Manuel Rabold$^{1,2}$,
 \newauthor
-Prasenjit Saha,$^1$ and Joachim Wambsgans$^2$\\
-$^1${Institute for Theoretical Physics, University of Z$\ddot{u}$rich, 8057 Z$\ddot{u}$rich Switzerland}\\
-$^2${University of Heidelberg, Germany}
+Prasenjit Saha$^1$ and Joachim Wambsganss$^3$\\
+$^1${Physik-Institut, University of Zurich, Winterthurerstrasse 190,
+  8057 Zurich, Switzerland} \\
+$^2${Institute for Computational Science, University of Zurich,
+  Winterthurerstrasse 190, 8057 Zurich, Switzerland} \\
+$^3${Zentrum f\"ur Astronomie der Universit\"at Heidelberg,
+  M\"onchhofstrasse 12--14, 69120 Heidelberg, Germany}
 }

### 2015-07-09 paper1/intro.tex
first wave corrections intro.tex

--- 2015-02-15 paper1/intro.tex
@@
 
-Through extensive observations done in the past 50 years, an important part of the assumptions and details  regarding the mechanism have been confirmed and uncovered. 
-Still, the black hole shadow and the surrounding luminous accreting material (the black hole silhouette) remains to be probed. 
+Through extensive observations done in the past 50 years, an important part of the set of assumptions and details  regarding the mechanism have been confirmed and uncovered. 
+Still, the black hole shadow and the surrounding luminous accreting material (the black hole silhouette) remain to be probed. 
 Advancements in the very long baseline interferometer observations have made possible since 2007, the detection of structures of a few Schwarzschild radii of Sgr~A* at the center of our Galaxy \citep{2008JPhCS.131a2055D}.
@@
 
-A whole range of models have been applied starting from simple geometric models to more complex, physically sensible models (\citep{2008Natur.455...78D}, \citep{2011ApJ...738...38B}, \citep{2009ApJ...706..497M}, \citep{2010ApJ...717.1092D}).
+A whole range of models have been applied starting from simple geometric models to more complex, physically sensible models (\cite{2008Natur.455...78D}, \cite{2011ApJ...738...38B}, \cite{2009ApJ...706..497M}, \cite{2010ApJ...717.1092D}).
 A significant fraction of the later models predict a crescent shaped silhouette of the black hole. 
 This motivated \cite{2013MNRAS.434..765K} to use a simple geometric crescent model to fit the data. 
-The succesful results encouraged the authors to speculate that the shadow of the black hole will be observed in the near fututre with the previously mentioned observation method. \\
+The succesful results encouraged the authors to speculate that the shadow of the black hole will be observed in the near future with the previously mentioned observation method. \\
 
@@
 Therefore the typical distance to the objects is in the order of gigaparsecs or above. 
-A value that is orders of magnitude larger than the distance to M87. Direct observation of the black hole silhouette of quasars require significant technological advances.
+A value that is orders of magnitude larger than the distance to M87. The direct observations of the black hole silhouettes of quasars require significant technological advances.\\
+
 We argue in the present paper that one can probe and study the black hole shadow and its proximal environment belonging to a quasar without having to resolve sub-event horizon scales. 
-We consider that this is possible by analyzing the data from microlensing-events. 
+We consider that this is possible by analysing the data from microlensing-events as suggested by \cite{1999ApJ...524...49A}.
+They argue that if the magnification function $\mu_t(p)$ from equation (6) corresponds to a perfect fold then from the function $F(t)$ present in the same equation one can extract the source profile.
+We argue in the current article that one can partially
+reconstruct the black hole silhouette up to a simplified crescent geometry. The crescent fit facilitates an estimation of the black hole mass.       
 Moreover, unlike with the few nearby AGNs, the study of the very large quasar population allows for the statistical refinement of the global properties of AGN by highlighting any particular
 characteristic that is present only in Sgr~A* and M87 central region. 
-Furthermore, the probing of the event-horizon-scale inner regions of a large number of objects will facilitate the study of the evolution of the previously mentioned region.
+Furthermore, the probing of the event-horizon-scale inner regions of a large number of objects will facilitate the study of the evolution of the previously mentioned region. \\
 
 Gravitational microlensing of QSOs have been predicted as early as 1981 when \cite{1981ApJ...243..140G} suggested that 
-``if haloes of galaxies were composed of stars with masses less than  0.1 $M_\odot$, then these stars acting as individual lenses whould produce fluctuations of the order of 
+``if haloes of galaxies were composed of stars with masses less than  0.1 $M_\odot$, then these stars acting as individual lenses would produce fluctuations of the order of 
 unity in the intensities of the QSO images on time scales of 1 -14 years '' . 
@@
 multiple microlensing events of the same source, the QSO. 
-The sequential alignment of the source and observer with different lensing objects will result in an aparently random increase of the QSO light flux analogous to the scintilation events in a meter of radioactivity.
+The sequential alignment of the source and observer with different lensing objects will result in an apparently random increase of the QSO light flux analogous to the scintillation events in a meter of radioactivity.\\
 
@@
 In continuation we introduce the commonly used result involving the magnification of a point-like source near a fold singularity. 
-Namely that the magnification is aproximately linear outside the caustic and decreases from infinity with the square root of the distance to the boundary when inside the caustic.  
+Namely that the magnification is approximately linear outside the caustic and decreases from infinity with the square root of the distance to the boundary when inside the caustic.  
 In the last section of the theory (II) we present the general equations for estimating the flux received from an extended source near a fold. 
@@
 One of the core assumptions made when developing the simple models is that the Einstein angle is orders of magnitude greater than the apparent size of the source. 
-Due to this assumption one can imagine the caustic as an infinite wall to be crossed by the source as presented in \ref{infinite_fild}. 
+Due to this assumption one can imagine the caustic as an infinite wall to be crossed by the source as presented in \ref{fig:infinite_fold}. 
 The consequence of the limit is that the information regarding the profiles of the source along any direction parallel to the boundary value is reduced to the corresponding integral value. 
-The effect allows the introduction of a one dimensional flux function, which is the two dimensional flux integrated along the axis paralel to the boundary. 
+The effect allows the introduction of a one dimensional flux function, which is the two dimensional flux integrated along the axis parallel to the boundary. 
 The one dimensional profile contains exhaustively all the information that can be extracted  from the lens generated light-curve \\
@@
 \\ The fourth section contains the results of our simplified models calculations, including the numerical equation for lightcurve calculations and the corresponding plots.
-In order to facilitate comparison bewteen the lightcurves of the three distinct sources, we constrain the luminosity and half-light radius of all sources to have the same value. 
+In order to facilitate comparison between the lightcurves of the three distinct sources, we constrain the luminosity and half-light radius of all sources to have the same value. 
 Furthermore, towards the end of the section, the microlensing light curves were generated for more complex images of black hole silhouettes. 
@@
 Using the Wambsganss gravitational lensing code \citep{1990LNP...360..186W} a realistic magnification map is generated.
-The numerical method, hierachical tree structure and backwards raytracing, underlying the microlensing code is treated in section five.
+The numerical method, hierarchical tree structure and backwards raytracing, underlying the microlensing code is treated in section five.
 Aside from a basic review of the physical and numerical principles of the code, the description of its generalization from disk shaped source images to crescent ones can be found in the Microlensing code and magnification map as well.
@@
 
-In section \ref{sec:fitting}, we carried out a likelihood analysis for the parameterisation of the different source shapes and study the degeneracies involved amongst them.
-To do so, we mocked the datasets corressponding to each source shape and contaminate them with different noise components.
+In section \ref{sec:fitting}, we carried out a likelihood analysis for the parametrisation of the different source shapes and study the degeneracies involved amongst them.
+To do so, we mocked the datasets corresponding to each source shape and contaminate them with different noise components.
 Finally we run MCMC to explore the parameter space of the source shape and some nuisance parameters.

### 2015-07-09 paper1/theory.tex
Irshad edits
editing
ch theory.tex, +text discussion.tex
first wave corrections theory.tex

@@
 
-We start by introducing in a succint manner the gravitational lensing theory that is relevant for microlensing in general and for the scope of the
+We start by introducing in a succinct manner the gravitational lensing theory that is relevant for microlensing in general and for the scope of the
 present paper in particular. For a detailed presentation of the theory the reader is referred to the books and articles that represent the references
-of the current section (book by Schneider and others). The general microlensing equation can be written as:
+of the current section (i.e. \cite{Schneider_book}). The general microlensing equation can be written as:
 \begin{equation}
@@
 According to this equation, from an infinitesimally small source positioned at $\vec{x}$ in the lens plane an observed will receive a flux $\mu(\vec{x}) dF$ instead of the flux $dF$ that would be
-observed in the absence of the gravitational lens. In other words the source will dim or brighten depending on whether $\mu(\vec{x})$ is subunitary or supraunitary.\\
+observed in the absence of the gravitational lens. In other words the source will dim or brighten depending on whether $\mu(\vec{x})$ is sub-unitary or supra-unitary.\\
 
@@
 by critical curves where the determinant vanishes. The lens equation maps the critical curves of the lens plane into caustics in the source plane. An example of the caustic 
-shape can be seen in figure \ref{fig:magnification}.\\
+shape can be seen in figure \ref{fig:magnification_map}.\\
 
@@
 \centering
-  \includegraphics[width=0.95\hsize]{plots/IRIS567_path_2.eps}
+  \includegraphics[width=.9\hsize]{plots/IRIS567_path_2.eps}
 \caption{\label{fig:magnification_map} Magnification map}
@@
 
-The matrix $A$ at the coordinates of the caustic can have rank 1 or 0. If the rank is 1 then the coordinates belong to {\em fold\/} or {\em cusp\/} catastrophes. For the purpose 
+The matrix $A$ at the coordinates of the caustic can have rank 1 or 0.
+If the rank is 1 then the coordinates belong to {\em fold\/} or {\em cusp\/} catastrophes (\cite{1992A&A...260....1S}, \cite{2008mmc..confE...8D}).
+For the purpose
 of this paper we are interested only in the behaviour around {\em fold\/} catastrophe. A condition necessary to distinguish between a fold and cusp singularity is that the
-eigenvector of $A$ corresponding to the null eigenvalue is not a tangent vector of the critical curve \citep{1992A&A...260....1S}(Dominik 2008).
+eigenvector of $A$ corresponding to the null eigenvalue is not a tangent vector of the critical curve (\cite{1992A&A...260....1S}, \cite{2008mmc..confE...8D}).
 
 \begin{figure}
-\includegraphics[width = .8\textwidth]{plots/infinite_fold.eps}
+\includegraphics[width = .49\textwidth]{plots/infinite_fold.eps}
 \caption{\label{fig:infinite_fold}  Source profiles for gaussian, disk, crescents (right) and magnification map for an infinite fold (left). Objects have the same $S_{0}$ and $r_{1/2}$. }
@@
 In the present case of a crescent source the effective radius is dependent on the parameters $R_p$, $R_n$ and $\sqrt{a^2+b^2}\equiv c $ exclusively.
-From symmetry considerations the centroid of the source is colinear with the centers of the two disks and it is situated at a distance $d_c$ from the center of the brigth disk.
+From symmetry considerations the centroid of the source is collinear with the centers of the two disks and it is situated at a distance $d_c$ from the center of the bright disk.
 $d_c$ can be computed numerically with the use of a variation of equation \ref{eqn:s1_d}:
@@
 \begin{equation}
-\frac{S_0^C}{2} =  \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{0}^{r_{1/2}} \left[ \Theta_2  - \Theta_1  \right] p dp;
-\end{equation}
-
-\begin{equation}
-\Theta_1 = \left[ \pi - \arccos \frac{R_p^2 + p^2 -d_c^2}{2 R_p p} - \arccos \frac{R_p^2 - p^2 + d_c^2}{2 R_p d_c} \right] \Theta \left( p - R_p + d_c \right); 
-\end{equation}
-
-\begin{equation}
-\Theta_2 =  \pi - \arccos \frac{ p^2  + (c+d_c)^2 - R_n^2}{2 p \left(c + d_c \right)} \Theta \left( p + R_n - d_c -c \right). 
+\begin{aligned}
+S_0^C &=  \frac{4S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{0}^{r_{1/2}} \left[ \Theta_2  - \Theta_1  \right] p dp; \\
+\Theta_1 &= \bigg[ \pi - \arccos \frac{R_p^2 + p^2 -d_c^2}{2 R_p p} \\
+         & \phantom{= \bigg[ \pi} - \arccos \frac{R_p^2 - p^2 + d_c^2}{2 R_p d_c}
+            \bigg] \Theta \left( p - R_p + d_c \right); \\
+\Theta_2 &=  \pi - \arccos \frac{ p^2  + (c+d_c)^2 - R_n^2}
+                                {2 p \left(c + d_c \right)} \\
+         & \phantom{= \pi - } \Theta \left( p + R_n - d_c -c \right). 
+\end{aligned}
 \end{equation}
@@
 \begin{figure}
-\includegraphics[width = .8\textwidth]{plots/figure_rhalf.eps}
+\includegraphics[width = .49\textwidth]{plots/figure_rhalf.eps}
 \caption{\label{fig:geom_crescent} Geometry of crescent sources.}
@@
 The function $p_s(t)$ can be chosen to be equal to $v_p(t-t_0) + p_{s0}$. Where $p_{s0}$ is the coordinate $p$ of the source at the initial time, and $v_p$ is the component of the velocity
-along the $p$ axis. Such a modeling of the motion of the object in the source plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
-chosing the function $a(t)$ to be constant in time.
+along the $p$ axis. Such a modelling of the motion of the object in the source plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
+choosing the function $a(t)$ to be constant in time.
 
@@
 corresponding to the disc and guasssian shape. There are three regions where $S_{1D}$ and by inheritance the lighcurve
- have distinguishable behaviour. The first region that would be recorded on a lightcurve plot represents the preiod
+ have distinguishable behaviour. The first region that would be recorded on a lightcurve plot represents the period
 of time when the bright disk begins to be overlapped by the caustic and stops when just before the dark disk reaches
@@
 the center of the dark disc reaches the caustic before the center of the bright disc are characterized by a smoother
-broad peak in contrast to the cases where the center of the dark disc reaches the caustic after the brigh disk.
+broad peak in contrast to the cases where the center of the dark disc reaches the caustic after the bright disk.
 In the case of the latter the instance when the dark disc reaches the caustic corresponds to larger and
-larger magnifications untill it becomes a local and even global peak. The effect of the $R_n$  parameter on the
+larger magnifications until it becomes a local and even global peak. The effect of the $R_n$  parameter on the
 lighcurve is presented in figures 6 and 7. For the particular set of parameters where $R_p = R_n +a$ the third
@@
 The amplification values of the flux of light corresponding to a microlensing event are presented in the right panel of figure 10.
-In general the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions surrounding the luminous parts of the image have nonzero flux and thus are more extended than the simplified source model.
+In general the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions surrounding the luminous parts of the image have non-zero flux and thus are more extended than the simplified source model.
 

### 2015-07-09 paper1/code.tex
first wave code.tex

@@
 \section{Microlensing code and magnification maps}
-For the numerical computations of lightcurves for the models a and b, the microlensing code by Joachim Wambsgans was used, which is described in \cite{1999A&A...346L...5W}.
-This code is based on the hierachical tree technique to calculate the gravitational effect of the lensing masses.
-The underlying idea is that a microlensing situation can be separated into three planes, the observer, the lens plane and the source plane (assuming the thin lens approximation to be valid).
-The actual effect, namely the distortion of lightrays emitted from a distant source, on the way to the observer, is numerically computed by treating this situation from the opposite side: lightrays are shot backward from the observer to lens plane, where they become deflected according to the angle, which is given by the 2-dim.
-mass distribution in the lens plane.
+For the numerical more complex computations of lightcurves present in the second part of the article , the Microlensing code by Joachim Wambsgans was used \cite{1999A&A...346L...5W}.
+The code is based on the hierarchical tree technique to calculate the gravitational effect of the lensing masses.
+The underlying idea is that when the thin lens approximation is valid, the a microlensing distribution can be reduced to three planes, the observer plane, the lens plane and the source plane.
+The actual effect, namely the distortion of light rays emitted from a distant source, on the path to the observer, is numerically computed by through back-tracing: lightrays are shot backward from the observer to lens plane, where they become deflected according to the angle dependent on the 2D mass distribution in the lens plane.
 \begin{equation}
@@
 \end{equation}  
-For the computation of the individual deflection angles the tree idea is used.
-This means that the positions of all lensing masses are sorted into a 2-dimensional grid (in the lens plane).
-This grid is subdivided into 4 smaller squares until every cell contains only one mass.
+The choice for reversing the photon motion is motivated by computational resources considerations. 
+For the computation of the individual deflection angles the tree concept is used.
+The positions of all lensing masses are sorted into a 2-dimensional grid (in the lens plane).
+Which is subdivided into 4 smaller squares recursively until every cell contains only one mass.
 For the actual calculations of the deflection angle not every mass gravitational influence is treated equally.
-Instead it is made use of the fact that the influence of lenses on the considered lightray falls off with $r^{-1}$ from the point where the ray hits the lens plane.
-Hence masses situated at higher distances, can be grouped together in the calculation of their gravitational potential, by multipole moments.
-This way the amount of computation time is reduced.
+Instead it is made use of the fact that the influence of lenses on the considered light ray falls off with $r^{-1}$ from the point where the ray hits the lens plane.
+Hence masses situated at higher distances, can be grouped together in the calculation of their gravitational potential, using multipole moments.
+With this approach the amount of computation time is reduced.
 \\
+
+%I don't think this part is necessary
+
 When the deflection angles are determined, the deflected lightrays are traced to the source plane and a magnification map is created by counting the number of rays which hit each of its pixels.
@@
 This is done by treating the problem from the opposite direction, namely by using the observers location as starting point and then shooting the lightrays backwards till they hit the source plane.
+% up to here (would remove it)
+
 The result is a pixel map of the number of lightrays which arrive at the source plane.
@@
 The code thereby convolves the function describing the shape of the source with the magnification pattern of the map for each time step (see equation \ref{eqn:ft2d}).
-Note that in the most observational cases in microlensing, both lens and source are moving and further that the lens configuration, and with it the magnification pattern, is changing with time.
+
+%this is problematic
+ Note that in most observational cases in microlensing, both lens and source are moving and further that the lens configuration, and with it the magnification pattern, is changing with time.
 While the first subtlety is taken care of by a coordinate transformation in this analysis, for the second one the lens configuration is assumed to be constant in time.
@@
 For the model a it was desirable to work with a caustic geometry as simple as possible. 
-Therefore only two equal point like masses were used as lenses, to produce the magnification map depicted in figure \ref{fig:magnification_map}.
-For a description of the parameter set which characterizes the map see (reference).
-The magnification maps calculated by the microlensing code are in units of Einstein radii, hence they are principally scale free.
+% up to here   (would remove it)
+
+When generating the caustic map used in our analysis presented in the second part of the article only two equal point like masses were used as lenses, to produce the magnification map depicted in figure \ref{fig:magnification_map}.
+The magnification maps calculated by the Microlensing code are in units of Einstein radii, hence they are scale free.
 Physical scale is introduced to them by multiplying the pixels by the correct distance factor.
@@
 \\
-For the purpose of this analysis it was desirable to mimic the analytical behavior of a simple fold as much as possible, for comparing the numerical result with the analytical one, therefore the path of the source was chosen, so that it intersects the border of the caustic perpendicularly, and on a point where the border is almost straight.
+For the purpose of this analysis it was desirable to mimic the analytical behaviour of a simple fold as much as possible, for comparing the numerical result with the analytical one, therefore the path of the source was chosen, so that it intersects the border of the caustic perpendicularly, and on a point where the border is a fold caustic.

### 2015-07-10 paper1/fitting.tex
editing

@@
 \section{Model fitting and likelihood analysis}\label{sec:fitting}
+
+In this section we simulate the fitting of source models to
+lightcurves in the presence of random and systematic errors.  We
+generate mock data for three source models --- a crescent, a uniform
+disc and a Gaussian disc --- and then fit each mock light curve to all
+three source models.  We denote the possible cases letter pairs, with
+the first letter denoting the model and the second letter representing
+the source: thus CG means a {\em C\/}rescent model was fitted to mock
+data from a {\em G\/}aussian source, DC stands for a uniform {\em
+  D\/}isc model fitted to mock data from a {\em C\/}rescent source,
+and so on.
 
@@
 
+\subsection{Mock datasets}
+
+Mock light curves were generated by running a source across a
+caustic, specifically along the line from point C to point B in
+Figure~\ref{fig:magnification_map}.  The model light curves, to which
+these are fitted, are made by running a source across a different
+caustic, along the line from point A to point B in the same figure.
+That is, the mock data are generated with a clean but not ideal fold
+caustic and then fitted with another such caustic.  This mimics the
+unavoidable systematic error of not knowing the caustic exactly.
+
+The crescent source had parameters (as explained in
+\S\ref{subsec:crescent}) $R_p$ and $R_n$ being the radii of the outer
+and inner circles and $(a,b)$ being the coordinates of the inner
+circle with respect to the centre of the outer circle.  The parameter
+values were
+\begin{equation}
+   (R_p, R_n, a, b) = (50.0, 30.0, 15.0, 10.0).
+\label{eqn:cp}
+\end{equation}
+The uniform disc had the same (outer) radius as the crescent.  In the
+Gaussian source, we set $3\sigma=50$.  For convenience, below we will
+refer to $R_p$ of the Gaussian source, by which we actually mean
+$3\sigma$.
+
+Each mock light curve also has three nuisance parameters, namely the
+beginning and end of the event and the brightness normalization.
+These are to be marginalised out by the MCMC.
+
+Figure~\ref{fig:datafitting} shows the three light curves.  Each has
+250 points regularly spaced in time, with Gaussian noise at the level
+of 10\% of the current brightness.  The effective signal to noise can
+be considered to be $0.1\times\sqrt{250}\simeq150$.
+
+\subsection{Fitting}
+
+Parameter fitting and marginalisation was done by Markov-chain Monte
+Carlo (MCMC).  Figure~\ref{fig:burnin} shows the progress of reduction
+of $\chi^2$.
+
 \begin{figure}
@@
 \includegraphics[width=0.9\hsize]{plots/burnin_cg.eps}
-\caption{\label{fig:datafitting} The $\chi^2$ of each step of the full
-  MCMC for one case.}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.9\hsize]{plots/Rhalf_RnRp.eps}
-  \includegraphics[width=0.9\hsize]{plots/aRp_bRp.eps}
-\caption{\label{fig:mcmc} The 2$\sigma$ contours (or error ellipses)
-  of the crescent model when fitting with the three different
-  datasets.}
+\caption{\label{fig:burnin} The $\chi^2$ of each step of the full MCMC
+  for one case.}
 \end{figure}
@@
 \end{figure}
-
-In this section, we attempt to mock three datasets, each for crescent, uniform-disk and Gaussian disk case for a given fiducial parameters and contaminate it with systematic and random noise, and fit \& recover the model parameters.
-Broadly we have 9 cases to discuss, we name these nine cases as different permutations (couplet) of three letters C, D and G for Crescent, Uniform-Disk and Gaussian Disk case respectively.
-The first letter of the couplet represents the fit model and the second represents the dataset.
-For example, CG represents the case where we fit Gaussian disk data to a crescent model.
-Similarly, DC respresents the case when crescent data is being fit with uniform-disk model.
-We explored the parameter space and carried out a likelihood analysis on all three parameter sets using Markov-chain Monte-Carlo (MCMC).
-
-\subsection{Mock datasets and model parameters}
-The datasets for each source model is in the form of light curves, the brightness of the source observed at different time.
-Figure \ref{fig:datafitting} shows the three datasets.
-These datasets are generated using the magnification map shown in figure \ref{fig:magnification_map} if the source of the respective model moves from point C to B.
-The data assumed was very  ideal, including 250 different data points at regular interval of time.
-We added Gaussian random noise to this dataset of width 10\% of the brightness at the respective time, which makes the signal to noise of about $\sim$150.
-After adding random noise, the three datasets are looking highly contaminated and no clear intuitive behaviour of fit can be seen.
-
-The fiducial model assumed to generate these mock datasets are:
-\begin{equation}
-	[R_p, R_n, a, b] = [50.0, 30.0, 15.0, 10.0]  \ \ \ \rm{for\ Crescent\ model},
-	\label{eqn:cp}
-\end{equation}
-\\
-where, $R_p$ and $R_n$ are the radius of outer and inner circles of the crescent respectively and $a$ and $b$ are the coordinates of the inner circle center with center of outer circle as the origin.
-
-\begin{equation}
-	R_p = 50.0 \ \ \ \rm{for\ Uniform\ disc\ model}
-\end{equation}
-
-\begin{equation}
-	R_p = 50.0 \ \ \ \rm{for\ Gaussian\ disk\ model}
-\end{equation}
-\\
-where, $R_p$ being the size of the uniform disc and Gaussian disc source model.
-We also have some nuisance parameters in all these cases: the beginning of the event (I), the end of the event (E), normalisation of the lightcurve (N).
-We marginalised over all these nuisance parameters.
-When we kept these three nuisance parameters free to vary in MCMC, it adapts to the amount of data for which the maximum likelihood is most optimal to find out.
-We talk more about this adaptaion in the next section.
-Secondly, We try to fit the these datasets for a crossing of the source from a different caustic, in this case, from point A to B in figure \ref{fig:magnification_map}.
-This adds a natural systematic in the dataset and the data is no more ideal.
-With these three characterstics -- random noise, systematic noise and nuisance parameters -- we make the dataset very general and highly contaminated to mock the real observations.
-
-The aim of this exercise was to explore the degeneracies in  various model parameters and recover the correct fiducial model assumed from a highly contaminated dataset.
-Also, if it is possible that one can simply distinguish between circularly symmetric and asymmetric sources or uniform and Gaussian sources on the basis of maximum likelihoods.
 
@@
 This will provide the accurate recovery of the model parameters, however, for more precise recovery, one rely on precise measurements of the lightcurve.
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\hsize]{plots/Rhalf_RnRp.eps}
+  \includegraphics[width=0.9\hsize]{plots/aRp_bRp.eps}
+\caption{\label{fig:mcmc} The 2$\sigma$ contours (or error ellipses)
+  of the crescent model when fitting with the three different
+  datasets.}
+\end{figure}

### 2015-07-10 paper1/code.tex
edited the mcmc part

@@
 \section{Microlensing code and magnification maps}
-For the numerical more complex computations of lightcurves present in the second part of the article , the Microlensing code by Joachim Wambsgans was used \cite{1999A&A...346L...5W}.
-The code is based on the hierarchical tree technique to calculate the gravitational effect of the lensing masses.
-The underlying idea is that when the thin lens approximation is valid, the a microlensing distribution can be reduced to three planes, the observer plane, the lens plane and the source plane.
-The actual effect, namely the distortion of light rays emitted from a distant source, on the path to the observer, is numerically computed by through back-tracing: lightrays are shot backward from the observer to lens plane, where they become deflected according to the angle dependent on the 2D mass distribution in the lens plane.
+
+We now go beyond the case of an ideal fold, and consider microlensing
+by random distributions of point lenses.  This requires numerical
+computation of lightcurves, for which we use the microlensing code by
+\cite{1999A&A...346L...5W}. The code is based on the hierarchical tree
+technique to calculate the gravitational effect of the lensing
+masses. The underlying idea is that when the thin lens approximation
+is valid, the a microlensing distribution can be reduced to three
+planes, the observer plane, the lens plane and the source plane. The
+actual effect, namely the distortion of light rays emitted from a
+distant source, on the path to the observer, is numerically computed
+by through back-tracing: lightrays are shot backward from the observer
+to lens plane, where they become deflected according to the angle
+dependent on the 2D mass distribution in the lens plane.
 \begin{equation}

### 2015-07-10 paper1/fitting.tex
edited the mcmc part

@@
 disc and a Gaussian disc --- and then fit each mock light curve to all
-three source models.  We denote the possible cases letter pairs, with
-the first letter denoting the model and the second letter representing
-the source: thus CG means a {\em C\/}rescent model was fitted to mock
-data from a {\em G\/}aussian source, DC stands for a uniform {\em
-  D\/}isc model fitted to mock data from a {\em C\/}rescent source,
-and so on.
+three source models. Parameter fitting and marginalisation was done by
+Markov-chain Monte Carlo (MCMC).
 
@@
 \includegraphics[width=0.9\hsize]{plots/data_gg.eps}
-\caption{\label{fig:datafitting} Three different mock datasets with
+\caption{\label{fig:mockdata} Three different mock datasets with
   three different parameterisations fitted to each --- crescent (red),
@@
 
-Figure~\ref{fig:datafitting} shows the three light curves.  Each has
+Figure~\ref{fig:mockdata} shows the three light curves.  Each has
 250 points regularly spaced in time, with Gaussian noise at the level
@@
 
-\subsection{Fitting}
+\subsection{Fitting to mock data sets}
 
-Parameter fitting and marginalisation was done by Markov-chain Monte
-Carlo (MCMC).  Figure~\ref{fig:burnin} shows the progress of reduction
-of $\chi^2$.
+Each of the three mock light curves was fitted to all three models.
+We denote the nine possible cases with letter pairs, with the first
+letter denoting the model fitted and the second letter representing
+the source: thus CG means a {\em C\/}rescent model was fitted to mock
+data from a {\em G\/}aussian source, DC stands for a uniform {\em
+  D\/}isc model fitted to mock data from a {\em C\/}rescent source,
+and so on.
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\hsize]{plots/Rp4all.eps}
+\caption{\label{fig:mcmc} Posterior probability distribution for the
+  source size $R_p$.  The vertical line is the correct value. Same
+  color represeents the same dataset whereas same line style
+  corresponds to same model fit.  The legend gives the reduced
+  $\chi^2$ of the best fit in each case.}
+\end{figure}
+
+Figure~\ref{fig:mcmc} shows the posterior probability distribution of
+$R_p$ for all nine cases, along with the minimum reduced $\chi^2$ for
+each case.  The area under each curve is unity. Note that the height
+of the curves are not the likelihood, they are probability densities
+in parameter space.
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\hsize]{plots/Rhalf_RnRp.eps}
+  \includegraphics[width=0.9\hsize]{plots/aRp_bRp.eps}
+\caption{\label{fig:crescentfit} The 2$\sigma$ contours (or error ellipses)
+  of the crescent model when fitting with the three different
+  datasets.}
+\end{figure}
+
+Let us first consider the three cases where a crescent model was
+fitted.  These are the solid curves in Figure~\ref{fig:mcmc}, with the
+colours of the curves indicating the source.  Meanwhile,
+Figure~\ref{fig:crescentfit} shows $2\sigma$ of the inferred parameter
+values.
+
+\begin{enumerate}
+
+\item[1) {\bf CC}:] solid red curve in Figure~\ref{fig:mcmc} and red
+  ellipses in Figure~\ref{fig:crescentfit}.  In this case, a light
+  curve from a crescent source was being fitted to a crescent model.
+  The fit gives reduced $\chi^2$ close to unity, as expected.  The
+  recovered parameter values are near or slightly outside the
+  $2\sigma$ ellipses.  This is expected, since we added a small
+  systematic error, by using different caustics (though both clean
+  folds) for the mock data and the model fit.  The apparent degeneracy
+  between $a$ and $b$, seen in Figure~\ref{fig:crescentfit}, is also
+  expected, since only the distance of the crescent's small circle
+  from the caustic influences the magnification.
+
+\item[2) {\bf CD}:] solid blue curve and blue ellipses.  Here a
+  crescent model was fitted to a light curve from a uniform disc.  The
+  situation is formally a crescent with $R_n=0$ and $a,b$ arbitrary,
+  and this shows in the blue ellipses in the recovered parameter
+  values.  The redundant parameters $a$ and $b$, in effect, allow the
+  model to partly fit the noise, and hence the $\chi^2$ is somewhat
+  lower than in the previous case.
 
@@
 
-\begin{figure}
-\centering
-  \includegraphics[width=0.9\hsize]{plots/Rp4all.eps}
-\caption{\label{fig:mcmc} Posterior probability distribution for size
-  of the source, $R_p$. Same color represeents the same dataset
-  whereas same line style corresponds to same model fit. In the legend
-  showing the reduced $\chi^2$ of the best fit in each case.}
-\end{figure}
+\item[3) {\bf CG}:] Solid green curve and green ellipses.
+  Additionally, Figure~\ref{fig:burnin} shows the progress of
+  reduction of $\chi^2$ in this case, where a crescent-source model is
+  used to fit data from a Gaussian source.  The parameter values are
+  seemingly tightly constrained, but the recovered $R_p$ is completely
+  wrong.  The reduced $\chi^2$ is only 0.85, indicative of
+  over-fitting.  We can see what has happened from the red curve in
+  the bottom panel of Figure~\ref{fig:mockdata}.  The best-fit model
+  covers only a small part of the light curve.  That is, the fitting
+  procedure has exploited the nuisance parameters to find a good fit
+  to mainly noise.
 
-\subsection{Fitting with crescent model}
+\end{enumerate}
 
-The main results of the MCMC are shown in figure \ref{fig:mcmc}.
-In the top row, we show the nine probability density function of recovered size of the source ($R_p$) and its half light radius ($R_{1/2}$) (left and right respectively).
-The legend of top-left panel has information about the datasets, fitting model and best fit also applicable to top-right panel for $R_{1/2}$.
-The name of the legend has the same convention as explained in the previous section.
-The number accompanying the name of the legend is the best fit reduced $\chi^2$.
-We follow the convention of red, blue and green color for crescent, uniform disc and Gaussian disk datasets respectively whereas the fitting model is represented by different kind of curves: solid, dashed and dot-dashed respectively.
-
-First we explored the parameter space for crescent source model and run its likelihood exploration for three different datasets.
-Here we have four interesting parameters as given by equation \ref{eqn:cp} and we marginalise over all other nuisance parameters.
-Three solid curves in top row of figure \ref{fig:mcmc} represents the resulting likelihood of $R_p$, the overall size of the crescent source fitted with three different datasets.
-We also show results for half light radius $R_{1/2}$.
+Next we consider the three cases where a disc model is fitted.  These
+correspond to he dashed curves in Figure~\ref{fig:mcmc}.  There is
+only one interesting parameter to fit, the disc radius $R_p$, and no
+equivalent of Figure~\ref{fig:crescentfit} is needed.
 
@@
 
-\item Crescent data: This is the true case for crescent model fitting.
-The red solid line represents this case in upper panels of figure \ref{fig:mcmc} as 1D probability density function marginalised over all other parameters.
-The true value of $R_p$ can be recovered within 2$\sigma$ level with an accuracy up to 10\%.
-Also the best fit to the data is shown in red line in upper-left panel of figure \ref{fig:datafitting}.
-It is a very good fit which also represents a reduced $\chi^2$ close to unity (1.04).
-The orange contour in the lower panels of figure \ref{fig:mcmc} showing the 2D marginalised error ellipses.
-There is a certain degeneracy in $R_{1/2}$ and $R_n/R_p$ parameters but all the values are well recovered.
+\item[4) {\bf DC}:] red dashed curve.  On fitting a disc model to a
+  crescent-source source, the recovered $R_p$ is incorrect, but the
+  reduced $\chi^2$ of 1.29 signals that the data reject the model.
 
-\item Uniform disk data: In this case, we tried to fit crescent model to disk data.
-The solid blue line in top row panels and cyan contours in bottom row panel represent this case.
-The best fit reduced $\chi^2$ (0.90) is better than the previous case as the data is generated by one parameter only and fitting with four parameter that can also model the noise.
-This can also be seen in the best fit to the data, figure \ref{fig:datafitting} red line in top-right panel, where the curve seems to fit the noise as well.
-The maximum likelihood (and also average likelihood) is achieved close to $R_n = 0$, as already mentioned, this is intuitive as well.
-The other paramters, $a$ and $b$ are quite arbitary and highly uncertain.
+\item[5) {\bf DD}:] red dashed curve.  When the correct model is
+  fitted to a disc-source, the best fit $\chi^2=0.91$ is good and
+  $R_p$ is recovered within the uncertainty estimate.
 
-\item Gaussian disk data: In this case, Gaussian disc data is fited with crescent model.
-The solid green line in top row panels and green contours in bottom row panels of figure \ref{fig:mcmc} represent this case.
-By looking at figure \ref{fig:mcmc}, it can be seen that all the parameters can be constrained really well, even the best fit $\chi^2$ is much better than other two cases (0.85) which also shows a bit of over-fitting problem.
-However, this is misleading.
-The reason for these strong constraints is that the full data is not at all fitted, in this case the nuisance parameters are rejecting most of the data during the progress of MCMC.
-This can be seen in figure \ref{fig:datafitting} bottom-left panel in red line, where the best fit model is fitted only in very small range of data, even before the actual incidence happened.
-In other words, MCMC is unable to converge in this case and the only way to achieve better likelihood is to reject data points.
-This can be seen in bottom-left panel of figure \ref{fig:datafitting}.
-The $\chi^2$ keep on reducing through the span of MCMC chain and not saturating anywhere.
+\item[6) {\bf DG}:] green dashed curve.  On fitting a disc model to a
+  Gaussian-source model, the recovered $R_p$ is incorrect, but the
+  reduced $\chi^2$ gives no signal that something is amiss.  It
+  appears that a Gaussian source could mimic a uniform disc.
 
@@
 
-Therefore we can conclude that by fitting with crescent source model, one can recover the parameters very well if the dataset is generated by crescent source model.
-The uniform disk source model can also be recovered very well for the overall shape of the disk, as the latter is the special case of the former, but other parameters like $a$ and $b$ are arbitary and $R_n \sim 0$.
-No convergence can be achieved for data generated by Gaussian disk model.
-
-Please notice that in top row of figure \ref{fig:mcmc}, the height of the peaks does not represents the likelihood, it is just ensuring that the area under the curve is unity.
-The likelihood can be estimated from the number in the legend which is the best fit $\chi^2$.
-
- In the bottom rows we show three contours, for three different datasets (same color convention) fitted with crescent parameters only because this is the only case with more than one relevant parameter.
-Note that we plotted the 2D ellipses with principal component analysis of the MCMC and this is the reason they are perfect ellipses and not arbitary shape.
-
-\subsection{Fitting with uniform disc model}
-In this case, we have only one interesting parameter to fit, the outer radius of the uniform disc with uniform intensity.
-When fitting with the true case, the uniform disk data itself (dashed blue line in upper panels of figure \ref{fig:mcmc}), the best fit $\chi^2$ is very good (0.91) and also it recovers the true value of $R_p=50$.
-When fitting with crescent data (dashed red line), the maximum likelihood shifts towards higher values of $R_p$ as compared to the fiducial value.
-This is justified and intuitive as $R_n \sim > 0$.
-Also the best fit reduced $\chi^2$ is 1.29 which is much worse than the true case.
-However, when fitting with Gaussian disc data (dashed green line), the best fit reduced $\chi^2$ is still very low but still worse than the true case.
-However, it can be seen in figure \ref{fig:datafitting} (blue line bottom-left panel) that in this case also many points are rejected in order to gain better likelihood.
-
-Therefore we conclude here that looking at the maximum likelihood one can identify the true dataset and also recover the true values of the parameters.
-
-\subsection{Fitting with Gaussian disc model}
-In the last case, we fit Gaussian disc model to the three datasets (dot-dashed curves in top panels of figure \ref{fig:mcmc}).
-In this case, when fitted with the true case, Gaussian disc dataset, the true parameters can be recovered at the correct value and also the best fit reduced $\chi^2$ is close to unity (0.98).
-In all other cases the maximum likelihood is much worse and also the recovered values of $R_p$ are many $\sigma$ away.
-
-Hence, in this case also by looking at the maximum likelihood true case can be identified and parameters can be restored correctly.
-
-\subsection{Conclusions and interpretations}
-The main results of this work is given in figure \ref{fig:mcmc} and described in the previous section. Here we try to interpret those results and conclude the take away message.
-
-Lets assume that an ideal telescope observes a quasar for a long time and we get the lightcurve  and one can clearly see the event of caustic crossing in the light curve.
-Now one wants to extract the true shape of the source.
-To comment more specifically, lets assume three cases when the true shape of the source is: crescent, uniform disc or Gaussian disc:
+Finally, we consider the three cases where a Gaussian model is fitted.
+These correspond to he dotted curves in Figure~\ref{fig:mcmc}.  Again,
+there is only one interesting parameter to fit, the Gaussian
+$3\sigma$-radius, which we have called $R_p$.
 
@@
 
-\item Crescent shape source: if one tries to fit this curve with crescent model, the maximum likelihood will be good and one can recover the true parameters with good precision.
-In the other two cases, the maximum likelihood will be bad.
-Hence this shape can be recovered very well by trusting the maximum likelihood analysis.
-\item Uniform disc shape source: if one tries to fit this curve with Gaussian disk model, the likelihood will be very little to disprove it, however, the other two models are inditinguishable with nearly the same likelihood.
-At this point it does not matter as one can obtain the true result from both uniform disk and crescent model.
-\item Gaussian disc shape source: if one tries to fit this curve with uniform disk model or crescent model, again likelihood will discard this choice.
-Again the maximum likelihood analysis will point towards the best fitted true model.
+\item[7) {\bf GC}:] red dotted curve.  When a Gaussian-source model is
+  fitted to data from a crescent source, the recovered $R_p$ is wrong
+  but the reduced $\chi^2=1.29$ shows the data rejecting the model.
+
+\item[8) {\bf GD}:] blue dotted curve.  When a Gaussian-source model
+  is fitted to data from a uniform disc, the recovered $R_p$ is wrong
+  but the model is rejected anyway.
+
+\item[9) {\bf GG}:] green dotted curve.  When data from a Gaussian is
+  fitted to the correct model, $R_p$ is recovered within its estimated
+  uncertainty, and the best fit reduced $\chi^2$ is close to unity.
+
 \end{enumerate}
 
-Therefor, we conclude that maximum likelihood analysis can be well trusted in order to recover true shape and recovering the corresponding parameters.
-This will provide the accurate recovery of the model parameters, however, for more precise recovery, one rely on precise measurements of the lightcurve.
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.9\hsize]{plots/Rhalf_RnRp.eps}
-  \includegraphics[width=0.9\hsize]{plots/aRp_bRp.eps}
-\caption{\label{fig:mcmc} The 2$\sigma$ contours (or error ellipses)
-  of the crescent model when fitting with the three different
-  datasets.}
-\end{figure}
+The above results suggest the following strategy for fitting a
+lightcurve from an unknown source profile: first try a Gaussian-source
+model; if the data reject that model, try a uniform disc; if the
+uniform disc is also rejected by the data, try a crescent model.  Our
+numerical experiments indicate --- assuming one of the three source
+models is correct --- that the reduced $\chi^2$ would unmask the
+correct model, and its parameters would be correctly recovered.

### 2015-07-11 paper1/fitting.tex
editing

@@
 \centering
-\includegraphics[width=0.9\hsize]{plots/data_cc.eps}
-\includegraphics[width=0.9\hsize]{plots/data_dd.eps}
-\includegraphics[width=0.9\hsize]{plots/data_gg.eps}
+\includegraphics[width=0.9\hsize, bb=0 0 576 432]{plots/data_cc.eps}
+\includegraphics[width=0.9\hsize, bb=0 0 576 432]{plots/data_dd.eps}
+\includegraphics[width=0.9\hsize, bb=0 0 576 432]{plots/data_gg.eps}
 \caption{\label{fig:mockdata} Three different mock datasets with
@@
 \centering
-  \includegraphics[width=0.9\hsize]{plots/Rp4all.eps}
+  \includegraphics[width=0.9\hsize, bb=0 0 576 432]{plots/Rp4all.eps}
 \caption{\label{fig:mcmc} Posterior probability distribution for the
@@
 \centering
-  \includegraphics[width=0.9\hsize]{plots/Rhalf_RnRp.eps}
-  \includegraphics[width=0.9\hsize]{plots/aRp_bRp.eps}
+  \includegraphics[width=0.9\hsize, bb=0 0 576 432]{plots/Rhalf_RnRp.eps}
+  \includegraphics[width=0.9\hsize, bb=0 0 576 432]{plots/aRp_bRp.eps}
 \caption{\label{fig:crescentfit} The 2$\sigma$ contours (or error ellipses)
@@
 \centering
-\includegraphics[width=0.9\hsize]{plots/burnin_cg.eps}
+\includegraphics[width=0.9\hsize, bb=0 0 576 432]{plots/burnin_cg.eps}
 \caption{\label{fig:burnin} The $\chi^2$ of each step of the full MCMC

### 2015-07-12 paper1/code.tex
editing

@@
 computation of lightcurves, for which we use the microlensing code by
-\cite{1999A&A...346L...5W}. The code is based on the hierarchical tree
-technique to calculate the gravitational effect of the lensing
-masses. The underlying idea is that when the thin lens approximation
-is valid, the a microlensing distribution can be reduced to three
-planes, the observer plane, the lens plane and the source plane. The
-actual effect, namely the distortion of light rays emitted from a
-distant source, on the path to the observer, is numerically computed
-by through back-tracing: lightrays are shot backward from the observer
-to lens plane, where they become deflected according to the angle
-dependent on the 2D mass distribution in the lens plane.
-\begin{equation}
-\vec{\alpha}_{i}=\frac{4G}{c^{2}} \frac{D_{LS}}{D_{S}}\sum_{j=1}^{n}M_j \frac{\vec{r}_{ij}}{r^2_{ij}} 
-\end{equation}  
-The choice for reversing the photon motion is motivated by computational resources considerations. 
-For the computation of the individual deflection angles the tree concept is used.
-The positions of all lensing masses are sorted into a 2-dimensional grid (in the lens plane).
-Which is subdivided into 4 smaller squares recursively until every cell contains only one mass.
-For the actual calculations of the deflection angle not every mass gravitational influence is treated equally.
-Instead it is made use of the fact that the influence of lenses on the considered light ray falls off with $r^{-1}$ from the point where the ray hits the lens plane.
-Hence masses situated at higher distances, can be grouped together in the calculation of their gravitational potential, using multipole moments.
-With this approach the amount of computation time is reduced.
-\\
+\cite{1999A&A...346L...5W}.
+
+The code is based on the hierarchical tree technique to calculate the
+gravitational effect of the lensing masses. The underlying idea is
+that when the thin lens approximation is valid, the a microlensing
+distribution can be reduced to three planes, the observer plane, the
+lens plane and the source plane. The actual effect, namely the
+distortion of light rays emitted from a distant source, on the path to
+the observer, is numerically computed by through back-tracing:
+lightrays are shot backward from the observer to lens plane, where
+they become deflected according to the angle dependent on the 2D mass
+distribution in the lens plane.  The choice for reversing the photon
+motion is motivated by computational resources considerations.  For
+the computation of the individual deflection angles the tree concept
+is used. The positions of all lensing masses are sorted into a
+2-dimensional grid (in the lens plane). Which is subdivided into 4
+smaller squares recursively until every cell contains only one
+mass. For the actual calculations of the deflection angle not every
+mass gravitational influence is treated equally. Instead it is made
+use of the fact that the influence of lenses on the considered light
+ray falls off with $r^{-1}$ from the point where the ray hits the lens
+plane. Hence masses situated at higher distances, can be grouped
+together in the calculation of their gravitational potential, using
+multipole moments. With this approach the amount of computation time
+is reduced. \\
 
@@
 
-When the deflection angles are determined, the deflected lightrays are traced to the source plane and a magnification map is created by counting the number of rays which hit each of its pixels.
-This is the inverse of what would happen in a real observation: here an observer on earth would detect light emitted by several sources with her/his telescope.
-Each of these sources emits lightrays which are subject to the deflection by the intervening gravitational potential of the lensing mass.
-However only a fraction of those will end in the observers telescope on earth.
-Interesting for the analysis are hence those lightrays which are originating by the source of interest and further the subset of these which actually converge at the location of the observer.
-So instead of calculating the trajectories of lightrays emitted by the source in all directions and then succeedingly pick out the ones which end in the telescope of the observer, the computational effort can be restricted to these lightrays of interest in the first place.
-This is done by treating the problem from the opposite direction, namely by using the observers location as starting point and then shooting the lightrays backwards till they hit the source plane.
+When the deflection angles are determined, the deflected lightrays are
+traced to the source plane and a magnification map is created by
+counting the number of rays which hit each of its pixels. This is the
+inverse of what would happen in a real observation: here an observer
+on earth would detect light emitted by several sources with her/his
+telescope. Each of these sources emits lightrays which are subject to
+the deflection by the intervening gravitational potential of the
+lensing mass. However only a fraction of those will end in the
+observers telescope on earth. Interesting for the analysis are hence
+those lightrays which are originating by the source of interest and
+further the subset of these which actually converge at the location of
+the observer. So instead of calculating the trajectories of lightrays
+emitted by the source in all directions and then succeedingly pick out
+the ones which end in the telescope of the observer, the computational
+effort can be restricted to these lightrays of interest in the first
+place. This is done by treating the problem from the opposite
+direction, namely by using the observers location as starting point
+and then shooting the lightrays backwards till they hit the source
+plane.
 % up to here (would remove it)
 
-The result is a pixel map of the number of lightrays which arrive at the source plane.
-This intermediate result is denoted as the magnification map.
-It characterizes the information about the mass distribution in the lens plane and its effect on lightrays, and projects this information onto the source plane.
-\\\\
-Once the map is created, the lightcurve can be obtained by specifying the transition path of the source within the map.
-The code thereby convolves the function describing the shape of the source with the magnification pattern of the map for each time step (see equation \ref{eqn:ft2d}).
+The result is a pixel map of the number of lightrays which arrive at
+the source plane. This intermediate result is denoted as the
+magnification map. It characterizes the information about the mass
+distribution in the lens plane and its effect on lightrays, and
+projects this information onto the source plane. \\\\ Once the map is
+created, the lightcurve can be obtained by specifying the transition
+path of the source within the map. The code thereby convolves the
+function describing the shape of the source with the magnification
+pattern of the map for each time step (see equation \ref{eqn:ft2d}).
 
 %this is problematic
- Note that in most observational cases in microlensing, both lens and source are moving and further that the lens configuration, and with it the magnification pattern, is changing with time.
-While the first subtlety is taken care of by a coordinate transformation in this analysis, for the second one the lens configuration is assumed to be constant in time.
-\\\\
-For the model a it was desirable to work with a caustic geometry as simple as possible. 
+ Note that in most observational cases in microlensing, both lens and
+ source are moving and further that the lens configuration, and with
+ it the magnification pattern, is changing with time. While the first
+ subtlety is taken care of by a coordinate transformation in this
+ analysis, for the second one the lens configuration is assumed to be
+ constant in time.  \\\\ For the model a it was desirable to work with
+ a caustic geometry as simple as possible.
 % up to here   (would remove it)
 
-When generating the caustic map used in our analysis presented in the second part of the article only two equal point like masses were used as lenses, to produce the magnification map depicted in figure \ref{fig:magnification_map}.
-The magnification maps calculated by the Microlensing code are in units of Einstein radii, hence they are scale free.
-Physical scale is introduced to them by multiplying the pixels by the correct distance factor.
-The Einstein angle is given by
+When generating the caustic map used in our analysis presented in the
+second part of the article only two equal point like masses were used
+as lenses, to produce the magnification map depicted in figure
+\ref{fig:magnification_map}. The magnification maps calculated by the
+Microlensing code are in units of Einstein radii, hence they are scale
+free. Physical scale is introduced to them by multiplying the pixels
+by the correct distance factor. The Einstein angle is given by
 \begin{equation}
@@
 \end{equation}  
-and therefore depends on the four parameters $M,D_{S},D_{L},D_{LS}$. For the cosmological application or quasar lensing $D_{L}$ is of order 100 Mps to Gpc and $D_{S}$ of order Gpc.\\   
-The scale of the magnification map can be converted into a physical length by multiplying it by a factor of $ \sqrt{\frac{D_{LS}}{D_{S}D_{L}}}$.\\\\
-For the next step, the computation of the actual lightcurve, the code was modified to also allow for crescent shaped images specified through the parameter set $R_p,R_n,a,b$.
-Here $R_p$ denotes the outer radius of the crescent and $R_n$ the inner one.
-The orientation of the source image with respect to the magnification map is specified by the parameters $a$ and $b$ as the shift of the center of the inner disk from the center of the outer disk in $x$- and $y$-direction respectively.
-In the original version of the code, gaussian and disk shaped images where already implemented.
-Those are completely characterized by the single parameter $R_p$.
-The values of the parameters are specified in pixel units corresponding to the magnification map.
-Further one needs to specify the start and end point coordinates of the path, which the center of the source image follows through the magnification map (see the depiction in figure \ref{fig:magnification_map}).
-Hence the points along this path are specified through the number of timesteps for which the computation of the brightness is to be carried out.
-Those points correspond to the actual measurement of the brightness of an object in the observational case.
-For each timestep the two-dimensional convolution of equation \ref{eqn:ft2d} is carried out numerically for the position of the source on the magnification map.
-\\
-For the purpose of this analysis it was desirable to mimic the analytical behaviour of a simple fold as much as possible, for comparing the numerical result with the analytical one, therefore the path of the source was chosen, so that it intersects the border of the caustic perpendicularly, and on a point where the border is a fold caustic.
+and therefore depends on the four parameters
+$M,D_{S},D_{L},D_{LS}$. For the cosmological application or quasar
+lensing $D_{L}$ is of order 100 Mps to Gpc and $D_{S}$ of order
+Gpc.\\ The scale of the magnification map can be converted into a
+physical length by multiplying it by a factor of
+$\sqrt{\frac{D_{LS}}{D_{S}D_{L}}}$.\\\\ For the next step, the
+computation of the actual lightcurve, the code was modified to also
+allow for crescent shaped images specified through the parameter set
+$R_p,R_n,a,b$. Here $R_p$ denotes the outer radius of the crescent and
+$R_n$ the inner one. The orientation of the source image with respect
+to the magnification map is specified by the parameters $a$ and $b$ as
+the shift of the center of the inner disk from the center of the outer
+disk in $x$- and $y$-direction respectively. In the original version
+of the code, gaussian and disk shaped images where already
+implemented. Those are completely characterized by the single
+parameter $R_p$. The values of the parameters are specified in pixel
+units corresponding to the magnification map. Further one needs to
+specify the start and end point coordinates of the path, which the
+center of the source image follows through the magnification map (see
+the depiction in figure \ref{fig:magnification_map}). Hence the points
+along this path are specified through the number of timesteps for
+which the computation of the brightness is to be carried out. Those
+points correspond to the actual measurement of the brightness of an
+object in the observational case. For each timestep the
+two-dimensional convolution of equation \ref{eqn:ft2d} is carried out
+numerically for the position of the source on the magnification
+map. \\ For the purpose of this analysis it was desirable to mimic the
+analytical behaviour of a simple fold as much as possible, for
+comparing the numerical result with the analytical one, therefore the
+path of the source was chosen, so that it intersects the border of the
+caustic perpendicularly, and on a point where the border is a fold
+caustic.

### 2015-07-12 paper1/fitting.tex
editing

--- 2015-07-10 paper1/fitting.tex
@@
 \centering
-\includegraphics[width=0.9\hsize]{plots/data_cc.eps}
-\includegraphics[width=0.9\hsize]{plots/data_dd.eps}
-\includegraphics[width=0.9\hsize]{plots/data_gg.eps}
+\includegraphics[width=0.9\hsize%,bb=0 0 576 432
+                ]{plots/data_cc.eps}
+\includegraphics[width=0.9\hsize%,bb=0 0 576 432
+                ]{plots/data_dd.eps}
+\includegraphics[width=0.9\hsize%,bb=0 0 576 432
+                ]{plots/data_gg.eps}
 \caption{\label{fig:mockdata} Three different mock datasets with
@@
 \centering
-  \includegraphics[width=0.9\hsize]{plots/Rp4all.eps}
+  \includegraphics[width=0.9\hsize%,bb=0 0 576 432
+                  ]{plots/Rp4all.eps}
 \caption{\label{fig:mcmc} Posterior probability distribution for the
@@
 \centering
-  \includegraphics[width=0.9\hsize]{plots/Rhalf_RnRp.eps}
-  \includegraphics[width=0.9\hsize]{plots/aRp_bRp.eps}
+  \includegraphics[width=0.9\hsize%,bb=0 0 576 432
+                  ]{plots/Rhalf_RnRp.eps}
+  \includegraphics[width=0.9\hsize%,bb=0 0 576 432
+                  ]{plots/aRp_bRp.eps}
 \caption{\label{fig:crescentfit} The 2$\sigma$ contours (or error ellipses)
@@
 \centering
-\includegraphics[width=0.9\hsize]{plots/burnin_cg.eps}
+\includegraphics[width=0.9\hsize%,bb=0 0 576 432
+                ]{plots/burnin_cg.eps}
 \caption{\label{fig:burnin} The $\chi^2$ of each step of the full MCMC

### 2015-07-12 paper1/theory.tex
editing

@@
 
-\subsection{Basic microlensing theory}
+We start by introducing in a succinct manner the gravitational lensing
+theory that is relevant for microlensing in general and for the scope
+of the present paper in particular.  More detailed presentation of the
+theory can be found in several references, such as
+\cite{2001stgl.book.....P}.
 
-We start by introducing in a succinct manner the gravitational lensing theory that is relevant for microlensing in general and for the scope of the
-present paper in particular. For a detailed presentation of the theory the reader is referred to the books and articles that represent the references
-of the current section (i.e. \cite{Schneider_book}). The general microlensing equation can be written as:
+\subsection{Microlensing essentials}
+
+The gravitational lens equation
 \begin{equation}
-\vec{y} =   
- \begin{pmatrix}
-  1-\gamma - k & 0 \\
-  0 & 1 + \gamma -k \\
- \end{pmatrix} 
-\vec{x} - \sum_{i=1}^{n} m_i \frac{\vec{x} - \vec{x_i}}{\left( \vec{x} - \vec{x_i} \right)^2}.
-\label{eqn:microlensingequation}
+\vec\beta = \vec\theta - \vec\alpha(\vec\theta)
+\label{eqn:lens}
 \end{equation}
+relates the apparent sky position $\vec\theta$ of a light source to
+its true but unobservable sky position $\vec\beta$ through the bending
+angle $\vec\alpha$.  The latter is an integral 
+\begin{equation}
+\vec\alpha = (1+z_{_L})\frac{D_S-D_L}{D_SD_L} \frac{4G}{c^2}
+\int \Sigma(\vec\theta')
+\frac{\vec\theta-\vec\theta'}{|\vec\theta-\vec\theta'|^2}\,
+d^2\vec\theta'
+\end{equation}
+depending on the projected density $\Sigma(\vec\theta)$ (kg/steradian)
+of the lens, the lens redshift $z_{_L}$ and the comoving distances
+$D_L$ and $D_S$ to the lens and source.  The derivative of the
+apparent position with respect to the source position
+\begin{equation}
+M(\vec\theta) =
+\left(\frac{\partial\beta}{\partial\theta}\right)^{-1}
+\label{eqn:magnif-matrix}
+\end{equation}
+is known as the magnification matrix, and its determinant
+\begin{equation}
+\mu(\vec\theta) = \det|M(\vec\theta)|
+\end{equation}
+is the brightness amplification of an image of a point source.  In
+other words, the source will brighten or dim according to whether
+$\mu(\vec\theta)$ is more or less than unity.  If there are multiple
+mages at distinct but not observationally resolved $\vec\theta_i$ from
+the same $\vec\beta$, a total brightness amplification
+of\begin{equation} \mu_{\rm total} = \sum_{i} \mu(\vec\theta_i)
+\end{equation}
+applies. It is possible for the magnification to become formally
+infinite, as a result of an eigenvalues of the magnification matrix
+(\ref{eqn:magnif-matrix}) becoming infinite.  This generically happens
+on curves on the $\vec\theta$ plane, known as {\em critical curves}.
+Mapping a critical curve to the source plane, through the lens
+equation (\ref{eqn:lens}), to the source plane $\vec\beta$ gives the
+so-called caustics.  Caustics can appear in optical system, not just
+gravitational lensing.  For a point source, caustics are singularities
+of the magnification; for finite-size sources caustics correspond to
+high and sharply changing magnification.
 
-The equation is valid for a set of $\it{n}$ point (Schwarzschild) lenses placed at coordinates $\vec{x_i}$.
-$\vec{y}$ represents the coordinates in the source plane for the respective ray, while $\vec{x}$ represents the coordinates in the lens plane.
-Furthermore, $\gamma$ and $k$ are the shear and surface mass density.
-For the lens mapping, a Jacobian can be defined:
-\begin{equation} 
-A(\vec{x}) = 
- \begin{pmatrix}
-  \frac{\partial{y_1}}{\partial{x_1}} & \frac{\partial{y_1}}{\partial{x_2}} \\
-  \frac{\partial{y_2}}{\partial{x_1}} & \frac{\partial{y_2}}{\partial{x_2}} \\
- \end{pmatrix}. 
-\end{equation}
-The determinant of $A(\vec{x})$ is related to the local magnification factor through the equation:
-
-\begin{equation}
-\mu(\vec{x}) = \left| \frac{1}{det(A(\vec{x}))} \right|
-\label{eqn:mu_x}
-\end{equation}
-
-According to this equation, from an infinitesimally small source positioned at $\vec{x}$ in the lens plane an observed will receive a flux $\mu(\vec{x}) dF$ instead of the flux $dF$ that would be
-observed in the absence of the gravitational lens. In other words the source will dim or brighten depending on whether $\mu(\vec{x})$ is sub-unitary or supra-unitary.\\
-
-In general, for a single infinitesimally small source more than one image can be observed, each with magnification $\mu_j$. For microlensing events the angular sizes of the
-images cannot be resolved. The only observable is the flux to which all nearby images contribute. A total magnification
-\begin{equation}
- \mu_t = \sum_{j} \mu_j
-\end{equation}
- can be defined and account for an increase/decrease of brightness from all images.
-
-\subsection{Magnification of a point source near a fold}
-
-The determinant of the Jacobian $A$ can have both positive and negative values depending in which region it is computed. Such regions are well defined and separated
-by critical curves where the determinant vanishes. The lens equation maps the critical curves of the lens plane into caustics in the source plane. An example of the caustic 
-shape can be seen in figure \ref{fig:magnification_map}.\\
+Caustics are important in all forms of lensing with multiple images,
+but they have a special significance for lens quasars, first pointed
+out by \cite{1979Natur.282..561C}.  The granularity of the mass
+distribution $\Sigma(\vec\theta)$ due to individual stars produces a
+caustic network on the scale of $\sqrt{GM_\odot D_L/c^2}$ or
+micro-arcsec.  Extended sources wash out this micro-caustic structure,
+but the optical-continuum source of quasars is even smaller.  Hence
+sharp changes in the brightness of lensed quasars are observable, as
+they cross the micro-caustics \citep[e.g.,][]{2012A&A...544A..62S}.
 
@@
 
-Along these critical curves in the lens plane or caustics in the source plane,  the magnification $\mu$ is infinite, a result that simply follows from equation \ref{eqn:mu_x}.
-This statement holds only for infinitesimally small sources.
-For extended sources the maximum amplification is finite since only an infinitesimally small area of the source
-overlaps with the caustics \citep{1986BAAS...18T.907S}.
+\subsection{Magnification near a fold caustic}
 
-The matrix $A$ at the coordinates of the caustic can have rank 1 or 0.
-If the rank is 1 then the coordinates belong to {\em fold\/} or {\em cusp\/} catastrophes (\cite{1992A&A...260....1S}, \cite{2008mmc..confE...8D}).
-For the purpose
-of this paper we are interested only in the behaviour around {\em fold\/} catastrophe. A condition necessary to distinguish between a fold and cusp singularity is that the
-eigenvector of $A$ corresponding to the null eigenvalue is not a tangent vector of the critical curve (\cite{1992A&A...260....1S}, \cite{2008mmc..confE...8D}).
+A simple example of a caustic network is shown in
+Figure~\ref{fig:magnification_map}.  There are two general categories
+of caustics in gravitational lensing, cusps and fold, and examples of
+both kinds can be seen in this figure.  Magnification near a caustic
+has universal properties, independent of the system and has been
+extensively studied
+\citep{1986ApJ...310..568B,1992A&A...260....1S,2002ApJ...574..970G,2002ApJ...580..468G}.
+In particular, at distance $p$ from a fold caustic
+\begin {equation}
+ \mu(p) = \mu_0 + C_0 \frac{1}{\sqrt{p}} \Theta(p).
+\end {equation}
+Here the magnification of a point source near a caustic is equal to
+the sum of the magnification due to other reasons $\mu_0$, assumed to
+be locally constant, and a decrease with the square root of the
+distance from the fold. The later term becomes activated only after
+the source enters the region interior to the caustic curves when the
+values of the stepfunction $\Theta(p)$ become unity. The
+proportionality constant $C_0$ depends on the local conditions in the
+vicinity of the caustic.
+
+A source of arbitrary shape can be described by a two dimensional brightness function $S_{2D}(p - p_s, q - q_s)$ defined for a coordinate system $p,q$ where $p_s, q_s$ denote the coordinates of the center of source.
 
@@
 \includegraphics[width = .49\textwidth]{plots/infinite_fold.eps}
-\caption{\label{fig:infinite_fold}  Source profiles for gaussian, disk, crescents (right) and magnification map for an infinite fold (left). Objects have the same $S_{0}$ and $r_{1/2}$. }
+\caption{\label{fig:infinite_fold} Source profiles for Gaussian, disk,
+  crescents (right) and magnification map for an infinite fold
+  (left). Objects have the same $S_{0}$ and $r_{1/2}$. }
 \end{figure}
 
-The behaviour of the images of a point source and their corresponding magnifications near such a caustic has been thoroughly studied in the past. The reader who is interested in 
-understanding the details related to this subject is referred to papers such as \cite{1986ApJ...310..568B}, \cite{1992A&A...260....1S}, \cite{2002ApJ...574..970G}, \cite{2002ApJ...580..468G}.
-For the present paper we only
-summarize the common result:
-\begin {equation}
- \mu_{t}(p) = \mu_0 + C_0 \frac{1}{\sqrt{p}} \Theta(p).
-\end {equation}
-As the previous equation describes the magnification of a point source near a caustic is equal to the sum of the magnification due to other reasons $\mu_0$, assumed to be locally constant,  and a decrease with
-the square root of the distance from the fold.
-The later term becomes activated only after the source enters the region interior to the caustic curves when the values of the stepfunction $\Theta(p)$ become unity.
-The proportionality
-constant $C_0$  is sensible only to the local behaviour of Fermat's potential and units used for the lens plane coordinates $\vec{y}(p,q)$.
-
-\subsection{Magnification of an extended source near a fold}
-
-A source of arbitrary shape can be described by a two dimensional brightness function $S_{2D}(p - p_s, q - q_s)$ defined for a coordinate system $p,q$ where $p_s, q_s$ denote the coordinates of the center of source.
-
-For a microlensing event the lightcurve can be written for an undefined source shape as:
-
+For a microlensing event the lightcurve can be written for an
+undefined source shape as:
 \begin{equation}
@@
 \end{equation}
-
-In order to build the previous equation we have considered that the time dependency of the flux $F$ is given only by the motion of the source with respect to a fixed caustic.
-Therefore the only time dependent quantities in the right hand side of the equation
-are the coordinates of the center of the source $p_s,q_s$ and by construct $S_{2D}$.
+In order to build the previous equation we have considered that the
+time dependency of the flux $F$ is given only by the motion of the
+source with respect to a fixed caustic. Therefore the only time
+dependent quantities in the right hand side of the equation are the
+coordinates of the center of the source $p_s,q_s$ and by construct
+$S_{2D}$.
 
@@
 
-\subsection{Crescent source with constant surface brightness distribution}
+\subsection{Crescent source with constant surface brightness distribution}\label{subsec:crescent}
 

### 2015-07-12 paper1/fitting.tex
fitting.tex small corrections

@@
 \caption{\label{fig:mockdata} Three different mock datasets with
-  three different parameterisations fitted to each --- crescent (red),
+  three different parametrisations fitted to each --- crescent (red),
   uniform-disk (blue) and Gaussian disk (green).}
@@
   source size $R_p$.  The vertical line is the correct value. Same
-  color represeents the same dataset whereas same line style
+  color represents the same dataset whereas same line style
   corresponds to same model fit.  The legend gives the reduced

### 2015-07-12 paper1/fitting.tex
editing

--- 2015-07-11 paper1/fitting.tex
@@
 \centering
-\includegraphics[width=0.9\hsize, bb=0 0 576 432]{plots/data_cc.eps}
-\includegraphics[width=0.9\hsize, bb=0 0 576 432]{plots/data_dd.eps}
-\includegraphics[width=0.9\hsize, bb=0 0 576 432]{plots/data_gg.eps}
+\includegraphics[width=0.9\hsize,bb=0 0 576 432
+                ]{plots/data_cc.eps}
+\includegraphics[width=0.9\hsize,bb=0 0 576 432
+                ]{plots/data_dd.eps}
+\includegraphics[width=0.9\hsize,bb=0 0 576 432
+                ]{plots/data_gg.eps}
 \caption{\label{fig:mockdata} Three different mock datasets with
@@
 \centering
-  \includegraphics[width=0.9\hsize, bb=0 0 576 432]{plots/Rp4all.eps}
+  \includegraphics[width=0.9\hsize,bb=0 0 576 432
+                  ]{plots/Rp4all.eps}
 \caption{\label{fig:mcmc} Posterior probability distribution for the
@@
 \centering
-  \includegraphics[width=0.9\hsize, bb=0 0 576 432]{plots/Rhalf_RnRp.eps}
-  \includegraphics[width=0.9\hsize, bb=0 0 576 432]{plots/aRp_bRp.eps}
+  \includegraphics[width=0.9\hsize,bb=0 0 576 432
+                  ]{plots/Rhalf_RnRp.eps}
+  \includegraphics[width=0.9\hsize,bb=0 0 576 432
+                  ]{plots/aRp_bRp.eps}
 \caption{\label{fig:crescentfit} The 2$\sigma$ contours (or error ellipses)
@@
 \centering
-\includegraphics[width=0.9\hsize, bb=0 0 576 432]{plots/burnin_cg.eps}
+\includegraphics[width=0.9\hsize,bb=0 0 576 432
+                ]{plots/burnin_cg.eps}
 \caption{\label{fig:burnin} The $\chi^2$ of each step of the full MCMC

### 2015-07-12 paper1/theory.tex
figure2 with arrow
editing
improved figures 2-10

@@
 \begin{figure}
-\includegraphics[width = .49\textwidth]{plots/infinite_fold.eps}
+\includegraphics[width = .49\textwidth]{plots/infinite_fold_ar.eps}
 \caption{\label{fig:infinite_fold} Source profiles for Gaussian, disk,
   crescents (right) and magnification map for an infinite fold
-  (left). Objects have the same $S_{0}$ and $r_{1/2}$. }
+  (left). Objects have the same $S_{0}$ and $r_{1/2}$. The black arrow marks the direction of motion of the
+sources relative to the caustic.}
 \end{figure}
@@
 \centering
-	\includegraphics[width = 0.48\textwidth]{plots/4source_profile.eps}
+	\includegraphics[width = 0.48\textwidth]{plots/S1D_all.eps}
 	\includegraphics[width = 0.48\textwidth]{plots/4source_magnification.eps}
-\caption{\label{fig:lightcurve_gauss} Lightcurves of crescents, gaussian and disk for sources with identical $S_0$ and $r_{1/2}$.
+\caption{\label{fig:lightcurve_gauss} Lightcurves of crescents, Gaussian and disk for sources with identical $S_0$ and $r_{1/2}$.
 Crescent source has $r_{1/2}$ = 0.72754 $R_p$, $R_n$ = 0.4 $R_p$, a = 0.3/-0.3 $R_p$ }
@@
 \centering
-	\includegraphics[width = 0.48\textwidth]{plots/4avar_profile.eps}
+	\includegraphics[width = 0.48\textwidth]{plots/S1D_var_a.eps}
 	\includegraphics[width = 0.48\textwidth]{plots/4avar_magnification.eps}
@@
 \centering
-	\includegraphics[width = 0.48\textwidth]{plots/4rn_profile.eps}
-	\includegraphics[width = 0.48\textwidth]{plots/4rn_magnification.eps}
+	\includegraphics[width = 0.48\textwidth]{plots/S1D_var_rn_a_poz.eps}
+	\includegraphics[width = 0.48\textwidth]{plots/5Rn_back_var_magnification.eps}
 \caption{\label{fig:lightcurve_crescent_back} Lightcurves of backwards crescents with  a = 0.3 $R_p$.}
@@
 \centering
-	\includegraphics[width = 0.48\textwidth]{plots/4rn2_profile.eps}
-	\includegraphics[width = 0.48\textwidth]{plots/4rn2_magnification.eps}
+	\includegraphics[width = 0.48\textwidth]{plots/S1D_var_rn_a_neg.eps}
+	\includegraphics[width = 0.48\textwidth]{plots/5Rn_forw_var_magnification.eps}
 \caption{\label{fig:lightcurve_crescent} Lightcurve of forwards crescents with a = -0.3 $R_p$.}

### 2015-07-12 paper1/code.tex

@@
-\section{Microlensing code and magnification maps}
+\section{Numerical microlensing magnification maps}
 
-We now go beyond the case of an ideal fold, and consider microlensing
-by random distributions of point lenses.  This requires numerical
-computation of lightcurves, for which we use the microlensing code by
-\cite{1999A&A...346L...5W}.
+Having used to simple model of an ideal fold to gain insight, we now
+move to the numerical computation of microlensing light curves with
+realistic mass distributions.  For this we we used the microlensing
+code by \cite{1999A&A...346L...5W}.
 
-The code is based on the hierarchical tree technique to calculate the
-gravitational effect of the lensing masses. The underlying idea is
-that when the thin lens approximation is valid, the a microlensing
-distribution can be reduced to three planes, the observer plane, the
-lens plane and the source plane. The actual effect, namely the
-distortion of light rays emitted from a distant source, on the path to
-the observer, is numerically computed by through back-tracing:
-lightrays are shot backward from the observer to lens plane, where
-they become deflected according to the angle dependent on the 2D mass
-distribution in the lens plane.  The choice for reversing the photon
-motion is motivated by computational resources considerations.  For
-the computation of the individual deflection angles the tree concept
-is used. The positions of all lensing masses are sorted into a
-2-dimensional grid (in the lens plane). Which is subdivided into 4
-smaller squares recursively until every cell contains only one
-mass. For the actual calculations of the deflection angle not every
-mass gravitational influence is treated equally. Instead it is made
-use of the fact that the influence of lenses on the considered light
-ray falls off with $r^{-1}$ from the point where the ray hits the lens
-plane. Hence masses situated at higher distances, can be grouped
-together in the calculation of their gravitational potential, using
-multipole moments. With this approach the amount of computation time
-is reduced. \\
+The code uses a ray-shooting technique to compute the gravitational
+lensing effect of a mass distribution consisting of (a)~a smooth
+component and (b)~a random distribution of point masses representing
+stars.  The ray shooting maps a grid of $\vec\theta$ to $\vec\beta$
+using the lens equation (\ref{eqn:lens}).  That is, the rays are shot
+from the observer back to the source.
 
-%I don't think this part is necessary
+For the computation of the individual deflection angles
+$\vec\alpha(\vec\theta)$ a heirarchical or tree method is used.  The
+positions of all lensing masses are put into a grid of $\vec\theta$.
+Each grid cell is subdivided into four smaller squares recursively
+until every cell contains only one mass.  Nearby masses are added
+individually, while distance masses are clumped into larger grid cells
+whose net contribution is approximated by its first few multipole
+moments.  Scaled units are used, with the constant pre-factor in the
+deflection angle (\ref{eq:alpha}) separated out.
 
-When the deflection angles are determined, the deflected lightrays are
-traced to the source plane and a magnification map is created by
-counting the number of rays which hit each of its pixels. This is the
-inverse of what would happen in a real observation: here an observer
-on earth would detect light emitted by several sources with her/his
-telescope. Each of these sources emits lightrays which are subject to
-the deflection by the intervening gravitational potential of the
-lensing mass. However only a fraction of those will end in the
-observers telescope on earth. Interesting for the analysis are hence
-those lightrays which are originating by the source of interest and
-further the subset of these which actually converge at the location of
-the observer. So instead of calculating the trajectories of lightrays
-emitted by the source in all directions and then succeedingly pick out
-the ones which end in the telescope of the observer, the computational
-effort can be restricted to these lightrays of interest in the first
-place. This is done by treating the problem from the opposite
-direction, namely by using the observers location as starting point
-and then shooting the lightrays backwards till they hit the source
-plane.
-% up to here (would remove it)
+The result of ray shooting is a pixel map on the $\vec\beta$ plane of
+the number of lightrays which arrive at the source plane from a
+particular observer.  This intermediate result is effectively a
+magnification map on the source plane.  Once the map is created, the
+lightcurve can be obtained by specifying the transit path of the
+source across the map.  At each point on this transit line, the code
+computes equation (\ref{eqn:ft2d}) convolving the brightness
+distribution of the source with the magnification pattern of the map.
+In real life, not only are both lens and source moving, but the lens
+lens configuration, and with it the magnification pattern, is also
+changing with time.  While the first subtlety is taken care of by a
+coordinate transformation in this analysis, for the second one the
+lens configuration is assumed to be constant in time.
 
-The result is a pixel map of the number of lightrays which arrive at
-the source plane. This intermediate result is denoted as the
-magnification map. It characterizes the information about the mass
-distribution in the lens plane and its effect on lightrays, and
-projects this information onto the source plane. \\\\ Once the map is
-created, the lightcurve can be obtained by specifying the transition
-path of the source within the map. The code thereby convolves the
-function describing the shape of the source with the magnification
-pattern of the map for each time step (see equation \ref{eqn:ft2d}).
+When generating the magnification map depicted in
+Figure~\ref{fig:magnification_map}, which is used in our analysis
+below, only two point masses were included.  This was done in order to
+have clean fold caustics.  For the computation of the actual
+lightcurve, the code was modified to also allow for crescent shaped
+images specified through the parameter set $R_p,R_n,a,b$. Here $R_p$
+denotes the outer radius of the crescent and $R_n$ the inner one. The
+orientation of the source image with respect to the magnification map
+is specified by the parameters $a$ and $b$ as the shift of the center
+of the inner disk from the center of the outer disk in $x$- and
+$y$-direction respectively. In the original version of the code,
+gaussian and disk shaped images where already implemented. Those are
+completely characterized by the single parameter $R_p$. The values of
+the parameters are specified in pixel units corresponding to the
+magnification map. Further one needs to specify the start and end
+point coordinates of the path, which the center of the source image
+follows through the magnification map (see the depiction in figure
+\ref{fig:magnification_map}). Hence the points along this path are
+specified through the number of timesteps for which the computation of
+the brightness is to be carried out. Those points correspond to the
+actual measurement of the brightness of an object in the observational
+case. For each timestep the two-dimensional convolution of equation
+\ref{eqn:ft2d} is carried out numerically for the position of the
+source on the magnification map.
 
-%this is problematic
- Note that in most observational cases in microlensing, both lens and
- source are moving and further that the lens configuration, and with
- it the magnification pattern, is changing with time. While the first
- subtlety is taken care of by a coordinate transformation in this
- analysis, for the second one the lens configuration is assumed to be
- constant in time.  \\\\ For the model a it was desirable to work with
- a caustic geometry as simple as possible.
-% up to here   (would remove it)
-
-When generating the caustic map used in our analysis presented in the
-second part of the article only two equal point like masses were used
-as lenses, to produce the magnification map depicted in figure
-\ref{fig:magnification_map}. The magnification maps calculated by the
-Microlensing code are in units of Einstein radii, hence they are scale
-free. Physical scale is introduced to them by multiplying the pixels
-by the correct distance factor. The Einstein angle is given by
-\begin{equation}
-\theta_{E}=\sqrt{\frac{4GM}{c^{2}} \frac{D_{LS}}{D_{S}D_{L}}} 
-\end{equation}  
-and therefore depends on the four parameters
-$M,D_{S},D_{L},D_{LS}$. For the cosmological application or quasar
-lensing $D_{L}$ is of order 100 Mps to Gpc and $D_{S}$ of order
-Gpc.\\ The scale of the magnification map can be converted into a
-physical length by multiplying it by a factor of
-$\sqrt{\frac{D_{LS}}{D_{S}D_{L}}}$.\\\\ For the next step, the
-computation of the actual lightcurve, the code was modified to also
-allow for crescent shaped images specified through the parameter set
-$R_p,R_n,a,b$. Here $R_p$ denotes the outer radius of the crescent and
-$R_n$ the inner one. The orientation of the source image with respect
-to the magnification map is specified by the parameters $a$ and $b$ as
-the shift of the center of the inner disk from the center of the outer
-disk in $x$- and $y$-direction respectively. In the original version
-of the code, gaussian and disk shaped images where already
-implemented. Those are completely characterized by the single
-parameter $R_p$. The values of the parameters are specified in pixel
-units corresponding to the magnification map. Further one needs to
-specify the start and end point coordinates of the path, which the
-center of the source image follows through the magnification map (see
-the depiction in figure \ref{fig:magnification_map}). Hence the points
-along this path are specified through the number of timesteps for
-which the computation of the brightness is to be carried out. Those
-points correspond to the actual measurement of the brightness of an
-object in the observational case. For each timestep the
-two-dimensional convolution of equation \ref{eqn:ft2d} is carried out
-numerically for the position of the source on the magnification
-map. \\ For the purpose of this analysis it was desirable to mimic the
+For the purpose of this analysis it was desirable to mimic the
 analytical behaviour of a simple fold as much as possible, for

### 2015-07-17 paper1/fitting.tex
fitting changes

@@
 disc and a Gaussian disc --- and then fit each mock light curve to all
-three source models. Parameter fitting and marginalisation was done by
+three source models. Parameter fitting and marginalising was done by
 Markov-chain Monte Carlo (MCMC).
@@
 \caption{\label{fig:mockdata} Three different mock datasets with
-  three different parametrisations fitted to each --- crescent (red),
+  three different parametrizations fitted to each --- crescent (red),
   uniform-disk (blue) and Gaussian disk (green).}
@@
 \S\ref{subsec:crescent}) $R_p$ and $R_n$ being the radii of the outer
-and inner circles and $(a,b)$ being the coordinates of the inner
-circle with respect to the centre of the outer circle.  The parameter
-values were
+and inner circles and $(\alpha,\beta)$ being the coordinates of the inner
+circle with respect to the centre of the outer circle in the coordinate system associated with
+the image in figure 1 and not associated with the caustic surface as defined in the previous sections.  The parameter values were
 \begin{equation}
-   (R_p, R_n, a, b) = (50.0, 30.0, 15.0, 10.0).
+   (R_p, R_n, $\alpha$, $\beta$) = (50.0, 30.0, 15.0, 10.0).
 \label{eqn:cp}
@@
 Each mock light curve also has three nuisance parameters, namely the
-beginning and end of the event and the brightness normalization.
+beginning and end of the event and the brightness normalisation.
 These are to be marginalised out by the MCMC.
@@
 of 10\% of the current brightness.  The effective signal to noise can
-be considered to be $0.1\times\sqrt{250}\simeq150$.
+be considered to be \textbf{ bad math here -- $0.1\times\sqrt{250}\simeq150$}.
 
@@
 We denote the nine possible cases with letter pairs, with the first
-letter denoting the model fitted and the second letter representing
+letter denoting the assumed model for the fitting procedure and 
+the second letter representing
 the source: thus CG means a {\em C\/}rescent model was fitted to mock
@@
 Figure~\ref{fig:crescentfit} shows $2\sigma$ of the inferred parameter
-values.
+values. \textbf{In the following enumerated conclusions we associate the word "wrong"
+with parameters for which the real value corresponds to a PDF of the fitted model smaller than 
+0.01. }
 
@@
   and this shows in the blue ellipses in the recovered parameter
-  values.  The redundant parameters $a$ and $b$, in effect, allow the
+  values.  The redundant parameters $\alpha$ and $\beta$, in effect, allow the
   model to partly fit the noise, and hence the $\chi^2$ is somewhat
@@
 Next we consider the three cases where a disc model is fitted.  These
-correspond to he dashed curves in Figure~\ref{fig:mcmc}.  There is
+correspond to the dashed curves in Figure~\ref{fig:mcmc}.  There is
 only one interesting parameter to fit, the disc radius $R_p$, and no
@@
 Finally, we consider the three cases where a Gaussian model is fitted.
-These correspond to he dotted curves in Figure~\ref{fig:mcmc}.  Again,
+These correspond to the dotted curves in Figure~\ref{fig:mcmc}.  Again,
 there is only one interesting parameter to fit, the Gaussian
@@
 models is correct --- that the reduced $\chi^2$ would unmask the
-correct model, and its parameters would be correctly recovered.
+correct model, \textbf{and its parameters would be correctly recovered}.

### 2015-07-17 paper1/fitting.tex
small correction

@@
 and inner circles and $(\alpha,\beta)$ being the coordinates of the inner
-circle with respect to the centre of the outer circle in the coordinate system associated with
-the image in figure 1 and not associated with the caustic surface as defined in the previous sections.  The parameter values were
+circle with respect to the centre of the outer circle in the coordinate system associated with the image in Figure 1 and not associated 
+with the caustic surface as defined in the previous sections.  The parameter values were
 \begin{equation}

### 2015-07-18 paper1/fitting.tex
small edits

@@
 \begin{equation}
-   (R_p, R_n, $\alpha$, $\beta$) = (50.0, 30.0, 15.0, 10.0).
+   (R_p, R_n, \alpha, \beta) = (50.0, 30.0, 15.0, 10.0).
 \label{eqn:cp}
@@
 
-Figure~\ref{fig:mockdata} shows the three light curves.  Each has
-250 points regularly spaced in time, with Gaussian noise at the level
-of 10\% of the current brightness.  The effective signal to noise can
-be considered to be \textbf{ bad math here -- $0.1\times\sqrt{250}\simeq150$}.
+Figure~\ref{fig:mockdata} shows the three light curves.  Each has 250
+points regularly spaced in time, with Gaussian noise at the level of
+10\% of the current brightness.  The effective signal to noise can be
+considered to be $0.1\times\sqrt{250}\simeq15$.
 
@@
 Figure~\ref{fig:crescentfit} shows $2\sigma$ of the inferred parameter
-values. \textbf{In the following enumerated conclusions we associate the word "wrong"
-with parameters for which the real value corresponds to a PDF of the fitted model smaller than 
-0.01. }
+values.
 
@@
   seemingly tightly constrained, but the recovered $R_p$ is completely
-  wrong.  The reduced $\chi^2$ is only 0.85, indicative of
+  wrong, since the correct value is far from the green curve and green
+  ellipses.  The reduced $\chi^2$ is only 0.85, indicative of
   over-fitting.  We can see what has happened from the red curve in

### 2015-07-20 paper1/chains/analysismcmc.py
plots changed

@@ -2,6 +2,7 @@ import numpy
 import pylab
 
 from circles import Rhalf
+pylab.figure(figsize=(8.7,7))
 
 def plottinghist(chain,specs,burn):
 	from numpy import genfromtxt,mean,std,log,max,linspace
@@ -51,9 +52,12 @@ plottinghist('GG.chain',':g',500)
 pylab.legend(loc=9,fontsize=12)
 pylab.xlim(25,102)
 #pylab.xlim(15,50)
-pylab.xlabel('$\mathtt{R_{p}}$',fontsize=20)
-pylab.ylabel('$\mathtt{Probability\ density\ function}$',fontsize=20)
+pylab.xlabel('$\mathrm{R_{p}}$',fontsize=20)
+pylab.ylabel('$\mathrm{Probability\ density\ function}$',fontsize=20)
 pylab.ylim(ymax=1.0)
 pylab.axvline(x=50,color='k',ls='-.',lw=2)
 pylab.savefig('Rp4all.eps')
+
+pylab.tick_params(axis='both', which='major', labelsize=20)
+pylab.tick_params(axis='both', which='minor', labelsize=20)
 pylab.show()

### 2015-07-20 paper1/chains/burn_in.py

@@ -1,16 +1,21 @@
 import numpy
 import pylab
 
+pylab.figure(figsize=(8.7,7))
+
+
 chain = ['CC','CD','CG','DC','DD','DG','GC','GD','GG']
 chain = ['CG']
 spec = ['k','b','g','k','m','g','k','y','c']
 for i in range(len(chain)):
 	data = numpy.genfromtxt('%s.chain'%chain[i])
 	x = numpy.linspace(1,len(data)+1,len(data))
-	pylab.semilogx(x,-2.0*data[:,0],spec[i],label='$\mathtt{CrescentFit\ to\ GaussianDiskData}$')
-pylab.xlabel('$\mathtt{Number\ of\ MCMC\ steps}$',fontsize=22)
-pylab.ylabel('$\mathtt{\chi^2}$',fontsize=22)
+	pylab.semilogx(x,-2.0*data[:,0],spec[i],label='$\mathrm{CrescentFit\ to\ GaussianDiskData}$')
+pylab.xlabel('$\mathrm{Number\ of\ MCMC\ steps}$',fontsize=22)
+pylab.ylabel('$\mathrm{\chi^2}$',fontsize=22)
 pylab.xlim(min(x),max(x))
 pylab.legend(loc=3,fontsize=20)
+pylab.tick_params(axis='both', which='major', labelsize=20)
+pylab.tick_params(axis='both', which='minor', labelsize=20)
 pylab.savefig('burnin_cg.eps')
 pylab.show()

### 2015-07-20 paper1/chains/plotcontours.py

@@ -1,6 +1,7 @@
 import numpy
 import pylab
 
+pylab.figure(figsize=(8.7,7))
 
 def plotcontours(array1,array2,CL,spec,opaque,description):
 	import numpy as np
@@ -95,8 +96,8 @@ files = files[::-1]
 spec = ['#999933','#009999','#cc6633']
 spec = ['g','b','r']
 spec = spec[::-1]
-labels = ['$\mathtt{GaussianDiskData}$',\
-'$\mathtt{UniformDiskData}$','$\mathtt{CrescentData}$']
+labels = ['$\mathrm{GaussianDiskData}$',\
+'$\mathrm{UniformDiskData}$','$\mathrm{CrescentData}$']
 labels = labels[::-1]
 for i in range(len(files)):
 	f1 = numpy.genfromtxt(files[i]+'.chain',skip_header=burn[i])
@@ -116,8 +117,12 @@ pylab.plot(50.0,30.0/50.0,'xk',ms=14)
 pylab.legend(loc=1)
 pylab.ylim(0.0,1.0)
 pylab.xlim(25,102)
-pylab.xlabel('$\mathtt{R_{P}}$',fontsize=22)
-pylab.ylabel('$\mathtt{R_n/R_p}$',fontsize=22)
+pylab.xlabel('$\mathrm{R_{P}}$',fontsize=20)
+pylab.ylabel('$\mathrm{R_n/R_p}$',fontsize=20)
+
+pylab.tick_params(axis='both', which='major', labelsize=20)
+pylab.tick_params(axis='both', which='minor', labelsize=20)
+pylab.tight_layout()
 pylab.savefig('Rhalf_RnRp.eps')
 pylab.show()
 
@@ -140,7 +145,22 @@ pylab.plot(15./50.,-10./50.,'xk',ms=14)
 pylab.legend(loc=1)
 pylab.ylim(-0.4,1.0)
 pylab.xlim(0.0,0.8)
-pylab.xlabel('$\mathtt{a/R_{p}}$',fontsize=22)
-pylab.ylabel('$\mathtt{b/R_p}$',fontsize=22)
+pylab.xlabel('$\mathrm{\\alpha/R_{p}}$',fontsize=20)
+pylab.ylabel('$\mathrm{\\beta/R_p}$',fontsize=20)
+
+pylab.tick_params(axis='both', which='major', labelsize=20)
[…]

### 2015-07-20 paper1/data/plotdata.py

@@ -1,6 +1,7 @@
 import numpy
 import pylab
 
+pylab.figure(figsize=(8.7,7))
 
 def plot(filename,filename_bestfit1,filename_bestfit2,filename_bestfit3,name):
 	x = numpy.genfromtxt(filename)
@@ -8,19 +9,24 @@ def plot(filename,filename_bestfit1,filename_bestfit2,filename_bestfit3,name):
 	y2 = numpy.genfromtxt(filename_bestfit2)
 	y3 = numpy.genfromtxt(filename_bestfit3)
 	pylab.errorbar(x[:,0]+850,x[:,1],x[:,1]*0.1,color='k',\
-		label='$\mathtt{%s}$'%name)
+		label='$\mathrm{%s}$'%name)
 	pylab.plot(y1[:,0]-y1[1,0]+y1[1,2]+850,y1[:,1],'r',lw=2)
 	pylab.plot(y2[:,0]-y2[1,0]+y2[1,2]+850,y2[:,1],'b',lw=2)
 	pylab.plot(y3[:,0]-y3[1,0]+y3[1,2]+850,y3[:,1],'g',lw=2)
 
 
-plot('data_gauss_reverse_noise.txt','cg.bestfit','dg.bestfit','gg.bestfit','Gaussian\ Disk\ data')
+plot('data_crescent_reverse_noise.txt','cc.bestfit','dc.bestfit','gc.bestfit','Crescent\ data')
+# plot('data_disk_reverse_noise.txt','cd.bestfit','dd.bestfit','gd.bestfit','Uniform\ Disk\ data')
+# plot('data_gauss_reverse_noise.txt','cg.bestfit','dg.bestfit','gg.bestfit','Gaussian\ Disk\ data')
+
 pylab.legend(loc=2,fontsize=22)
 #pylab.axes().set_aspect(20.0)
 pylab.xlim(0,250)
 pylab.ylim(ymin=2)
-pylab.xlabel('$\mathtt{Time}$',fontsize=22)
-pylab.ylabel('$\mathtt{Brightness}$',fontsize=22)
-pylab.savefig('data_gg.eps')
+pylab.xlabel('$\mathrm{Time}$',fontsize=20)
+pylab.ylabel('$\mathrm{Brightness}$',fontsize=20)
+pylab.tick_params(axis='both', which='major', labelsize=20)
+pylab.tick_params(axis='both', which='minor', labelsize=20)
+pylab.savefig('data_cc.eps')
 pylab.show()
 

### 2015-07-20 paper1/code.tex
small edits
Discussion expanded

@@
 
-Having used to simple model of an ideal fold to gain insight, we now
+Having used the simple model of an ideal fold to gain insight, we now
 move to the numerical computation of microlensing light curves with
@@
 moments.  Scaled units are used, with the constant pre-factor in the
-deflection angle (\ref{eq:alpha}) separated out.
+deflection angle (\ref{eqn:alpha}) separated out.
 
@@
 gaussian and disk shaped images where already implemented. Those are
-completely characterized by the single parameter $R_p$. The values of
+completely characterised by the single parameter $R_p$. The values of
 the parameters are specified in pixel units corresponding to the

### 2015-07-20 paper1/discussion.tex
fitting changes
Discussion expanded

@@
 and compare them with the microlensing lightcurves of other simple mathematically describable source profiles. 
-In order to mimic the behaviour of flux of light from the source in the proximity of a fold caustic we make use of the simple approximation described in equation (5). 
-The equation would exhaustively describe the magnification map and offer a good universal aproximation for the 
+In order to mimic the behaviour of the flux of light from the source in the proximity of a fold caustic we make use of the simple approximation described in equation (5). 
+The equation would exhaustively describe the magnification map and offer a good universal approximation for the 
 particular microlensing regime that we consider. 
@@
 lightcurves is not evenly distributed. The shape of the lightcurve will be maintained. 
-The datapoints corresponding to the source position before and during the overlapping of the caustic will be affected by
- smaller errors than the datapoints corresponding to later times. \\
+The data points corresponding to the source position before and during the overlapping of the caustic will be affected by
+ smaller errors than the data points corresponding to later times. \\
 
-The first two source profiles that we consider are the uniform disk and symmetric gaussian source. 
+The first two source profiles that we consider are the uniform disk and symmetric Gaussian source. 
 Both of them can be described by a half-light radius $r_{1/2}$ and a total unlensed light flux $S_0$. 
@@
 Two of the parameters can be reduced by expressing the results in terms of $S_0$ and $r_{1/2}$. 
-The later being determinable for any set of parameters $R_p, R_n$ and $a^2+b^2$. Moreover, one of the displacemement 
+The later being determinable for any set of parameters $R_p, R_n$ and $a^2+b^2$. Moreover, one of the displacement 
 parameters $b$ has no impact on the one dimensional profile of the source that results from the projection of the source image on an axis perpendicular to the caustic.
@@
 Since the one dimensional source profile that corresponds to an axis perpendicular to the caustic contains exhaustively all the information regarding the source that can be revealed by the lightcurve,
-the value of the parameter $b$ does not effect the shape of the lighcurve.
+the value of the parameter $b$ does not have an effect on the shape of the lighcurve.
 
-Nevertheless the $b$ parameter is relevant to the calculation of $r_{1/2}$. 
+Nevertheless the $b$ parameter is relevant for the calculation of $r_{1/2}$. 
 It's qualitative effect is to decrease the value of the half-light radius when the absolute value of the parameter is increased.  
@@
 Figure 4 reveals that the lighcurve of a crescent source has more visible features than the other two light-curves 
-corresponding to the disc and guasssian shape. The parameters $R_n$ and $a$ have strong influences on the shape of 
+corresponding to the disc and Guasssian shape. The parameters $R_n$ and $a$ have strong influences on the shape of 
 the microlensed lightcurve, as can be seen in figures 5, 6 and 7.
@@
 The observer would basically detect only the lightcurve of such as source.
-As described in section (Lightcurve of a crescent source) the timing of the onset and offset of the previously described periods can be used to estimate the values of the radii and one of the displacement parameters when assuming a geometric crescent, the halfwidths and one displacement parameters when assuming a constant brightness elliptical crescent and the halfwidth of the dark ellipse when assuming an elliptical gaussian crescent.
+As described in section 4.3 (Lightcurve of a crescent shaped source) the timing of the onset and offset of the previously described periods can be used to estimate the values of the radii and one of the displacement parameters when assuming a geometric crescent.
 All quantities can be estimated in terms of the relative velocity of the source in a direction perpendicular to the caustic.\\
@@
 study of the dynamics of the stellar structure which contains the gravitational lens.
-A better estimate of the relative velocity would facillitate a better estimate of the effective non-rotating black hole mass associated to the black hole shadow.
+A better estimate of the relative velocity would facilitate a better estimate of the effective non-rotating black hole mass associated to the black hole shadow.
 
@@
 In case of a high quality lightcurve with insignificant noise and measurement errors, the parameters can be obtained 
-by simply identifying the characterstic instances of time withought making use of the actual values of the magnification
+by simply identifying the characteristic instances of time without making use of the actual values of the magnification
  map. If the effect of the errors and the noise distorts the magnification time function enough so that the 
-characteristic points are not identifiable with the characteristic periods still visible, the boundaries of the periods 
-can be roughly estimated. To study if good estimates of the parameters can still be obtained when datapoints are
- strongly affected by the noise we propose ....
-
-[Irshad's contribution]\\
-
-This work presents a toy model to put useful constraints on the interior structure of quasars, particularly the environment geometry of the event horizon.
-To carry out this analysis on real data, one needs two situations: First, a quasar which can be observed with a reasonably good telescope with high signal to noise ($\sim 150$) and second, a caustic crossing event of the quasar.
-These events are not rare, however, to get an isolated caustic crossing might be.
-As the size of the source is negligible as compared to big caustic structures in the sky, even the single fold caustic crossing is not that rare.
-
-In this toy analysis, we used very ideal data, which might not be the case with real source but we contaminated it to match a reasonable signal to noise.
-
-The main emphasis of this work was the crescent model parameters, which we show can be well recovered by the maximum likelihood analysis, apart from highly degenerate $a$ and $b$ parameters.
-The reason of this degeneracy is also very intuitive, if one integrates parallel to the caustic to get the integrated 1D brightness profile, so if the motion of the source is either parallel or perpendicular to the caustic, one of the two parameters ($a$ and $b$) is indistinguishable and in case of angled motion of the source, these two are highly degenerate.
-This degeneracy can be broken if multiple crossing events of the same quasar is available.
-So for example, one can use the first crossing to estimate the likelihood of the four parameters and use them as priors in the second crossing and so on.
-In this way, one can accurately and precisely recover all four parameters.
-Using our toy analysis we conclude that it is feasible, in practice, to recover 2D shape of the source using a microlensing event and distinguish between symmetric and asymmetric sources.
+characteristic epochs are not identifiable with the characteristic periods still visible, the boundaries of the periods 
+can be roughly estimated.
+Furthermore, if direct estimates of the parameters cannot be obtained we propose the use of a strong statistical tool such as Markov-Chain Monte Carlo.
+In Section 6  we have studied the possibility of identifying a crescent source and  the possibility of recovering the respective parameters.
+As magnification map we have used a complex numerical one generated with the microlensing code by \cite{1999A&A...346L...5W}.
+In addition we have added to the signal a Gaussian  noise with a SNR of 1.6.
+In the experiment we have considered all nine combinations of original source profiles and assumed fitting source models.
+Effectively we have fitted using MCMC all three sources with all three fitting assumed models.
+The results of the experiment allowed us build a procedure for distinguishing the shape of the source assuming the one of the three models we have considered are a good approximation.
+The procedure would be useful for observers that endeavour to gain more information about an unresolved source for which they can study the microlensing lightcurve.
+As a first step in the procedure, one should first attempt to use MCMC with a Gaussian model assumption.
+If either the value of the $\chi^2$ or the number of rejected datapoints is large then the next step is to change the assumed source model to an uniform disc and redo the fitting.
+Finally if the uniform disc assumption is rejected as well then the fitting should be done with a crescent source model assumption.
+If at each of the three steps the data rejects the model than the source cannot be approximated by none of the three models.
+Otherwise if at one of the steps the data does not reject the model then the respective model is a good approximation.
 
@@
 Mihai Tomozeiu simulated and studied the ideal behaviour of the microlensing lightcurves for the different source profiles discussed.
-Joachim Wambsganss provided the numerical code used by Manuel Rabold to create the magnification map and corresponding lighcurves used in the MCMC analysis, analisis that was performed by Irshad Mohammed in the last part of the presented work.
+Joachim Wambsganss provided the numerical code used by Manuel Rabold to create the magnification map and corresponding lighcurves used in the MCMC analysis, analysis that was performed by Irshad Mohammed in the last part of the presented work.

### 2015-07-20 paper1/ms.tex
title and authors edited

--- 2014-04-15 paper1/ms.tex
@@
 \usepackage{graphicx}
+\usepackage{epstopdf}
+\usepackage{multicol}
 
-\title{Microlensing as a possible probe of the internal structure of quasars}
+\title[Quasar microlensing and its internal structure]{Properties of quasar microlensing light curves: probing its internal structure}
+% \title{Microlensing as a possible probe of the internal structure of quasars}
 
-\author[I. Mohammed et al.] {Irshad Mohammed\thanks{irshad@physik.uzh.ch}$^1$,
-Mihai Tomozeiu$^1$, Manuel Rabold$^1$,
+\author[Tomozeiu et al]{Mihai Tomozeiu\thanks{mihai@physik.uzh.ch},$^{1,2}$ 
+Irshad Mohammed,$^{1,2}$ 
+Manuel Rabold,$^{1,2}$
+Prasenjit Saha,$^{1,2}$  
 \newauthor
-Prasenjit Saha,$^1$ and Joachim Wambsgans$^2$\\
-$^1${Institute for Theoretical Physics, University of Z$\ddot{u}$rich, 8057 Z$\ddot{u}$rich Switzerland}\\
-$^2${University of Heidelberg, Germany}
+and Joachim Wambsganss$^3$\\
+$^1${Physik-Institut, University of Zurich, Winterthurerstrasse 190,
+  8057 Zurich, Switzerland} \\
+$^2${Institute for Computational Science, University of Zurich,
+  Winterthurerstrasse 190, 8057 Zurich, Switzerland} \\
+$^3${Zentrum f\"ur Astronomie der Universit\"at Heidelberg,
+  M\"onchhofstrasse 12--14, 69120 Heidelberg, Germany}
 }

### 2015-08-22 paper1/discussion.tex
another set of corrections

@@
 characteristic epochs are not identifiable with the characteristic periods still visible, the boundaries of the periods 
-can be roughly estimated.
-Furthermore, if direct estimates of the parameters cannot be obtained we propose the use of a strong statistical tool such as Markov-Chain Monte Carlo.
-In Section 6  we have studied the possibility of identifying a crescent source and  the possibility of recovering the respective parameters.
-As magnification map we have used a complex numerical one generated with the microlensing code by \cite{1999A&A...346L...5W}.
-In addition we have added to the signal a Gaussian  noise with a SNR of 1.6.
-In the experiment we have considered all nine combinations of original source profiles and assumed fitting source models.
-Effectively we have fitted using MCMC all three sources with all three fitting assumed models.
-The results of the experiment allowed us build a procedure for distinguishing the shape of the source assuming the one of the three models we have considered are a good approximation.
-The procedure would be useful for observers that endeavour to gain more information about an unresolved source for which they can study the microlensing lightcurve.
-As a first step in the procedure, one should first attempt to use MCMC with a Gaussian model assumption.
-If either the value of the $\chi^2$ or the number of rejected datapoints is large then the next step is to change the assumed source model to an uniform disc and redo the fitting.
-Finally if the uniform disc assumption is rejected as well then the fitting should be done with a crescent source model assumption.
-If at each of the three steps the data rejects the model than the source cannot be approximated by none of the three models.
+can be roughly estimated. Furthermore, if direct estimates of the parameters cannot be obtained we propose the use of a strong
+statistical tool such as Markov-Chain Monte Carlo. In Section 6  we have studied the possibility of identifying a crescent 
+source and  the possibility of recovering the respective parameters. As magnification map we have used a complex numerical 
+one generated with the microlensing code by \cite{1999A&A...346L...5W}. In addition we have added to the signal a Gaussian 
+noise with a SNR of 1.6.  In the experiment we have considered all nine combinations of original source profiles and assumed fitting source models.
+Effectively we have fitted using MCMC all three sources with all three assumed fitting models. The results of the experiment allowed us build a 
+procedure for distinguishing the shape of the source assuming that one of the three models we have considered is a good approximation.  
+The procedure would be useful for observers that endeavour to gain more information about an unresolved source for which they can study 
+the microlensing lightcurve.  As a first step in the procedure, one should first attempt to use MCMC with a Gaussian model assumption. 
+If either the value of the $\chi^2$ or the number of rejected datapoints is large then the next step is to change the assumed source model to 
+an uniform disc and redo the fitting. Finally if the uniform disc assumption is rejected as well then the fitting should be done with a 
+crescent source model assumption. If at each of the three steps the data rejects the model then the source cannot be approximated by any of the three models. 
 Otherwise if at one of the steps the data does not reject the model then the respective model is a good approximation.
 
-[Contributions from authors]\\
-Prasenjit Saha provided the original idea and plan for the research project as well as multiple contributions to the analysis.
-Mihai Tomozeiu simulated and studied the ideal behaviour of the microlensing lightcurves for the different source profiles discussed.
-Joachim Wambsganss provided the numerical code used by Manuel Rabold to create the magnification map and corresponding lighcurves used in the MCMC analysis, analysis that was performed by Irshad Mohammed in the last part of the presented work.
+[Contributions from the authors]\\
+P.S. provided the original idea and plan for the research project as well as multiple contributions to the analysis. M.T. simulated and studied the microlensing 
+lightcurves for the different source profiles discussed for clean data and approximated fold magnification amplification. J.W. provided the training and the
+numerical code used by M. R. to create the magnification map and corresponding lighcurves used in the MCMC analysis, analysis that was performed by 
+I. M. in the last part of the presented work.

### 2015-08-22 paper1/theory.tex
Discussion expanded
another set of corrections

@@
 d^2\vec\theta'
+\label{eqn:alpha}
 \end{equation}
@@
 \centering
-  \includegraphics[width=.9\hsize]{plots/IRIS567_path_2.eps}
-\caption{\label{fig:magnification_map} Magnification map}
+\includegraphics[width=0.9\hsize,bb=0 0 576 432
+		]{plots/IRIS567_path_2.eps}
+\caption{\label{fig:magnification_map} Magnification map used in the second part of the presented work.
+The white line marks the trajectory of the center of the sources for which the lightcurves presented in figure 12 are generated.}
 \end{figure}
@@
 \begin{equation}
- S_{2D}^C =  S_{2D}^{Dp} -  S_{2D}^{Dn}   (n+1)
+ S_{2D}^C =  S_{2D}^{Dp} -  S_{2D}^{Dn}  
  \label{eqn:s2d}
@@
 \begin{equation}
-\frac{S_0^C}{2} =  \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{d_c}^{R_p} \left[ \sqrt{R_p^2 - p^2} - \sqrt{R_p^2 - \left(p-c\right)^2} \Theta \left(R_p^2 - \left(p-c\right)^2 \right) \right] dp.
+\begin{aligned}
+\frac{S_0^C}{2} & = \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{d_c}^{R_p} \bigg[ \sqrt{R_p^2 - p^2} \\ 
+		& - \sqrt{R_p^2 - \left(p-c\right)^2} \Theta \left(R_p^2 - \left(p-c\right)^2 \right) \bigg] dp. 
+\end{aligned}
 \end{equation}
@@
 \end{equation}
+
 which can be simplified to:
@@
 \centering
-	\includegraphics[width = 0.48\textwidth]{plots/S1D_all.eps}
-	\includegraphics[width = 0.48\textwidth]{plots/4source_magnification.eps}
+	\includegraphics[width = 0.48\textwidth ,bb=0 0 576 520
+	]{plots/S1D_all.eps}
+	\includegraphics[width = 0.48\textwidth ,bb=0 0 576 520
+	]{plots/4source_magnification.eps}
 \caption{\label{fig:lightcurve_gauss} Lightcurves of crescents, Gaussian and disk for sources with identical $S_0$ and $r_{1/2}$.
@@
 \begin{equation}
- F^D(t) = \int_{-\infty}^\infty  \left( \mu_0 + \frac{C_0}{\sqrt{p}} \Theta \left( p \right) \right) \left( \frac{2 S_0^D}{ \pi R} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}} \Theta \left(R^2 - \left(p-p_s(t) \right)^2 \right) \right) \mathrm{d}p.
+\begin{aligned}
+ F^D(t) &= \int_{-\infty}^\infty  \left( \mu_0 + \frac{C_0}{\sqrt{p}} \Theta \left( p \right) \right) \\
+	& \bigg[ \frac{2 S_0^D}{ \pi R} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}}  \Theta \left(R^2 - \left(p-p_s(t) \right)^2 \right) \bigg] \mathrm{d}p.
+\end{aligned}
 \end{equation}
@@
 \centering
-	\includegraphics[width = 0.48\textwidth]{plots/S1D_var_a.eps}
-	\includegraphics[width = 0.48\textwidth]{plots/4avar_magnification.eps}
+	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
+	]{plots/S1D_var_a.eps}
+	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
+	]{plots/4avar_magnification.eps}
 \caption{\label{fig:lightcurve_disk} Lightcurves of crescent sources with $R_n$ = 0.4 $R_p$.}
@@
 The lightcurve of a crescent shaped source with unamplified flux $S_0^C$, radii $R_p$, $R_n$ and center displacement $a(t)$ is:
+
 \begin{equation}
- F^c(t) = \mu_0 S_0^C + C_0 \frac{2 S_0^C}{\pi \left( R_p^2 -R_n^2 \right) } 
-\left(\int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) \right)^2 }{p}} \mathrm{d}p 
-  -  \int_{max(0, p_s(t) - a(t) - R)}^{max(0, p_s(t) -a(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) +a(t) \right)^2 }{p}}  \mathrm{d}p \right)
+\begin{aligned}
+ F^c(t) &= \mu_0 S_0^C + C_0 \frac{2 S_0^C}{\pi \left( R_p^2 -R_n^2 \right) } \\
+	&\bigg[ \int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) \right)^2 }{p}} \mathrm{d}p \\
+	&  -  \int_{max(0, p_s(t) - a(t) - R)}^{max(0, p_s(t) -a(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) +a(t) \right)^2 }{p}} \mathrm{d}p  \bigg] .
+\end{aligned}
 \end{equation}
@@
 \centering
-	\includegraphics[width = 0.48\textwidth]{plots/S1D_var_rn_a_poz.eps}
-	\includegraphics[width = 0.48\textwidth]{plots/5Rn_back_var_magnification.eps}
+	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
+	]{plots/S1D_var_rn_a_poz.eps}
+	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
+	]{plots/5Rn_back_var_magnification.eps}
 \caption{\label{fig:lightcurve_crescent_back} Lightcurves of backwards crescents with  a = 0.3 $R_p$.}
@@
 \centering
-	\includegraphics[width = 0.48\textwidth]{plots/S1D_var_rn_a_neg.eps}
-	\includegraphics[width = 0.48\textwidth]{plots/5Rn_forw_var_magnification.eps}
+	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
+	]{plots/S1D_var_rn_a_neg.eps}
+	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
+	]{plots/5Rn_forw_var_magnification.eps}
 \caption{\label{fig:lightcurve_crescent} Lightcurve of forwards crescents with a = -0.3 $R_p$.}
@@
 \centering
-	\includegraphics[width = 0.48\textwidth]{plots/ch_points.eps}
-        \includegraphics[width = 0.48\textwidth]{plots/ch_instances.eps}
+	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
+	]{plots/ch_points.eps}
+        \includegraphics[width = 0.48\textwidth,bb=0 0 576 520
+	]{plots/ch_instances.eps}
 \caption{\label{fig:char_points} The characteristic points and instances of a crescent source's one dimensional profile and microlensing lightcurve. }
@@
 
-\citep{2012MNRAS.421.1517D} have created a radiative image of M87 based on the GRMHD simulations presented in \citep{2009MNRAS.394L.126M}.
+\cite{2012MNRAS.421.1517D} have created a radiative image of M87 based on the GRMHD simulations presented in \citep{2009MNRAS.394L.126M}.
 The respective image, reproduced in figure \ref{fig:M87_image}, has been projected to a 1D profile associated to the perpendicular direction to a fold caustic approaching the image from the right.
@@
 \centering
-	\includegraphics[width = 0.48\textwidth]{plots/M87_shape.eps}
-        \includegraphics[width = 0.48\textwidth]{plots/M87_lc.eps}
+	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
+	]{plots/M87_shape.eps}
+        \includegraphics[width = 0.48\textwidth,bb=0 0 576 520
+	]{plots/M87_lc.eps}
 \caption{\label{fig:M87_plots} The one dimensional profile obtained from the numerical integration along the ordonata axis of the source image (figure 9) is presented in the left subplot.

### 2016-02-22 paper1/intro.tex
another set of corrections
a new ref

@@
 
-A whole range of models have been applied starting from simple geometric models to more complex, physically sensible models (\cite{2008Natur.455...78D}, \cite{2011ApJ...738...38B}, \cite{2009ApJ...706..497M}, \cite{2010ApJ...717.1092D}).
+A whole range of models have been applied starting from simple geometric models to more complex, physically sensible models (\cite{2008Natur.455...78D}, \cite{2011ApJ...738...38B}, \cite{2009ApJ...706..497M}, \cite{2010ApJ...717.1092D}, \cite{2016arXiv160205527F}).
 A significant fraction of the later models predict a crescent shaped silhouette of the black hole. 
@@
 
-%Irshad - you can add here your paragraphs
-
 In section \ref{sec:fitting}, we carried out a likelihood analysis for the parametrisation of the different source shapes and study the degeneracies involved amongst them.

### 2016-03-09 paper1/fitting.tex
fixed one typo
reprocessed eps files

@@
 \centering
-\includegraphics[width=0.9\hsize,bb=0 0 576 432
-                ]{plots/data_cc.eps}
-\includegraphics[width=0.9\hsize,bb=0 0 576 432
-                ]{plots/data_dd.eps}
-\includegraphics[width=0.9\hsize,bb=0 0 576 432
-                ]{plots/data_gg.eps}
+\includegraphics[width=0.9\hsize]{figures/data_cc.eps}
+\includegraphics[width=0.9\hsize]{figures/data_dd.eps}
+\includegraphics[width=0.9\hsize]{figures/data_gg.eps}
 \caption{\label{fig:mockdata} Three different mock datasets with
@@
 \centering
-  \includegraphics[width=0.9\hsize,bb=0 0 576 432
-                  ]{plots/Rp4all.eps}
+  \includegraphics[width=0.9\hsize]{figures/Rp4all.eps}
 \caption{\label{fig:mcmc} Posterior probability distribution for the
@@
 \centering
-  \includegraphics[width=0.9\hsize,bb=0 0 576 432
-                  ]{plots/Rhalf_RnRp.eps}
-  \includegraphics[width=0.9\hsize,bb=0 0 576 432
-                  ]{plots/aRp_bRp.eps}
+  \includegraphics[width=0.9\hsize]{figures/Rhalf_RnRp.eps}
+  \includegraphics[width=0.9\hsize]{figures/aRp_bRp.eps}
 \caption{\label{fig:crescentfit} The 2$\sigma$ contours (or error ellipses)
@@
 \centering
-\includegraphics[width=0.9\hsize,bb=0 0 576 432
-                ]{plots/burnin_cg.eps}
+\includegraphics[width=0.9\hsize]{figures/burnin_cg.eps}
 \caption{\label{fig:burnin} The $\chi^2$ of each step of the full MCMC
@@
 
-\item[5) {\bf DD}:] red dashed curve.  When the correct model is
+\item[5) {\bf DD}:] blue dashed curve.  When the correct model is
   fitted to a disc-source, the best fit $\chi^2=0.91$ is good and

### 2016-03-09 paper1/theory.tex
reprocessed eps files

@@
 \centering
-\includegraphics[width=0.9\hsize,bb=0 0 576 432
-		]{plots/IRIS567_path_2.eps}
+\includegraphics[width=0.9\hsize]{figures/IRIS567_path.eps}
 \caption{\label{fig:magnification_map} Magnification map used in the second part of the presented work.
@@
 \begin{figure}
-\includegraphics[width = .49\textwidth]{plots/infinite_fold_ar.eps}
+\includegraphics[width = .49\textwidth]{figures/infinite_fold_ar.eps}
 \caption{\label{fig:infinite_fold} Source profiles for Gaussian, disk,
@@
 \begin{figure}
-\includegraphics[width = .49\textwidth]{plots/figure_rhalf.eps}
+\includegraphics[width = .49\textwidth]{figures/figure_rhalf.eps}
 \caption{\label{fig:geom_crescent} Geometry of crescent sources.}
@@
 \centering
-	\includegraphics[width = 0.48\textwidth ,bb=0 0 576 520
-	]{plots/S1D_all.eps}
-	\includegraphics[width = 0.48\textwidth ,bb=0 0 576 520
-	]{plots/4source_magnification.eps}
+	\includegraphics[width = 0.48\textwidth]{figures/S1D_all.eps}
+	\includegraphics[width = 0.48\textwidth]{figures/4source_magnification.eps}
 \caption{\label{fig:lightcurve_gauss} Lightcurves of crescents, Gaussian and disk for sources with identical $S_0$ and $r_{1/2}$.
@@
 \centering
-	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
-	]{plots/S1D_var_a.eps}
-	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
-	]{plots/4avar_magnification.eps}
+	\includegraphics[width = 0.48\textwidth]{figures/S1D_var_a.eps}
+	\includegraphics[width = 0.48\textwidth]{figures/4avar_magnification.eps}
 \caption{\label{fig:lightcurve_disk} Lightcurves of crescent sources with $R_n$ = 0.4 $R_p$.}
@@
 \centering
-	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
-	]{plots/S1D_var_rn_a_poz.eps}
-	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
-	]{plots/5Rn_back_var_magnification.eps}
+	\includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_poz.eps}
+	\includegraphics[width = 0.48\textwidth]{figures/5Rn_back_var_magnification.eps}
 \caption{\label{fig:lightcurve_crescent_back} Lightcurves of backwards crescents with  a = 0.3 $R_p$.}
@@
 \centering
-	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
-	]{plots/S1D_var_rn_a_neg.eps}
-	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
-	]{plots/5Rn_forw_var_magnification.eps}
+	\includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_neg.eps}
+	\includegraphics[width = 0.48\textwidth]{figures/5Rn_forw_var_magnification.eps}
 \caption{\label{fig:lightcurve_crescent} Lightcurve of forwards crescents with a = -0.3 $R_p$.}
@@
 \centering
-	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
-	]{plots/ch_points.eps}
-        \includegraphics[width = 0.48\textwidth,bb=0 0 576 520
-	]{plots/ch_instances.eps}
+	\includegraphics[width = 0.48\textwidth]{figures/ch_points.eps}
+        \includegraphics[width = 0.48\textwidth]{figures/ch_instances.eps}
 \caption{\label{fig:char_points} The characteristic points and instances of a crescent source's one dimensional profile and microlensing lightcurve. }
@@
 \centering
-        \includegraphics[width = 0.48\textwidth]{plots/M87_image.eps}
+        \includegraphics[width = 0.48\textwidth]{figures/M87_image.eps}
 \caption{\label{fig:M87_image} The simulated image of M87's black hole sillouette obtained by \citep{2012MNRAS.421.1517D} from a GRMHD simulation presented in \citep{2009MNRAS.394L.126M}.
@@
 \centering
-	\includegraphics[width = 0.48\textwidth,bb=0 0 576 520
-	]{plots/M87_shape.eps}
-        \includegraphics[width = 0.48\textwidth,bb=0 0 576 520
-	]{plots/M87_lc.eps}
+	\includegraphics[width = 0.48\textwidth]{figures/M87_shape.eps}
+        \includegraphics[width = 0.48\textwidth]{figures/M87_lc.eps}
 \caption{\label{fig:M87_plots} The one dimensional profile obtained from the numerical integration along the ordonata axis of the source image (figure 9) is presented in the left subplot.

### 2016-03-10 paper1/intro.tex
editing

@@
 \section{introduction}
-More than half a century after their discovery, quasars still maintain the interest of a productive part of the astrophysical community. 
-Although several mechanisms have been proposed to explain how such systems function, a single mechanism is widely accepted at the present time. 
-The model was first proposed by \cite{1964ApJ...140..796S}, followed by \cite{1964SPhD....9..246Z} in the same year and \cite{1969Natur.223..690L} five years later.
-The present name of the model is the Super-Massive Black Hole (SMBH) paradigm.\\
 
-According to the paradigm, the object is powered through the accretion of matter from the proximal environment into the black hole. 
-The radiation emitted excites the surrounding medium which becomes detectable as the narrow line regions, broad band regions and torus. 
-Moreover, in the direction perpendicular to the accretion disk, where the medium is more transparent, jets will appear.
-When the medium between and observer and such an object is transparent a quasar can be observed \citep[e.g.,][]{1984RvMP...56..255B}. 
-In other words an active galactic nucleus (AGN) with one of the poles orientated towards Earth would be detected as a quasar by an observer. \\
+Active galactic nucleii are thought to be powered through the
+accretion of matter from the proximal environment into a supermassive
+black hole.  The radiation emitted excites the surrounding medium
+which becomes detectable as narrow line regions, broad line regions
+and optical continuum.  Moreover, in the direction perpendicular to
+the accretion disk, where the medium is more transparent, jets will
+appear.  If a jet is oriented towards the Earth, a quasar is observed
+\citep[e.g.,][]{1984RvMP...56..255B}.  While the basic mechanism
+\citep[originating in the work
+  of][]{1964ApJ...140..796S,1964SPhD....9..246Z,1969Natur.223..690L}
+is not in doubt, the central engines, near the event horizons of the
+black holes, remain to be probed.
 
-Through extensive observations done in the past 50 years, an important part of the set of assumptions and details  regarding the mechanism have been confirmed and uncovered. 
-Still, the black hole shadow and the surrounding luminous accreting material (the black hole silhouette) remain to be probed. 
-Advancements in the very long baseline interferometer observations have made possible since 2007, the detection of structures of a few Schwarzschild radii of Sgr~A* at the center of our Galaxy \citep{2008JPhCS.131a2055D}.
-Furthermore, jet launching structure near the supermasve black hole in M87 were resolved with the same observation tool \citep{2012Sci...338..355D}.
-With all the advancement, the current number of baselines does not allow the direct generation of images from the observations still it does allow the fitting of the data to predefined models \citep{2013MNRAS.434..765K}.
+For the black holes in the Galactic centre and the centre of M87
+---the two nearest objects that barely qualify as AGN--- the central
+engines (which are $<0.1\rm\,mas$ on the sky) are close to being
+resolved through very long baseline interferometry, which shows
+preliminary indications of the jet-launching structures
+\citep{2008JPhCS.131a2055D,2012Sci...338..355D,2013MNRAS.434..765K,2016arXiv160205527F}.
+Current data do not deliver images, but require fitting to to
+predefined models for the images.  A whole range of models have been
+applied, starting from simple geometric models to more complex
+physical models
+\citep{2008Natur.455...78D,2011ApJ...738...38B,2009ApJ...706..497M,2010ApJ...717.1092D}.
+The more complicated models nonetheless tend to predict a crescent
+shaped silhouette of the black hole.  This motivated
+\cite{2013MNRAS.434..765K} to use a simple geometric crescent model to
+fit the data, and argued that the crescent is nothing but the
+silhouette of the event horizon.
 
-A whole range of models have been applied starting from simple geometric models to more complex, physically sensible models (\cite{2008Natur.455...78D}, \cite{2011ApJ...738...38B}, \cite{2009ApJ...706..497M}, \cite{2010ApJ...717.1092D}, \cite{2016arXiv160205527F}).
-A significant fraction of the later models predict a crescent shaped silhouette of the black hole. 
-This motivated \cite{2013MNRAS.434..765K} to use a simple geometric crescent model to fit the data. 
-The succesful results encouraged the authors to speculate that the shadow of the black hole will be observed in the near future with the previously mentioned observation method. \\
+The great majority of quasars, however, lie at redshifts beyond 2
+\citep{2014A&A...563A..54P} and their central engines would be orders
+of magnitude smaller on the sky. The direct observations of the black
+hole silhouettes of quasars is far beyond foreseeable instrumentation.
+In the present paper we consider a possible indirect method,
+previously investigated by \cite{1999ApJ...524...49A}, through which
+caustics in gravitational microlensing could probe the black hole
+shadow and its proximal quasar environment.
 
-Hundreds of thousands of quasars have been detected up to date, mainly from the Sloan Digital Sky Survey.
-The great majority of them present a redshift larger than 2.15 of their spectra (\citep{2014A&A...563A..54P}).
-Therefore the typical distance to the objects is in the order of gigaparsecs or above. 
-A value that is orders of magnitude larger than the distance to M87. The direct observations of the black hole silhouettes of quasars require significant technological advances.\\
-
-We argue in the present paper that one can probe and study the black hole shadow and its proximal environment belonging to a quasar without having to resolve sub-event horizon scales. 
-We consider that this is possible by analysing the data from microlensing-events as suggested by \cite{1999ApJ...524...49A}.
 They argue that if the magnification function $\mu_t(p)$ from equation (6) corresponds to a perfect fold then from the function $F(t)$ present in the same equation one can extract the source profile.

### 2016-03-10 paper1/intro.tex
editing

@@
 previously investigated by \cite{1999ApJ...524...49A}, through which
-caustics in gravitational microlensing could probe the black hole
-shadow and its proximal quasar environment.
+gravitational microlensing could probe the black hole shadow and its
+proximal quasar environment.
 
-They argue that if the magnification function $\mu_t(p)$ from equation (6) corresponds to a perfect fold then from the function $F(t)$ present in the same equation one can extract the source profile.
-We argue in the current article that one can partially
-reconstruct the black hole silhouette up to a simplified crescent geometry. The crescent fit facilitates an estimation of the black hole mass.       
-Moreover, unlike with the few nearby AGNs, the study of the very large quasar population allows for the statistical refinement of the global properties of AGN by highlighting any particular
-characteristic that is present only in Sgr~A* and M87 central region. 
-Furthermore, the probing of the event-horizon-scale inner regions of a large number of objects will facilitate the study of the evolution of the previously mentioned region. \\
+Gravitational microlensing of quasars, reviewed in
+Section~\ref{sec:microlensing} below, refers to sharp changes in the
+observed brightness of quasars that have been lensed by intervening
+galaxies, without any changes in the intrinsic luminosity.
+Microlensing affects only the light from the innermost part of the
+quasar, such as the optical continuum
+\citep[e.g.,][]{2012A&A...544A..62S} and is a consequence of two
+things: the very small size of the central engine, and granularity of
+the mass distribution of a lensing galaxy due to stars.  The latter
+means that lensed brightness is not a smooth function of source
+position, but a complicated network of singular curves known as
+caustics.  Most astrophysical sources straddle several caustics, and
+hence their net brightness varies smoothly with location.  The central
+engine of quasars, however, is smaller.  Hence the lensed brightness
+undergoes sudden changes as a quasar crosses a caustic.
+Figure~\ref{fig:magnification_map} is a simple illustration.  The
+effect supplies an upper limit on the size of the central engine, and
+can also be used to study the mass distribution and kinematics of
+stars in the lensing galaxy as well
+\citep[e.g.,][]{2012ApJ...744..111P}.  Caustics have an additional
+remarkable feature: though they can be very complicated, they have
+some universal properties well-known from catastrophe theory.  In
+particular, very close to the simplest caustics (known as folds), the
+observed brightness is simple $\propto1/\sqrt p$ where $p$ is the
+transverse distance of the source from the caustic.  Since the
+observed brightness is a convolution of the magnification map with the
+source brightness, the lightcurve provides tomographic information
+about the source profile if the magnification map is known.
+\cite{1999ApJ...524...49A} showed that if the latter is the
+magnification near a pure fold, the brightness distribution of the
+source can in principle be deconvolved out of the light curve.
+Applied by \cite{2012MNRAS.423..676A} Pure folds, however, are not to
+be found in real systems.  In this paper we argue a simpler form of
+the idea holds more promise.  We show that if a quasar crosses not a
+pure but still fairly clean fold, such as in
+Figure~\ref{fig:magnification_map}, the simple crescent model of
+\cite{2013MNRAS.434..765K} approximating a black-hole silhouette could
+be recovered from the light curve.
 
@@
 
-Estimates of the Einstein radius for the particular case of a QSO at redshift 2 lensed by a compact stellar mass object at redshift 0.5 was found to be $10^{-6} \sqrt{M/M_\odot}\rm\,arcsec$ \citep{2001PASA...18..207W}.
 The Schwarzschild radius of the Sgr A* black hole, situated at 8~kpc from Earth \citep{1993ARA&A..31..345R} spans an angle of 10 microarcsec on the sky \citep{2008JPhCS.131a2055D}.

### 2016-03-10 paper1/theory.tex
editing

@@
-\section{Theory}
+\section{Microlensing}\label{sec:microlensing}
 
@@
 
-\subsection{Microlensing essentials}
+\subsection{Magnification}
 
@@
 distribution $\Sigma(\vec\theta)$ due to individual stars produces a
-caustic network on the scale of $\sqrt{GM_\odot D_L/c^2}$ or
-micro-arcsec.  Extended sources wash out this micro-caustic structure,
-but the optical-continuum source of quasars is even smaller.  Hence
-sharp changes in the brightness of lensed quasars are observable, as
-they cross the micro-caustics \citep[e.g.,][]{2012A&A...544A..62S}.
+caustic network on the scale of $\sqrt{GM_\odot D_L/c^2}$ or $10^{-6}
+\sqrt{M/M_\odot}\rm\,arcsec$ for typical lens and source redshifst
+\citep{2001PASA...18..207W}.  Extended sources wash out this
+micro-caustic structure, but the optical-continuum source of quasars
+is even smaller.
 

### 2016-03-11 paper1/code.tex
edited introduction

@@
-\section{Numerical microlensing magnification maps}
+\section{Numerical microlensing magnification maps}\label{sec:numerics}
 
@@
 caustic.
+
+\subsection{Microlensing a simulated image of M87}
+
+\begin{figure}
+\centering
+        \includegraphics[width = 0.48\textwidth]{figures/M87_image.eps}
+\caption{\label{fig:M87_image} The simulated image of M87's black hole
+  sillouette obtained by \citep{2012MNRAS.421.1517D} from a GRMHD
+  simulation presented in \citep{2009MNRAS.394L.126M}. The color bar
+  is linear and scaled to reach value 1 for the maximum
+  brightness. Points p2 and p3, mark the start and end of the
+  equivalent dark disk and are associated to the moment of their
+  overlap with the caustic t2 and t3.}
+\end{figure}
+
+\cite{2012MNRAS.421.1517D} have created a radiative image of M87 based on the GRMHD simulations presented in \citep{2009MNRAS.394L.126M}.
+The respective image, reproduced in figure \ref{fig:M87_image}, has been projected to a 1D profile associated to the perpendicular direction to a fold caustic approaching the image from the right.
+The projection is presented in the left panel of figure 10.
+The amplification values of the flux of light corresponding to a microlensing event are presented in the right panel of figure 10.
+In general the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions surrounding the luminous parts of the image have non-zero flux and thus are more extended than the simplified source model.
+
+\begin{figure}
+\centering
+	\includegraphics[width = 0.48\textwidth]{figures/M87_shape.eps}
+        \includegraphics[width = 0.48\textwidth]{figures/M87_lc.eps}
+\caption{\label{fig:M87_plots} The one dimensional profile obtained
+  from the numerical integration along the ordonata axis of the source
+  image (figure 9) is presented in the lower panel. The corresponding
+  lightcurve associated to the 1D profile is presented in the right
+  panel. The points p2 and p3 mark the start and end of the equivalent
+  dark disk and are associated to the moments of their overlap with
+  the caustic: t2 and t3.}
+\end{figure}

### 2016-03-11 paper1/intro.tex
edited introduction

@@
 hole silhouettes of quasars is far beyond foreseeable instrumentation.
-In the present paper we consider a possible indirect method,
-previously investigated by \cite{1999ApJ...524...49A}, through which
-gravitational microlensing could probe the black hole shadow and its
-proximal quasar environment.
+In the present paper we consider a possible indirect method, related
+to \cite{1999ApJ...524...49A}, through which gravitational
+microlensing could probe the black hole shadow and its proximal quasar
+environment.
 
@@
 the mass distribution of a lensing galaxy due to stars.  The latter
-means that lensed brightness is not a smooth function of source
-position, but a complicated network of singular curves known as
-caustics.  Most astrophysical sources straddle several caustics, and
-hence their net brightness varies smoothly with location.  The central
-engine of quasars, however, is smaller.  Hence the lensed brightness
-undergoes sudden changes as a quasar crosses a caustic.
-Figure~\ref{fig:magnification_map} is a simple illustration.  The
-effect supplies an upper limit on the size of the central engine, and
-can also be used to study the mass distribution and kinematics of
-stars in the lensing galaxy as well
+means that the brightness of a lensed object (or lensing
+magnification) is not a smooth function of source position, but a
+contains a complicated network of singular curves, known as caustics.
+Figure~\ref{fig:magnification_map} shows part of a magnification map
+with a few caustics.  The lensed brightness would be given by placing
+the source on such a magnification map and integrating the surface
+brightness weighted by the magnification.  Most astrophysical sources
+straddle several caustics, and hence their net brightness varies
+smoothly with location.  The central engine of quasars, however, is
+smaller than the typical spacing between caustics.  As a result, the
+lensed brightness undergoes sudden changes as a quasar crosses a
+caustic.  The effect supplies an upper limit on the size of the
+central engine, and can also be used to study the mass distribution
+and kinematics of stars in the lensing galaxy as well
 \citep[e.g.,][]{2012ApJ...744..111P}.  Caustics have an additional
@@
 particular, very close to the simplest caustics (known as folds), the
-observed brightness is simple $\propto1/\sqrt p$ where $p$ is the
-transverse distance of the source from the caustic.  Since the
-observed brightness is a convolution of the magnification map with the
-source brightness, the lightcurve provides tomographic information
-about the source profile if the magnification map is known.
-\cite{1999ApJ...524...49A} showed that if the latter is the
-magnification near a pure fold, the brightness distribution of the
-source can in principle be deconvolved out of the light curve.
-Applied by \cite{2012MNRAS.423..676A} Pure folds, however, are not to
-be found in real systems.  In this paper we argue a simpler form of
-the idea holds more promise.  We show that if a quasar crosses not a
-pure but still fairly clean fold, such as in
-Figure~\ref{fig:magnification_map}, the simple crescent model of
-\cite{2013MNRAS.434..765K} approximating a black-hole silhouette could
-be recovered from the light curve.
+magnification is approximately constant on one side and
+$\propto1/\sqrt p$ where $p$ is the transverse distance of the source
+from the caustic.  This property will be exploited later.
 
-Gravitational microlensing of QSOs have been predicted as early as 1981 when \cite{1981ApJ...243..140G} suggested that 
-``if haloes of galaxies were composed of stars with masses less than  0.1 $M_\odot$, then these stars acting as individual lenses would produce fluctuations of the order of 
-unity in the intensities of the QSO images on time scales of 1 -14 years '' . 
-What creates the magnification event is the relative and independent motion of the lens and source with respect to the observer.
-Since the respective lens is surrounded by similar objects one would expect to observe during a long enough time interval
-multiple microlensing events of the same source, the QSO. 
-The sequential alignment of the source and observer with different lensing objects will result in an apparently random increase of the QSO light flux analogous to the scintillation events in a meter of radioactivity.\\
+In Section~\ref{sec:source-models} we introduce the three source
+profiles used in our subsequent models and simulations: a
+constant-brightness disc, a circular Gaussian, and the crescent source
+introduced by \citep{2013MNRAS.434..765K}.  The latter is simply a
+constant-brightness disc with a smaller, non-concentric disc cut out
+of it.  We also derive the half-light radius for a crescent.  The
+half-light radius can characterise the source size for all three types
+of source.
 
-The Schwarzschild radius of the Sgr A* black hole, situated at 8~kpc from Earth \citep{1993ARA&A..31..345R} spans an angle of 10 microarcsec on the sky \citep{2008JPhCS.131a2055D}.
-If the quasars observed at $z=2$ would have similar linear sizes, the aparent size of the luminous part of the QSO would be many orders of magnitude below the 1 microarcsec level.
-Therefore the aparent size would be significantly smaller than the corresponding Einstein radius of a solar mass lens.
-The magnitude differences between the Einstein angle and the aparrent size of the bright part of the QSO motivates the introduction in the first part of the paper of simplified models for     
-calculating the lightcurves of QSO magnification events.
-The simplified models are introduced in the theory section of the present paper.
-The theory (II) section of the present article starts with an introduction to the basic microlensing theory in which we present the general microlensing equation followed by the concept of magnification.
-We underline that for the particular case of micro gravitational lensing,
-the images are not resolved and thus the magnifications of each images contribute to a total magnification. 
-Furthermore, the concept of critical curves and caustics are introduced. Afterwards we focus on a particular catastrophe, namely the fold. 
-We choose to study events related to the previously mentioned singularity since the magnification events caused by the crossing of such boundaries are most probable.
-A quick inspection of figure \ref{fig:magnification_map} by the reader should motivate the previous statement.
-In continuation we introduce the commonly used result involving the magnification of a point-like source near a fold singularity. 
-Namely that the magnification is approximately linear outside the caustic and decreases from infinity with the square root of the distance to the boundary when inside the caustic.  
-In the last section of the theory (II) we present the general equations for estimating the flux received from an extended source near a fold. 
-The only elements required to compute the flux as a function of time are the two dimensional brightness function of the object and the magnification of a point-like source. 
-One of the core assumptions made when developing the simple models is that the Einstein angle is orders of magnitude greater than the apparent size of the source. 
-Due to this assumption one can imagine the caustic as an infinite wall to be crossed by the source as presented in \ref{fig:infinite_fold}. 
-The consequence of the limit is that the information regarding the profiles of the source along any direction parallel to the boundary value is reduced to the corresponding integral value. 
-The effect allows the introduction of a one dimensional flux function, which is the two dimensional flux integrated along the axis parallel to the boundary. 
-The one dimensional profile contains exhaustively all the information that can be extracted  from the lens generated light-curve \\
+Section~\ref{sec:fold-crossing} shows the light curves that result
+when each of the model sources crosses an ideal fold.  This would
+apply in Figure~\ref{fig:magnification_map} to sources along the path
+AB or BC, for sources small enough that the curvature of the caustics
+is negligible.  With this assumption one can imagine the caustic as an
+infinite wall to be crossed by the source as presented in
+Figure~\ref{fig:infinite_fold}.  The source brightness distribution
+parallel to the caustic naturally makes no difference to the
+observable brightness; each source can be replaced by an effective
+one-dimensional source profile, by flattening the source so it becomes
+perpendicular to the caustic.  In principle, the effective
+one-dimensional brightness profile could be recovered from the light
+curve by deconvolution.  \cite{1999ApJ...524...49A} modelled this
+profile as the result of a circular accretion disc seen through the
+spacetime around a Kerr black hole, and \cite{2012MNRAS.423..676A}
+have applied the idea to observed light curves to infer properties of
+quasar accretion discs.  In this work we take a simpler but arguably
+more robust approach: we study features in the light curves
+characteristic of a crescent-like source which in turn would indicate
+a black-hole silhouette.  Figure~\ref{fig:char_points} shows the
+qualitative features: there is a period during which the dark cutout
+disc is crossing the caustic, and before and after there are periods
+when the only the bright parts of the crescent are in transit across
+the caustic.  The details depend on the orientation of the crescent,
+but basically the dark disc causes a rising light curve to plateau or
+dip.
 
-In the third section we introduce the three source profiles used in our models and simulations.
-We start with two commonly used and simple sources.
-Namely the gaussian brightness distribution and the constant intensity disk.
-In addition we use a geometric crescent-shaped source effectively the same as the one introduced by \citep{2013MNRAS.434..765K} with constant surface brightness.
-\\ The fourth section contains the results of our simplified models calculations, including the numerical equation for lightcurve calculations and the corresponding plots.
-In order to facilitate comparison between the lightcurves of the three distinct sources, we constrain the luminosity and half-light radius of all sources to have the same value. 
-Furthermore, towards the end of the section, the microlensing light curves were generated for more complex images of black hole silhouettes. 
-Namely the images of Sgr A* and M87 compiled in the paper \citep{2015MNRAS.446.1973R} and based on simulations presented in \citep{2009MNRAS.394L.126M} article were used as sources.
+The crescent characteristics identified in
+Figure~\ref{fig:char_points} are from an idealised source crossing an
+idealised caustic.  In Section~\ref{sec:numerics} we see what happens
+when we replace both of these with numerical simulations.  The ideal
+fold-transit is replaced by the path AB in
+Figure~\ref{fig:magnification_map}.  The simple crescent is replaced
+by a source (Figure~\ref{fig:M87_image}) from published simulations of
+the black-hole environment of M87
+\citep{2009MNRAS.394L.126M,2012MNRAS.421.1517D}.  The simulated light
+curve so obtained (lower panel of Figure~\ref{fig:M87_plots}) still
+shows, albeit faintly, features from the crescent.
 
-We continue to the second part of the paper, where we have advanced the level of complexity of the amplification map.
-Using the Wambsganss gravitational lensing code \citep{1990LNP...360..186W} a realistic magnification map is generated.
-The numerical method, hierarchical tree structure and backwards raytracing, underlying the microlensing code is treated in section five.
-Aside from a basic review of the physical and numerical principles of the code, the description of its generalization from disk shaped source images to crescent ones can be found in the Microlensing code and magnification map as well.
-Furthermore, the mass distribution in the lens plane, chosen to produce the magnification map used in this analysis, is introduced and motivated.
-
-In section \ref{sec:fitting}, we carried out a likelihood analysis for the parametrisation of the different source shapes and study the degeneracies involved amongst them.
-To do so, we mocked the datasets corresponding to each source shape and contaminate them with different noise components.
-Finally we run MCMC to explore the parameter space of the source shape and some nuisance parameters.
-The motivation of this part is to explore the possibility if using a simple toy model is it possible to differentiate between symmetric and asymmetric sources based on maximum likelihood analysis.
-
-Finally, in the Discussion (VII) section, the main results of the current work and their implications are discussed.
+In Section~\ref{sec:fitting} we carry out source fitting to
+lightcurves, with both noise and systematic errors are present.  We
+generate three lightcurves by taking a uniform disc, a circular
+Gaussian, and a crescent across the path AB in
+Figure~\ref{fig:magnification}, and then adding noise.  The path
+simulates crossing a clean but not ideal fold.  In addition, we
+generate templates lightcurves by running the three source types, with
+various parameter values, across the path CB.  That is, the templates
+come from a similar but not identical caustic, thus deliberately
+generating a systematic error.  We then fit the noisy lightcurves to
+the templates using Markov chain Monte-Carlo.  We find that the
+correct source type can be inferred from the $\chi^2$ values.  The
+parameter values can also be inferred. The fitting errors are larger
+than the formal uncertainties, which is expected in the presence of
+systematic errors, but still appear acceptable.

### 2016-03-11 paper1/theory.tex
edited introduction

@@
 \includegraphics[width=0.9\hsize]{figures/IRIS567_path.eps}
-\caption{\label{fig:magnification_map} Magnification map used in the second part of the presented work.
-The white line marks the trajectory of the center of the sources for which the lightcurves presented in figure 12 are generated.}
+\caption{\label{fig:magnification_map} Magnification map used later in
+  this paper (Section~\ref{sec:fitting}). The white line marks the
+  trajectory of the center of the sources for which the lightcurves
+  presented in figure 12 are generated.  Note that this represents an
+  atypically simple region of any realistic magnification map.}
 \end{figure}
@@
 
-\section{Models for extended sources}
+\section{Models for extended sources}\label{sec:source-models}
 
@@
 
-\section{Lightcurves of the extended sources during fold crossing}
+\section{Lightcurves of the extended sources during fold crossing}\label{sec:fold-crossing}
 
-Using equation \ref{eqn:ft} and the one-dimensional flux function presented in the previous section one can compute numerically the lightcurves of the three extented sources for the simplified 
+Using equation \ref{eqn:ft} and the one-dimensional flux function
+presented in the previous section one can compute numerically the
+lightcurves of the three extented sources for the simplified
 infinite-wall-caustics model.
@@
 In the opposite case the peak appears before the end of the bright disc reaches the caustic and shape is smoother. \\
-
-\subsection{Microlensing a simulated image of M87}
-
-\begin{figure}
-\centering
-        \includegraphics[width = 0.48\textwidth]{figures/M87_image.eps}
-\caption{\label{fig:M87_image} The simulated image of M87's black hole sillouette obtained by \citep{2012MNRAS.421.1517D} from a GRMHD simulation presented in \citep{2009MNRAS.394L.126M}.
-The color bar is linear and scaled to reach value 1 for the maximum brightness.
-Points p2 and p3, mark the start and end of the equivalent dark disk and are associated to the moment of their overlap with the caustic t2 and t3.}
-\end{figure}
-
-\cite{2012MNRAS.421.1517D} have created a radiative image of M87 based on the GRMHD simulations presented in \citep{2009MNRAS.394L.126M}.
-The respective image, reproduced in figure \ref{fig:M87_image}, has been projected to a 1D profile associated to the perpendicular direction to a fold caustic approaching the image from the right.
-The projection is presented in the left panel of figure 10.
-The amplification values of the flux of light corresponding to a microlensing event are presented in the right panel of figure 10.
-In general the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions surrounding the luminous parts of the image have non-zero flux and thus are more extended than the simplified source model.
-
-\begin{figure}
-\centering
-	\includegraphics[width = 0.48\textwidth]{figures/M87_shape.eps}
-        \includegraphics[width = 0.48\textwidth]{figures/M87_lc.eps}
-\caption{\label{fig:M87_plots} The one dimensional profile obtained from the numerical integration along the ordonata axis of the source image (figure 9) is presented in the left subplot.
-The corresponding lightcurve associated to the 1D profile is presented in the right subplot.
-The points p2 and p3 mark the start and end of the equivalent dark disk and are associated to the moments of their overlap with the caustic: t2 and t3.}
-\end{figure}

### 2016-03-11 paper1/intro.tex
typo fix

@@
 Gaussian, and a crescent across the path AB in
-Figure~\ref{fig:magnification}, and then adding noise.  The path
+Figure~\ref{fig:magnification_map}, and then adding noise.  The path
 simulates crossing a clean but not ideal fold.  In addition, we

### 2016-03-13 paper1/discussion.tex
Changes

@@
 \section{Discussion}
-[Mihai's contribution]\\
+
 In the current paper we simulate and study the resulting microlensing lightcurves of geometric crescent-shaped sources 
@@
 The later's gravity would cause a shadow in the bright region limited by the extent of the event horizon of the black hole.
-Therefore the radius of the bright disk would provide an estimate of the size of the accretion disk and the radius of the dark disc would provide and estimate of the Schwarzschild radius of the black hole.
-Moreover a value of the Schwarzschild radius
-can be used to estimate the mass of the black hole $M \approx \frac{\Delta t_{dark}  v_p c^2}{2G}$. In the previous 
+\textbf{Therefore the radius of the bright disk would provide an estimate of the size of the accretion disk and the radius of the dark disc would provide and estimate of the gravitationally magnified Schwarzschild radius of the black hole $R_{S}^{magnified} = \Delta t_{dark}*v_p$.
+By $R_{S}^{magnified}$ we refer to the apparent Schwarzschild radius which is larger than the real value at large distances due to the black hole's own gravity.
+Moreover the gravitationally magnified value of the Schwarzschild radius is a monotonic function of the black hole's mass.
+Therefore it
+can be used to estimate the mass of the black hole if it was not rotating.} In the previous 
 expression the $\Delta t_{dark}$ denotes the period of overlap between the black hole shadow and fold caustic. 
@@
 
-Furthermore, a simulated image of M87 presented in \citep{2012MNRAS.421.1517D} (figure 9) has been microlensed.
-On the resulting lightcurve the instances corresponding to the start and end of the black hole shadow and caustic overlap were distinguishable as can be seen in figure 10.
-
-The parameters whose values cannot be determined due to the loss of information from the directions parallel to the 
-caustic could be obtained in the eventuality in which the same source crosses multiple caustics that are not parallel. 
-Multiple crossing of caustics can reveal details of the one dimensional flux profile corresponding to multiple distinct 
-directions which would allow the reconstruction of the two dimensional profile analogous to the process through which an image of a CT scan is obtained.
-
-In case of a high quality lightcurve with insignificant noise and measurement errors, the parameters can be obtained 
+\textbf{In case of a high quality lightcurve with insignificant noise and measurement errors, the parameters can be obtained 
 by simply identifying the characteristic instances of time without making use of the actual values of the magnification
@@
 crescent source model assumption. If at each of the three steps the data rejects the model then the source cannot be approximated by any of the three models. 
-Otherwise if at one of the steps the data does not reject the model then the respective model is a good approximation.
+Otherwise if at one of the steps the data does not reject the model then the respective model is a good approximation.   }
 
-[Contributions from the authors]\\
-P.S. provided the original idea and plan for the research project as well as multiple contributions to the analysis. M.T. simulated and studied the microlensing 
-lightcurves for the different source profiles discussed for clean data and approximated fold magnification amplification. J.W. provided the training and the
-numerical code used by M. R. to create the magnification map and corresponding lighcurves used in the MCMC analysis, analysis that was performed by 
-I. M. in the last part of the presented work.
+\textbf{Furthermore, a simulated image of M87 presented in \citep{2012MNRAS.421.1517D} (figure 9) has been microlensed.
+On the resulting lightcurve the instances corresponding to the start and end of the black hole shadow and caustic overlap were distinguishable as can be seen in figure 10.}
+
+The parameters whose values cannot be determined due to the loss of information from the directions parallel to the 
+caustic could be obtained in the eventuality in which the same source crosses multiple caustics that are not parallel. 
+Multiple crossing of caustics can reveal details of the one dimensional flux profile corresponding to multiple distinct 
+directions which would allow the reconstruction of the two dimensional profile analogous to the process through which an image of a CT scan is obtained.

### 2016-03-13 paper1/intro.tex
Changes

@@
 
-Active galactic nucleii are thought to be powered through the
+Active galactic nuclei are thought to be powered through the
 accretion of matter from the proximal environment into a supermassive
@@
 ---the two nearest objects that barely qualify as AGN--- the central
-engines (which are $<0.1\rm\,mas$ on the sky) are close to being
+engines (\textbf{which are $<0.1\rm\,mas$ on the sky}) are close to being
 resolved through very long baseline interferometry, which shows
@@
 \citep{2008JPhCS.131a2055D,2012Sci...338..355D,2013MNRAS.434..765K,2016arXiv160205527F}.
-Current data do not deliver images, but require fitting to to
+Current data do not deliver images, but require fitting to
 predefined models for the images.  A whole range of models have been
@@
 things: the very small size of the central engine, and granularity of
-the mass distribution of a lensing galaxy due to stars.  The latter
-means that the brightness of a lensed object (or lensing
-magnification) is not a smooth function of source position, but a
-contains a complicated network of singular curves, known as caustics.
+the mass distribution of a lensing galaxy due to stars.  \textbf{The latter
+means that the local magnification is not a smooth function of source 
+position. It contains a complicated network of singular curves, known as caustics.}
 Figure~\ref{fig:magnification_map} shows part of a magnification map
@@
 systematic errors, but still appear acceptable.
+
+\textbf{Finally in Section 7 we discuss in more depth the 
+implications of the results presented in the previously mentioned sections.
+One of the most interesting implication is the possibility to estimate the black hole
+mass from the reconstructed parameters. This further requires good approximations of the 
+relative transversal velocity between the quasar and the lens. Proper constraints can be set
+with independent observations of the stellar structure that contains the gravitational lens.}
+
+\textbf{The thorough study of the possibility to reconstruct the quasar's structural 
+parameters from light-curves containing multiple microlensing events represents 
+the target of future work. This will most probably require the use of powerful
+statistical tools. Another path for future work is the designing of an observation regime 
+best suited for acquiring the necessary microlensed light-curve .    }

### 2016-03-13 paper1/theory.tex
Changes

--- 2016-03-10 paper1/theory.tex
@@
 The other two, p2 and p3, mark the boundaries of the dark disc component.
-Since $S_1D$ is a projection of $S_2D$ on a line perpendicular to the caustic, the following relation holds:
+Since $S_{1D}$ is a projection of $S_{2D}$ on a line perpendicular to the caustic, the following relation holds:
 \begin{equation}
@@
         \includegraphics[width = 0.48\textwidth]{figures/M87_image.eps}
-\caption{\label{fig:M87_image} The simulated image of M87's black hole sillouette obtained by \citep{2012MNRAS.421.1517D} from a GRMHD simulation presented in \citep{2009MNRAS.394L.126M}.
-The color bar is linear and scaled to reach value 1 for the maximum brightness.
-Points p2 and p3, mark the start and end of the equivalent dark disk and are associated to the moment of their overlap with the caustic t2 and t3.}
+\caption{\label{fig:M87_image} The simulated image of M87's black hole
+  sillouette obtained by \citep{2012MNRAS.421.1517D} from a GRMHD
+  simulation presented in \citep{2009MNRAS.394L.126M}. The color bar
+  is linear and scaled to reach value 1 for the maximum
+  brightness.}
 \end{figure}
@@
         \includegraphics[width = 0.48\textwidth]{figures/M87_lc.eps}
-\caption{\label{fig:M87_plots} The one dimensional profile obtained from the numerical integration along the ordonata axis of the source image (figure 9) is presented in the left subplot.
-The corresponding lightcurve associated to the 1D profile is presented in the right subplot.
-The points p2 and p3 mark the start and end of the equivalent dark disk and are associated to the moments of their overlap with the caustic: t2 and t3.}
+\caption{\label{fig:M87_plots} The one dimensional profile obtained
+  from the numerical integration along the ordonata axis of the source
+  image (figure 9) is presented in the lower panel. The corresponding
+  lightcurve associated to the 1D profile is presented in the right
+  panel. The points p2 and p3 mark the start and end of the equivalent
+  dark disk and are associated to the moments of their overlap with
+  the caustic: t2 and t3.}
 \end{figure}

### 2016-03-14 paper1/discussion.tex
Discussion paragraph switching

@@
 With the different source profiles and their corresponding lightcurves studied we can change our point of view of the system to that of an observer. 
-The observer would basically detect only the lightcurve of such as source.
-As described in section 4.3 (Lightcurve of a crescent shaped source) the timing of the onset and offset of the previously described periods can be used to estimate the values of the radii and one of the displacement parameters when assuming a geometric crescent.
-All quantities can be estimated in terms of the relative velocity of the source in a direction perpendicular to the caustic.\\
+The observer would basically detect only the lightcurve of such a source. As described in section 4.3 the timing of the onset and offset of the previously 
+described periods can be used to estimate the values of the radii and one of the displacement parameters when assuming a geometric crescent. 
+All quantities can be estimated in terms of the relative velocity of the source in a direction perpendicular to the caustic.
 
-The previously mentioned abstract parameters can be related to physical quantities specific to the central region of a quasar.
-Thus the luminous region would correspond to the bright accretion disc that surrounds the black hole.
-The later's gravity would cause a shadow in the bright region limited by the extent of the event horizon of the black hole.
-\textbf{Therefore the radius of the bright disk would provide an estimate of the size of the accretion disk and the radius of the dark disc would provide and estimate of the gravitationally magnified Schwarzschild radius of the black hole $R_{S}^{magnified} = \Delta t_{dark}*v_p$.
-By $R_{S}^{magnified}$ we refer to the apparent Schwarzschild radius which is larger than the real value at large distances due to the black hole's own gravity.
-Moreover the gravitationally magnified value of the Schwarzschild radius is a monotonic function of the black hole's mass.
-Therefore it
-can be used to estimate the mass of the black hole if it was not rotating.} In the previous 
-expression the $\Delta t_{dark}$ denotes the period of overlap between the black hole shadow and fold caustic. 
-$v_p$ denotes the component of the relative velocity of the source and fold which is perpendicular to the caustic.  
-The respective velocity is an unknown, though it can be constrained on a case by case basis to an order of magnitude or even better. This would require the 
-study of the dynamics of the stellar structure which contains the gravitational lens.
-A better estimate of the relative velocity would facilitate a better estimate of the effective non-rotating black hole mass associated to the black hole shadow.
+Furthermore, a simulated image of M87 presented in \citep{2012MNRAS.421.1517D} (figure 9) has been microlensed. On the resulting lightcurve the instances 
+corresponding to the start and end of the black hole shadow and caustic overlap were distinguishable as can be seen in figure 10.
 
-\textbf{In case of a high quality lightcurve with insignificant noise and measurement errors, the parameters can be obtained 
+In case of a high quality lightcurve with insignificant noise and measurement errors, the parameters can be obtained 
 by simply identifying the characteristic instances of time without making use of the actual values of the magnification
@@
 crescent source model assumption. If at each of the three steps the data rejects the model then the source cannot be approximated by any of the three models. 
-Otherwise if at one of the steps the data does not reject the model then the respective model is a good approximation.   }
+Otherwise if at one of the steps the data does not reject the model then the respective model is a good approximation.
 
-\textbf{Furthermore, a simulated image of M87 presented in \citep{2012MNRAS.421.1517D} (figure 9) has been microlensed.
-On the resulting lightcurve the instances corresponding to the start and end of the black hole shadow and caustic overlap were distinguishable as can be seen in figure 10.}
+The previously mentioned abstract parameters can be related to physical quantities specific to the central region of a quasar. 
+As such the luminous region would correspond to the bright accretion disc that surrounds the black hole. The later's gravity would 
+cause a shadow in the bright region limited by the extent of the event horizon of the black hole. \textbf{Therefore the radius 
+of the bright disk would provide an estimate of the size of the accretion disk and the radius of the dark disc would provide
+ and estimate of the gravitationally magnified Schwarzschild radius of the black hole $R_{S}^{magnified} = \Delta t_{dark} \cdot v_p$.
+By $R_{S}^{magnified}$ we refer to the apparent Schwarzschild radius which is larger than the real value at large distances due to the
+black hole's own gravity.  Moreover the gravitationally magnified value of the Schwarzschild radius is a monotonic function of the black 
+hole's mass. Therefore it can be used to estimate the mass of the black hole if it was not rotating.} In the previous 
+expression the $\Delta t_{dark}$ denotes the period of overlap between the black hole shadow and fold caustic. 
+$v_p$ denotes the component of the relative velocity of the source and fold which is perpendicular to the caustic.  
+The respective velocity is an unknown, though it can be constrained on a case by case basis to an order of magnitude 
+or even better. This would require the study of the dynamics of the stellar structure which contains the gravitational lens. 
+A better estimate of the relative velocity would facilitate a better estimate of the effective non-rotating black hole 
+mass associated to the black hole shadow.
 

### 2016-03-14 paper1/intro.tex
edits

--- 2016-03-11 paper1/intro.tex
@@
 the mass distribution of a lensing galaxy due to stars.  The latter
-means that the brightness of a lensed object (or lensing
-magnification) is not a smooth function of source position, but a
-contains a complicated network of singular curves, known as caustics.
-Figure~\ref{fig:magnification_map} shows part of a magnification map
-with a few caustics.  The lensed brightness would be given by placing
-the source on such a magnification map and integrating the surface
-brightness weighted by the magnification.  Most astrophysical sources
-straddle several caustics, and hence their net brightness varies
-smoothly with location.  The central engine of quasars, however, is
-smaller than the typical spacing between caustics.  As a result, the
-lensed brightness undergoes sudden changes as a quasar crosses a
-caustic.  The effect supplies an upper limit on the size of the
-central engine, and can also be used to study the mass distribution
-and kinematics of stars in the lensing galaxy as well
+means that the local magnification is not a smooth function of source
+position. It contains a complicated network of singular curves, known
+as caustics.  Figure~\ref{fig:magnification_map} shows part of a
+magnification map with a few caustics.  The lensed brightness would be
+given by placing the source on such a magnification map and
+integrating the surface brightness weighted by the magnification.
+Most astrophysical sources straddle several caustics, and hence their
+net brightness varies smoothly with location.  The central engine of
+quasars, however, is smaller than the typical spacing between
+caustics.  As a result, the lensed brightness undergoes sudden changes
+as a quasar crosses a caustic.  The effect supplies an upper limit on
+the size of the central engine, and can also be used to study the mass
+distribution and kinematics of stars in the lensing galaxy as well
 \citep[e.g.,][]{2012ApJ...744..111P}.  Caustics have an additional
@@
 
-The crescent characteristics identified in
-Figure~\ref{fig:char_points} are from an idealised source crossing an
-idealised caustic.  In Section~\ref{sec:numerics} we see what happens
-when we replace both of these with numerical simulations.  The ideal
-fold-transit is replaced by the path AB in
-Figure~\ref{fig:magnification_map}.  The simple crescent is replaced
+The simple crescent is replaced
 by a source (Figure~\ref{fig:M87_image}) from published simulations of
@@
 shows, albeit faintly, features from the crescent.
+The crescent characteristics identified in
+Figure~\ref{fig:char_points} are from an idealised source crossing an
+idealised caustic.
+
+In Section~\ref{sec:numerics} we see what happens
+when we replace both of these with numerical simulations.  The ideal
+fold-transit is replaced by the path AB in
+Figure~\ref{fig:magnification_map}.
 
@@
 systematic errors, but still appear acceptable.
+
+Finally in Section~\ref{sec:discussion} we discuss in more depth the
+implications of the results presented in the previously mentioned
+sections.  One of the most interesting implication is the possibility
+to estimate the black hole mass from the reconstructed
+parameters. This further requires good approximations of the relative
+transversal velocity between the quasar and the lens. Proper
+constraints can be set with independent observations of the stellar
+structure that contains the gravitational lens.
+
+The thorough study of the possibility to reconstruct the quasar's
+structural parameters from light-curves containing multiple
+microlensing events represents the target of future work. This will
+most probably require the use of powerful statistical tools. Another
+path for future work is the designing of an observation regime best
+suited for acquiring the necessary microlensed light-curve.

### 2016-03-14 paper1/ms.tex
another set of corrections
Small title typo
edits

--- 2015-07-09 paper1/ms.tex
@@
 
-\title{Microlensing as a possible probe of the internal structure of quasars}
+%\title[Quasar microlensing and its internal structure]{Properties of quasar microlensing light curves: probing its internal structure}
+\title{Probing the internal structures of quasars using microlensing events}
 
-\author[Tomozeiu et al]{Mihai Tomozeiu$^{1,2}$,Irshad Mohammed$^1$, Manuel Rabold$^{1,2}$,
+\author[Tomozeiu et al]{Mihai Tomozeiu\thanks{mihai@physik.uzh.ch},$^{1,2}$ 
+Irshad Mohammed,$^{1,2,3}$ 
+Manuel Rabold,$^{1,2}$
+Prasenjit Saha,$^{1,2}$  
 \newauthor
-Prasenjit Saha$^1$ and Joachim Wambsganss$^3$\\
+and Joachim Wambsganss$^4$\\
 $^1${Physik-Institut, University of Zurich, Winterthurerstrasse 190,
@@
   Winterthurerstrasse 190, 8057 Zurich, Switzerland} \\
-$^3${Zentrum f\"ur Astronomie der Universit\"at Heidelberg,
+$^3${Theoretical Astrophysics Group, Fermi National Accelerator Laboratory, Batavia, IL 60510, USA}\\
+$^4${Zentrum f\"ur Astronomie der Universit\"at Heidelberg,
   M\"onchhofstrasse 12--14, 69120 Heidelberg, Germany}
@@
 
-\input{code}
-
-\input{fitting}
+\input{numerics}
 

### 2016-03-14 paper1/numerics.tex
edits

--- 2016-03-09 paper1/fitting.tex
@@
-\section{Model fitting and likelihood analysis}\label{sec:fitting}
+\section{Fitting mock data}\label{sec:numerics}
 
-In this section we simulate the fitting of source models to
-lightcurves in the presence of random and systematic errors.  We
-generate mock data for three source models --- a crescent, a uniform
+Having used the simple model of an ideal fold to gain insight, we now
+move to the numerical computation of microlensing light curves with
+realistic mass distributions, and then fit source models to
+lightcurves in the presence of random and systematic errors.
+
+First we generate microlensing magnification maps using the
+microlensing code by \cite{1999A&A...346L...5W}.  We then generate
+mock light curves for three source models --- a crescent, a uniform
 disc and a Gaussian disc --- and then fit each mock light curve to all
@@
 Markov-chain Monte Carlo (MCMC).
+
+\subsection{Numerical microlensing magnification maps}
+
+The code uses a ray-shooting technique to compute the gravitational
+lensing effect of a mass distribution consisting of (a)~a smooth
+component and (b)~a random distribution of point masses representing
+stars.  The ray shooting maps a grid of $\vec\theta$ to $\vec\beta$
+using the lens equation (\ref{eqn:lens}).  That is, the rays are shot
+from the observer back to the source.
+
+For the computation of the individual deflection angles
+$\vec\alpha(\vec\theta)$ a heirarchical or tree method is used.  The
+positions of all lensing masses are put into a grid of $\vec\theta$.
+Each grid cell is subdivided into four smaller squares recursively
+until every cell contains only one mass.  Nearby masses are added
+individually, while distance masses are clumped into larger grid cells
+whose net contribution is approximated by its first few multipole
+moments.  Scaled units are used, with the constant pre-factor in the
+deflection angle (\ref{eqn:alpha}) separated out.
+
+The result of ray shooting is a pixel map on the $\vec\beta$ plane of
+the number of lightrays which arrive at the source plane from a
+particular observer.  This intermediate result is effectively a
+magnification map on the source plane.  Once the map is created, the
+lightcurve can be obtained by specifying the transit path of the
+source across the map.  At each point on this transit line, the code
+computes equation (\ref{eqn:ft2d}) convolving the brightness
+distribution of the source with the magnification pattern of the map.
+In real life, not only are both lens and source moving, but the lens
+lens configuration, and with it the magnification pattern, is also
+changing with time.  While the first subtlety is taken care of by a
+coordinate transformation in this analysis, for the second one the
+lens configuration is assumed to be constant in time.
+
+When generating the magnification map depicted in
+Figure~\ref{fig:magnification_map}, which is used in our analysis
+below, only two point masses were included.  This was done in order to
+have clean fold caustics.  For the computation of the actual
+lightcurve, the code was modified to also allow for crescent shaped
+images specified through the parameter set $R_p,R_n,a,b$. Here $R_p$
+denotes the outer radius of the crescent and $R_n$ the inner one. The
+orientation of the source image with respect to the magnification map
+is specified by the parameters $a$ and $b$ as the shift of the center
+of the inner disk from the center of the outer disk in $x$- and
+$y$-direction respectively. In the original version of the code,
+gaussian and disk shaped images where already implemented. Those are
+completely characterised by the single parameter $R_p$. The values of
+the parameters are specified in pixel units corresponding to the
+magnification map. Further one needs to specify the start and end
+point coordinates of the path, which the center of the source image
+follows through the magnification map (see the depiction in figure
+\ref{fig:magnification_map}). Hence the points along this path are
+specified through the number of timesteps for which the computation of
+the brightness is to be carried out. Those points correspond to the
+actual measurement of the brightness of an object in the observational
+case. For each timestep the two-dimensional convolution of equation
+\ref{eqn:ft2d} is carried out numerically for the position of the
+source on the magnification map.
+
+For the purpose of this analysis it was desirable to mimic the
+analytical behaviour of a simple fold as much as possible, for
+comparing the numerical result with the analytical one, therefore the
+path of the source was chosen, so that it intersects the border of the
+caustic perpendicularly, and on a point where the border is a fold
+caustic.
 
@@
 
-\subsection{Mock datasets}
+\subsection{Mock light curves}
 
@@
 \S\ref{subsec:crescent}) $R_p$ and $R_n$ being the radii of the outer
-and inner circles and $(\alpha,\beta)$ being the coordinates of the inner
-circle with respect to the centre of the outer circle in the coordinate system associated with the image in Figure 1 and not associated 
-with the caustic surface as defined in the previous sections.  The parameter values were
+and inner circles and $(\alpha,\beta)$ being the coordinates of the
+inner circle with respect to the centre of the outer circle in the
+coordinate system associated with the image in Figure 1 and not
+associated with the caustic surface as defined in the previous
+sections.  The parameter values were
 \begin{equation}
@@
 points regularly spaced in time, with Gaussian noise at the level of
-10\% of the current brightness.  The effective signal to noise can be
-considered to be $0.1\times\sqrt{250}\simeq15$.
+10\% of the current brightness.  This amounts formally to a single
+summary data point with a signal to noise of
+$10\times\sqrt{250}\simeq160$.
 
-\subsection{Fitting to mock data sets}
+\subsection{Model fitting and likelihood analysis}
 

### 2016-03-14 paper1/discussion.tex
edits

@@
-\section{Discussion}
+\section{Discussion}\label{sec:discussion}
 
@@
 
-The previously mentioned abstract parameters can be related to physical quantities specific to the central region of a quasar. 
-As such the luminous region would correspond to the bright accretion disc that surrounds the black hole. The later's gravity would 
-cause a shadow in the bright region limited by the extent of the event horizon of the black hole. \textbf{Therefore the radius 
-of the bright disk would provide an estimate of the size of the accretion disk and the radius of the dark disc would provide
- and estimate of the gravitationally magnified Schwarzschild radius of the black hole $R_{S}^{magnified} = \Delta t_{dark} \cdot v_p$.
-By $R_{S}^{magnified}$ we refer to the apparent Schwarzschild radius which is larger than the real value at large distances due to the
-black hole's own gravity.  Moreover the gravitationally magnified value of the Schwarzschild radius is a monotonic function of the black 
-hole's mass. Therefore it can be used to estimate the mass of the black hole if it was not rotating.} In the previous 
-expression the $\Delta t_{dark}$ denotes the period of overlap between the black hole shadow and fold caustic. 
-$v_p$ denotes the component of the relative velocity of the source and fold which is perpendicular to the caustic.  
-The respective velocity is an unknown, though it can be constrained on a case by case basis to an order of magnitude 
-or even better. This would require the study of the dynamics of the stellar structure which contains the gravitational lens. 
-A better estimate of the relative velocity would facilitate a better estimate of the effective non-rotating black hole 
-mass associated to the black hole shadow.
+The previously mentioned abstract parameters can be related to
+physical quantities specific to the central region of a quasar.  As
+such the luminous region would correspond to the bright accretion disc
+that surrounds the black hole. The later's gravity would cause a
+shadow in the bright region limited by the extent of the event horizon
+of the black hole. Therefore the radius of the bright disk would
+provide an estimate of the size of the accretion disk and the radius
+of the dark disc would provide and estimate of the gravitationally
+magnified Schwarzschild radius of the black hole $R_{S}^{magnified} =
+\Delta t_{dark} \cdot v_p$.  By $R_{S}^{magnified}$ we refer to the
+apparent Schwarzschild radius which is larger than the real value at
+large distances due to the black hole's own gravity.  Moreover the
+gravitationally magnified value of the Schwarzschild radius is a
+monotonic function of the black hole's mass. Therefore it can be used
+to estimate the mass of the black hole if it was not rotating. In the
+previous expression the $\Delta t_{dark}$ denotes the period of
+overlap between the black hole shadow and fold caustic.  $v_p$ denotes
+the component of the relative velocity of the source and fold which is
+perpendicular to the caustic.  The respective velocity is an unknown,
+though it can be constrained on a case by case basis to an order of
+magnitude or even better. This would require the study of the dynamics
+of the stellar structure which contains the gravitational lens.  A
+better estimate of the relative velocity would facilitate a better
+estimate of the effective non-rotating black hole mass associated to
+the black hole shadow.
 

### 2016-03-14 paper1/intro.tex
edits

@@
 but basically the dark disc causes a rising light curve to plateau or
-dip.
+dip.  These features are still present, albeit faintly, if the simple
+crescent is replaced by a source based on an accretion-disc simulation
+of a black-hole environment
+(Figures~\ref{fig:M87_plots}--\ref{fig:char_points}).
 
-The simple crescent is replaced
-by a source (Figure~\ref{fig:M87_image}) from published simulations of
-the black-hole environment of M87
-\citep{2009MNRAS.394L.126M,2012MNRAS.421.1517D}.  The simulated light
-curve so obtained (lower panel of Figure~\ref{fig:M87_plots}) still
-shows, albeit faintly, features from the crescent.
-The crescent characteristics identified in
-Figure~\ref{fig:char_points} are from an idealised source crossing an
-idealised caustic.
-
-In Section~\ref{sec:numerics} we see what happens
-when we replace both of these with numerical simulations.  The ideal
-fold-transit is replaced by the path AB in
-Figure~\ref{fig:magnification_map}.
-
-In Section~\ref{sec:fitting} we carry out source fitting to
+In Section~\ref{sec:numerics} we carry out source fitting to
 lightcurves, with both noise and systematic errors are present.  We

### 2016-03-14 paper1/ms.tex
edits
merged text into ms.tex

@@
 
-%\title[Quasar microlensing and its internal structure]{Properties of quasar microlensing light curves: probing its internal structure}
 \title{Probing the internal structures of quasars using microlensing events}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{intro}
+\section{introduction}
 
-\input{theory}
+Active galactic nuclei are thought to be powered through the
+accretion of matter from the proximal environment into a supermassive
+black hole.  The radiation emitted excites the surrounding medium
+which becomes detectable as narrow line regions, broad line regions
+and optical continuum.  Moreover, in the direction perpendicular to
+the accretion disk, where the medium is more transparent, jets will
+appear.  If a jet is oriented towards the Earth, a quasar is observed
+\citep[e.g.,][]{1984RvMP...56..255B}.  While the basic mechanism
+\citep[originating in the work
+  of][]{1964ApJ...140..796S,1964SPhD....9..246Z,1969Natur.223..690L}
+is not in doubt, the central engines, near the event horizons of the
+black holes, remain to be probed.
 
-\input{numerics}
+For the black holes in the Galactic centre and the centre of M87
+---the two nearest objects that barely qualify as AGN--- the central
+engines (which are $<0.1\rm\,mas$ on the sky) are close to being
+resolved through very long baseline interferometry, which shows
+preliminary indications of the jet-launching structures
+\citep{2008JPhCS.131a2055D,2012Sci...338..355D,2013MNRAS.434..765K,2016arXiv160205527F}.
+Current data do not deliver images, but require fitting to
+predefined models for the images.  A whole range of models have been
+applied, starting from simple geometric models to more complex
+physical models
+\citep{2008Natur.455...78D,2011ApJ...738...38B,2009ApJ...706..497M,2010ApJ...717.1092D}.
+The more complicated models nonetheless tend to predict a crescent
+shaped silhouette of the black hole.  This motivated
+\cite{2013MNRAS.434..765K} to use a simple geometric crescent model to
+fit the data, and argued that the crescent is nothing but the
+silhouette of the event horizon.
 
-\input{discussion}
+The great majority of quasars, however, lie at redshifts beyond 2
+\citep{2014A&A...563A..54P} and their central engines would be orders
+of magnitude smaller on the sky. The direct observations of the black
+hole silhouettes of quasars is far beyond foreseeable instrumentation.
+In the present paper we consider a possible indirect method, related
+to \cite{1999ApJ...524...49A}, through which gravitational
+microlensing could probe the black hole shadow and its proximal quasar
+environment.
+
+Gravitational microlensing of quasars, reviewed in
+Section~\ref{sec:microlensing} below, refers to sharp changes in the
+observed brightness of quasars that have been lensed by intervening
+galaxies, without any changes in the intrinsic luminosity.
+Microlensing affects only the light from the innermost part of the
+quasar, such as the optical continuum
+\citep[e.g.,][]{2012A&A...544A..62S} and is a consequence of two
+things: the very small size of the central engine, and granularity of
+the mass distribution of a lensing galaxy due to stars.  The latter
+means that the local magnification is not a smooth function of source
+position. It contains a complicated network of singular curves, known
+as caustics.  Figure~\ref{fig:magnification_map} shows part of a
+magnification map with a few caustics.  The lensed brightness would be
+given by placing the source on such a magnification map and
+integrating the surface brightness weighted by the magnification.
+Most astrophysical sources straddle several caustics, and hence their
+net brightness varies smoothly with location.  The central engine of
+quasars, however, is smaller than the typical spacing between
+caustics.  As a result, the lensed brightness undergoes sudden changes
+as a quasar crosses a caustic.  The effect supplies an upper limit on
+the size of the central engine, and can also be used to study the mass
+distribution and kinematics of stars in the lensing galaxy as well
+\citep[e.g.,][]{2012ApJ...744..111P}.  Caustics have an additional
+remarkable feature: though they can be very complicated, they have
+some universal properties well-known from catastrophe theory.  In
+particular, very close to the simplest caustics (known as folds), the
+magnification is approximately constant on one side and
+$\propto1/\sqrt p$ where $p$ is the transverse distance of the source
+from the caustic.  This property will be exploited later.
+
+In Section~\ref{sec:source-models} we introduce the three source
+profiles used in our subsequent models and simulations: a
+constant-brightness disc, a circular Gaussian, and the crescent source
+introduced by \citep{2013MNRAS.434..765K}.  The latter is simply a
+constant-brightness disc with a smaller, non-concentric disc cut out
+of it.  We also derive the half-light radius for a crescent.  The
+half-light radius can characterise the source size for all three types
+of source.
+
+Section~\ref{sec:fold-crossing} shows the light curves that result
+when each of the model sources crosses an ideal fold.  This would
+apply in Figure~\ref{fig:magnification_map} to sources along the path
+AB or BC, for sources small enough that the curvature of the caustics
+is negligible.  With this assumption one can imagine the caustic as an
+infinite wall to be crossed by the source as presented in
+Figure~\ref{fig:infinite_fold}.  The source brightness distribution
+parallel to the caustic naturally makes no difference to the
+observable brightness; each source can be replaced by an effective
+one-dimensional source profile, by flattening the source so it becomes
+perpendicular to the caustic.  In principle, the effective
+one-dimensional brightness profile could be recovered from the light
+curve by deconvolution.  \cite{1999ApJ...524...49A} modelled this
+profile as the result of a circular accretion disc seen through the
+spacetime around a Kerr black hole, and \cite{2012MNRAS.423..676A}
+have applied the idea to observed light curves to infer properties of
+quasar accretion discs.  In this work we take a simpler but arguably
+more robust approach: we study features in the light curves
+characteristic of a crescent-like source which in turn would indicate
+a black-hole silhouette.  Figure~\ref{fig:char_points} shows the
+qualitative features: there is a period during which the dark cutout
+disc is crossing the caustic, and before and after there are periods
+when the only the bright parts of the crescent are in transit across
+the caustic.  The details depend on the orientation of the crescent,
+but basically the dark disc causes a rising light curve to plateau or
+dip.  These features are still present, albeit faintly, if the simple
+crescent is replaced by a source based on an accretion-disc simulation
+of a black-hole environment
+(Figures~\ref{fig:M87_plots}--\ref{fig:char_points}).
+
+In Section~\ref{sec:numerics} we carry out source fitting to
+lightcurves, with both noise and systematic errors are present.  We
+generate three lightcurves by taking a uniform disc, a circular
+Gaussian, and a crescent across the path AB in
+Figure~\ref{fig:magnification_map}, and then adding noise.  The path
+simulates crossing a clean but not ideal fold.  In addition, we
+generate templates lightcurves by running the three source types, with
+various parameter values, across the path CB.  That is, the templates
+come from a similar but not identical caustic, thus deliberately
+generating a systematic error.  We then fit the noisy lightcurves to
+the templates using Markov chain Monte-Carlo.  We find that the
+correct source type can be inferred from the $\chi^2$ values.  The
+parameter values can also be inferred. The fitting errors are larger
+than the formal uncertainties, which is expected in the presence of
+systematic errors, but still appear acceptable.
+
+Finally in Section~\ref{sec:discussion} we discuss in more depth the
+implications of the results presented in the previously mentioned
+sections.  One of the most interesting implication is the possibility
+to estimate the black hole mass from the reconstructed
+parameters. This further requires good approximations of the relative
+transversal velocity between the quasar and the lens. Proper
+constraints can be set with independent observations of the stellar
+structure that contains the gravitational lens.
+
+The thorough study of the possibility to reconstruct the quasar's
+structural parameters from light-curves containing multiple
+microlensing events represents the target of future work. This will
+most probably require the use of powerful statistical tools. Another
+path for future work is the designing of an observation regime best
+suited for acquiring the necessary microlensed light-curve.
+\section{Microlensing}\label{sec:microlensing}
+
+We start by introducing in a succinct manner the gravitational lensing
+theory that is relevant for microlensing in general and for the scope
+of the present paper in particular.  More detailed presentation of the
+theory can be found in several references, such as
+\cite{2001stgl.book.....P}.
+
+\subsection{Magnification}
+
+The gravitational lens equation
+\begin{equation}
+\vec\beta = \vec\theta - \vec\alpha(\vec\theta)
+\label{eqn:lens}
+\end{equation}
+relates the apparent sky position $\vec\theta$ of a light source to
+its true but unobservable sky position $\vec\beta$ through the bending
+angle $\vec\alpha$.  The latter is an integral 
+\begin{equation}
+\vec\alpha = (1+z_{_L})\frac{D_S-D_L}{D_SD_L} \frac{4G}{c^2}
+\int \Sigma(\vec\theta')
+\frac{\vec\theta-\vec\theta'}{|\vec\theta-\vec\theta'|^2}\,
+d^2\vec\theta'
+\label{eqn:alpha}
+\end{equation}
+depending on the projected density $\Sigma(\vec\theta)$ (kg/steradian)
+of the lens, the lens redshift $z_{_L}$ and the comoving distances
+$D_L$ and $D_S$ to the lens and source.  The derivative of the
+apparent position with respect to the source position
+\begin{equation}
+M(\vec\theta) =
+\left(\frac{\partial\beta}{\partial\theta}\right)^{-1}
+\label{eqn:magnif-matrix}
+\end{equation}
+is known as the magnification matrix, and its determinant
+\begin{equation}
+\mu(\vec\theta) = \det|M(\vec\theta)|
+\end{equation}
+is the brightness amplification of an image of a point source.  In
+other words, the source will brighten or dim according to whether
+$\mu(\vec\theta)$ is more or less than unity.  If there are multiple
+mages at distinct but not observationally resolved $\vec\theta_i$ from
+the same $\vec\beta$, a total brightness amplification
+of\begin{equation} \mu_{\rm total} = \sum_{i} \mu(\vec\theta_i)
+\end{equation}
+applies. It is possible for the magnification to become formally
+infinite, as a result of an eigenvalues of the magnification matrix
+(\ref{eqn:magnif-matrix}) becoming infinite.  This generically happens
+on curves on the $\vec\theta$ plane, known as {\em critical curves}.
+Mapping a critical curve to the source plane, through the lens
+equation (\ref{eqn:lens}), to the source plane $\vec\beta$ gives the
+so-called caustics.  Caustics can appear in optical system, not just
+gravitational lensing.  For a point source, caustics are singularities
+of the magnification; for finite-size sources caustics correspond to
+high and sharply changing magnification.
+
+Caustics are important in all forms of lensing with multiple images,
+but they have a special significance for lens quasars, first pointed
+out by \cite{1979Natur.282..561C}.  The granularity of the mass
+distribution $\Sigma(\vec\theta)$ due to individual stars produces a
+caustic network on the scale of $\sqrt{GM_\odot D_L/c^2}$ or $10^{-6}
+\sqrt{M/M_\odot}\rm\,arcsec$ for typical lens and source redshifst
+\citep{2001PASA...18..207W}.  Extended sources wash out this
+micro-caustic structure, but the optical-continuum source of quasars
+is even smaller.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\hsize]{figures/IRIS567_path.eps}
+\caption{\label{fig:magnification_map} Magnification map used later in
+  this paper (Section~\ref{sec:numerics}). The white line marks the
+  trajectory of the center of the sources for which the lightcurves
+  presented in figure 12 are generated.  Note that this represents an
+  atypically simple region of any realistic magnification map.}
+\end{figure}
+
+\subsection{Magnification near a fold caustic}
+
+A simple example of a caustic network is shown in
+Figure~\ref{fig:magnification_map}.  There are two general categories
+of caustics in gravitational lensing, cusps and fold, and examples of
+both kinds can be seen in this figure.  Magnification near a caustic
+has universal properties, independent of the system and has been
+extensively studied
+\citep{1986ApJ...310..568B,1992A&A...260....1S,2002ApJ...574..970G,2002ApJ...580..468G}.
+In particular, at distance $p$ from a fold caustic
+\begin {equation}
+ \mu(p) = \mu_0 + C_0 \frac{1}{\sqrt{p}} \Theta(p).
+\end {equation}
+Here the magnification of a point source near a caustic is equal to
+the sum of the magnification due to other reasons $\mu_0$, assumed to
+be locally constant, and a decrease with the square root of the
+distance from the fold. The later term becomes activated only after
+the source enters the region interior to the caustic curves when the
+values of the stepfunction $\Theta(p)$ become unity. The
+proportionality constant $C_0$ depends on the local conditions in the
+vicinity of the caustic.
+
+A source of arbitrary shape can be described by a two dimensional brightness function $S_{2D}(p - p_s, q - q_s)$ defined for a coordinate system $p,q$ where $p_s, q_s$ denote the coordinates of the center of source.
+
+\begin{figure}
+\includegraphics[width = .49\textwidth]{figures/infinite_fold_ar.eps}
+\caption{\label{fig:infinite_fold} Source profiles for Gaussian, disk,
+  crescents (right) and magnification map for an infinite fold
+  (left). Objects have the same $S_{0}$ and $r_{1/2}$. The black arrow marks the direction of motion of the
+sources relative to the caustic.}
+\end{figure}
+
+For a microlensing event the lightcurve can be written for an
+undefined source shape as:
+\begin{equation}
+ F(t) = \int_{-\infty}^\infty \int_{-\infty}^\infty S_{2D}(p-p_s(t), q-q_s(t)) \mu_t(p) \mathrm{d}q \mathrm{d}p
+ \label{eqn:ft2d}
+\end{equation}
+In order to build the previous equation we have considered that the
+time dependency of the flux $F$ is given only by the motion of the
+source with respect to a fixed caustic. Therefore the only time
+dependent quantities in the right hand side of the equation are the
+coordinates of the center of the source $p_s,q_s$ and by construct
+$S_{2D}$.
+
+Due to the choice of the coordinate system the amplification factor has no dependence on the $q$ coordinate. The previous equation can be rewritten as:
+
+\begin{equation}
+ F(t) 
+= \int_{-\infty}^\infty  \mu_t(p) S_{1D}\left(p-p_s(t)\right) \mathrm{d}p,
+\label{eqn:ft}
+\end{equation}
+\\
+where we have defined the one dimensional flux function as:
+\begin{equation}
+ S_{1D}(p-p_s(t)) = \int_{-\infty}^\infty S_{2D}(p-p_s(t), q-q_s(t)) \mathrm{d}q
+\end{equation}
+
+This representation is a valid approximation only when the apparent size of the source is much smaller than the corresponding Einstein angle of the lens. In this context 
+all the information about the source shape and brightness that can be contained in the lightcurve is exhaustively given by the 1D flux function.
+In other words, if two sources with different $S_{2D}$ have the same $S_{1D}$ they cannot be distinguished by studying their lightcurves.
+
+\section{Models for extended sources}\label{sec:source-models}
+
+In the present study we are analysing three types of sources with different surface brightness: 
+\begin{enumerate}
+ \renewcommand{\theenumi}{(\arabic{enumi})}
+  \item a rotationally symmetric source with a bivariate gaussian surface brightness distribution,
+  \item a disk source with constant surface brightness distribution,
+  \item a crescent shaped source with constant surface brightness distribution.
+\end{enumerate}
+The first two sources are the typical choices used in the literature to describe the luminous parts of a quasar (Prasenjit should give some citations here). The third one is a recently proposed
+variant \citep{2013MNRAS.434..765K}.
+
+\subsection{Rotationally symmetric source with a bivariate gaussian surface brightness distribution}
+
+A symmetric 2D gaussian can be described mathematically as:
+
+\begin{equation}
+ S_{2D}^G(p-p_s, q-q_s) = \frac{S_0^G}{2 \pi \sigma^2} e^{-\frac{(p-p_s)^2}{2 \sigma^2}} e^{-\frac{(q-q_s)^2}{2 \sigma^2}}.
+\end{equation}
+\\
+The corresponding 1D brightness is:
+
+\begin{equation}
+ S_{1D}^G(p-p_s) = \frac{S_0^G}{\sqrt{2 \pi} \sigma} e^{-\frac{(p-p_s)^2}{2 \sigma^2}}.
+\end{equation}
+\\
+Other parameters of the model are the total flux $S_0^G$ and $\sigma$.
+
+Although such a definition for a source would have non-zero surface/linear brightness for any coordinate $p,q$, the amount of light received by a detector from outside a $3 \sigma$ disk centered at $p_s, q_s$
+would be insignificant. For a gaussian distributed surface brightness source the half-light radius is directly proportional to the parameter $\sigma$ according to the equation:
+\begin{equation}
+r_{1/2} = \sqrt{ln(4)} \sigma.
+\end{equation}
+
+\subsection{Disk source with constant surface brightness distribution}
+
+One can construct mathematically a disk source with constant surface brightness and radius $R$ using a stepfunction:
+\begin{equation}
+ S_{2D}^D(p-p_s, q-q_s) = \frac{S_0^D}{\pi R^2} \Theta \left( R^2 - \left( p-p_s \right)^2 - \left( q-q_s \right)^2 \right).
+\end{equation}
+By integrating over $q$ coordinate the linear brightness function is obtained:
+
+\begin{equation}
+ S_{1D}^D(p-p_s) = \frac{2 S_0^D}{\pi R}  \sqrt{1 - \frac{(p-p_s)^2}{R^2} }    \Theta \left( R^2 - \left( p-p_s \right)^2 \right).
+\end{equation}
+\\
+The half-light radius of a uniform disk source is $R/\sqrt{2}$.
+
+\subsection{Crescent source with constant surface brightness distribution}\label{subsec:crescent}
+
+The surface brightness distribution of a geometric crescent can be built by considering two disk sources of constant brightness.
+One larger disk will contribute positively to the total flux, while one smaller disk
+that is interior to the large one will contribute negatively. This superposition can be written for 2D as:\\
+
+\begin{equation}
+ S_{2D}^C =  S_{2D}^{Dp} -  S_{2D}^{Dn}  
+ \label{eqn:s2d}
+\end{equation}
+with\\
+
+\begin{equation}
+ S_{2D}^{Dp}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dp}}{\pi R_p^2} \Theta \left( R_p^2 - \left( p-p_{sp} \right)^2 - \left( q-q_{sp} \right)^2 \right)
+\end{equation}
+\\
+and
+\begin{equation}
+ S_{2D}^{Dn}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dn}}{\pi R_n^2} \Theta \left( R_n^2 - \left( p-p_{sn} \right)^2 - \left( q-q_{sn} \right)^2 \right).
+\end{equation}
+\\
+The following notations were used: $R_p, (p_{sp}, q_{sp}), R_n, (p_{sn},q_{sn})$ are the radii and coordinate of the center for the larger positive disk and smaller negative disk respectively.
+$S_0^{Dp},S_0^{Dn}$ represent the total flux of radiation received from the large and small disk.
+From this point forward we will not use the total flux from each source.
+Instead we will
+use the difference which in this case is th total flux from the crescent-shaped source $S_0^C$. \\
+Equation \ref{eqn:s2d} can be written as:\\
+
+\begin{align}
+ S_{2D}^C &= \frac{S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \left\{ \Theta \left[ R_p^2 - \left( p-p_{sp} \right)^2 - \left( q-q_{sp} \right)^2 \right] \right.\nonumber\\
+ &\qquad \left. {} -  \Theta \left[ R_n^2 - \left( p-p_{sn} \right)^2 - \left( q-q_{sn} \right)^2 \right] \right\}.
+\end{align}
+\\
+Analogous for the linear brightness function:
+
+\begin{align}
+ S_{1D}^C &= \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \left\{ \sqrt{R_p^2 - (p-p_{sp})^2}  \Theta \left[ R_p^2 - \left( p-p_{sp} \right)^2 \right] \right.\nonumber\\
+ &\qquad \left. {} - \sqrt{R_n^2 - (p-p_{sn})^2 } \Theta \left[ R_n^2 - \left( p-p_{sn} \right)^2 \right] \right\}.
+\label{eqn:s1_d}
+\end{align}
+
+There are some constraints on the parameters used to define a crescent in the previously presented manner that need to be stated. First, we must impose the obvious $R_p > R_n$ relation. Secondly, 
+the small disk must always be interior to the large disk:
+\begin{equation}
+ R_p \ge R_n + \sqrt{\left(p_{sp} - p_{sn} \right)^2 + \left(q_{sp} - q_{sn} \right)^2}
+\end{equation}
+
+For the distances between the centers of the two disks we will use the same notations as the one found in the paper \citep{2013MNRAS.434..765K}, $a \equiv p_{sn} - p_{sp}$ and $b \equiv q_{sn} - q_{sp}$.
+
+The half-light radius of any source is invariant to any rotational transformation.
+In the present case of a crescent source the effective radius is dependent on the parameters $R_p$, $R_n$ and $\sqrt{a^2+b^2}\equiv c $ exclusively.
+From symmetry considerations the centroid of the source is collinear with the centers of the two disks and it is situated at a distance $d_c$ from the center of the bright disk.
+$d_c$ can be computed numerically with the use of a variation of equation \ref{eqn:s1_d}:
+
+\begin{equation}
+\begin{aligned}
+\frac{S_0^C}{2} & = \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{d_c}^{R_p} \bigg[ \sqrt{R_p^2 - p^2} \\ 
+		& - \sqrt{R_p^2 - \left(p-c\right)^2} \Theta \left(R_p^2 - \left(p-c\right)^2 \right) \bigg] dp. 
+\end{aligned}
+\end{equation}
+
+With the position of the centroid determined, the half-light radius can be also be computed numerically:
+
+\begin{equation}
+\begin{aligned}
+S_0^C &=  \frac{4S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{0}^{r_{1/2}} \left[ \Theta_2  - \Theta_1  \right] p dp; \\
+\Theta_1 &= \bigg[ \pi - \arccos \frac{R_p^2 + p^2 -d_c^2}{2 R_p p} \\
+         & \phantom{= \bigg[ \pi} - \arccos \frac{R_p^2 - p^2 + d_c^2}{2 R_p d_c}
+            \bigg] \Theta \left( p - R_p + d_c \right); \\
+\Theta_2 &=  \pi - \arccos \frac{ p^2  + (c+d_c)^2 - R_n^2}
+                                {2 p \left(c + d_c \right)} \\
+         & \phantom{= \pi - } \Theta \left( p + R_n - d_c -c \right). 
+\end{aligned}
+\end{equation}
+
+\begin{figure}
+\includegraphics[width = .49\textwidth]{figures/figure_rhalf.eps}
+\caption{\label{fig:geom_crescent} Geometry of crescent sources.}
+\end{figure}
+
+\section{Lightcurves of the extended sources during fold crossing}\label{sec:fold-crossing}
+
+Using equation \ref{eqn:ft} and the one-dimensional flux function
+presented in the previous section one can compute numerically the
+lightcurves of the three extented sources for the simplified
+infinite-wall-caustics model.
+
+\subsection{Lightcurve of the gaussian source}
+
+The amount of light received by an observer from a source with a gaussian distributed brightness with $\sigma$ and total flux $S_0^G$ in the absence of any gravitational lensing is:
+\begin{equation}
+ F^G(t) = \int_{-\infty}^\infty  \left( \mu_0 + \frac{C_0}{\sqrt{p}} \Theta \left( p \right) \right) \left( \frac{S_0^G}{\sqrt{2 \pi} \sigma} e^{-\frac{(p-p_s(t))^2}{2 \sigma^2}} \right) \mathrm{d}p.
+\end{equation}
+
+which can be simplified to:
+\begin{equation}
+ F^G(t) = \mu_0 S_0^G + \frac{C_0 S_0^G}{\sqrt{2\pi} \sigma} \int_{0}^\infty \frac{e^{-\frac{(p-p_s(t))^2}{2 \sigma^2}}}{\sqrt{p}} \mathrm{d}p.
+\end{equation}
+
+\begin{figure}
+\centering
+	\includegraphics[width = 0.48\textwidth]{figures/S1D_all.eps}
+	\includegraphics[width = 0.48\textwidth]{figures/4source_magnification.eps}
+\caption{\label{fig:lightcurve_gauss} Lightcurves of crescents, Gaussian and disk for sources with identical $S_0$ and $r_{1/2}$.
+Crescent source has $r_{1/2}$ = 0.72754 $R_p$, $R_n$ = 0.4 $R_p$, a = 0.3/-0.3 $R_p$ }
+\end{figure}
+
+\subsection{Lightcurve of the disk shaped source}
+
+Analogous to the gaussian shaped source, the disk source with uniform brightness, radius $R$ and unmagnified flux $S_0^D$ has a lightcurve described by the equation:
+
+\begin{equation}
+\begin{aligned}
+ F^D(t) &= \int_{-\infty}^\infty  \left( \mu_0 + \frac{C_0}{\sqrt{p}} \Theta \left( p \right) \right) \\
+	& \bigg[ \frac{2 S_0^D}{ \pi R} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}}  \Theta \left(R^2 - \left(p-p_s(t) \right)^2 \right) \bigg] \mathrm{d}p.
+\end{aligned}
+\end{equation}
+
+which is equivalent to:
+\begin{equation}
+ F^D(t) = \mu_0 S_0^D + \frac{2 C_0 S_0^D}{\pi R} \int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \frac{1}{\sqrt{p}} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}} \mathrm{d}p.
+\end{equation}
+
+\begin{figure}
+\centering
+	\includegraphics[width = 0.48\textwidth]{figures/S1D_var_a.eps}
+	\includegraphics[width = 0.48\textwidth]{figures/4avar_magnification.eps}
+\caption{\label{fig:lightcurve_disk} Lightcurves of crescent sources with $R_n$ = 0.4 $R_p$.}
+\end{figure}
+
+\subsection{Lightcurve of the crescent shaped source}
+
+The lightcurve of a crescent shaped source with unamplified flux $S_0^C$, radii $R_p$, $R_n$ and center displacement $a(t)$ is:
+
+\begin{equation}
+\begin{aligned}
+ F^c(t) &= \mu_0 S_0^C + C_0 \frac{2 S_0^C}{\pi \left( R_p^2 -R_n^2 \right) } \\
+	&\bigg[ \int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) \right)^2 }{p}} \mathrm{d}p \\
+	&  -  \int_{max(0, p_s(t) - a(t) - R)}^{max(0, p_s(t) -a(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) +a(t) \right)^2 }{p}} \mathrm{d}p  \bigg] .
+\end{aligned}
+\end{equation}
+
+\begin{figure}
+\centering
+	\includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_poz.eps}
+	\includegraphics[width = 0.48\textwidth]{figures/5Rn_back_var_magnification.eps}
+\caption{\label{fig:lightcurve_crescent_back} Lightcurves of backwards crescents with  a = 0.3 $R_p$.}
+\end{figure}
+
+The function $p_s(t)$ can be chosen to be equal to $v_p(t-t_0) + p_{s0}$. Where $p_{s0}$ is the coordinate $p$ of the source at the initial time, and $v_p$ is the component of the velocity
+along the $p$ axis. Such a modelling of the motion of the object in the source plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
+choosing the function $a(t)$ to be constant in time.
+
+\begin{figure}
+\centering
+	\includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_neg.eps}
+	\includegraphics[width = 0.48\textwidth]{figures/5Rn_forw_var_magnification.eps}
+\caption{\label{fig:lightcurve_crescent} Lightcurve of forwards crescents with a = -0.3 $R_p$.}
+\end{figure}
+
+There are four characteristic points visible on the resulting one dimensional light profile of the crescent shaped source.
+Two of the points, p1 and p4, mark the outer boundaries of the luminous disc component.
+The other two, p2 and p3, mark the boundaries of the dark disc component.
+Since $S_{1D}$ is a projection of $S_{2D}$ on a line perpendicular to the caustic, the following relation holds:
+\begin{equation}
+	p_2-p_1 = R_p -R_n - a.
+\end{equation}
+In addition there are two other obvious relations the characteristic points which are independent of the projection:
+\begin{equation}
+	p_4 -p_1 = 2 R_p,
+\end{equation}
+
+\begin{equation}
+        p_3 -p_2 = 2 R_n.
+\end{equation}
+All four points mark the positions where derivative $\frac{dS_{1D}}{dp}$ is discontinuous.
+The points can be used to define three regions: $p_1 - p_2$ where $S_1D$ is convex, $p_2 - p_3$ where $S_1D$ is concave, and $p_3 - p_4$ where $S_1D$ is convex again.
+\\
+
+Due to the nature of the caustic and the monotonic behaviour of the magnification map on both sides of the caustic.
+The previously mentioned characteristic points are inherited by the microlensing lightcurve.
+The points on the temporal dimension $t_1, t_2, t_3$ and $t_4$ correspond to instances in time when the fold is aligned with $p_1, p_2, p_3$ and $p_4$, respectively.
+For a constant relative velocity $v_p$ between the source and the caustic there is a simple relation between the points $p_i$ and instances $t_i$:
+
+\begin{equation}
+	t_j - t_i = \frac{p_j - p_i}{v_p}.
+\end{equation}
+
+With the use of the previous four equations the following identities can be written:
+
+\begin{equation}
+	R_p = \frac{v_p \left( t_4 -t_1 \right)}{2}, 
+\end{equation}
+
+\begin{equation}
+        R_n = \frac{v_p \left( t_3 -t_2 \right)}{2}, 
+\end{equation}
+
+\begin{equation}
+        a = \frac{v_p \left( t_4 +t_1 - t_3 - t_2 \right)}{2}. 
+\end{equation}
+
+\begin{figure}
+\centering
+	\includegraphics[width = 0.48\textwidth]{figures/ch_points.eps}
+        \includegraphics[width = 0.48\textwidth]{figures/ch_instances.eps}
+\caption{\label{fig:char_points} The characteristic points and instances of a crescent source's one dimensional profile and microlensing lightcurve. }
+\end{figure}
+
+Figure 4 reveals that the lighcurve of a crescent source has more visible features than the other two light-curves
+corresponding to the disc and guasssian shape. There are three regions where $S_{1D}$ and by inheritance the lighcurve
+ have distinguishable behaviour. The first region that would be recorded on a lightcurve plot represents the period
+of time when the bright disk begins to be overlapped by the caustic and stops when just before the dark disk reaches
+the caustic. During this period of time the flux of light from the source is increasingly magnified. The second
+period starts and ends with the overlapping of the dark disc. As a boundary of the two regions there is a
+distinguishable point where slope of the magnification is drastically changed. This apparent discontinuity in the
+first derivative of the magnification function is caused by the caustic amplification of the sudden drop in the $S_{1d}$
+function. During the respective period the magnification growth slows or even reverses during the first part of the
+period and starts to grow faster again as the dark disk ends its overlap with the caustic. At the point where
+the dark disc clears the caustic the growth of the magnification is infinite, which appears as a saddle point
+on the lightcurve. Next, the final period correspond to the case when the dark disc has cleared the caustic
+and the bright disc continues to overlap with the caustic. During this period the lighcurve reaches a peak
+that for most of the parameter space is global and for the rest of the parameter space local.
+At the end of the period the growth of the magnification is negative infinite. the respective point appears 
+as a second saddle point on the lightcurve. Past this point the magnification of any finite source will 
+decay in roughly the same manner. \\
+
+The impact of parameter $a$ on the shape of the lightcurve can be observed in figure 5 (a-var). Sources where 
+the center of the dark disc reaches the caustic before the center of the bright disc are characterized by a smoother
+broad peak in contrast to the cases where the center of the dark disc reaches the caustic after the bright disk.
+In the case of the latter the instance when the dark disc reaches the caustic corresponds to larger and
+larger magnifications until it becomes a local and even global peak. The effect of the $R_n$  parameter on the
+lighcurve is presented in figures 6 and 7. For the particular set of parameters where $R_p = R_n +a$ the third
+period of time discussed previously does not exist. Particularizing further, if the value of the radius of the dark
+disc is comparable to the value of the radius of the bright one then the position and shape of the maximum magnification
+ are strikingly different. In case the crescent reaches the caustic with the bright region first, the peak magnification
+ happens when the dark region reaches the caustic and it is characterized by a sharp variation in magnification growth,
+In the opposite case the peak appears before the end of the bright disc reaches the caustic and shape is smoother. \\
+
+\subsection{Microlensing a simulated image of M87}
+
+\begin{figure}
+\centering
+        \includegraphics[width = 0.48\textwidth]{figures/M87_image.eps}
+\caption{\label{fig:M87_image} The simulated image of M87's black hole
+  sillouette obtained by \citep{2012MNRAS.421.1517D} from a GRMHD
+  simulation presented in \citep{2009MNRAS.394L.126M}. The color bar
+  is linear and scaled to reach value 1 for the maximum
+  brightness.}
+\end{figure}
+
+\cite{2012MNRAS.421.1517D} have created a radiative image of M87 based on the GRMHD simulations presented in \citep{2009MNRAS.394L.126M}.
+The respective image, reproduced in figure \ref{fig:M87_image}, has been projected to a 1D profile associated to the perpendicular direction to a fold caustic approaching the image from the right.
+The projection is presented in the left panel of figure 10.
+The amplification values of the flux of light corresponding to a microlensing event are presented in the right panel of figure 10.
+In general the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions surrounding the luminous parts of the image have non-zero flux and thus are more extended than the simplified source model.
+
+\begin{figure}
+\centering
+	\includegraphics[width = 0.48\textwidth]{figures/M87_shape.eps}
+        \includegraphics[width = 0.48\textwidth]{figures/M87_lc.eps}
+\caption{\label{fig:M87_plots} The one dimensional profile obtained
+  from the numerical integration along the ordonata axis of the source
+  image (figure 9) is presented in the lower panel. The corresponding
+  lightcurve associated to the 1D profile is presented in the right
+  panel. The points p2 and p3 mark the start and end of the equivalent
+  dark disk and are associated to the moments of their overlap with
+  the caustic: t2 and t3.}
+\end{figure}
+
+\section{Fitting mock data}\label{sec:numerics}
+
+Having used the simple model of an ideal fold to gain insight, we now
+move to the numerical computation of microlensing light curves with
+realistic mass distributions, and then fit source models to
+lightcurves in the presence of random and systematic errors.
+
+First we generate microlensing magnification maps using the
+microlensing code by \cite{1999A&A...346L...5W}.  We then generate
+mock light curves for three source models --- a crescent, a uniform
+disc and a Gaussian disc --- and then fit each mock light curve to all
+three source models. Parameter fitting and marginalising was done by
+Markov-chain Monte Carlo (MCMC).
+
+\subsection{Numerical microlensing magnification maps}
+
+The code uses a ray-shooting technique to compute the gravitational
+lensing effect of a mass distribution consisting of (a)~a smooth
+component and (b)~a random distribution of point masses representing
+stars.  The ray shooting maps a grid of $\vec\theta$ to $\vec\beta$
+using the lens equation (\ref{eqn:lens}).  That is, the rays are shot
+from the observer back to the source.
+
+For the computation of the individual deflection angles
+$\vec\alpha(\vec\theta)$ a heirarchical or tree method is used.  The
+positions of all lensing masses are put into a grid of $\vec\theta$.
+Each grid cell is subdivided into four smaller squares recursively
+until every cell contains only one mass.  Nearby masses are added
+individually, while distance masses are clumped into larger grid cells
+whose net contribution is approximated by its first few multipole
+moments.  Scaled units are used, with the constant pre-factor in the
+deflection angle (\ref{eqn:alpha}) separated out.
+
+The result of ray shooting is a pixel map on the $\vec\beta$ plane of
+the number of lightrays which arrive at the source plane from a
+particular observer.  This intermediate result is effectively a
+magnification map on the source plane.  Once the map is created, the
+lightcurve can be obtained by specifying the transit path of the
+source across the map.  At each point on this transit line, the code
+computes equation (\ref{eqn:ft2d}) convolving the brightness
+distribution of the source with the magnification pattern of the map.
+In real life, not only are both lens and source moving, but the lens
+lens configuration, and with it the magnification pattern, is also
+changing with time.  While the first subtlety is taken care of by a
+coordinate transformation in this analysis, for the second one the
+lens configuration is assumed to be constant in time.
+
+When generating the magnification map depicted in
+Figure~\ref{fig:magnification_map}, which is used in our analysis
+below, only two point masses were included.  This was done in order to
+have clean fold caustics.  For the computation of the actual
+lightcurve, the code was modified to also allow for crescent shaped
+images specified through the parameter set $R_p,R_n,a,b$. Here $R_p$
+denotes the outer radius of the crescent and $R_n$ the inner one. The
+orientation of the source image with respect to the magnification map
+is specified by the parameters $a$ and $b$ as the shift of the center
+of the inner disk from the center of the outer disk in $x$- and
+$y$-direction respectively. In the original version of the code,
+gaussian and disk shaped images where already implemented. Those are
+completely characterised by the single parameter $R_p$. The values of
+the parameters are specified in pixel units corresponding to the
+magnification map. Further one needs to specify the start and end
+point coordinates of the path, which the center of the source image
+follows through the magnification map (see the depiction in figure
+\ref{fig:magnification_map}). Hence the points along this path are
+specified through the number of timesteps for which the computation of
+the brightness is to be carried out. Those points correspond to the
+actual measurement of the brightness of an object in the observational
+case. For each timestep the two-dimensional convolution of equation
+\ref{eqn:ft2d} is carried out numerically for the position of the
+source on the magnification map.
+
+For the purpose of this analysis it was desirable to mimic the
+analytical behaviour of a simple fold as much as possible, for
+comparing the numerical result with the analytical one, therefore the
+path of the source was chosen, so that it intersects the border of the
+caustic perpendicularly, and on a point where the border is a fold
+caustic.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\hsize]{figures/data_cc.eps}
+\includegraphics[width=0.9\hsize]{figures/data_dd.eps}
+\includegraphics[width=0.9\hsize]{figures/data_gg.eps}
+\caption{\label{fig:mockdata} Three different mock datasets with
+  three different parametrizations fitted to each --- crescent (red),
+  uniform-disk (blue) and Gaussian disk (green).}
+\end{figure}
+
+\subsection{Mock light curves}
+
+Mock light curves were generated by running a source across a
+caustic, specifically along the line from point C to point B in
+Figure~\ref{fig:magnification_map}.  The model light curves, to which
+these are fitted, are made by running a source across a different
+caustic, along the line from point A to point B in the same figure.
+That is, the mock data are generated with a clean but not ideal fold
+caustic and then fitted with another such caustic.  This mimics the
+unavoidable systematic error of not knowing the caustic exactly.
+
+The crescent source had parameters (as explained in
+\S\ref{subsec:crescent}) $R_p$ and $R_n$ being the radii of the outer
+and inner circles and $(\alpha,\beta)$ being the coordinates of the
+inner circle with respect to the centre of the outer circle in the
+coordinate system associated with the image in Figure 1 and not
+associated with the caustic surface as defined in the previous
+sections.  The parameter values were
+\begin{equation}
+   (R_p, R_n, \alpha, \beta) = (50.0, 30.0, 15.0, 10.0).
+\label{eqn:cp}
+\end{equation}
+The uniform disc had the same (outer) radius as the crescent.  In the
+Gaussian source, we set $3\sigma=50$.  For convenience, below we will
+refer to $R_p$ of the Gaussian source, by which we actually mean
+$3\sigma$.
+
+Each mock light curve also has three nuisance parameters, namely the
+beginning and end of the event and the brightness normalisation.
+These are to be marginalised out by the MCMC.
+
+Figure~\ref{fig:mockdata} shows the three light curves.  Each has 250
+points regularly spaced in time, with Gaussian noise at the level of
+10\% of the current brightness.  This amounts formally to a single
+summary data point with a signal to noise of
+$10\times\sqrt{250}\simeq160$.
+
+\subsection{Model fitting and likelihood analysis}
+
+Each of the three mock light curves was fitted to all three models.
+We denote the nine possible cases with letter pairs, with the first
+letter denoting the assumed model for the fitting procedure and 
+the second letter representing
+the source: thus CG means a {\em C\/}rescent model was fitted to mock
+data from a {\em G\/}aussian source, DC stands for a uniform {\em
+  D\/}isc model fitted to mock data from a {\em C\/}rescent source,
+and so on.
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\hsize]{figures/Rp4all.eps}
+\caption{\label{fig:mcmc} Posterior probability distribution for the
+  source size $R_p$.  The vertical line is the correct value. Same
+  color represents the same dataset whereas same line style
+  corresponds to same model fit.  The legend gives the reduced
+  $\chi^2$ of the best fit in each case.}
+\end{figure}
+
+Figure~\ref{fig:mcmc} shows the posterior probability distribution of
+$R_p$ for all nine cases, along with the minimum reduced $\chi^2$ for
+each case.  The area under each curve is unity. Note that the height
+of the curves are not the likelihood, they are probability densities
+in parameter space.
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\hsize]{figures/Rhalf_RnRp.eps}
+  \includegraphics[width=0.9\hsize]{figures/aRp_bRp.eps}
+\caption{\label{fig:crescentfit} The 2$\sigma$ contours (or error ellipses)
+  of the crescent model when fitting with the three different
+  datasets.}
+\end{figure}
+
+Let us first consider the three cases where a crescent model was
+fitted.  These are the solid curves in Figure~\ref{fig:mcmc}, with the
+colours of the curves indicating the source.  Meanwhile,
+Figure~\ref{fig:crescentfit} shows $2\sigma$ of the inferred parameter
+values.
+
+\begin{enumerate}
+
+\item[1) {\bf CC}:] solid red curve in Figure~\ref{fig:mcmc} and red
+  ellipses in Figure~\ref{fig:crescentfit}.  In this case, a light
+  curve from a crescent source was being fitted to a crescent model.
+  The fit gives reduced $\chi^2$ close to unity, as expected.  The
+  recovered parameter values are near or slightly outside the
+  $2\sigma$ ellipses.  This is expected, since we added a small
+  systematic error, by using different caustics (though both clean
+  folds) for the mock data and the model fit.  The apparent degeneracy
+  between $a$ and $b$, seen in Figure~\ref{fig:crescentfit}, is also
+  expected, since only the distance of the crescent's small circle
+  from the caustic influences the magnification.
+
+\item[2) {\bf CD}:] solid blue curve and blue ellipses.  Here a
+  crescent model was fitted to a light curve from a uniform disc.  The
+  situation is formally a crescent with $R_n=0$ and $a,b$ arbitrary,
+  and this shows in the blue ellipses in the recovered parameter
+  values.  The redundant parameters $\alpha$ and $\beta$, in effect, allow the
+  model to partly fit the noise, and hence the $\chi^2$ is somewhat
+  lower than in the previous case.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\hsize]{figures/burnin_cg.eps}
+\caption{\label{fig:burnin} The $\chi^2$ of each step of the full MCMC
+  for one case.}
+\end{figure}
+
+\item[3) {\bf CG}:] Solid green curve and green ellipses.
+  Additionally, Figure~\ref{fig:burnin} shows the progress of
+  reduction of $\chi^2$ in this case, where a crescent-source model is
+  used to fit data from a Gaussian source.  The parameter values are
+  seemingly tightly constrained, but the recovered $R_p$ is completely
+  wrong, since the correct value is far from the green curve and green
+  ellipses.  The reduced $\chi^2$ is only 0.85, indicative of
+  over-fitting.  We can see what has happened from the red curve in
+  the bottom panel of Figure~\ref{fig:mockdata}.  The best-fit model
+  covers only a small part of the light curve.  That is, the fitting
+  procedure has exploited the nuisance parameters to find a good fit
+  to mainly noise.
+
+\end{enumerate}
+
+Next we consider the three cases where a disc model is fitted.  These
+correspond to the dashed curves in Figure~\ref{fig:mcmc}.  There is
+only one interesting parameter to fit, the disc radius $R_p$, and no
+equivalent of Figure~\ref{fig:crescentfit} is needed.
+
+\begin{enumerate}
+
+\item[4) {\bf DC}:] red dashed curve.  On fitting a disc model to a
+  crescent-source source, the recovered $R_p$ is incorrect, but the
+  reduced $\chi^2$ of 1.29 signals that the data reject the model.
+
+\item[5) {\bf DD}:] blue dashed curve.  When the correct model is
+  fitted to a disc-source, the best fit $\chi^2=0.91$ is good and
+  $R_p$ is recovered within the uncertainty estimate.
+
+\item[6) {\bf DG}:] green dashed curve.  On fitting a disc model to a
+  Gaussian-source model, the recovered $R_p$ is incorrect, but the
+  reduced $\chi^2$ gives no signal that something is amiss.  It
+  appears that a Gaussian source could mimic a uniform disc.
+
+\end{enumerate}
+
+Finally, we consider the three cases where a Gaussian model is fitted.
+These correspond to the dotted curves in Figure~\ref{fig:mcmc}.  Again,
+there is only one interesting parameter to fit, the Gaussian
+$3\sigma$-radius, which we have called $R_p$.
+
+\begin{enumerate}
+
+\item[7) {\bf GC}:] red dotted curve.  When a Gaussian-source model is
+  fitted to data from a crescent source, the recovered $R_p$ is wrong
+  but the reduced $\chi^2=1.29$ shows the data rejecting the model.
+
+\item[8) {\bf GD}:] blue dotted curve.  When a Gaussian-source model
+  is fitted to data from a uniform disc, the recovered $R_p$ is wrong
+  but the model is rejected anyway.
+
+\item[9) {\bf GG}:] green dotted curve.  When data from a Gaussian is
+  fitted to the correct model, $R_p$ is recovered within its estimated
+  uncertainty, and the best fit reduced $\chi^2$ is close to unity.
+
+\end{enumerate}
+
+The above results suggest the following strategy for fitting a
+lightcurve from an unknown source profile: first try a Gaussian-source
+model; if the data reject that model, try a uniform disc; if the
+uniform disc is also rejected by the data, try a crescent model.  Our
+numerical experiments indicate --- assuming one of the three source
+models is correct --- that the reduced $\chi^2$ would unmask the
+correct model, \textbf{and its parameters would be correctly recovered}.
+
+\section{Discussion}\label{sec:discussion}
+
+In the current paper we simulate and study the resulting microlensing lightcurves of geometric crescent-shaped sources 
+and compare them with the microlensing lightcurves of other simple mathematically describable source profiles. 
+In order to mimic the behaviour of the flux of light from the source in the proximity of a fold caustic we make use of the simple approximation described in equation (5). 
+The equation would exhaustively describe the magnification map and offer a good universal approximation for the 
+particular microlensing regime that we consider. 
+Namely, the shape of the caustic boundary in the proximity of the source in the respective plane can be approximated with
+ a line due to reason that the local radius of curvature of caustic is orders of magnitude greater than the half-light
+ radius of the studied source. 
+In particular cases in which the previously mentioned approximation loses its validity the impact on the quality of the 
+lightcurves is not evenly distributed. The shape of the lightcurve will be maintained. 
+The data points corresponding to the source position before and during the overlapping of the caustic will be affected by
+ smaller errors than the data points corresponding to later times. \\
+
+The first two source profiles that we consider are the uniform disk and symmetric Gaussian source. 
+Both of them can be described by a half-light radius $r_{1/2}$ and a total unlensed light flux $S_0$. 
+With the two parameters constrained the one dimensional profiles as well as the lightcurves of the two source are completely determined, since no free parameter remains.
+
+The previous statement does not hold for a crescent source.
+In the case of the crescent source there are in total five parameters: the integrated flux of the source $S_0$,
+ the radii of the bright/dark disk $R_p$/$R_n$ and the displacement of the centers of the two disks on the axes perpendicular and parallel to the caustic $a$ and $b$.
+
+Two of the parameters can be reduced by expressing the results in terms of $S_0$ and $r_{1/2}$. 
+The later being determinable for any set of parameters $R_p, R_n$ and $a^2+b^2$. Moreover, one of the displacement 
+parameters $b$ has no impact on the one dimensional profile of the source that results from the projection of the source image on an axis perpendicular to the caustic.
+
+Since the one dimensional source profile that corresponds to an axis perpendicular to the caustic contains exhaustively all the information regarding the source that can be revealed by the lightcurve,
+the value of the parameter $b$ does not have an effect on the shape of the lighcurve.
+
+Nevertheless the $b$ parameter is relevant for the calculation of $r_{1/2}$. 
+It's qualitative effect is to decrease the value of the half-light radius when the absolute value of the parameter is increased.  
+With two parameters constrained and another irrelevant to the shape of the lightcurve, two free parameters remain $R_n$ and $a$. 
+Figure 4 reveals that the lighcurve of a crescent source has more visible features than the other two light-curves 
+corresponding to the disc and Guasssian shape. The parameters $R_n$ and $a$ have strong influences on the shape of 
+the microlensed lightcurve, as can be seen in figures 5, 6 and 7.
+Moreover, the one dimensional source profile corresponding to the direction perpendicular to the caustic reveal four characteristic points.
+The overlap of each of these points with the caustic leaves visible features on the lightcurve at the corresponding instances of time.
+In timely order the instances correspond to the start of the overlap between the caustic and the bright disk, the start of the overlap between the caustic and the dark disk, the end of the overlap between the caustic and dark disk and finally the end of the overlap between the caustic and the bright disk.
+
+With the different source profiles and their corresponding lightcurves studied we can change our point of view of the system to that of an observer. 
+The observer would basically detect only the lightcurve of such a source. As described in section 4.3 the timing of the onset and offset of the previously 
+described periods can be used to estimate the values of the radii and one of the displacement parameters when assuming a geometric crescent. 
+All quantities can be estimated in terms of the relative velocity of the source in a direction perpendicular to the caustic.
+
+Furthermore, a simulated image of M87 presented in \citep{2012MNRAS.421.1517D} (figure 9) has been microlensed. On the resulting lightcurve the instances 
+corresponding to the start and end of the black hole shadow and caustic overlap were distinguishable as can be seen in figure 10.
+
+In case of a high quality lightcurve with insignificant noise and measurement errors, the parameters can be obtained 
+by simply identifying the characteristic instances of time without making use of the actual values of the magnification
+ map. If the effect of the errors and the noise distorts the magnification time function enough so that the 
+characteristic epochs are not identifiable with the characteristic periods still visible, the boundaries of the periods 
+can be roughly estimated. Furthermore, if direct estimates of the parameters cannot be obtained we propose the use of a strong
+statistical tool such as Markov-Chain Monte Carlo. In Section 6  we have studied the possibility of identifying a crescent 
+source and  the possibility of recovering the respective parameters. As magnification map we have used a complex numerical 
+one generated with the microlensing code by \cite{1999A&A...346L...5W}. In addition we have added to the signal a Gaussian 
+noise with a SNR of 1.6.  In the experiment we have considered all nine combinations of original source profiles and assumed fitting source models.
+Effectively we have fitted using MCMC all three sources with all three assumed fitting models. The results of the experiment allowed us build a 
+procedure for distinguishing the shape of the source assuming that one of the three models we have considered is a good approximation.  
+The procedure would be useful for observers that endeavour to gain more information about an unresolved source for which they can study 
+the microlensing lightcurve.  As a first step in the procedure, one should first attempt to use MCMC with a Gaussian model assumption. 
+If either the value of the $\chi^2$ or the number of rejected datapoints is large then the next step is to change the assumed source model to 
+an uniform disc and redo the fitting. Finally if the uniform disc assumption is rejected as well then the fitting should be done with a 
+crescent source model assumption. If at each of the three steps the data rejects the model then the source cannot be approximated by any of the three models. 
+Otherwise if at one of the steps the data does not reject the model then the respective model is a good approximation.
+
+The previously mentioned abstract parameters can be related to
+physical quantities specific to the central region of a quasar.  As
+such the luminous region would correspond to the bright accretion disc
+that surrounds the black hole. The later's gravity would cause a
+shadow in the bright region limited by the extent of the event horizon
+of the black hole. Therefore the radius of the bright disk would
+provide an estimate of the size of the accretion disk and the radius
+of the dark disc would provide and estimate of the gravitationally
+magnified Schwarzschild radius of the black hole $R_{S}^{magnified} =
+\Delta t_{dark} \cdot v_p$.  By $R_{S}^{magnified}$ we refer to the
+apparent Schwarzschild radius which is larger than the real value at
+large distances due to the black hole's own gravity.  Moreover the
+gravitationally magnified value of the Schwarzschild radius is a
+monotonic function of the black hole's mass. Therefore it can be used
+to estimate the mass of the black hole if it was not rotating. In the
+previous expression the $\Delta t_{dark}$ denotes the period of
+overlap between the black hole shadow and fold caustic.  $v_p$ denotes
+the component of the relative velocity of the source and fold which is
+perpendicular to the caustic.  The respective velocity is an unknown,
+though it can be constrained on a case by case basis to an order of
+magnitude or even better. This would require the study of the dynamics
+of the stellar structure which contains the gravitational lens.  A
+better estimate of the relative velocity would facilitate a better
+estimate of the effective non-rotating black hole mass associated to
+the black hole shadow.
+
+The parameters whose values cannot be determined due to the loss of information from the directions parallel to the 
+caustic could be obtained in the eventuality in which the same source crosses multiple caustics that are not parallel. 
+Multiple crossing of caustics can reveal details of the one dimensional flux profile corresponding to multiple distinct 
+directions which would allow the reconstruction of the two dimensional profile analogous to the process through which an image of a CT scan is obtained.
 

### 2016-03-14 paper1/ms.tex
Title change
Changed figure 2

@@
 
-\title{Probing the internal structures of quasars using microlensing events}
-
+\title{Microlensing as a possible probe of event-horizon structure in quasars}
 \author[Tomozeiu et al]{Mihai Tomozeiu\thanks{mihai@physik.uzh.ch},$^{1,2}$ 
@@
 
+\textit{Author contributions}\\
+Prasenjit Saha provided the original idea and plan for the research project as well as multiple contributions to the analysis.
+ Mihai Tomozeiu simulated and studied the ideal behaviour of the microlensing lightcurves for the different source profiles discussed.
+ Joachim Wambsganss provided the numerical code used by Manuel Rabold to create the magnification map and corresponding lighcurves 
+used in the MCMC analysis performed by Irshad Mohammed in the last part of the presented work.
+
 \bibliographystyle{mn2e}

### 2016-03-14 paper1/ms.tex
Changes

@@
 \textit{Author contributions}\\
-Prasenjit Saha provided the original idea and plan for the research project as well as multiple contributions to the analysis.
- Mihai Tomozeiu simulated and studied the ideal behaviour of the microlensing lightcurves for the different source profiles discussed.
+Prasenjit Saha provided the original idea and plan for the research project as well as multiple contributions to the analysis and manuscript preparation.
+Mihai Tomozeiu simulated and studied the ideal behaviour of the microlensing lightcurves for the different source profiles discussed and prepared the manuscript.
  Joachim Wambsganss provided the numerical code used by Manuel Rabold to create the magnification map and corresponding lighcurves 
 used in the MCMC analysis performed by Irshad Mohammed in the last part of the presented work.
+Both Manuel Rabold and Irshad Mohammed had large contributions in writting the "Fitting Mock Data" section.
 

### 2016-03-14 paper1/ms.tex
typo fix

@@
 of a black-hole environment
-(Figures~\ref{fig:M87_plots}--\ref{fig:char_points}).
+(Figures~\ref{fig:M87_image}--\ref{fig:M87_plots}).
 

### 2016-03-14 paper1/ms.tex
few grammar and typos are fixed

@@
 
-Active galactic nuclei are thought to be powered through the
+Active galactic nuclei are thought to be powered by the
 accretion of matter from the proximal environment into a supermassive
@@
 and optical continuum.  Moreover, in the direction perpendicular to
-the accretion disk, where the medium is more transparent, jets will
+the accretion disc, where the medium is more transparent, jets will
 appear.  If a jet is oriented towards the Earth, a quasar is observed
@@
 \citep{2008JPhCS.131a2055D,2012Sci...338..355D,2013MNRAS.434..765K,2016arXiv160205527F}.
-Current data do not deliver images, but require fitting to
+Current data do not deliver images but require fitting to
 predefined models for the images.  A whole range of models have been
@@
 of magnitude smaller on the sky. The direct observations of the black
-hole silhouettes of quasars is far beyond foreseeable instrumentation.
-In the present paper we consider a possible indirect method, related
+hole silhouettes of quasars are far beyond foreseeable instrumentation.
+In the present paper, we consider a possible indirect method, related
 to \cite{1999ApJ...524...49A}, through which gravitational
@@
 
-Gravitational microlensing of quasars, reviewed in
+Gravitational microlensing of quasars reviewed in
 Section~\ref{sec:microlensing} below, refers to sharp changes in the
-observed brightness of quasars that have been lensed by intervening
+observed brightness of quasars that have been lensed by an intervening
 galaxies, without any changes in the intrinsic luminosity.
@@
 integrating the surface brightness weighted by the magnification.
-Most astrophysical sources straddle several caustics, and hence their
+Most astrophysical sources straddle several caustics, and hence, their
 net brightness varies smoothly with location.  The central engine of
@@
 have applied the idea to observed light curves to infer properties of
-quasar accretion discs.  In this work we take a simpler but arguably
+quasar accretion discs.  In this work, we take a simpler but arguably
 more robust approach: we study features in the light curves
@@
 constraints can be set with independent observations of the stellar
-structure that contains the gravitational lens.
+structure that contain the gravitational lens.
 
@@
 equation (\ref{eqn:lens}), to the source plane $\vec\beta$ gives the
-so-called caustics.  Caustics can appear in optical system, not just
+so-called caustics.  Caustics can appear in the optical system, not just
 gravitational lensing.  For a point source, caustics are singularities
@@
 caustic network on the scale of $\sqrt{GM_\odot D_L/c^2}$ or $10^{-6}
-\sqrt{M/M_\odot}\rm\,arcsec$ for typical lens and source redshifst
+\sqrt{M/M_\odot}\rm\,arcsec$ for typical lens and source redshift
 \citep{2001PASA...18..207W}.  Extended sources wash out this
@@
 be locally constant, and a decrease with the square root of the
-distance from the fold. The later term becomes activated only after
+distance from the fold. The latter term becomes activated only after
 the source enters the region interior to the caustic curves when the
-values of the stepfunction $\Theta(p)$ become unity. The
+values of the step function $\Theta(p)$ become unity. The
 proportionality constant $C_0$ depends on the local conditions in the
@@
 
-A source of arbitrary shape can be described by a two dimensional brightness function $S_{2D}(p - p_s, q - q_s)$ defined for a coordinate system $p,q$ where $p_s, q_s$ denote the coordinates of the center of source.
+A source of arbitrary shape can be described by a two-dimensional brightness function $S_{2D}(p - p_s, q - q_s)$ defined for a coordinate system $p,q$ where $p_s, q_s$ denote the coordinates of the center of a source.
 
@@
 The surface brightness distribution of a geometric crescent can be built by considering two disk sources of constant brightness.
-One larger disk will contribute positively to the total flux, while one smaller disk
+One larger disk will contribute positively to the total flux while one smaller disk
 that is interior to the large one will contribute negatively. This superposition can be written for 2D as:\\
@@
 presented in the previous section one can compute numerically the
-lightcurves of the three extented sources for the simplified
+lightcurves of the three extended sources for the simplified
 infinite-wall-caustics model.
@@
 
-There are four characteristic points visible on the resulting one dimensional light profile of the crescent shaped source.
+There are four characteristic points visible on the resulting one-dimensional light profile of the crescent shaped source.
 Two of the points, p1 and p4, mark the outer boundaries of the luminous disc component.
@@
 
-Figure 4 reveals that the lighcurve of a crescent source has more visible features than the other two light-curves
-corresponding to the disc and guasssian shape. There are three regions where $S_{1D}$ and by inheritance the lighcurve
- have distinguishable behaviour. The first region that would be recorded on a lightcurve plot represents the period
+Figure 4 reveals that the lightcurve of a crescent source has more visible features than the other two light-curves
+corresponding to the disc and gaussian shape. There are three regions where $S_{1D}$ and by inheritance, the lightcurve
+ has distinguishable behaviour. The first region that would be recorded on a lightcurve plot represents the period
 of time when the bright disk begins to be overlapped by the caustic and stops when just before the dark disk reaches
-the caustic. During this period of time the flux of light from the source is increasingly magnified. The second
-period starts and ends with the overlapping of the dark disc. As a boundary of the two regions there is a
-distinguishable point where slope of the magnification is drastically changed. This apparent discontinuity in the
+the caustic. During this period of time, the flux of light from the source is increasingly magnified. The second
+period starts and ends with the overlapping of the dark disc. As a boundary of the two regions, there is a
+distinguishable point where the slope of the magnification is drastically changed. This apparent discontinuity in the
 first derivative of the magnification function is caused by the caustic amplification of the sudden drop in the $S_{1d}$
-function. During the respective period the magnification growth slows or even reverses during the first part of the
+function. During the respective period, the magnification growth slows or even reverses during the first part of the
 period and starts to grow faster again as the dark disk ends its overlap with the caustic. At the point where
-the dark disc clears the caustic the growth of the magnification is infinite, which appears as a saddle point
-on the lightcurve. Next, the final period correspond to the case when the dark disc has cleared the caustic
-and the bright disc continues to overlap with the caustic. During this period the lighcurve reaches a peak
+the dark disc clears the caustic, the growth of the magnification is infinite, which appears as a saddle point
+on the lightcurve. Next, the final period corresponds to the case when the dark disc has cleared the caustic
+and the bright disc continues to overlap with the caustic. During this period, the lightcurve reaches a peak
 that for most of the parameter space is global and for the rest of the parameter space local.
-At the end of the period the growth of the magnification is negative infinite. the respective point appears 
+At the end of the period, the growth of the magnification is negative infinite. the respective point appears 
 as a second saddle point on the lightcurve. Past this point the magnification of any finite source will 
@@
 larger magnifications until it becomes a local and even global peak. The effect of the $R_n$  parameter on the
-lighcurve is presented in figures 6 and 7. For the particular set of parameters where $R_p = R_n +a$ the third
+light curve is presented in figures 6 and 7. For the particular set of parameters where $R_p = R_n +a$ the third
 period of time discussed previously does not exist. Particularizing further, if the value of the radius of the dark
@@
  happens when the dark region reaches the caustic and it is characterized by a sharp variation in magnification growth,
-In the opposite case the peak appears before the end of the bright disc reaches the caustic and shape is smoother. \\
+In the opposite case, the peak appears before the end of the bright disc reaches the caustic and shape is smoother. \\
 
@@
 \caption{\label{fig:M87_image} The simulated image of M87's black hole
-  sillouette obtained by \citep{2012MNRAS.421.1517D} from a GRMHD
+  silhouette obtained by \citep{2012MNRAS.421.1517D} from a GRMHD
   simulation presented in \citep{2009MNRAS.394L.126M}. The color bar
@@
 \cite{2012MNRAS.421.1517D} have created a radiative image of M87 based on the GRMHD simulations presented in \citep{2009MNRAS.394L.126M}.
-The respective image, reproduced in figure \ref{fig:M87_image}, has been projected to a 1D profile associated to the perpendicular direction to a fold caustic approaching the image from the right.
+The respective image reproduced in figure \ref{fig:M87_image}, has been projected to a 1D profile associated to the perpendicular direction to a fold caustic approaching the image from the right.
 The projection is presented in the left panel of figure 10.
 The amplification values of the flux of light corresponding to a microlensing event are presented in the right panel of figure 10.
-In general the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions surrounding the luminous parts of the image have non-zero flux and thus are more extended than the simplified source model.
+In general, the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions surrounding the luminous parts of the image have non-zero flux and thus are more extended than the simplified source model.
 
@@
   panel. The points p2 and p3 mark the start and end of the equivalent
-  dark disk and are associated to the moments of their overlap with
+  dark disk and are associated with the moments of their overlap with
   the caustic: t2 and t3.}
@@
 For the computation of the individual deflection angles
-$\vec\alpha(\vec\theta)$ a heirarchical or tree method is used.  The
+$\vec\alpha(\vec\theta)$ a hierarchical or tree method is used.  The
 positions of all lensing masses are put into a grid of $\vec\theta$.
@@
 until every cell contains only one mass.  Nearby masses are added
-individually, while distance masses are clumped into larger grid cells
+individually while distance masses are clumped into larger grid cells
 whose net contribution is approximated by its first few multipole
@@
 distribution of the source with the magnification pattern of the map.
-In real life, not only are both lens and source moving, but the lens
+In real life, not only are both lens and source moving but the lens
 lens configuration, and with it the magnification pattern, is also
@@
 $y$-direction respectively. In the original version of the code,
-gaussian and disk shaped images where already implemented. Those are
+gaussian and disk-shaped images were already implemented. Those are
 completely characterised by the single parameter $R_p$. The values of
 the parameters are specified in pixel units corresponding to the
-magnification map. Further one needs to specify the start and end
+magnification map. Further, one needs to specify the start and end
 point coordinates of the path, which the center of the source image
 follows through the magnification map (see the depiction in figure
-\ref{fig:magnification_map}). Hence the points along this path are
-specified through the number of timesteps for which the computation of
+\ref{fig:magnification_map}). Hence, the points along this path are
+specified through the number of time steps for which the computation of
 the brightness is to be carried out. Those points correspond to the
@@
 
-For the purpose of this analysis it was desirable to mimic the
+For the purpose of this analysis, it was desirable to mimic the
 analytical behaviour of a simple fold as much as possible, for
-comparing the numerical result with the analytical one, therefore the
+comparing the numerical result with the analytical one, therefore, the
 path of the source was chosen, so that it intersects the border of the
@@
   recovered parameter values are near or slightly outside the
-  $2\sigma$ ellipses.  This is expected, since we added a small
+  $2\sigma$ ellipses.  This is expected since we added a small
   systematic error, by using different caustics (though both clean
@@
   values.  The redundant parameters $\alpha$ and $\beta$, in effect, allow the
-  model to partly fit the noise, and hence the $\chi^2$ is somewhat
+  model to partly fit the noise, and hence, the $\chi^2$ is somewhat
   lower than in the previous case.
@@
   seemingly tightly constrained, but the recovered $R_p$ is completely
-  wrong, since the correct value is far from the green curve and green
+  wrong since the correct value is far from the green curve and green
   ellipses.  The reduced $\chi^2$ is only 0.85, indicative of
@@
 \item[9) {\bf GG}:] green dotted curve.  When data from a Gaussian is
-  fitted to the correct model, $R_p$ is recovered within its estimated
-  uncertainty, and the best fit reduced $\chi^2$ is close to unity.
+  fitted to the correct model, $R_p$ is recovered within it's estimated
+  uncertainty and the best fit reduced $\chi^2$ is close to unity.
 
@@
 uniform disc is also rejected by the data, try a crescent model.  Our
-numerical experiments indicate --- assuming one of the three source
-models is correct --- that the reduced $\chi^2$ would unmask the
+numerical experiments indicate --- assuming one of the three sources
+models are correct --- that the reduced $\chi^2$ would unmask the
 correct model, \textbf{and its parameters would be correctly recovered}.
@@
 
-In the current paper we simulate and study the resulting microlensing lightcurves of geometric crescent-shaped sources 
+In the current paper, we simulate and study the resulting microlensing lightcurves of geometric crescent-shaped sources 
 and compare them with the microlensing lightcurves of other simple mathematically describable source profiles. 
-In order to mimic the behaviour of the flux of light from the source in the proximity of a fold caustic we make use of the simple approximation described in equation (5). 
+In order to mimic the behaviour of the flux of light from the source in the proximity of a fold caustic, we make use of the simple approximation described in equation (5). 
 The equation would exhaustively describe the magnification map and offer a good universal approximation for the 
 particular microlensing regime that we consider. 
-Namely, the shape of the caustic boundary in the proximity of the source in the respective plane can be approximated with
+Namely, the shape of the caustic boundary in the proximity of the source in the respective plane can be approximated by
  a line due to reason that the local radius of curvature of caustic is orders of magnitude greater than the half-light
  radius of the studied source. 
-In particular cases in which the previously mentioned approximation loses its validity the impact on the quality of the 
+In particular cases in which the previously mentioned approximation loses its validity, the impact on the quality of the 
 lightcurves is not evenly distributed. The shape of the lightcurve will be maintained. 
@@
 Both of them can be described by a half-light radius $r_{1/2}$ and a total unlensed light flux $S_0$. 
-With the two parameters constrained the one dimensional profiles as well as the lightcurves of the two source are completely determined, since no free parameter remains.
+With the two parameters constrained the one-dimensional profiles, as well as the lightcurves of the two source, are completely determined, since no free parameter remains.
 
@@
 The later being determinable for any set of parameters $R_p, R_n$ and $a^2+b^2$. Moreover, one of the displacement 
-parameters $b$ has no impact on the one dimensional profile of the source that results from the projection of the source image on an axis perpendicular to the caustic.
-
-Since the one dimensional source profile that corresponds to an axis perpendicular to the caustic contains exhaustively all the information regarding the source that can be revealed by the lightcurve,
-the value of the parameter $b$ does not have an effect on the shape of the lighcurve.
-
-Nevertheless the $b$ parameter is relevant for the calculation of $r_{1/2}$. 
+parameters $b$ has no impact on the one-dimensional profile of the source that results from the projection of the source image on an axis perpendicular to the caustic.
+
+Since the one-dimensional source profile that corresponds to an axis perpendicular to the caustic contains exhaustively all the information regarding the source that can be revealed by the lightcurve,
+the value of the parameter $b$ does not have an effect on the shape of the lightcurve.
+
+Nevertheless, the $b$ parameter is relevant for the calculation of $r_{1/2}$. 
 It's qualitative effect is to decrease the value of the half-light radius when the absolute value of the parameter is increased.  
 With two parameters constrained and another irrelevant to the shape of the lightcurve, two free parameters remain $R_n$ and $a$. 
-Figure 4 reveals that the lighcurve of a crescent source has more visible features than the other two light-curves 
-corresponding to the disc and Guasssian shape. The parameters $R_n$ and $a$ have strong influences on the shape of 
-the microlensed lightcurve, as can be seen in figures 5, 6 and 7.
-Moreover, the one dimensional source profile corresponding to the direction perpendicular to the caustic reveal four characteristic points.
+Figure 4 reveals that the lightcurve of a crescent source has more visible features than the other two light-curves 
+corresponding to the disc and Gaussian shape. The parameters $R_n$ and $a$ have strong influences on the shape of 
+the microlensed lightcurve as can be seen in figures 5, 6 and 7.
+Moreover, the one-dimensional source profile corresponding to the direction perpendicular to the caustic reveal four characteristic points.
 The overlap of each of these points with the caustic leaves visible features on the lightcurve at the corresponding instances of time.
-In timely order the instances correspond to the start of the overlap between the caustic and the bright disk, the start of the overlap between the caustic and the dark disk, the end of the overlap between the caustic and dark disk and finally the end of the overlap between the caustic and the bright disk.
+In timely order, the instances correspond to the start of the overlap between the caustic and the bright disk, the start of the overlap between the caustic and the dark disk, the end of the overlap between the caustic and dark disk and finally the end of the overlap between the caustic and the bright disk.
 
@@
 
-In case of a high quality lightcurve with insignificant noise and measurement errors, the parameters can be obtained 
+In the case of a high-quality lightcurve with insignificant noise and measurement errors, the parameters can be obtained 
 by simply identifying the characteristic instances of time without making use of the actual values of the magnification
@@
 statistical tool such as Markov-Chain Monte Carlo. In Section 6  we have studied the possibility of identifying a crescent 
-source and  the possibility of recovering the respective parameters. As magnification map we have used a complex numerical 
-one generated with the microlensing code by \cite{1999A&A...346L...5W}. In addition we have added to the signal a Gaussian 
-noise with a SNR of 1.6.  In the experiment we have considered all nine combinations of original source profiles and assumed fitting source models.
-Effectively we have fitted using MCMC all three sources with all three assumed fitting models. The results of the experiment allowed us build a 
+source and  the possibility of recovering the respective parameters. As magnification map, we have used a complex numerical 
+one generated with the microlensing code by \cite{1999A&A...346L...5W}. In addition, we have added to the signal a Gaussian 
+noise with an SNR of 1.6.  In the experiment, we have considered all nine combinations of original source profiles and assumed fitting source models.
+Effectively we have fitted using MCMC all three sources with all three assumed fitting models. The results of the experiment allowed us to build a 
 procedure for distinguishing the shape of the source assuming that one of the three models we have considered is a good approximation.  
@@
 If either the value of the $\chi^2$ or the number of rejected datapoints is large then the next step is to change the assumed source model to 
-an uniform disc and redo the fitting. Finally if the uniform disc assumption is rejected as well then the fitting should be done with a 
+a uniform disc and redo the fitting. Finally, if the uniform disc assumption is rejected as well then the fitting should be done with a 
 crescent source model assumption. If at each of the three steps the data rejects the model then the source cannot be approximated by any of the three models. 
-Otherwise if at one of the steps the data does not reject the model then the respective model is a good approximation.
+Otherwise, if at one of the steps the data does not reject the model then the respective model is a good approximation.
 
@@
 shadow in the bright region limited by the extent of the event horizon
-of the black hole. Therefore the radius of the bright disk would
+of the black hole. Therefore, the radius of the bright disk would
 provide an estimate of the size of the accretion disk and the radius
@@
 apparent Schwarzschild radius which is larger than the real value at
-large distances due to the black hole's own gravity.  Moreover the
+large distances due to the black hole's own gravity.  Moreover, the
 gravitationally magnified value of the Schwarzschild radius is a
-monotonic function of the black hole's mass. Therefore it can be used
+monotonic function of the black hole's mass. Therefore, it can be used
 to estimate the mass of the black hole if it was not rotating. In the
-previous expression the $\Delta t_{dark}$ denotes the period of
+previous expression, the $\Delta t_{dark}$ denotes the period of
 overlap between the black hole shadow and fold caustic.  $v_p$ denotes
@@
 better estimate of the relative velocity would facilitate a better
-estimate of the effective non-rotating black hole mass associated to
+estimate of the effective non-rotating black hole mass associated with
 the black hole shadow.
@@
 caustic could be obtained in the eventuality in which the same source crosses multiple caustics that are not parallel. 
-Multiple crossing of caustics can reveal details of the one dimensional flux profile corresponding to multiple distinct 
-directions which would allow the reconstruction of the two dimensional profile analogous to the process through which an image of a CT scan is obtained.
+Multiple crossing of caustics can reveal details of the one-dimensional flux profile corresponding to multiple distinct 
+directions which would allow the reconstruction of the two-dimensional profile analogous to the process through which an image of a CT scan is obtained.
 
@@
 Mihai Tomozeiu simulated and studied the ideal behaviour of the microlensing lightcurves for the different source profiles discussed and prepared the manuscript.
- Joachim Wambsganss provided the numerical code used by Manuel Rabold to create the magnification map and corresponding lighcurves 
+ Joachim Wambsganss provided the numerical code used by Manuel Rabold to create the magnification map and corresponding lightcurves 
 used in the MCMC analysis performed by Irshad Mohammed in the last part of the presented work.
-Both Manuel Rabold and Irshad Mohammed had large contributions in writting the "Fitting Mock Data" section.
+Both Manuel Rabold and Irshad Mohammed had large contributions in writing the "Fitting Mock Data" section.
 

### 2016-07-04 fitting/overlays.py
three panels

@@ -1,0 +1,72 @@
+import numpy as np
+import lcurv
+import sys
+import scipy.stats as st
+import matplotlib.pyplot as pl
+
+def mockdata():
+    global pars
+    lcurv.readmap()
+    Rp = 39.4
+    sig = 20/(np.log(4))**.5
+    Rn,a,b = 0.4, 0, 0.3
+    pars = [ 51, 290, 1.0, Rp,  Rn, a, b ]
+    pars = np.array(pars)
+    print(pars)
+    errlev = 0.04
+    lcurv.mockdata(pars,errlev)
+
+
+def gencurves(fname):
+    chain = np.genfromtxt(fname+'.chain', skip_header=0, skip_footer=1)
+    print(chain.shape)
+    minchis=1e30
+    for i in range(chain.shape[0]):
+        x = chain[i,:]
+        if x[1] < minchis:
+            q = x[3:]
+            minchis = x[1]
+    print(q)
+    print(minchis,lcurv.chis(q),1-st.chi2.cdf(minchis,pars[1]-pars[0]+1-7))
+    lcurv.writecurves(q)
+
+
+def plotcurves(p):
+    fil = open('curves.txt')
+    all = fil.readlines()
+    t = []
+    m = []
+    b = []
+    db = []
+    for l in all:
+        s = l.split()
+        t.append(int(s[0])/30.)
+        m.append(float(s[1]))
+        b.append(float(s[2]))
+        db.append(float(s[3]))
+    p.plot(t,m)
+    p.axis([0,6,0,7])
+    p.errorbar(t,b,yerr=db,linestyle='none',color='gray')
+    p.set_yticks(np.arange(0,7,2))
+
+
+fig = pl.figure()
+
+mockdata()
+
+gencurves('cc_forward')
+p = fig.add_subplot(311)
+p.get_xaxis().set_visible(False)
+plotcurves(p)
+
+gencurves('cc_backward')
+p = fig.add_subplot(312)
+p.get_xaxis().set_visible(False)
+plotcurves(p)
+
+gencurves('gc_forward')
+p = fig.add_subplot(313)
+plotcurves(p)
+
+pl.show()
+

### 2016-07-05 fitting/rhalf.py
summary fitting figure

@@ -22,12 +22,12 @@ def Rhalf(Rp,Rn,a,b):
 		if area[i-1] < 0.5*area[-1] and area[i] > 0.5*area[-1]:
 			return r[i]
 
-Rp = float(argv[1])
-Rn = 0.4*Rp
-a = 0
-b = 0.3*Rp
-
-print Rhalf(Rp,Rn,a,b)
+if __name__=="__main__":
+	Rp = float(argv[1])
+	Rn = 0.4*Rp
+	a = 0
+	b = 0.3*Rp
+	print(Rhalf(Rp,Rn,a,b))
 
 
 

### 2016-07-05 fitting/summaryfig.py

@@ -1,14 +1,15 @@
 import numpy as np
-import lcurv
-import sys
 import scipy.stats as st
 import matplotlib.pyplot as pl
+import lcurv
+import sys
+from rhalf import Rhalf
+
 
 def mockdata():
     global pars
     lcurv.readmap()
     Rp = 39.4
-    sig = 20/(np.log(4))**.5
     Rn,a,b = 0.4, 0, 0.3
     pars = [ 51, 290, 1.0, Rp,  Rn, a, b ]
     pars = np.array(pars)
@@ -18,14 +19,30 @@ def mockdata():
 
 
 def gencurves(fname):
+    global rh,wt
     chain = np.genfromtxt(fname+'.chain', skip_header=0, skip_footer=1)
     print(chain.shape)
+    rh = []
+    wt = []
     minchis=1e30
     for i in range(chain.shape[0]):
         x = chain[i,:]
         if x[1] < minchis:
             q = x[3:]
             minchis = x[1]
+        if chain[i,0] > 100:
+            rp,rn,a,b = chain[i,6],chain[i,7],chain[i,8],chain[i,9]
+            if rn < 0:
+                v = rp*(np.log(4))**.5
+            else:
+                rn *= rp
+                a *= rp
+                b *= rp
+                v = Rhalf(rp,rn,a,b)
+            rh.append(v/30)
+            w = chain[i,0] - chain[i-1,0]
+            w /= chain[-1,0]
+            wt.append(w)
     print(q)
     print(minchis,lcurv.chis(q),1-st.chi2.cdf(minchis,pars[1]-pars[0]+1-7))
     lcurv.writecurves(q)
@@ -48,25 +65,27 @@ def plotcurves(p):
     p.axis([0,6,0,7])
     p.errorbar(t,b,yerr=db,linestyle='none',color='gray')
     p.set_yticks(np.arange(0,7,2))
+    p.set_xlabel('$t [r_{1/2}/v]$')
+    p.set_ylabel('magnification')
 
+def plothist(p):
+    p.hist(rh,np.linspace(0.75,1.25,26),weights=wt)
+    p.set_xlabel('inferred:actual $r_{1/2}$')
 
 fig = pl.figure()
 
 mockdata()
 
-gencurves('cc_forward')
-p = fig.add_subplot(311)
-p.get_xaxis().set_visible(False)
-plotcurves(p)
-
-gencurves('cc_backward')
[…]

### 2016-07-05 paper1/ms.tex
Acknowledgement added
new figures for data-fitting section

@@
 \newauthor
-and Joachim Wambsganss$^4$\\
+and Joachim Wambsganss$^{1,4}$\\
 $^1${Physik-Institut, University of Zurich, Winterthurerstrasse 190,
@@
 
-In quasars that have been lensed by galaxies, the point-like images
-sometimes show sharp brightness variations (microlensing).  These
-brightness changes are associated with the innermost region of the
-quasar passing behind a complicated pattern of caustics due to the
-stars in the lensing galaxy.  In this paper, we study whether the
-universal properties of optical caustics could enable extraction of
-shape information about the central engine of quasars.  We present a
-toy model with a crescent-shaped source crossing a fold caustic.  The
-silhouette of a black hole over an accretion disk tends to produce
-roughly crescent sources.  When a crescent source crosses a fold
-caustic, the resulting light curve is noticeably different from the
-case of a disc or Gaussian source.  The crescent parameters, apart
-from one degeneracy, can be recovered.
+In quasars which are lensed by galaxies, the point-like images sometimes show sharp and uncorrelated brightness 
+variations (microlensing). These brightness changes are associated with the innermost region of the quasar passing 
+through a complicated pattern of caustics produced by the stars in the lensing galaxy. In this paper, we study 
+whether the universal properties of optical caustics could enable extraction of shape information about the central 
+engine of quasars. We present a toy model with a crescent-shaped source crossing a fold caustic. The silhouette 
+of a black hole over an accretion disk tends to produce roughly crescent sources. When a crescent-shaped source 
+crosses a fold caustic, the resulting light curve is noticeably different from the case of a circular luminosity 
+profile or Gaussian source. With good enough monitoring data, the crescent parameters, apart from one degeneracy, 
+can be recovered.
+
 \end{abstract}
@@
   trajectory of the center of the sources for which the lightcurves
-  presented in figure 12 are generated.  Note that this represents an
+  presented in figure 11 are generated.  Note that this represents an
   atypically simple region of any realistic magnification map.}
@@
 
-\begin{figure}
-\centering
-        \includegraphics[width = 0.48\textwidth]{figures/M87_image.eps}
-\caption{\label{fig:M87_image} The simulated image of M87's black hole
-  silhouette obtained by \citep{2012MNRAS.421.1517D} from a GRMHD
-  simulation presented in \citep{2009MNRAS.394L.126M}. The color bar
-  is linear and scaled to reach value 1 for the maximum
-  brightness.}
-\end{figure}
-
 \cite{2012MNRAS.421.1517D} have created a radiative image of M87 based on the GRMHD simulations presented in \citep{2009MNRAS.394L.126M}. 
-The respective image reproduced in figure \ref{fig:M87_image}, has been projected to a 1D profile associated to the perpendicular direction to a fold caustic approaching the image from the right.
-The projection is presented in the left panel of figure 10.
-The amplification values of the flux of light corresponding to a microlensing event are presented in the right panel of figure 10.
-In general, the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions surrounding the luminous parts of the image have non-zero flux and thus are more extended than the simplified source model.
+The top right-most image in figure 5 of \cite{2012MNRAS.421.1517D} has been projected to a 1D profile associated to the perpendicular 
+direction to a fold caustic approaching the image from the right. The projection is presented in the upper panel of figure 9. 
+The amplification values of the flux of light corresponding to a microlensing event are presented in the lower panel of figure 9. 
+In general, the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions 
+surrounding the luminous parts of the image have non-zero flux and thus are more extended than the simplified source model.
 
@@
   from the numerical integration along the ordonata axis of the source
-  image (figure 9) is presented in the lower panel. The corresponding
-  lightcurve associated to the 1D profile is presented in the right
+  image is presented in the upper panel. The corresponding
+  lightcurve associated to the 1D profile is presented in the lower
   panel. The points p2 and p3 mark the start and end of the equivalent
@@
 First we generate microlensing magnification maps using the
-microlensing code by \cite{1999A&A...346L...5W}.  We then generate
+microlensing code created by Joachim Wambsgans (\citep{1999A&A...346L...5W, 1990PhDT.......180W, 1999JCoAM.109..353W}, 
+\cite{1990PhDT.......180W}, \cite{1999JCoAM.109..353W}). We then generate
 mock light curves for three source models --- a crescent, a uniform
@@
 Markov-chain Monte Carlo (MCMC).
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\hsize]{figures/cc_forward.eps}
+\caption{\label{fig:cc_forward} $\chi^2=243, P=0.32$.}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\hsize]{figures/cc_backward.eps}
+\caption{\label{fig:cc_backward} $\chi^2=284, P=0.01$.}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\hsize]{figures/gc_forward.eps}
+\caption{\label{fig:gc_forward} $\chi^2=344, P=3\times10^{-6}$.}
+\end{figure}
 
@@
 models are correct --- that the reduced $\chi^2$ would unmask the
-correct model, \textbf{and its parameters would be correctly recovered}.
+correct model, and its parameters would be correctly recovered.
 
@@
 
-Furthermore, a simulated image of M87 presented in \citep{2012MNRAS.421.1517D} (figure 9) has been microlensed. On the resulting lightcurve the instances 
-corresponding to the start and end of the black hole shadow and caustic overlap were distinguishable as can be seen in figure 10.
+Furthermore, a simulated image of M87 presented in \citep{2012MNRAS.421.1517D} has been microlensed (figure 9). On the resulting lightcurve the instances 
+corresponding to the start and end of the black hole shadow and caustic overlap were distinguishable.
 
@@
 \textit{Author contributions}\\
-Prasenjit Saha provided the original idea and plan for the research project as well as multiple contributions to the analysis and manuscript preparation.
-Mihai Tomozeiu simulated and studied the ideal behaviour of the microlensing lightcurves for the different source profiles discussed and prepared the manuscript.
- Joachim Wambsganss provided the numerical code used by Manuel Rabold to create the magnification map and corresponding lightcurves 
-used in the MCMC analysis performed by Irshad Mohammed in the last part of the presented work.
-Both Manuel Rabold and Irshad Mohammed had large contributions in writing the "Fitting Mock Data" section.
+Prasenjit Saha provided the original idea and plan for the research project as well as multiple contributions to the analysis 
+and manuscript preparation. Mihai Tomozeiu simulated and studied the ideal behaviour of the microlensing lightcurves for the 
+different source profiles discussed and prepared the manuscript.
+Joachim Wambsganss contributed to the research planning and provided the numerical code used by Manuel Rabold to create 
+the magnification map and corresponding lightcurves used in the MCMC analysis performed by Irshad Mohammed in the last 
+part of the presented work. Both Manuel Rabold and Irshad Mohammed had large contributions in writing the "Fitting Mock Data" section.
+
+\section{Acknowledgement}
+IM is supported by Fermi Research Alliance, LLC under Contract No. De-AC02-07CH11359 with the United States Department of Energy.
+J.W. would like to acknowledge and thank the Pauli Center for Theoretical Studies of ETH Zurich and University of Zurich for 
+generous support during the Schrödinger visiting professorship in 2013.
 

### 2016-07-05 fitting/summaryfig.py

@@ -61,31 +61,34 @@ def plotcurves(p):
         m.append(float(s[1]))
         b.append(float(s[2]))
         db.append(float(s[3]))
-    p.plot(t,m)
     p.axis([0,6,0,7])
     p.errorbar(t,b,yerr=db,linestyle='none',color='gray')
+    p.plot(t,m,color='black')
     p.set_yticks(np.arange(0,7,2))
     p.set_xlabel('$t [r_{1/2}/v]$')
     p.set_ylabel('magnification')
 
 def plothist(p):
-    p.hist(rh,np.linspace(0.75,1.25,26),weights=wt)
+    p.hist(rh,np.linspace(0.75,1.25,26),weights=wt,
+           histtype='step',lw=2,color='black')
+    p.axis([0.75,1.25,0,0.6])
     p.set_xlabel('inferred:actual $r_{1/2}$')
 
-fig = pl.figure()
 
 mockdata()
 
-#gencurves('cc_forward')
-#gencurves('cc_backward')
-gencurves('gc_forward')
+for fname in ['gc_forward','cc_forward','cc_backward']:
+    gencurves(fname)
 
-p = fig.add_subplot(211)
-plotcurves(p)
+    fig = pl.figure()
+    fig.subplots_adjust(hspace=0.3)
 
-p = fig.add_subplot(212)
-plothist(p)
-p.get_yaxis().set_visible(False)
+    p = fig.add_subplot(211)
+    plotcurves(p)
 
-pl.show()
+    p = fig.add_subplot(212)
+    plothist(p)
+    p.get_yaxis().set_visible(False)
+
+    pl.show()
 

### 2016-07-18 paper1/ms.tex
editing ... unstable

@@
 
-\subsection{Rotationally symmetric source with a bivariate gaussian surface brightness distribution}
+\subsection{Rotationally symmetric source with a bivariate gaussian surface brightness distribution}\label{subsec:gaussian}
 
@@
 Having used the simple model of an ideal fold to gain insight, we now
-move to the numerical computation of microlensing light curves with
-realistic mass distributions, and then fit source models to
-lightcurves in the presence of random and systematic errors.
-
-First we generate microlensing magnification maps using the
-microlensing code created by Joachim Wambsgans (\citep{1999A&A...346L...5W, 1990PhDT.......180W, 1999JCoAM.109..353W}, 
-\cite{1990PhDT.......180W}, \cite{1999JCoAM.109..353W}). We then generate
-mock light curves for three source models --- a crescent, a uniform
-disc and a Gaussian disc --- and then fit each mock light curve to all
-three source models. Parameter fitting and marginalising was done by
+consider the question of whether the characteristics of a crescent
+source could be discerned in real data having noise and systematic
+error as well.  Our method is to generate microlensing lightcurves
+\citep[using the code from][with a small extension to provide the
+  option of crescent sources]
+      {1990PhDT.......180W,1999A&A...346L...5W,1999JCoAM.109..353W}
+      and then fitting these mock lightcurves to source models using
       Markov-chain Monte Carlo (MCMC).
 
-\begin{figure}
-\centering
-  \includegraphics[width=0.9\hsize]{figures/cc_forward.eps}
-\caption{\label{fig:cc_forward} $\chi^2=243, P=0.32$.}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.9\hsize]{figures/cc_backward.eps}
-\caption{\label{fig:cc_backward} $\chi^2=284, P=0.01$.}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.9\hsize]{figures/gc_forward.eps}
-\caption{\label{fig:gc_forward} $\chi^2=344, P=3\times10^{-6}$.}
-\end{figure}
-
-\subsection{Numerical microlensing magnification maps}
-
-The code uses a ray-shooting technique to compute the gravitational
-lensing effect of a mass distribution consisting of (a)~a smooth
-component and (b)~a random distribution of point masses representing
-stars.  The ray shooting maps a grid of $\vec\theta$ to $\vec\beta$
-using the lens equation (\ref{eqn:lens}).  That is, the rays are shot
-from the observer back to the source.
-
-For the computation of the individual deflection angles
+\subsection{Numerical microlensing lightcurves}
+
+The microlensing code uses a ray-shooting technique to compute the
+gravitational lensing effect of a mass distribution consisting of
+(a)~a smooth component and (b)~a random distribution of point masses
+representing stars.  The ray shooting maps a grid of $\vec\theta$ to
+$\vec\beta$ using the lens equation (\ref{eqn:lens}).  That is, the
+rays are shot from the observer back to the source.  For the
+computation of the individual deflection angles
 $\vec\alpha(\vec\theta)$ a hierarchical or tree method is used.  The
@@
 moments.  Scaled units are used, with the constant pre-factor in the
-deflection angle (\ref{eqn:alpha}) separated out.
-
-The result of ray shooting is a pixel map on the $\vec\beta$ plane of
-the number of lightrays which arrive at the source plane from a
-particular observer.  This intermediate result is effectively a
-magnification map on the source plane.  Once the map is created, the
-lightcurve can be obtained by specifying the transit path of the
-source across the map.  At each point on this transit line, the code
-computes equation (\ref{eqn:ft2d}) convolving the brightness
-distribution of the source with the magnification pattern of the map.
-In real life, not only are both lens and source moving but the lens
-lens configuration, and with it the magnification pattern, is also
-changing with time.  While the first subtlety is taken care of by a
-coordinate transformation in this analysis, for the second one the
-lens configuration is assumed to be constant in time.
+deflection angle (\ref{eqn:alpha}) separated out.  The result of ray
+shooting is a pixel map on the $\vec\beta$ plane of the number of
+lightrays which arrive at the source plane from a particular observer.
+This intermediate result is effectively a magnification map on the
+source plane.
 
@@
 below, only two point masses were included.  This was done in order to
-have clean fold caustics.  For the computation of the actual
-lightcurve, the code was modified to also allow for crescent shaped
-images specified through the parameter set $R_p,R_n,a,b$. Here $R_p$
-denotes the outer radius of the crescent and $R_n$ the inner one. The
-orientation of the source image with respect to the magnification map
-is specified by the parameters $a$ and $b$ as the shift of the center
-of the inner disk from the center of the outer disk in $x$- and
-$y$-direction respectively. In the original version of the code,
-gaussian and disk-shaped images were already implemented. Those are
-completely characterised by the single parameter $R_p$. The values of
-the parameters are specified in pixel units corresponding to the
-magnification map. Further, one needs to specify the start and end
-point coordinates of the path, which the center of the source image
-follows through the magnification map (see the depiction in figure
-\ref{fig:magnification_map}). Hence, the points along this path are
-specified through the number of time steps for which the computation of
-the brightness is to be carried out. Those points correspond to the
-actual measurement of the brightness of an object in the observational
-case. For each timestep the two-dimensional convolution of equation
-\ref{eqn:ft2d} is carried out numerically for the position of the
-source on the magnification map.
+have clean fold caustics.  There was no smooth component.
+
+Once the map is created, the lightcurve can be obtained by specifying
+the transit path of the source across the map.  The source is moved
+along the transit line, one pixel at a time, and at each step the
+brightness distribution of the source is convolved
+(equation~\ref{eqn:ft2d}) with the magnification map to give the
+brightness.  In real life, not only are both lens and source moving
+but the lens lens configuration, and with it the magnification
+pattern, is also changing with time.  While the first subtlety is
+taken care of by a coordinate transformation in this analysis, for the
+second one the lens configuration is assumed to be constant in time.
 
@@
 caustic perpendicularly, and on a point where the border is a fold
-caustic.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\hsize]{figures/data_cc.eps}
-\includegraphics[width=0.9\hsize]{figures/data_dd.eps}
-\includegraphics[width=0.9\hsize]{figures/data_gg.eps}
-\caption{\label{fig:mockdata} Three different mock datasets with
-  three different parametrizations fitted to each --- crescent (red),
-  uniform-disk (blue) and Gaussian disk (green).}
-\end{figure}
-
-\subsection{Mock light curves}
-
-Mock light curves were generated by running a source across a
-caustic, specifically along the line from point C to point B in
-Figure~\ref{fig:magnification_map}.  The model light curves, to which
-these are fitted, are made by running a source across a different
-caustic, along the line from point A to point B in the same figure.
-That is, the mock data are generated with a clean but not ideal fold
-caustic and then fitted with another such caustic.  This mimics the
-unavoidable systematic error of not knowing the caustic exactly.
-
-The crescent source had parameters (as explained in
-\S\ref{subsec:crescent}) $R_p$ and $R_n$ being the radii of the outer
-and inner circles and $(\alpha,\beta)$ being the coordinates of the
-inner circle with respect to the centre of the outer circle in the
-coordinate system associated with the image in Figure 1 and not
-associated with the caustic surface as defined in the previous
-sections.  The parameter values were
-\begin{equation}
-   (R_p, R_n, \alpha, \beta) = (50.0, 30.0, 15.0, 10.0).
+caustic.  We choose the line AC in Figure~\ref{fig:magnification_map}
+and consider its two caustic-crossing segments separately:
+\begin{itemize}
+\item ``path~I'' runs from C to B;
+\item ``path II'' runs from A to B.
+\end{itemize}
+The magnification along paths I and II are very similar, but not
+identical.
+
+We considered two sources, as follows.
+\begin{itemize}
+\item The crescent source (as explained in \S\ref{subsec:crescent})
+  with parameters
+\begin{equation}
+   (R_p, R_n/Rp, a/Rp, b/Rp) = (1.31, 0.4, 0, 0.3)
 \label{eqn:cp}
 \end{equation}
-The uniform disc had the same (outer) radius as the crescent.  In the
-Gaussian source, we set $3\sigma=50$.  For convenience, below we will
-refer to $R_p$ of the Gaussian source, by which we actually mean
-$3\sigma$.
+corresponding to $r_{1/2}=1$.
+\item A Gaussian source (as explained in \S\ref{subsec:crescent}),
+  also with $r_{1/2}=1$.
+\end{itemize}
+Unit length has 30 pixels on the magnification map and 30 points on a
+light curve.
+
+Each lightcurve has 251 points regularly spaced in time, with Gaussian
+noise at the level of 4\% of the current brightness.
+
+Mock data were generated by taking path-I lightcurves for
 
@@
 
-Figure~\ref{fig:mockdata} shows the three light curves.  Each has 250
-points regularly spaced in time, with Gaussian noise at the level of
-10\% of the current brightness.  This amounts formally to a single
-summary data point with a signal to noise of
-$10\times\sqrt{250}\simeq160$.
-
-\subsection{Model fitting and likelihood analysis}
+This mimics the unavoidable systematic error of not knowing the
+caustic exactly.
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\hsize]{figures/cc_forward.eps}
+\caption{\label{fig:cc_forward} $\chi^2=243, P=0.32$.}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\hsize]{figures/cc_backward.eps}
+\caption{\label{fig:cc_backward} $\chi^2=284, P=0.01$.}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\hsize]{figures/gc_forward.eps}
+\caption{\label{fig:gc_forward} $\chi^2=344, P=3\times10^{-6}$.}
+\end{figure}
+
+\subsection{Model-fitting results}
 

### 2016-07-18 fitting/summaryfig.py

@@ -19,10 +19,11 @@ def mockdata():
 
 
 def gencurves(fname):
-    global rh,wt
+    global rh,rns,wt
     chain = np.genfromtxt(fname+'.chain', skip_header=0, skip_footer=1)
     print(chain.shape)
     rh = []
+    rns = []
     wt = []
     minchis=1e30
     for i in range(chain.shape[0]):
@@ -40,9 +41,14 @@ def gencurves(fname):
                 b *= rp
                 v = Rhalf(rp,rn,a,b)
             rh.append(v/30)
+            rns.append(rn/rp)
             w = chain[i,0] - chain[i-1,0]
-            w /= chain[-1,0]
+            w /= chain[-1,0] - 100
             wt.append(w)
+    rh = np.array(rh)
+    rns = np.array(rns)
+    wt = np.array(wt)
+    print(np.sum(wt))
     print(q)
     print(minchis,lcurv.chis(q),1-st.chi2.cdf(minchis,pars[1]-pars[0]+1-7))
     lcurv.writecurves(q)
@@ -65,14 +71,22 @@ def plotcurves(p):
     p.errorbar(t,b,yerr=db,linestyle='none',color='gray')
     p.plot(t,m,color='black')
     p.set_yticks(np.arange(0,7,2))
-    p.set_xlabel('$t [r_{1/2}/v]$')
+    p.set_xlabel('$t [r_{1/2}/v]$',labelpad=10)
     p.set_ylabel('magnification')
 
 def plothist(p):
-    p.hist(rh,np.linspace(0.75,1.25,26),weights=wt,
+    rmin,rmax,nr = 0,1.2,61
+    p.hist(rh,np.linspace(rmin,rmax,nr),weights=wt,
            histtype='step',lw=2,color='black')
-    p.axis([0.75,1.25,0,0.6])
-    p.set_xlabel('inferred:actual $r_{1/2}$')
+    fl = max(rns)
+    if fl > 0:
+        p.hist(rns,np.linspace(rmin,rmax,nr),weights=wt,
+               histtype='step',lw=2,color='black',linestyle='dashed')
+        p.set_xlabel('inferred $R_n/R_p$ and $r_{1/2}$',labelpad=10)
+    else:
+        p.set_xlabel('inferred $r_{1/2}$',labelpad=10)
+    p.axis([rmin,rmax,0,0.5])
+
 
 
 mockdata()

### 2016-07-19 paper1/ms.tex
editing

@@
 of a black-hole environment
-(Figures~\ref{fig:M87_image}--\ref{fig:M87_plots}).
+(Figure~\ref{fig:M87_plots}).
 
@@
 
-We considered two sources, as follows.
-\begin{itemize}
-\item The crescent source (as explained in \S\ref{subsec:crescent})
-  with parameters
+The mock lightcurve used a crescent source (as explained in
+\S\ref{subsec:crescent}) with parameters
 \begin{equation}
-   (R_p, R_n/Rp, a/Rp, b/Rp) = (1.31, 0.4, 0, 0.3)
+   \left(R_p, \frac{R_n}{R_p}, \frac{a}{R_p}, \frac{b}{R_p}\right) =
+   (1.31, 0.4, 0, 0.3)
 \label{eqn:cp}
 \end{equation}
-corresponding to $r_{1/2}=1$.
-\item A Gaussian source (as explained in \S\ref{subsec:crescent}),
-  also with $r_{1/2}=1$.
-\end{itemize}
-Unit length has 30 pixels on the magnification map and 30 points on a
-light curve.
+corresponding to $r_{1/2}=1$.  Unit length has 30 pixels on the
+magnification map.  The source was carried for 250 pixels along part
+of path~I, giving 251 brightness values.  Gaussian noise at the level
+of 4\% of the current brightness was then added to give the mock
+lightcurve.
 
-Each lightcurve has 251 points regularly spaced in time, with Gaussian
-noise at the level of 4\% of the current brightness.
+\subsection{Model-fitting old results}
 
-Mock data were generated by taking path-I lightcurves for
-
-Each mock light curve also has three nuisance parameters, namely the
-beginning and end of the event and the brightness normalisation.
-These are to be marginalised out by the MCMC.
-
-This mimics the unavoidable systematic error of not knowing the
-caustic exactly.
+Note that the mock lightcurve has three nuisance parameters, namely
+the brightness normalisation and the start- and endpoints on path~I.
+The two latter can also be thought of as the location of the caustic
+with respect to the starting point and the speed of the source.
 
@@
   \includegraphics[width=0.9\hsize]{figures/cc_forward.eps}
-\caption{\label{fig:cc_forward} $\chi^2=243, P=0.32$.}
+\caption{\label{fig:cc_forward} Mock lightcurve fitted with the
+  correct source type and caustic.  The upper panel shows the mock
+  data points, and running through them is a seven-parameter best fit:
+  these being the crescent parameters $(R_p,R_n,a,b)$ and additionally
+  the start and end times and the overall normalisation. The curve has
+  $\chi^2=243, P=0.32$ indicating a good fit.  The lower panel shows
+  the posterior distributions for $r_{1/2}$ (solid histogram) and
+  $R_n/R_p$ (dashed histogram).}
 \end{figure}
@@
   \includegraphics[width=0.9\hsize]{figures/cc_backward.eps}
-\caption{\label{fig:cc_backward} $\chi^2=284, P=0.01$.}
+\caption{\label{fig:cc_backward} Like Figure~\ref{fig:cc_backward},
+  except that the fit uses path~II whereas the mock data still come
+  from path~I. The curve has $\chi^2=284, P=0.01$ indicating a
+  tolerable fit.}
 \end{figure}
@@
   \includegraphics[width=0.9\hsize]{figures/gc_forward.eps}
-\caption{\label{fig:gc_forward} $\chi^2=344, P=3\times10^{-6}$.}
+\caption{\label{fig:gc_forward} Like Figure~\ref{fig:cc_backward}, but
+  showing the best fit to a Gaussian source.  The histogram shows the
+  posterior distribution of $r_{1/2}$ (there is no $R_n$).  The curve
+  has $\chi^2=344, P=3\times10^{-6}$ indicating rejection of the
+  Gaussian source.}
 \end{figure}
 
-\subsection{Model-fitting results}
-
-Each of the three mock light curves was fitted to all three models.
-We denote the nine possible cases with letter pairs, with the first
-letter denoting the assumed model for the fitting procedure and 
-the second letter representing
-the source: thus CG means a {\em C\/}rescent model was fitted to mock
-data from a {\em G\/}aussian source, DC stands for a uniform {\em
-  D\/}isc model fitted to mock data from a {\em C\/}rescent source,
-and so on.
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.9\hsize]{figures/Rp4all.eps}
-\caption{\label{fig:mcmc} Posterior probability distribution for the
-  source size $R_p$.  The vertical line is the correct value. Same
-  color represents the same dataset whereas same line style
-  corresponds to same model fit.  The legend gives the reduced
-  $\chi^2$ of the best fit in each case.}
-\end{figure}
-
-Figure~\ref{fig:mcmc} shows the posterior probability distribution of
-$R_p$ for all nine cases, along with the minimum reduced $\chi^2$ for
-each case.  The area under each curve is unity. Note that the height
-of the curves are not the likelihood, they are probability densities
-in parameter space.
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.9\hsize]{figures/Rhalf_RnRp.eps}
-  \includegraphics[width=0.9\hsize]{figures/aRp_bRp.eps}
-\caption{\label{fig:crescentfit} The 2$\sigma$ contours (or error ellipses)
-  of the crescent model when fitting with the three different
-  datasets.}
-\end{figure}
-
-Let us first consider the three cases where a crescent model was
-fitted.  These are the solid curves in Figure~\ref{fig:mcmc}, with the
-colours of the curves indicating the source.  Meanwhile,
-Figure~\ref{fig:crescentfit} shows $2\sigma$ of the inferred parameter
-values.
-
-\begin{enumerate}
-
-\item[1) {\bf CC}:] solid red curve in Figure~\ref{fig:mcmc} and red
-  ellipses in Figure~\ref{fig:crescentfit}.  In this case, a light
-  curve from a crescent source was being fitted to a crescent model.
-  The fit gives reduced $\chi^2$ close to unity, as expected.  The
-  recovered parameter values are near or slightly outside the
-  $2\sigma$ ellipses.  This is expected since we added a small
-  systematic error, by using different caustics (though both clean
-  folds) for the mock data and the model fit.  The apparent degeneracy
-  between $a$ and $b$, seen in Figure~\ref{fig:crescentfit}, is also
-  expected, since only the distance of the crescent's small circle
-  from the caustic influences the magnification.
-
-\item[2) {\bf CD}:] solid blue curve and blue ellipses.  Here a
-  crescent model was fitted to a light curve from a uniform disc.  The
-  situation is formally a crescent with $R_n=0$ and $a,b$ arbitrary,
-  and this shows in the blue ellipses in the recovered parameter
-  values.  The redundant parameters $\alpha$ and $\beta$, in effect, allow the
-  model to partly fit the noise, and hence, the $\chi^2$ is somewhat
-  lower than in the previous case.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\hsize]{figures/burnin_cg.eps}
-\caption{\label{fig:burnin} The $\chi^2$ of each step of the full MCMC
-  for one case.}
-\end{figure}
-
-\item[3) {\bf CG}:] Solid green curve and green ellipses.
-  Additionally, Figure~\ref{fig:burnin} shows the progress of
-  reduction of $\chi^2$ in this case, where a crescent-source model is
-  used to fit data from a Gaussian source.  The parameter values are
-  seemingly tightly constrained, but the recovered $R_p$ is completely
-  wrong since the correct value is far from the green curve and green
-  ellipses.  The reduced $\chi^2$ is only 0.85, indicative of
-  over-fitting.  We can see what has happened from the red curve in
-  the bottom panel of Figure~\ref{fig:mockdata}.  The best-fit model
-  covers only a small part of the light curve.  That is, the fitting
-  procedure has exploited the nuisance parameters to find a good fit
-  to mainly noise.
-
-\end{enumerate}
-
-Next we consider the three cases where a disc model is fitted.  These
-correspond to the dashed curves in Figure~\ref{fig:mcmc}.  There is
-only one interesting parameter to fit, the disc radius $R_p$, and no
-equivalent of Figure~\ref{fig:crescentfit} is needed.
-
-\begin{enumerate}
-
-\item[4) {\bf DC}:] red dashed curve.  On fitting a disc model to a
-  crescent-source source, the recovered $R_p$ is incorrect, but the
-  reduced $\chi^2$ of 1.29 signals that the data reject the model.
-
-\item[5) {\bf DD}:] blue dashed curve.  When the correct model is
-  fitted to a disc-source, the best fit $\chi^2=0.91$ is good and
-  $R_p$ is recovered within the uncertainty estimate.
-
-\item[6) {\bf DG}:] green dashed curve.  On fitting a disc model to a
-  Gaussian-source model, the recovered $R_p$ is incorrect, but the
-  reduced $\chi^2$ gives no signal that something is amiss.  It
-  appears that a Gaussian source could mimic a uniform disc.
-
-\end{enumerate}
-
-Finally, we consider the three cases where a Gaussian model is fitted.
-These correspond to the dotted curves in Figure~\ref{fig:mcmc}.  Again,
-there is only one interesting parameter to fit, the Gaussian
-$3\sigma$-radius, which we have called $R_p$.
-
-\begin{enumerate}
-
-\item[7) {\bf GC}:] red dotted curve.  When a Gaussian-source model is
-  fitted to data from a crescent source, the recovered $R_p$ is wrong
-  but the reduced $\chi^2=1.29$ shows the data rejecting the model.
-
-\item[8) {\bf GD}:] blue dotted curve.  When a Gaussian-source model
-  is fitted to data from a uniform disc, the recovered $R_p$ is wrong
-  but the model is rejected anyway.
-
-\item[9) {\bf GG}:] green dotted curve.  When data from a Gaussian is
-  fitted to the correct model, $R_p$ is recovered within it's estimated
-  uncertainty and the best fit reduced $\chi^2$ is close to unity.
-
-\end{enumerate}
-
-The above results suggest the following strategy for fitting a
-lightcurve from an unknown source profile: first try a Gaussian-source
-model; if the data reject that model, try a uniform disc; if the
-uniform disc is also rejected by the data, try a crescent model.  Our
-numerical experiments indicate --- assuming one of the three sources
-models are correct --- that the reduced $\chi^2$ would unmask the
-correct model, and its parameters would be correctly recovered.
+This mimics the unavoidable systematic error of not knowing the
+caustic exactly.
 

### 2016-07-19 paper1/ms.tex
revised sec 5 in place

@@
 of 4\% of the current brightness was then added to give the mock
-lightcurve.
+lightcurve.  If we assume the brightness measurements are nightly, and
+the transverse speed is 600~km/s, $r_{1/2}$ comes to 24~light-minutes,
+which is plausible for a quasar.
 
-\subsection{Model-fitting old results}
+\subsection{Model-fitting results}
 
-Note that the mock lightcurve has three nuisance parameters, namely
-the brightness normalisation and the start- and endpoints on path~I.
-The two latter can also be thought of as the location of the caustic
-with respect to the starting point and the speed of the source.
+We now proceed to fit the mock lightcurve using MCMC, to recover the
+source parameters.
 
@@
   these being the crescent parameters $(R_p,R_n,a,b)$ and additionally
-  the start and end times and the overall normalisation. The curve has
-  $\chi^2=243, P=0.32$ indicating a good fit.  The lower panel shows
-  the posterior distributions for $r_{1/2}$ (solid histogram) and
-  $R_n/R_p$ (dashed histogram).}
+  the start and end times and the overall normalisation.  (Ideal
+  lightcurves for the same crescent parameters appear above in
+  Figures~\ref{fig:lightcurve_gauss}, \ref{fig:lightcurve_disk},
+  \ref{fig:lightcurve_crescent} and \ref{fig:char_points}.)  The curve
+  has $\chi^2=243, P=0.32$ indicating a good fit.  The lower panel
+  shows the posterior distributions for $r_{1/2}$ (solid histogram)
+  and $R_n/R_p$ (dashed histogram).}
 \end{figure}
@@
   from path~I. The curve has $\chi^2=284, P=0.01$ indicating a
-  tolerable fit.}
+  marginally acceptable fit.}
 \end{figure}
@@
   showing the best fit to a Gaussian source.  The histogram shows the
-  posterior distribution of $r_{1/2}$ (there is no $R_n$).  The curve
-  has $\chi^2=344, P=3\times10^{-6}$ indicating rejection of the
-  Gaussian source.}
+  posterior distribution of $r_{1/2}$ (there is no $R_n$).  It appears
+  to show $r_{1/2}$ as accurately recovered, but is not to be trusted,
+  because the fit has $\chi^2=344, P=3\times10^{-6}$ indicating
+  rejection of the Gaussian source.  Looking closely, the curve is a
+  little too shallow near $t=3$, and it is unable to fit the second
+  sub-peak near $t=5$.}
 \end{figure}
 
-This mimics the unavoidable systematic error of not knowing the
-caustic exactly.
+Note that fitting involves three nuisance parameters, namely the
+brightness normalisation and the start- and endpoints on the path.
+The latter two are equivalent to considering the location of the
+caustic with respect to the starting point and the speed of the
+source.  The fitting method automatically marginalises over all the
+nuisance parameters.
+
+Figure \ref{fig:cc_forward} shows a fit to the mock lightcurve and
+recovery of $r_{1/2}$ and $R_n/R_p$.
+
+Figure \ref{fig:cc_backward} shows the result when a different caustic
+is used to fit, than was used to generate the mock lightcurve.  This
+mimics the unavoidable systematic error of not knowing the caustic
+exactly.  The fit is slightly degraded, but still acceptable.
+
+Figure \ref{fig:gc_forward} shows the result of attempting to a
+Gaussian source.  There is a small but statistically very significant
+misfit because a Gaussian curve cannot reproduce the part of the
+lightcurve with the slight flattening and then a second sub-peak
+(cf.~Figure~\ref{fig:char_points}) as the dark part of the crescent
+crosses the caustic.
 

### 2016-07-19 fitting/summaryfig.py
figures improvement

@@ -71,8 +71,8 @@ def plotcurves(p):
     p.errorbar(t,b,yerr=db,linestyle='none',color='gray')
     p.plot(t,m,color='black')
     p.set_yticks(np.arange(0,7,2))
-    p.set_xlabel('$t [r_{1/2}/v]$',labelpad=10)
-    p.set_ylabel('magnification')
+    p.set_xlabel('$t [r_{1/2}/v]$',labelpad=10, fontsize=16)
+    p.set_ylabel('magnification', fontsize=16)
 
 def plothist(p):
     rmin,rmax,nr = 0,1.2,61
@@ -84,7 +84,7 @@ def plothist(p):
                histtype='step',lw=2,color='black',linestyle='dashed')
         p.set_xlabel('inferred $R_n/R_p$ and $r_{1/2}$',labelpad=10)
     else:
-        p.set_xlabel('inferred $r_{1/2}$',labelpad=10)
+        p.set_xlabel('inferred $r_{1/2}$',labelpad=10, fontsize=16)
     p.axis([rmin,rmax,0,0.5])
 
 
@@ -94,8 +94,8 @@ mockdata()
 for fname in ['gc_forward','cc_forward','cc_backward']:
     gencurves(fname)
 
-    fig = pl.figure()
-    fig.subplots_adjust(hspace=0.3)
+    fig = pl.figure(figsize=(5,10))
+    fig.subplots_adjust(hspace=0.2)
 
     p = fig.add_subplot(211)
     plotcurves(p)
@@ -104,5 +104,6 @@ for fname in ['gc_forward','cc_forward','cc_backward']:
     plothist(p)
     p.get_yaxis().set_visible(False)
 
-    pl.show()
+    fig.savefig('../paper1/figures/%s.eps'%fname)
 
+    pl.show()

### 2016-07-20 paper1/ms.tex
figures improvement
editing

@@
 
-Active galactic nuclei are thought to be powered by the
-accretion of matter from the proximal environment into a supermassive
-black hole.  The radiation emitted excites the surrounding medium
-which becomes detectable as narrow line regions, broad line regions
-and optical continuum.  Moreover, in the direction perpendicular to
-the accretion disc, where the medium is more transparent, jets will
-appear.  If a jet is oriented towards the Earth, a quasar is observed
-\citep[e.g.,][]{1984RvMP...56..255B}.  While the basic mechanism
-\citep[originating in the work
+Active galactic nuclei are thought to be powered by the accretion of
+matter from the proximal environment into a supermassive black hole.
+The radiation emitted excites the surrounding medium which becomes
+detectable as narrow line regions, broad line regions and optical
+continuum.  In the direction perpendicular to the accretion disc,
+where the medium is more transparent, the outflow tends to be
+collimated to jets.  If the orientation is such that the observer's
+view of the central region is not blocked by the accretion disc or by
+a jet, a quasar is seen \citep[e.g.,][]{1984RvMP...56..255B}.  While
+the basic mechanism \citep[originating in the work
   of][]{1964ApJ...140..796S,1964SPhD....9..246Z,1969Natur.223..690L}
@@
 
-For the black holes in the Galactic centre and the centre of M87
----the two nearest objects that barely qualify as AGN--- the central
-engines (which are $<0.1\rm\,mas$ on the sky) are close to being
-resolved through very long baseline interferometry, which shows
-preliminary indications of the jet-launching structures
+Modestly active galactic nucleii are present in the Galactic centre
+and in M87.  In these two cases, the central engines ($<0.1\rm\,mas$
+on the sky) are close to being resolved through very long baseline
+interferometry, which shows preliminary indications of the
+jet-launching structures
 \citep{2008JPhCS.131a2055D,2012Sci...338..355D,2013MNRAS.434..765K,2016arXiv160205527F}.
-Current data do not deliver images but require fitting to
-predefined models for the images.  A whole range of models have been
-applied, starting from simple geometric models to more complex
-physical models
+Current data do not deliver images but require fitting to predefined
+models for the images.  A whole range of models have been applied,
+starting from simple geometric models to more complex physical models
 \citep{2008Natur.455...78D,2011ApJ...738...38B,2009ApJ...706..497M,2010ApJ...717.1092D}.
@@
 crescent is replaced by a source based on an accretion-disc simulation
-of a black-hole environment
-(Figure~\ref{fig:M87_plots}).
+of a black-hole environment (Figure~\ref{fig:M87_plots}).
 
@@
 
+\subsection{Magnification}
+
 We start by introducing in a succinct manner the gravitational lensing
@@
 \cite{2001stgl.book.....P}.
-
-\subsection{Magnification}
 
@@
 \begin{figure}
-\centering
-  \includegraphics[width=0.9\hsize]{figures/cc_forward.eps}
+% \centering
+  \includegraphics[width=0.5\textwidth]{figures/cc_forward.eps}
 \caption{\label{fig:cc_forward} Mock lightcurve fitted with the
@@
 \centering
-  \includegraphics[width=0.9\hsize]{figures/cc_backward.eps}
+  \includegraphics[width=0.5\textwidth]{figures/cc_backward.eps}
 \caption{\label{fig:cc_backward} Like Figure~\ref{fig:cc_backward},
@@
 \centering
-  \includegraphics[width=0.9\hsize]{figures/gc_forward.eps}
+  \includegraphics[width=0.5\textwidth]{figures/gc_forward.eps}
 \caption{\label{fig:gc_forward} Like Figure~\ref{fig:cc_backward}, but

### 2016-07-20 paper1/ms.tex
editing intro

@@
 jet-launching structures
-\citep{2008JPhCS.131a2055D,2012Sci...338..355D,2013MNRAS.434..765K,2016arXiv160205527F}.
+\citep{2008JPhCS.131a2055D,2012Sci...338..355D,2016ApJ...820...90F}.
 Current data do not deliver images but require fitting to predefined
@@
 \cite{2013MNRAS.434..765K} to use a simple geometric crescent model to
-fit the data, and argued that the crescent is nothing but the
-silhouette of the event horizon.
+fit the available sub-mm interferometric data, They argued that the
+crescent is nothing but the silhouette of the event horizon.
 
@@
 of magnitude smaller on the sky. The direct observations of the black
-hole silhouettes of quasars are far beyond foreseeable instrumentation.
-In the present paper, we consider a possible indirect method, related
-to \cite{1999ApJ...524...49A}, through which gravitational
-microlensing could probe the black hole shadow and its proximal quasar
-environment.
+hole silhouettes of quasars are far beyond foreseeable
+instrumentation.  In the present paper, we consider a possible
+indirect method, related to \cite{1999ApJ...524...49A} and
+\cite{2013ApJ...769..131C}, through which gravitational microlensing
+could probe the black hole shadow and its proximal quasar environment.
 
-Gravitational microlensing of quasars reviewed in
+Gravitational microlensing of quasars, reviewed in
 Section~\ref{sec:microlensing} below, refers to sharp changes in the
 observed brightness of quasars that have been lensed by an intervening
-galaxies, without any changes in the intrinsic luminosity.
-Microlensing affects only the light from the innermost part of the
-quasar, such as the optical continuum
-\citep[e.g.,][]{2012A&A...544A..62S} and is a consequence of two
-things: the very small size of the central engine, and granularity of
-the mass distribution of a lensing galaxy due to stars.  The latter
-means that the local magnification is not a smooth function of source
-position. It contains a complicated network of singular curves, known
-as caustics.  Figure~\ref{fig:magnification_map} shows part of a
-magnification map with a few caustics.  The lensed brightness would be
-given by placing the source on such a magnification map and
-integrating the surface brightness weighted by the magnification.
-Most astrophysical sources straddle several caustics, and hence, their
-net brightness varies smoothly with location.  The central engine of
-quasars, however, is smaller than the typical spacing between
-caustics.  As a result, the lensed brightness undergoes sudden changes
-as a quasar crosses a caustic.  The effect supplies an upper limit on
-the size of the central engine, and can also be used to study the mass
+galaxies, without any changes in the intrinsic luminosity.  It is a
+consequence of two things: the very small size of the central engine,
+and granularity of the mass distribution of a lensing galaxy due to
+stars.  The latter means that the local magnification is not a smooth
+function of source position. It contains a complicated network of
+singular curves, known as caustics.
+Figure~\ref{fig:magnification_map} shows part of a magnification map
+with a few caustics.  The lensed brightness would be given by placing
+the source on such a magnification map and integrating the surface
+brightness weighted by the magnification.  Most astrophysical sources
+straddle several caustics, and hence, their net brightness varies
+smoothly with location.  The central engine of quasars, however, is
+smaller than the typical spacing between caustics.  As a result, the
+lensed brightness undergoes sudden changes as a quasar crosses a
+caustic.  Quasar microlensing is mostly observed in the optical
+continuum \citep[e.g.,][]{2012A&A...544A..62S} but is also known in
+x-rays \citep[e.g.,][]{2016arXiv160207601N} and has been argued for at
+cm wavelengths \citep{2000A&A...358..793K}.  The effect supplies an
+upper limit on the size of the central engine
+\citep[e.g.,][]{2015ApJ...814L..26M} and can be used to study the mass
 distribution and kinematics of stars in the lensing galaxy as well
-\citep[e.g.,][]{2012ApJ...744..111P}.  Caustics have an additional
-remarkable feature: though they can be very complicated, they have
-some universal properties well-known from catastrophe theory.  In
-particular, very close to the simplest caustics (known as folds), the
-magnification is approximately constant on one side and
+\citep[e.g.,][]{2012ApJ...744..111P}.  Microlensing caustics have an
+additional remarkable feature: though they can be very complicated,
+they have some universal properties well-known from catastrophe
+theory.  In particular, very close to the simplest caustics (known as
+folds), the magnification is approximately constant on one side and
 $\propto1/\sqrt p$ where $p$ is the transverse distance of the source
@@
     \includegraphics[width = 0.48\textwidth]{figures/4avar_magnification.eps}
-\caption{\label{fig:lightcurve_disk} Lightcurves of crescent sources with $R_n$ = 0.4 $R_p$.}
+\caption{\label{fig:a_var} Lightcurves of crescent sources with $R_n$
+  = 0.4 $R_p$.}
 \end{figure}
@@
 
-Figure 4 reveals that the lightcurve of a crescent source has more visible features than the other two light-curves
-corresponding to the disc and gaussian shape. There are three regions where $S_{1D}$ and by inheritance, the lightcurve
- has distinguishable behaviour. The first region that would be recorded on a lightcurve plot represents the period
-of time when the bright disk begins to be overlapped by the caustic and stops when just before the dark disk reaches
-the caustic. During this period of time, the flux of light from the source is increasingly magnified. The second
-period starts and ends with the overlapping of the dark disc. As a boundary of the two regions, there is a
-distinguishable point where the slope of the magnification is drastically changed. This apparent discontinuity in the
-first derivative of the magnification function is caused by the caustic amplification of the sudden drop in the $S_{1d}$
-function. During the respective period, the magnification growth slows or even reverses during the first part of the
-period and starts to grow faster again as the dark disk ends its overlap with the caustic. At the point where
-the dark disc clears the caustic, the growth of the magnification is infinite, which appears as a saddle point
-on the lightcurve. Next, the final period corresponds to the case when the dark disc has cleared the caustic
-and the bright disc continues to overlap with the caustic. During this period, the lightcurve reaches a peak
-that for most of the parameter space is global and for the rest of the parameter space local.
-At the end of the period, the growth of the magnification is negative infinite. the respective point appears 
-as a second saddle point on the lightcurve. Past this point the magnification of any finite source will 
-decay in roughly the same manner. \\
+Figure 4 reveals that the lightcurve of a crescent source has more
+visible features than the other two light-curves corresponding to the
+disc and gaussian shape. There are three regions where $S_{1D}$ and by
+inheritance, the lightcurve has distinguishable behaviour. The first
+region that would be recorded on a lightcurve plot represents the
+period of time when the bright disk begins to be overlapped by the
+caustic and stops when just before the dark disk reaches the
+caustic. During this period of time, the flux of light from the source
+is increasingly magnified. The second period starts and ends with the
+overlapping of the dark disc. As a boundary of the two regions, there
+is a distinguishable point where the slope of the magnification is
+drastically changed. This apparent discontinuity in the first
+derivative of the magnification function is caused by the caustic
+amplification of the sudden drop in the $S_{1d}$ function. During the
+respective period, the magnification growth slows or even reverses
+during the first part of the period and starts to grow faster again as
+the dark disk ends its overlap with the caustic. At the point where
+the dark disc clears the caustic, the growth of the magnification is
+infinite, which appears as a saddle point on the lightcurve. Next, the
+final period corresponds to the case when the dark disc has cleared
+the caustic and the bright disc continues to overlap with the
+caustic. During this period, the lightcurve reaches a peak that for
+most of the parameter space is global and for the rest of the
+parameter space local.  At the end of the period, the growth of the
+magnification is negative infinite. the respective point appears as a
+second saddle point on the lightcurve. Past this point the
+magnification of any finite source will decay in roughly the same
+manner.
 
-The impact of parameter $a$ on the shape of the lightcurve can be observed in figure 5 (a-var). Sources where 
-the center of the dark disc reaches the caustic before the center of the bright disc are characterized by a smoother
-broad peak in contrast to the cases where the center of the dark disc reaches the caustic after the bright disk.
-In the case of the latter the instance when the dark disc reaches the caustic corresponds to larger and
-larger magnifications until it becomes a local and even global peak. The effect of the $R_n$  parameter on the
-light curve is presented in figures 6 and 7. For the particular set of parameters where $R_p = R_n +a$ the third
-period of time discussed previously does not exist. Particularizing further, if the value of the radius of the dark
-disc is comparable to the value of the radius of the bright one then the position and shape of the maximum magnification
- are strikingly different. In case the crescent reaches the caustic with the bright region first, the peak magnification
- happens when the dark region reaches the caustic and it is characterized by a sharp variation in magnification growth,
-In the opposite case, the peak appears before the end of the bright disc reaches the caustic and shape is smoother. \\
-
-\subsection{Microlensing a simulated image of M87}
-
-\cite{2012MNRAS.421.1517D} have created a radiative image of M87 based on the GRMHD simulations presented in \citep{2009MNRAS.394L.126M}. 
-The top right-most image in figure 5 of \cite{2012MNRAS.421.1517D} has been projected to a 1D profile associated to the perpendicular 
-direction to a fold caustic approaching the image from the right. The projection is presented in the upper panel of figure 9. 
-The amplification values of the flux of light corresponding to a microlensing event are presented in the lower panel of figure 9. 
-In general, the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions 
-surrounding the luminous parts of the image have non-zero flux and thus are more extended than the simplified source model.
+The impact of parameter $a$ on the shape of the lightcurve can be
+observed in Figure~\ref{fig:a_var}. Sources where the center of the
+dark disc reaches the caustic before the center of the bright disc are
+characterized by a smoother broad peak in contrast to the cases where
+the center of the dark disc reaches the caustic after the bright disk.
+In the case of the latter the instance when the dark disc reaches the
+caustic corresponds to larger and larger magnifications until it
+becomes a local and even global peak. The effect of the $R_n$
+parameter on the light curve is presented in
+Figures~\ref{fig:lightcurve_crescent} and
+\ref{fig:lightcurve_crescent_back}. For the particular set of
+parameters where $R_p = R_n +a$ the third period of time discussed
+previously does not exist. Particularizing further, if the value of
+the radius of the dark disc is comparable to the value of the radius
+of the bright one then the position and shape of the maximum
+magnification are strikingly different. In case the crescent reaches
+the caustic with the bright region first, the peak magnification
+happens when the dark region reaches the caustic and it is
+characterized by a sharp variation in magnification growth, In the
+opposite case, the peak appears before the end of the bright disc
+reaches the caustic and shape is smoother.
 
@@
 
+\subsection{Microlensing a simulated image of M87}
+
+\cite{2012MNRAS.421.1517D} have created a radiative image of M87 based on the GRMHD simulations presented in \citep{2009MNRAS.394L.126M}. 
+The top right-most image in figure 5 of \cite{2012MNRAS.421.1517D} has been projected to a 1D profile associated to the perpendicular 
+direction to a fold caustic approaching the image from the right. The projection is presented in the upper panel of figure 9. 
+The amplification values of the flux of light corresponding to a microlensing event are presented in the lower panel of figure 9. 
+In general, the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions 
+surrounding the luminous parts of the image have non-zero flux and thus are more extended than the simplified source model.
+
 \section{Fitting mock data}\label{sec:numerics}
@@
       Markov-chain Monte Carlo (MCMC).
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.5\textwidth]{figures/cc_forward.eps}
+\caption{\label{fig:cc_forward} Mock lightcurve fitted with the
+  correct source type and caustic.  The upper panel shows the mock
+  data points, and running through them is a seven-parameter best fit:
+  these being the crescent parameters $(R_p,R_n,a,b)$ and additionally
+  the start and end times and the overall normalisation.  (Ideal
+  lightcurves for the same crescent parameters appear above in
+  Figures~\ref{fig:lightcurve_gauss}, \ref{fig:a_var},
+  \ref{fig:lightcurve_crescent} and \ref{fig:char_points}.)  The curve
+  has $\chi^2=243, P=0.32$ indicating a good fit.  The lower panel
+  shows the posterior distributions for $r_{1/2}$ (solid histogram)
+  and $R_n/R_p$ (dashed histogram).}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.5\textwidth]{figures/cc_backward.eps}
+\caption{\label{fig:cc_backward} Like Figure~\ref{fig:cc_backward},
+  except that the fit uses path~II whereas the mock data still come
+  from path~I. The curve has $\chi^2=284, P=0.01$ indicating a
+  marginally acceptable fit.}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.5\textwidth]{figures/gc_forward.eps}
+\caption{\label{fig:gc_forward} Like Figure~\ref{fig:cc_backward}, but
+  showing the best fit to a Gaussian source.  The histogram shows the
+  posterior distribution of $r_{1/2}$ (there is no $R_n$).  It appears
+  to show $r_{1/2}$ as accurately recovered, but is not to be trusted,
+  because the fit has $\chi^2=344, P=3\times10^{-6}$ indicating
+  rejection of the Gaussian source.  Looking closely, the curve is a
+  little too shallow near $t=3$, and it is unable to fit the second
+  sub-peak near $t=5$.}
+\end{figure}
 
@@
 
-\begin{figure}
-% \centering
-  \includegraphics[width=0.5\textwidth]{figures/cc_forward.eps}
-\caption{\label{fig:cc_forward} Mock lightcurve fitted with the
-  correct source type and caustic.  The upper panel shows the mock
-  data points, and running through them is a seven-parameter best fit:
-  these being the crescent parameters $(R_p,R_n,a,b)$ and additionally
-  the start and end times and the overall normalisation.  (Ideal
-  lightcurves for the same crescent parameters appear above in
-  Figures~\ref{fig:lightcurve_gauss}, \ref{fig:lightcurve_disk},
-  \ref{fig:lightcurve_crescent} and \ref{fig:char_points}.)  The curve
-  has $\chi^2=243, P=0.32$ indicating a good fit.  The lower panel
-  shows the posterior distributions for $r_{1/2}$ (solid histogram)
-  and $R_n/R_p$ (dashed histogram).}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.5\textwidth]{figures/cc_backward.eps}
-\caption{\label{fig:cc_backward} Like Figure~\ref{fig:cc_backward},
-  except that the fit uses path~II whereas the mock data still come
-  from path~I. The curve has $\chi^2=284, P=0.01$ indicating a
-  marginally acceptable fit.}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.5\textwidth]{figures/gc_forward.eps}
-\caption{\label{fig:gc_forward} Like Figure~\ref{fig:cc_backward}, but
-  showing the best fit to a Gaussian source.  The histogram shows the
-  posterior distribution of $r_{1/2}$ (there is no $R_n$).  It appears
-  to show $r_{1/2}$ as accurately recovered, but is not to be trusted,
-  because the fit has $\chi^2=344, P=3\times10^{-6}$ indicating
-  rejection of the Gaussian source.  Looking closely, the curve is a
-  little too shallow near $t=3$, and it is unable to fit the second
-  sub-peak near $t=5$.}
-\end{figure}
-
 Note that fitting involves three nuisance parameters, namely the
@@
 
-Nevertheless, the $b$ parameter is relevant for the calculation of $r_{1/2}$. 
-It's qualitative effect is to decrease the value of the half-light radius when the absolute value of the parameter is increased.  
-With two parameters constrained and another irrelevant to the shape of the lightcurve, two free parameters remain $R_n$ and $a$. 
-Figure 4 reveals that the lightcurve of a crescent source has more visible features than the other two light-curves 
-corresponding to the disc and Gaussian shape. The parameters $R_n$ and $a$ have strong influences on the shape of 
-the microlensed lightcurve as can be seen in figures 5, 6 and 7.
-Moreover, the one-dimensional source profile corresponding to the direction perpendicular to the caustic reveal four characteristic points.
-The overlap of each of these points with the caustic leaves visible features on the lightcurve at the corresponding instances of time.
-In timely order, the instances correspond to the start of the overlap between the caustic and the bright disk, the start of the overlap between the caustic and the dark disk, the end of the overlap between the caustic and dark disk and finally the end of the overlap between the caustic and the bright disk.
+Nevertheless, the $b$ parameter is relevant for the calculation of
+$r_{1/2}$.  It's qualitative effect is to decrease the value of the
+half-light radius when the absolute value of the parameter is
+increased.  With two parameters constrained and another irrelevant to
+the shape of the lightcurve, two free parameters remain $R_n$ and $a$.
+Figure 4 reveals that the lightcurve of a crescent source has more
+visible features than the other two light-curves corresponding to the
+disc and Gaussian shape. The parameters $R_n$ and $a$ have strong
+influences on the shape of the microlensed lightcurve as can be seen
+in Figures~\ref{fig:a_var}, \ref{fig:lightcurve_crescent} and
+\ref{fig:lightcurve_crescent_back}.  Moreover, the one-dimensional
+source profile corresponding to the direction perpendicular to the
+caustic reveal four characteristic points. The overlap of each of
+these points with the caustic leaves visible features on the
+lightcurve at the corresponding instances of time. In timely order,
+the instances correspond to the start of the overlap between the
+caustic and the bright disk, the start of the overlap between the
+caustic and the dark disk, the end of the overlap between the caustic
+and dark disk and finally the end of the overlap between the caustic
+and the bright disk.
 

### 2016-07-21 paper1/ms.tex
editing

@@
 crescent is nothing but the silhouette of the event horizon.
-
 The great majority of quasars, however, lie at redshifts beyond 2
@@
 hole silhouettes of quasars are far beyond foreseeable
-instrumentation.  In the present paper, we consider a possible
-indirect method, related to \cite{1999ApJ...524...49A} and
-\cite{2013ApJ...769..131C}, through which gravitational microlensing
-could probe the black hole shadow and its proximal quasar environment.
+instrumentation.
+
+In the present paper, we consider a possible indirect method, related
+to \cite{1999ApJ...524...49A}, through which gravitational
+microlensing could probe the black hole shadow and its proximal quasar
+environment.
 
@@
 have applied the idea to observed light curves to infer properties of
-quasar accretion discs.  In this work, we take a simpler but arguably
-more robust approach: we study features in the light curves
+quasar accretion discs.  A similar idea appears in
+\cite{2013ApJ...769..131C}.  In this work, we take a simpler but
+arguably more robust approach: we study features in the light curves
 characteristic of a crescent-like source which in turn would indicate

### 2016-09-22 paper1/ms.tex
Corrections
other corrections

@@
 
-\section{introduction}
+\section{The General Picture }
 
-Active galactic nuclei are thought to be powered by the accretion of
+Active galactic nuclei (AGN) are thought to be powered by the accretion of
 matter from the proximal environment into a supermassive black hole.
@@
 In the present paper, we consider a possible indirect method, related
-to \cite{1999ApJ...524...49A}, through which gravitational
+to \cite{1999ApJ...524...49A} and \cite{2015ApJ...814L..26M}, 
+through which gravitational
 microlensing could probe the black hole shadow and its proximal quasar
@@
 Here the magnification of a point source near a caustic is equal to
-the sum of the magnification due to other reasons $\mu_0$, assumed to
+the sum of the magnification due to other reasons $\mu_0$ 
+(such as zero point magnification and the effect of other weaker and 
+distant lenses), assumed to
 be locally constant, and a decrease with the square root of the
@@
 along the $p$ axis. Such a modelling of the motion of the object in the source plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
-choosing the function $a(t)$ to be constant in time.
+choosing the function $a$ to be constant in time.
 
@@
 
-There are four characteristic points visible on the resulting one-dimensional light profile of the crescent shaped source.
+In figure \ref{fig:char_points} There are four characteristic points visible on the resulting one-dimensional light profile of the crescent shaped source.
 Two of the points, p1 and p4, mark the outer boundaries of the luminous disc component.
@@
 \begin{equation}
-    R_p = \frac{v_p \left( t_4 -t_1 \right)}{2}, 
+    R_p = \frac{v_p \times  \left( t_4 -t_1 \right)}{2}, 
 \end{equation}
@@
 \begin{equation}
-        R_n = \frac{v_p \left( t_3 -t_2 \right)}{2}, 
+        R_n = \frac{v_p \times \left( t_3 -t_2 \right)}{2}, 
 \end{equation}
@@
 \begin{equation}
-        a = \frac{v_p \left( t_4 +t_1 - t_3 - t_2 \right)}{2}. 
+        a = \frac{v_p \times \left( t_4 +t_1 - t_3 - t_2 \right)}{2}. 
 \end{equation}
@@
 
-Figure 4 reveals that the lightcurve of a crescent source has more
+Figure \ref{fig:lightcurve_gauss} reveals that the 
+lightcurve of a crescent source has more
 visible features than the other two light-curves corresponding to the
@@
   lightcurve associated to the 1D profile is presented in the lower
-  panel. The points p2 and p3 mark the start and end of the equivalent
+  panel. The points $\rm p_2$ and $\rm p_3$ mark the start and 
+  end of the equivalent
   dark disk and are associated with the moments of their overlap with
-  the caustic: t2 and t3.}
+  the caustic: $\rm t_2$ and $\rm t_3$. $\rm h_{w}$ denotes the maximum width of the crescent.}
 \end{figure}

### 2016-09-28 paper1/ms.tex
small edits

@@
 suited for acquiring the necessary microlensed light-curve.
+
 \section{Microlensing}\label{sec:microlensing}
-
-\subsection{Magnification}
 
@@
 \cite{2001stgl.book.....P}.
+
+\subsection{Magnification}
 
@@
 error as well.  Our method is to generate microlensing lightcurves
-\citep[using the code from][with a small extension to provide the
-  option of crescent sources]
-      {1990PhDT.......180W,1999A&A...346L...5W,1999JCoAM.109..353W}
-      and then fitting these mock lightcurves to source models using
-      Markov-chain Monte Carlo (MCMC).
+using a widely-used microlensing code
+\citep{1990PhDT.......180W,1999A&A...346L...5W,1999JCoAM.109..353W} to
+which we have added a small extension to provide the option of
+crescent sources, and then fit these mock lightcurves to source models
+using Markov-chain Monte Carlo (MCMC).
 

### 2016-10-03 paper1/ms.tex
changes

@@
 \includegraphics[width = .49\textwidth]{figures/figure_rhalf.eps}
-\caption{\label{fig:geom_crescent} Geometry of crescent sources.}
+\caption{\label{fig:geom_crescent} Geometry of crescent sources: the source can be seen as a superposition of a smaller dark disk over and inside a larger brighter disk.
+$R_p$ and $R_n$ mark the radii of the bright and dark disk.
+$d_c$ and $c$ mark the distance from the center of the bright disk to the centroid and from the dark disk to the centroid.
+}
 \end{figure}
@@
     \includegraphics[width = 0.48\textwidth]{figures/4source_magnification.eps}
-\caption{\label{fig:lightcurve_gauss} Lightcurves of crescents, Gaussian and disk for sources with identical $S_0$ and $r_{1/2}$.
-Crescent source has $r_{1/2}$ = 0.72754 $R_p$, $R_n$ = 0.4 $R_p$, a = 0.3/-0.3 $R_p$ }
+\caption{\label{fig:lightcurve_gauss} 1D profiles (top panel) and simulated lightcurves (bottom) of crescent, Gaussian and disk sources with identical total flux ($S_0$) and half-light radius($r_{1/2}$).
+Crescent source has $r_{1/2}$ = 0.72754 $R_p$, $R_n$ = 0.4 $R_p$, a = 0.3/-0.3 $R_p$.
+}
 \end{figure}
@@
     \includegraphics[width = 0.48\textwidth]{figures/4avar_magnification.eps}
-\caption{\label{fig:a_var} Lightcurves of crescent sources with $R_n$
-  = 0.4 $R_p$.}
+\caption{\label{fig:a_var} 1D profiles (top panel) and simulated lightcurves (bottom) of crescent sources with identical $R_p$ and $R_n$ ($R_n$ = 0.4 $R_p$) and different disk centers displacement (a).}
 \end{figure}
@@
     \includegraphics[width = 0.48\textwidth]{figures/5Rn_back_var_magnification.eps}
-\caption{\label{fig:lightcurve_crescent_back} Lightcurves of backwards crescents with  a = 0.3 $R_p$.}
+\caption{\label{fig:lightcurve_crescent_back} 1D profiles (top panel) and simulated lightcurves (bottom) of backwards crescent sources with identical $R_p$ and $a$ ($a$ = 0.3 $R_p$) and different dark disk radii ($R_n$).}
 \end{figure}
@@
     \includegraphics[width = 0.48\textwidth]{figures/5Rn_forw_var_magnification.eps}
-\caption{\label{fig:lightcurve_crescent} Lightcurve of forwards crescents with a = -0.3 $R_p$.}
+\caption{\label{fig:lightcurve_crescent} 1D profiles (top panel) and simulated lightcurves (bottom) of forwards crescent sources with identical $R_p$ and $a$ ($a$ = - 0.3 $R_p$) and different dark disk radii ($R_n$).
+}
 \end{figure}

### 2016-10-03 paper1/ms.tex
changes

@@
         \includegraphics[width = 0.48\textwidth]{figures/ch_instances.eps}
-\caption{\label{fig:char_points} The characteristic points and instances of a crescent source's one dimensional profile and microlensing lightcurve. }
+\caption{\label{fig:char_points} The characteristic points and instances 
+of a crescent source's one dimensional profile and microlensing lightcurve.
+In the top panel a typical 1D profile  of a crescent source is presented. 
+There are two light gray regions that mark the regions of the bright disk 
+that are not overlapped by the dark disk. The dark gray band marks the 
+region where the two disks overlap. $p_1$ and $p_4$ mark the start and 
+end of the bright disk in the 
+1D representation. Analogous, $p_2$ and $p_3$ mark the start and 
+end of the dark disk.
+The bottom panel represents the associated lightcurve with the points 
+$t_i$ marking the moment when point $p_i$ overlaps with the caustic.
+The gray shades are analogous to the ones in the top panel.
+ }
 \end{figure}
@@
 \caption{\label{fig:M87_plots} The one dimensional profile obtained
-  from the numerical integration along the ordonata axis of the source
-  image is presented in the upper panel. The corresponding
+  from the numerical integration along the ordonata axis of a simulated source
+  image is presented in the upper panel. The source image was presented in 
+ \cite{2012MNRAS.421.1517D} in the top right-most panel of figure 5 
+(of the respective paper) as a result of the simulations 
+presented in \cite{2009MNRAS.394L.126M}. The corresponding
   lightcurve associated to the 1D profile is presented in the lower

### 2016-10-03 paper1/ms.tex
changes

@@
   image is presented in the upper panel. The source image was presented in 
- \cite{2012MNRAS.421.1517D} in the top right-most panel of figure 5 
+ Dexter et al 2012 in the top right-most panel of figure 5 
 (of the respective paper) as a result of the simulations 
-presented in \cite{2009MNRAS.394L.126M}. The corresponding
+presented in McKinney \& Blandford 2009. The corresponding
   lightcurve associated to the 1D profile is presented in the lower

### 2016-10-03 paper1/ms.tex
ms2.tex:

@@
  radius of the studied source. 
-In particular cases in which the previously mentioned approximation loses its validity, the impact on the quality of the 
-lightcurves is not evenly distributed. The shape of the lightcurve will be maintained. 
-The data points corresponding to the source position before and during the overlapping of the caustic will be affected by
- smaller errors than the data points corresponding to later times. \\
+%In particular cases in which the previously mentioned approximation loses its validity, the impact on the quality of the 
+%lightcurves is not evenly distributed. The shape of the lightcurve will be maintained. 
+%The data points corresponding to the source position before and during the overlapping of the caustic will be affected by
+% smaller errors than the data points corresponding to later times. \\
 

### 2016-10-03 paper1/ms2.tex
ms2.tex:

--- 2016-10-03 paper1/ms.tex
@@
-\documentclass[usenatbib]{mn2e}
-
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{epstopdf}
-\usepackage{multicol}
-
-\title{Microlensing as a possible probe of event-horizon structure in quasars}
-\author[Tomozeiu et al]{Mihai Tomozeiu\thanks{mihai@physik.uzh.ch},$^{1,2}$ 
-Irshad Mohammed,$^{1,2,3}$ 
-Manuel Rabold,$^{1,2}$
-Prasenjit Saha,$^{1,2}$  
-\newauthor
-and Joachim Wambsganss$^{1,4}$\\
-$^1${Physik-Institut, University of Zurich, Winterthurerstrasse 190,
-  8057 Zurich, Switzerland} \\
-$^2${Institute for Computational Science, University of Zurich,
-  Winterthurerstrasse 190, 8057 Zurich, Switzerland} \\
-$^3${Theoretical Astrophysics Group, Fermi National Accelerator Laboratory, Batavia, IL 60510, USA}\\
-$^4${Zentrum f\"ur Astronomie der Universit\"at Heidelberg,
-  M\"onchhofstrasse 12--14, 69120 Heidelberg, Germany}
-}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-
-In quasars which are lensed by galaxies, the point-like images sometimes show sharp and uncorrelated brightness 
-variations (microlensing). These brightness changes are associated with the innermost region of the quasar passing 
-through a complicated pattern of caustics produced by the stars in the lensing galaxy. In this paper, we study 
-whether the universal properties of optical caustics could enable extraction of shape information about the central 
-engine of quasars. We present a toy model with a crescent-shaped source crossing a fold caustic. The silhouette 
-of a black hole over an accretion disk tends to produce roughly crescent sources. When a crescent-shaped source 
-crosses a fold caustic, the resulting light curve is noticeably different from the case of a circular luminosity 
-profile or Gaussian source. With good enough monitoring data, the crescent parameters, apart from one degeneracy, 
-can be recovered.
-
-\end{abstract}
-
-\begin{keywords}
-Supermassive black holes, microlensing, quasars.
-\end{keywords}
-
-\section{The General Picture }
-
-Active galactic nuclei (AGN) are thought to be powered by the accretion of
-matter from the proximal environment into a supermassive black hole.
-The radiation emitted excites the surrounding medium which becomes
-detectable as narrow line regions, broad line regions and optical
-continuum.  In the direction perpendicular to the accretion disc,
-where the medium is more transparent, the outflow tends to be
-collimated to jets.  If the orientation is such that the observer's
-view of the central region is not blocked by the accretion disc or by
-a jet, a quasar is seen \citep[e.g.,][]{1984RvMP...56..255B}.  While
-the basic mechanism \citep[originating in the work
-  of][]{1964ApJ...140..796S,1964SPhD....9..246Z,1969Natur.223..690L}
-is not in doubt, the central engines, near the event horizons of the
-black holes, remain to be probed.
-
-Modestly active galactic nucleii are present in the Galactic centre
-and in M87.  In these two cases, the central engines ($<0.1\rm\,mas$
-on the sky) are close to being resolved through very long baseline
-interferometry, which shows preliminary indications of the
-jet-launching structures
-\citep{2008JPhCS.131a2055D,2012Sci...338..355D,2016ApJ...820...90F}.
-Current data do not deliver images but require fitting to predefined
-models for the images.  A whole range of models have been applied,
-starting from simple geometric models to more complex physical models
-\citep{2008Natur.455...78D,2011ApJ...738...38B,2009ApJ...706..497M,2010ApJ...717.1092D}.
-The more complicated models nonetheless tend to predict a crescent
-shaped silhouette of the black hole.  This motivated
-\cite{2013MNRAS.434..765K} to use a simple geometric crescent model to
-fit the available sub-mm interferometric data, They argued that the
-crescent is nothing but the silhouette of the event horizon.
-The great majority of quasars, however, lie at redshifts beyond 2
-\citep{2014A&A...563A..54P} and their central engines would be orders
-of magnitude smaller on the sky. The direct observations of the black
-hole silhouettes of quasars are far beyond foreseeable
-instrumentation.
-
-In the present paper, we consider a possible indirect method, related
-to \cite{1999ApJ...524...49A} and \cite{2015ApJ...814L..26M}, 
-through which gravitational
-microlensing could probe the black hole shadow and its proximal quasar
-environment.
-
-Gravitational microlensing of quasars, reviewed in
-Section~\ref{sec:microlensing} below, refers to sharp changes in the
-observed brightness of quasars that have been lensed by an intervening
-galaxies, without any changes in the intrinsic luminosity.  It is a
-consequence of two things: the very small size of the central engine,
-and granularity of the mass distribution of a lensing galaxy due to
-stars.  The latter means that the local magnification is not a smooth
-function of source position. It contains a complicated network of
-singular curves, known as caustics.
-Figure~\ref{fig:magnification_map} shows part of a magnification map
-with a few caustics.  The lensed brightness would be given by placing
-the source on such a magnification map and integrating the surface
-brightness weighted by the magnification.  Most astrophysical sources
-straddle several caustics, and hence, their net brightness varies
-smoothly with location.  The central engine of quasars, however, is
-smaller than the typical spacing between caustics.  As a result, the
-lensed brightness undergoes sudden changes as a quasar crosses a
-caustic.  Quasar microlensing is mostly observed in the optical
-continuum \citep[e.g.,][]{2012A&A...544A..62S} but is also known in
-x-rays \citep[e.g.,][]{2016arXiv160207601N} and has been argued for at
-cm wavelengths \citep{2000A&A...358..793K}.  The effect supplies an
-upper limit on the size of the central engine
-\citep[e.g.,][]{2015ApJ...814L..26M} and can be used to study the mass
-distribution and kinematics of stars in the lensing galaxy as well
-\citep[e.g.,][]{2012ApJ...744..111P}.  Microlensing caustics have an
-additional remarkable feature: though they can be very complicated,
-they have some universal properties well-known from catastrophe
-theory.  In particular, very close to the simplest caustics (known as
-folds), the magnification is approximately constant on one side and
-$\propto1/\sqrt p$ where $p$ is the transverse distance of the source
-from the caustic.  This property will be exploited later.
-
-In Section~\ref{sec:source-models} we introduce the three source
-profiles used in our subsequent models and simulations: a
-constant-brightness disc, a circular Gaussian, and the crescent source
-introduced by \citep{2013MNRAS.434..765K}.  The latter is simply a
-constant-brightness disc with a smaller, non-concentric disc cut out
-of it.  We also derive the half-light radius for a crescent.  The
-half-light radius can characterise the source size for all three types
-of source.
-
-Section~\ref{sec:fold-crossing} shows the light curves that result
-when each of the model sources crosses an ideal fold.  This would
-apply in Figure~\ref{fig:magnification_map} to sources along the path
-AB or BC, for sources small enough that the curvature of the caustics
-is negligible.  With this assumption one can imagine the caustic as an
-infinite wall to be crossed by the source as presented in
-Figure~\ref{fig:infinite_fold}.  The source brightness distribution
-parallel to the caustic naturally makes no difference to the
-observable brightness; each source can be replaced by an effective
-one-dimensional source profile, by flattening the source so it becomes
-perpendicular to the caustic.  In principle, the effective
-one-dimensional brightness profile could be recovered from the light
-curve by deconvolution.  \cite{1999ApJ...524...49A} modelled this
-profile as the result of a circular accretion disc seen through the
-spacetime around a Kerr black hole, and \cite{2012MNRAS.423..676A}
-have applied the idea to observed light curves to infer properties of
-quasar accretion discs.  A similar idea appears in
-\cite{2013ApJ...769..131C}.  In this work, we take a simpler but
-arguably more robust approach: we study features in the light curves
-characteristic of a crescent-like source which in turn would indicate
-a black-hole silhouette.  Figure~\ref{fig:char_points} shows the
-qualitative features: there is a period during which the dark cutout
-disc is crossing the caustic, and before and after there are periods
-when the only the bright parts of the crescent are in transit across
-the caustic.  The details depend on the orientation of the crescent,
-but basically the dark disc causes a rising light curve to plateau or
-dip.  These features are still present, albeit faintly, if the simple
-crescent is replaced by a source based on an accretion-disc simulation
-of a black-hole environment (Figure~\ref{fig:M87_plots}).
-
-In Section~\ref{sec:numerics} we carry out source fitting to
-lightcurves, with both noise and systematic errors are present.  We
-generate three lightcurves by taking a uniform disc, a circular
-Gaussian, and a crescent across the path AB in
-Figure~\ref{fig:magnification_map}, and then adding noise.  The path
-simulates crossing a clean but not ideal fold.  In addition, we
-generate templates lightcurves by running the three source types, with
-various parameter values, across the path CB.  That is, the templates
-come from a similar but not identical caustic, thus deliberately
-generating a systematic error.  We then fit the noisy lightcurves to
-the templates using Markov chain Monte-Carlo.  We find that the
-correct source type can be inferred from the $\chi^2$ values.  The
-parameter values can also be inferred. The fitting errors are larger
-than the formal uncertainties, which is expected in the presence of
-systematic errors, but still appear acceptable.
-
-Finally in Section~\ref{sec:discussion} we discuss in more depth the
-implications of the results presented in the previously mentioned
-sections.  One of the most interesting implication is the possibility
-to estimate the black hole mass from the reconstructed
-parameters. This further requires good approximations of the relative
-transversal velocity between the quasar and the lens. Proper
-constraints can be set with independent observations of the stellar
-structure that contain the gravitational lens.
-
-The thorough study of the possibility to reconstruct the quasar's
-structural parameters from light-curves containing multiple
-microlensing events represents the target of future work. This will
-most probably require the use of powerful statistical tools. Another
-path for future work is the designing of an observation regime best
-suited for acquiring the necessary microlensed light-curve.
-
-\section{Microlensing}\label{sec:microlensing}
-
-We start by introducing in a succinct manner the gravitational lensing
-theory that is relevant for microlensing in general and for the scope
-of the present paper in particular.  More detailed presentation of the
-theory can be found in several references, such as
-\cite{2001stgl.book.....P}.
-
-\subsection{Magnification}
-
-The gravitational lens equation
-\begin{equation}
-\vec\beta = \vec\theta - \vec\alpha(\vec\theta)
-\label{eqn:lens}
-\end{equation}
-relates the apparent sky position $\vec\theta$ of a light source to
-its true but unobservable sky position $\vec\beta$ through the bending
-angle $\vec\alpha$.  The latter is an integral 
-\begin{equation}
-\vec\alpha = (1+z_{_L})\frac{D_S-D_L}{D_SD_L} \frac{4G}{c^2}
-\int \Sigma(\vec\theta')
-\frac{\vec\theta-\vec\theta'}{|\vec\theta-\vec\theta'|^2}\,
-d^2\vec\theta'
-\label{eqn:alpha}
-\end{equation}
-depending on the projected density $\Sigma(\vec\theta)$ (kg/steradian)
-of the lens, the lens redshift $z_{_L}$ and the comoving distances
-$D_L$ and $D_S$ to the lens and source.  The derivative of the
-apparent position with respect to the source position
-\begin{equation}
-M(\vec\theta) =
-\left(\frac{\partial\beta}{\partial\theta}\right)^{-1}
-\label{eqn:magnif-matrix}
-\end{equation}
-is known as the magnification matrix, and its determinant
-\begin{equation}
-\mu(\vec\theta) = \det|M(\vec\theta)|
-\end{equation}
-is the brightness amplification of an image of a point source.  In
-other words, the source will brighten or dim according to whether
-$\mu(\vec\theta)$ is more or less than unity.  If there are multiple
-mages at distinct but not observationally resolved $\vec\theta_i$ from
-the same $\vec\beta$, a total brightness amplification
-of\begin{equation} \mu_{\rm total} = \sum_{i} \mu(\vec\theta_i)
-\end{equation}
-applies. It is possible for the magnification to become formally
-infinite, as a result of an eigenvalues of the magnification matrix
-(\ref{eqn:magnif-matrix}) becoming infinite.  This generically happens
-on curves on the $\vec\theta$ plane, known as {\em critical curves}.
-Mapping a critical curve to the source plane, through the lens
-equation (\ref{eqn:lens}), to the source plane $\vec\beta$ gives the
-so-called caustics.  Caustics can appear in the optical system, not just
-gravitational lensing.  For a point source, caustics are singularities
-of the magnification; for finite-size sources caustics correspond to
-high and sharply changing magnification.
-
-Caustics are important in all forms of lensing with multiple images,
-but they have a special significance for lens quasars, first pointed
-out by \cite{1979Natur.282..561C}.  The granularity of the mass
-distribution $\Sigma(\vec\theta)$ due to individual stars produces a
-caustic network on the scale of $\sqrt{GM_\odot D_L/c^2}$ or $10^{-6}
-\sqrt{M/M_\odot}\rm\,arcsec$ for typical lens and source redshift
-\citep{2001PASA...18..207W}.  Extended sources wash out this
-micro-caustic structure, but the optical-continuum source of quasars
-is even smaller.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\hsize]{figures/IRIS567_path.eps}
-\caption{\label{fig:magnification_map} Magnification map used later in
-  this paper (Section~\ref{sec:numerics}). The white line marks the
-  trajectory of the center of the sources for which the lightcurves
-  presented in figure 11 are generated.  Note that this represents an
-  atypically simple region of any realistic magnification map.}
-\end{figure}
-
-\subsection{Magnification near a fold caustic}
-
-A simple example of a caustic network is shown in
-Figure~\ref{fig:magnification_map}.  There are two general categories
-of caustics in gravitational lensing, cusps and fold, and examples of
-both kinds can be seen in this figure.  Magnification near a caustic
-has universal properties, independent of the system and has been
-extensively studied
-\citep{1986ApJ...310..568B,1992A&A...260....1S,2002ApJ...574..970G,2002ApJ...580..468G}.
-In particular, at distance $p$ from a fold caustic
-\begin {equation}
- \mu(p) = \mu_0 + C_0 \frac{1}{\sqrt{p}} \Theta(p).
-\end {equation}
-Here the magnification of a point source near a caustic is equal to
-the sum of the magnification due to other reasons $\mu_0$ 
-(such as zero point magnification and the effect of other weaker and 
-distant lenses), assumed to
-be locally constant, and a decrease with the square root of the
-distance from the fold. The latter term becomes activated only after
-the source enters the region interior to the caustic curves when the
-values of the step function $\Theta(p)$ become unity. The
-proportionality constant $C_0$ depends on the local conditions in the
-vicinity of the caustic.
-
-A source of arbitrary shape can be described by a two-dimensional brightness function $S_{2D}(p - p_s, q - q_s)$ defined for a coordinate system $p,q$ where $p_s, q_s$ denote the coordinates of the center of a source.
-
-\begin{figure}
-\includegraphics[width = .49\textwidth]{figures/infinite_fold_ar.eps}
-\caption{\label{fig:infinite_fold} Source profiles for Gaussian, disk,
-  crescents (right) and magnification map for an infinite fold
-  (left). Objects have the same $S_{0}$ and $r_{1/2}$. The black arrow marks the direction of motion of the
-sources relative to the caustic.}
-\end{figure}
-
-For a microlensing event the lightcurve can be written for an
-undefined source shape as:
-\begin{equation}
- F(t) = \int_{-\infty}^\infty \int_{-\infty}^\infty S_{2D}(p-p_s(t), q-q_s(t)) \mu_t(p) \mathrm{d}q \mathrm{d}p
- \label{eqn:ft2d}
-\end{equation}
-In order to build the previous equation we have considered that the
-time dependency of the flux $F$ is given only by the motion of the
-source with respect to a fixed caustic. Therefore the only time
-dependent quantities in the right hand side of the equation are the
-coordinates of the center of the source $p_s,q_s$ and by construct
-$S_{2D}$.
-
-Due to the choice of the coordinate system the amplification factor has no dependence on the $q$ coordinate. The previous equation can be rewritten as:
-
-\begin{equation}
- F(t) 
-= \int_{-\infty}^\infty  \mu_t(p) S_{1D}\left(p-p_s(t)\right) \mathrm{d}p,
-\label{eqn:ft}
-\end{equation}
-\\
-where we have defined the one dimensional flux function as:
-\begin{equation}
- S_{1D}(p-p_s(t)) = \int_{-\infty}^\infty S_{2D}(p-p_s(t), q-q_s(t)) \mathrm{d}q
-\end{equation}
-
-This representation is a valid approximation only when the apparent size of the source is much smaller than the corresponding Einstein angle of the lens. In this context 
-all the information about the source shape and brightness that can be contained in the lightcurve is exhaustively given by the 1D flux function.
-In other words, if two sources with different $S_{2D}$ have the same $S_{1D}$ they cannot be distinguished by studying their lightcurves.
-
-\section{Models for extended sources}\label{sec:source-models}
-
-In the present study we are analysing three types of sources with different surface brightness: 
-\begin{enumerate}
- \renewcommand{\theenumi}{(\arabic{enumi})}
-  \item a rotationally symmetric source with a bivariate gaussian surface brightness distribution,
-  \item a disk source with constant surface brightness distribution,
-  \item a crescent shaped source with constant surface brightness distribution.
-\end{enumerate}
-The first two sources are the typical choices used in the literature to describe the luminous parts of a quasar (Prasenjit should give some citations here). The third one is a recently proposed
-variant \citep{2013MNRAS.434..765K}.
-
-\subsection{Rotationally symmetric source with a bivariate gaussian surface brightness distribution}\label{subsec:gaussian}
-
-A symmetric 2D gaussian can be described mathematically as:
-
-\begin{equation}
- S_{2D}^G(p-p_s, q-q_s) = \frac{S_0^G}{2 \pi \sigma^2} e^{-\frac{(p-p_s)^2}{2 \sigma^2}} e^{-\frac{(q-q_s)^2}{2 \sigma^2}}.
-\end{equation}
-\\
-The corresponding 1D brightness is:
-
-\begin{equation}
- S_{1D}^G(p-p_s) = \frac{S_0^G}{\sqrt{2 \pi} \sigma} e^{-\frac{(p-p_s)^2}{2 \sigma^2}}.
-\end{equation}
-\\
-Other parameters of the model are the total flux $S_0^G$ and $\sigma$.
-
-Although such a definition for a source would have non-zero surface/linear brightness for any coordinate $p,q$, the amount of light received by a detector from outside a $3 \sigma$ disk centered at $p_s, q_s$
-would be insignificant. For a gaussian distributed surface brightness source the half-light radius is directly proportional to the parameter $\sigma$ according to the equation:
-\begin{equation}
-r_{1/2} = \sqrt{ln(4)} \sigma.
-\end{equation}
-
-\subsection{Disk source with constant surface brightness distribution}
-
-One can construct mathematically a disk source with constant surface brightness and radius $R$ using a stepfunction:
-\begin{equation}
- S_{2D}^D(p-p_s, q-q_s) = \frac{S_0^D}{\pi R^2} \Theta \left( R^2 - \left( p-p_s \right)^2 - \left( q-q_s \right)^2 \right).
-\end{equation}
-By integrating over $q$ coordinate the linear brightness function is obtained:
-
-\begin{equation}
- S_{1D}^D(p-p_s) = \frac{2 S_0^D}{\pi R}  \sqrt{1 - \frac{(p-p_s)^2}{R^2} }    \Theta \left( R^2 - \left( p-p_s \right)^2 \right).
-\end{equation}
-\\
-The half-light radius of a uniform disk source is $R/\sqrt{2}$.
-
-\subsection{Crescent source with constant surface brightness distribution}\label{subsec:crescent}
-
-The surface brightness distribution of a geometric crescent can be built by considering two disk sources of constant brightness.
-One larger disk will contribute positively to the total flux while one smaller disk
-that is interior to the large one will contribute negatively. This superposition can be written for 2D as:\\
-
-\begin{equation}
- S_{2D}^C =  S_{2D}^{Dp} -  S_{2D}^{Dn}  
- \label{eqn:s2d}
-\end{equation}
-with\\
-
-\begin{equation}
- S_{2D}^{Dp}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dp}}{\pi R_p^2} \Theta \left( R_p^2 - \left( p-p_{sp} \right)^2 - \left( q-q_{sp} \right)^2 \right)
-\end{equation}
-\\
-and
-\begin{equation}
- S_{2D}^{Dn}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dn}}{\pi R_n^2} \Theta \left( R_n^2 - \left( p-p_{sn} \right)^2 - \left( q-q_{sn} \right)^2 \right).
-\end{equation}
-\\
-The following notations were used: $R_p, (p_{sp}, q_{sp}), R_n, (p_{sn},q_{sn})$ are the radii and coordinate of the center for the larger positive disk and smaller negative disk respectively.
-$S_0^{Dp},S_0^{Dn}$ represent the total flux of radiation received from the large and small disk.
-From this point forward we will not use the total flux from each source.
-Instead we will
-use the difference which in this case is th total flux from the crescent-shaped source $S_0^C$. \\
-Equation \ref{eqn:s2d} can be written as:\\
-
-\begin{align}
- S_{2D}^C &= \frac{S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \left\{ \Theta \left[ R_p^2 - \left( p-p_{sp} \right)^2 - \left( q-q_{sp} \right)^2 \right] \right.\nonumber\\
- &\qquad \left. {} -  \Theta \left[ R_n^2 - \left( p-p_{sn} \right)^2 - \left( q-q_{sn} \right)^2 \right] \right\}.
-\end{align}
-\\
-Analogous for the linear brightness function:
-
-\begin{align}
- S_{1D}^C &= \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \left\{ \sqrt{R_p^2 - (p-p_{sp})^2}  \Theta \left[ R_p^2 - \left( p-p_{sp} \right)^2 \right] \right.\nonumber\\
- &\qquad \left. {} - \sqrt{R_n^2 - (p-p_{sn})^2 } \Theta \left[ R_n^2 - \left( p-p_{sn} \right)^2 \right] \right\}.
-\label{eqn:s1_d}
-\end{align}
-
-There are some constraints on the parameters used to define a crescent in the previously presented manner that need to be stated. First, we must impose the obvious $R_p > R_n$ relation. Secondly, 
-the small disk must always be interior to the large disk:
-\begin{equation}
- R_p \ge R_n + \sqrt{\left(p_{sp} - p_{sn} \right)^2 + \left(q_{sp} - q_{sn} \right)^2}
-\end{equation}
-
-For the distances between the centers of the two disks we will use the same notations as the one found in the paper \citep{2013MNRAS.434..765K}, $a \equiv p_{sn} - p_{sp}$ and $b \equiv q_{sn} - q_{sp}$.
-
-The half-light radius of any source is invariant to any rotational transformation.
-In the present case of a crescent source the effective radius is dependent on the parameters $R_p$, $R_n$ and $\sqrt{a^2+b^2}\equiv c $ exclusively.
-From symmetry considerations the centroid of the source is collinear with the centers of the two disks and it is situated at a distance $d_c$ from the center of the bright disk.
-$d_c$ can be computed numerically with the use of a variation of equation \ref{eqn:s1_d}:
-
-\begin{equation}
-\begin{aligned}
-\frac{S_0^C}{2} & = \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{d_c}^{R_p} \bigg[ \sqrt{R_p^2 - p^2} \\ 
-        & - \sqrt{R_p^2 - \left(p-c\right)^2} \Theta \left(R_p^2 - \left(p-c\right)^2 \right) \bigg] dp. 
-\end{aligned}
-\end{equation}
-
-With the position of the centroid determined, the half-light radius can be also be computed numerically:
-
-\begin{equation}
-\begin{aligned}
-S_0^C &=  \frac{4S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{0}^{r_{1/2}} \left[ \Theta_2  - \Theta_1  \right] p dp; \\
-\Theta_1 &= \bigg[ \pi - \arccos \frac{R_p^2 + p^2 -d_c^2}{2 R_p p} \\
-         & \phantom{= \bigg[ \pi} - \arccos \frac{R_p^2 - p^2 + d_c^2}{2 R_p d_c}
-            \bigg] \Theta \left( p - R_p + d_c \right); \\
-\Theta_2 &=  \pi - \arccos \frac{ p^2  + (c+d_c)^2 - R_n^2}
-                                {2 p \left(c + d_c \right)} \\
-         & \phantom{= \pi - } \Theta \left( p + R_n - d_c -c \right). 
-\end{aligned}
-\end{equation}
-
-\begin{figure}
-\includegraphics[width = .49\textwidth]{figures/figure_rhalf.eps}
-\caption{\label{fig:geom_crescent} Geometry of crescent sources: the source can be seen as a superposition of a smaller dark disk over and inside a larger brighter disk.
-$R_p$ and $R_n$ mark the radii of the bright and dark disk.
-$d_c$ and $c$ mark the distance from the center of the bright disk to the centroid and from the dark disk to the centroid.
-}
-\end{figure}
-
-\section{Lightcurves of the extended sources during fold crossing}\label{sec:fold-crossing}
-
-Using equation \ref{eqn:ft} and the one-dimensional flux function
-presented in the previous section one can compute numerically the
-lightcurves of the three extended sources for the simplified
-infinite-wall-caustics model.
-
-\subsection{Lightcurve of the gaussian source}
-
-The amount of light received by an observer from a source with a gaussian distributed brightness with $\sigma$ and total flux $S_0^G$ in the absence of any gravitational lensing is:
-\begin{equation}
- F^G(t) = \int_{-\infty}^\infty  \left( \mu_0 + \frac{C_0}{\sqrt{p}} \Theta \left( p \right) \right) \left( \frac{S_0^G}{\sqrt{2 \pi} \sigma} e^{-\frac{(p-p_s(t))^2}{2 \sigma^2}} \right) \mathrm{d}p.
-\end{equation}
-
-which can be simplified to:
-\begin{equation}
- F^G(t) = \mu_0 S_0^G + \frac{C_0 S_0^G}{\sqrt{2\pi} \sigma} \int_{0}^\infty \frac{e^{-\frac{(p-p_s(t))^2}{2 \sigma^2}}}{\sqrt{p}} \mathrm{d}p.
-\end{equation}
-
-\begin{figure}
-\centering
-    \includegraphics[width = 0.48\textwidth]{figures/S1D_all.eps}
-    \includegraphics[width = 0.48\textwidth]{figures/4source_magnification.eps}
-\caption{\label{fig:lightcurve_gauss} 1D profiles (top panel) and simulated lightcurves (bottom) of crescent, Gaussian and disk sources with identical total flux ($S_0$) and half-light radius($r_{1/2}$).
-Crescent source has $r_{1/2}$ = 0.72754 $R_p$, $R_n$ = 0.4 $R_p$, a = 0.3/-0.3 $R_p$.
-}
-\end{figure}
-
-\subsection{Lightcurve of the disk shaped source}
-
-Analogous to the gaussian shaped source, the disk source with uniform brightness, radius $R$ and unmagnified flux $S_0^D$ has a lightcurve described by the equation:
-
-\begin{equation}
-\begin{aligned}
- F^D(t) &= \int_{-\infty}^\infty  \left( \mu_0 + \frac{C_0}{\sqrt{p}} \Theta \left( p \right) \right) \\
-    & \bigg[ \frac{2 S_0^D}{ \pi R} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}}  \Theta \left(R^2 - \left(p-p_s(t) \right)^2 \right) \bigg] \mathrm{d}p.
-\end{aligned}
-\end{equation}
-
-which is equivalent to:
-\begin{equation}
- F^D(t) = \mu_0 S_0^D + \frac{2 C_0 S_0^D}{\pi R} \int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \frac{1}{\sqrt{p}} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}} \mathrm{d}p.
-\end{equation}
-
-\begin{figure}
-\centering
-    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_a.eps}
-    \includegraphics[width = 0.48\textwidth]{figures/4avar_magnification.eps}
-\caption{\label{fig:a_var} 1D profiles (top panel) and simulated lightcurves (bottom) of crescent sources with identical $R_p$ and $R_n$ ($R_n$ = 0.4 $R_p$) and different disk centers displacement (a).}
-\end{figure}
-
-\subsection{Lightcurve of the crescent shaped source}
-
-The lightcurve of a crescent shaped source with unamplified flux $S_0^C$, radii $R_p$, $R_n$ and center displacement $a(t)$ is:
-
-\begin{equation}
-\begin{aligned}
- F^c(t) &= \mu_0 S_0^C + C_0 \frac{2 S_0^C}{\pi \left( R_p^2 -R_n^2 \right) } \\
-    &\bigg[ \int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) \right)^2 }{p}} \mathrm{d}p \\
-    &  -  \int_{max(0, p_s(t) - a(t) - R)}^{max(0, p_s(t) -a(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) +a(t) \right)^2 }{p}} \mathrm{d}p  \bigg] .
-\end{aligned}
-\end{equation}
-
-\begin{figure}
-\centering
-    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_poz.eps}
-    \includegraphics[width = 0.48\textwidth]{figures/5Rn_back_var_magnification.eps}
-\caption{\label{fig:lightcurve_crescent_back} 1D profiles (top panel) and simulated lightcurves (bottom) of backwards crescent sources with identical $R_p$ and $a$ ($a$ = 0.3 $R_p$) and different dark disk radii ($R_n$).}
-\end{figure}
-
-The function $p_s(t)$ can be chosen to be equal to $v_p(t-t_0) + p_{s0}$. Where $p_{s0}$ is the coordinate $p$ of the source at the initial time, and $v_p$ is the component of the velocity
-along the $p$ axis. Such a modelling of the motion of the object in the source plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
-choosing the function $a$ to be constant in time.
-
-\begin{figure}
-\centering
-    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_neg.eps}
-    \includegraphics[width = 0.48\textwidth]{figures/5Rn_forw_var_magnification.eps}
-\caption{\label{fig:lightcurve_crescent} 1D profiles (top panel) and simulated lightcurves (bottom) of forwards crescent sources with identical $R_p$ and $a$ ($a$ = - 0.3 $R_p$) and different dark disk radii ($R_n$).
-}
-\end{figure}
-
-In figure \ref{fig:char_points} There are four characteristic points visible on the resulting one-dimensional light profile of the crescent shaped source.
-Two of the points, p1 and p4, mark the outer boundaries of the luminous disc component.
-The other two, p2 and p3, mark the boundaries of the dark disc component.
-Since $S_{1D}$ is a projection of $S_{2D}$ on a line perpendicular to the caustic, the following relation holds:
-\begin{equation}
-    p_2-p_1 = R_p -R_n - a.
-\end{equation}
-In addition there are two other obvious relations the characteristic points which are independent of the projection:
-\begin{equation}
-    p_4 -p_1 = 2 R_p,
-\end{equation}
-
-\begin{equation}
-        p_3 -p_2 = 2 R_n.
-\end{equation}
-All four points mark the positions where derivative $\frac{dS_{1D}}{dp}$ is discontinuous.
-The points can be used to define three regions: $p_1 - p_2$ where $S_1D$ is convex, $p_2 - p_3$ where $S_1D$ is concave, and $p_3 - p_4$ where $S_1D$ is convex again.
-\\
-
-Due to the nature of the caustic and the monotonic behaviour of the magnification map on both sides of the caustic.
-The previously mentioned characteristic points are inherited by the microlensing lightcurve.
-The points on the temporal dimension $t_1, t_2, t_3$ and $t_4$ correspond to instances in time when the fold is aligned with $p_1, p_2, p_3$ and $p_4$, respectively.
-For a constant relative velocity $v_p$ between the source and the caustic there is a simple relation between the points $p_i$ and instances $t_i$:
-
-\begin{equation}
-    t_j - t_i = \frac{p_j - p_i}{v_p}.
-\end{equation}
-
-With the use of the previous four equations the following identities can be written:
-
-\begin{equation}
-    R_p = \frac{v_p \times  \left( t_4 -t_1 \right)}{2}, 
-\end{equation}
-
-\begin{equation}
-        R_n = \frac{v_p \times \left( t_3 -t_2 \right)}{2}, 
-\end{equation}
-
-\begin{equation}
-        a = \frac{v_p \times \left( t_4 +t_1 - t_3 - t_2 \right)}{2}. 
-\end{equation}
-
-\begin{figure}
-\centering
-	\includegraphics[width = 0.48\textwidth]{figures/ch_points.eps}
-        \includegraphics[width = 0.48\textwidth]{figures/ch_instances.eps}
-\caption{\label{fig:char_points} The characteristic points and instances 
-of a crescent source's one dimensional profile and microlensing lightcurve.
-In the top panel a typical 1D profile  of a crescent source is presented. 
-There are two light gray regions that mark the regions of the bright disk 
-that are not overlapped by the dark disk. The dark gray band marks the 
-region where the two disks overlap. $p_1$ and $p_4$ mark the start and 
-end of the bright disk in the 
-1D representation. Analogous, $p_2$ and $p_3$ mark the start and 
-end of the dark disk.
-The bottom panel represents the associated lightcurve with the points 
-$t_i$ marking the moment when point $p_i$ overlaps with the caustic.
-The gray shades are analogous to the ones in the top panel.
- }
-\end{figure}
-
-Figure \ref{fig:lightcurve_gauss} reveals that the 
-lightcurve of a crescent source has more
-visible features than the other two light-curves corresponding to the
-disc and gaussian shape. There are three regions where $S_{1D}$ and by
-inheritance, the lightcurve has distinguishable behaviour. The first
-region that would be recorded on a lightcurve plot represents the
-period of time when the bright disk begins to be overlapped by the
-caustic and stops when just before the dark disk reaches the
-caustic. During this period of time, the flux of light from the source
-is increasingly magnified. The second period starts and ends with the
-overlapping of the dark disc. As a boundary of the two regions, there
-is a distinguishable point where the slope of the magnification is
-drastically changed. This apparent discontinuity in the first
-derivative of the magnification function is caused by the caustic
-amplification of the sudden drop in the $S_{1d}$ function. During the
-respective period, the magnification growth slows or even reverses
-during the first part of the period and starts to grow faster again as
-the dark disk ends its overlap with the caustic. At the point where
-the dark disc clears the caustic, the growth of the magnification is
-infinite, which appears as a saddle point on the lightcurve. Next, the
-final period corresponds to the case when the dark disc has cleared
-the caustic and the bright disc continues to overlap with the
-caustic. During this period, the lightcurve reaches a peak that for
-most of the parameter space is global and for the rest of the
-parameter space local.  At the end of the period, the growth of the
-magnification is negative infinite. the respective point appears as a
-second saddle point on the lightcurve. Past this point the
-magnification of any finite source will decay in roughly the same
-manner.
-
-The impact of parameter $a$ on the shape of the lightcurve can be
-observed in Figure~\ref{fig:a_var}. Sources where the center of the
-dark disc reaches the caustic before the center of the bright disc are
-characterized by a smoother broad peak in contrast to the cases where
-the center of the dark disc reaches the caustic after the bright disk.
-In the case of the latter the instance when the dark disc reaches the
-caustic corresponds to larger and larger magnifications until it
-becomes a local and even global peak. The effect of the $R_n$
-parameter on the light curve is presented in
-Figures~\ref{fig:lightcurve_crescent} and
-\ref{fig:lightcurve_crescent_back}. For the particular set of
-parameters where $R_p = R_n +a$ the third period of time discussed
-previously does not exist. Particularizing further, if the value of
-the radius of the dark disc is comparable to the value of the radius
-of the bright one then the position and shape of the maximum
-magnification are strikingly different. In case the crescent reaches
-the caustic with the bright region first, the peak magnification
-happens when the dark region reaches the caustic and it is
-characterized by a sharp variation in magnification growth, In the
-opposite case, the peak appears before the end of the bright disc
-reaches the caustic and shape is smoother.
-
-\begin{figure}
-\centering
-    \includegraphics[width = 0.48\textwidth]{figures/M87_shape.eps}
-        \includegraphics[width = 0.48\textwidth]{figures/M87_lc.eps}
-\caption{\label{fig:M87_plots} The one dimensional profile obtained
-  from the numerical integration along the ordonata axis of a simulated source
-  image is presented in the upper panel. The source image was presented in 
- Dexter et al 2012 in the top right-most panel of figure 5 
-(of the respective paper) as a result of the simulations 
-presented in McKinney \& Blandford 2009. The corresponding
-  lightcurve associated to the 1D profile is presented in the lower
-  panel. The points $\rm p_2$ and $\rm p_3$ mark the start and 
-  end of the equivalent
-  dark disk and are associated with the moments of their overlap with
-  the caustic: $\rm t_2$ and $\rm t_3$. $\rm h_{w}$ denotes the maximum width of the crescent.}
-\end{figure}
-
-\subsection{Microlensing a simulated image of M87}
-
-\cite{2012MNRAS.421.1517D} have created a radiative image of M87 based on the GRMHD simulations presented in \citep{2009MNRAS.394L.126M}. 
-The top right-most image in figure 5 of \cite{2012MNRAS.421.1517D} has been projected to a 1D profile associated to the perpendicular 
-direction to a fold caustic approaching the image from the right. The projection is presented in the upper panel of figure 9. 
-The amplification values of the flux of light corresponding to a microlensing event are presented in the lower panel of figure 9. 
-In general, the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions 
-surrounding the luminous parts of the image have non-zero flux and thus are more extended than the simplified source model.
-
-\section{Fitting mock data}\label{sec:numerics}
-
-Having used the simple model of an ideal fold to gain insight, we now
-consider the question of whether the characteristics of a crescent
-source could be discerned in real data having noise and systematic
-error as well.  Our method is to generate microlensing lightcurves
-using a widely-used microlensing code
-\citep{1990PhDT.......180W,1999A&A...346L...5W,1999JCoAM.109..353W} to
-which we have added a small extension to provide the option of
-crescent sources, and then fit these mock lightcurves to source models
-using Markov-chain Monte Carlo (MCMC).
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.5\textwidth]{figures/cc_forward.eps}
-\caption{\label{fig:cc_forward} Mock lightcurve fitted with the
-  correct source type and caustic.  The upper panel shows the mock
-  data points, and running through them is a seven-parameter best fit:
-  these being the crescent parameters $(R_p,R_n,a,b)$ and additionally
-  the start and end times and the overall normalisation.  (Ideal
-  lightcurves for the same crescent parameters appear above in
-  Figures~\ref{fig:lightcurve_gauss}, \ref{fig:a_var},
-  \ref{fig:lightcurve_crescent} and \ref{fig:char_points}.)  The curve
-  has $\chi^2=243, P=0.32$ indicating a good fit.  The lower panel
-  shows the posterior distributions for $r_{1/2}$ (solid histogram)
-  and $R_n/R_p$ (dashed histogram).}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.5\textwidth]{figures/cc_backward.eps}
-\caption{\label{fig:cc_backward} Like Figure~\ref{fig:cc_backward},
-  except that the fit uses path~II whereas the mock data still come
-  from path~I. The curve has $\chi^2=284, P=0.01$ indicating a
-  marginally acceptable fit.}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.5\textwidth]{figures/gc_forward.eps}
-\caption{\label{fig:gc_forward} Like Figure~\ref{fig:cc_backward}, but
-  showing the best fit to a Gaussian source.  The histogram shows the
-  posterior distribution of $r_{1/2}$ (there is no $R_n$).  It appears
-  to show $r_{1/2}$ as accurately recovered, but is not to be trusted,
-  because the fit has $\chi^2=344, P=3\times10^{-6}$ indicating
-  rejection of the Gaussian source.  Looking closely, the curve is a
-  little too shallow near $t=3$, and it is unable to fit the second
-  sub-peak near $t=5$.}
-\end{figure}
-
-\subsection{Numerical microlensing lightcurves}
-
-The microlensing code uses a ray-shooting technique to compute the
-gravitational lensing effect of a mass distribution consisting of
-(a)~a smooth component and (b)~a random distribution of point masses
-representing stars.  The ray shooting maps a grid of $\vec\theta$ to
-$\vec\beta$ using the lens equation (\ref{eqn:lens}).  That is, the
-rays are shot from the observer back to the source.  For the
-computation of the individual deflection angles
-$\vec\alpha(\vec\theta)$ a hierarchical or tree method is used.  The
-positions of all lensing masses are put into a grid of $\vec\theta$.
-Each grid cell is subdivided into four smaller squares recursively
-until every cell contains only one mass.  Nearby masses are added
-individually while distance masses are clumped into larger grid cells
-whose net contribution is approximated by its first few multipole
-moments.  Scaled units are used, with the constant pre-factor in the
-deflection angle (\ref{eqn:alpha}) separated out.  The result of ray
-shooting is a pixel map on the $\vec\beta$ plane of the number of
-lightrays which arrive at the source plane from a particular observer.
-This intermediate result is effectively a magnification map on the
-source plane.
-
-When generating the magnification map depicted in
-Figure~\ref{fig:magnification_map}, which is used in our analysis
-below, only two point masses were included.  This was done in order to
-have clean fold caustics.  There was no smooth component.
-
-Once the map is created, the lightcurve can be obtained by specifying
-the transit path of the source across the map.  The source is moved
-along the transit line, one pixel at a time, and at each step the
-brightness distribution of the source is convolved
-(equation~\ref{eqn:ft2d}) with the magnification map to give the
-brightness.  In real life, not only are both lens and source moving
-but the lens lens configuration, and with it the magnification
-pattern, is also changing with time.  While the first subtlety is
-taken care of by a coordinate transformation in this analysis, for the
-second one the lens configuration is assumed to be constant in time.
-
-For the purpose of this analysis, it was desirable to mimic the
-analytical behaviour of a simple fold as much as possible, for
-comparing the numerical result with the analytical one, therefore, the
-path of the source was chosen, so that it intersects the border of the
-caustic perpendicularly, and on a point where the border is a fold
-caustic.  We choose the line AC in Figure~\ref{fig:magnification_map}
-and consider its two caustic-crossing segments separately:
-\begin{itemize}
-\item ``path~I'' runs from C to B;
-\item ``path II'' runs from A to B.
-\end{itemize}
-The magnification along paths I and II are very similar, but not
-identical.
-
-The mock lightcurve used a crescent source (as explained in
-\S\ref{subsec:crescent}) with parameters
-\begin{equation}
-   \left(R_p, \frac{R_n}{R_p}, \frac{a}{R_p}, \frac{b}{R_p}\right) =
-   (1.31, 0.4, 0, 0.3)
-\label{eqn:cp}
-\end{equation}
-corresponding to $r_{1/2}=1$.  Unit length has 30 pixels on the
-magnification map.  The source was carried for 250 pixels along part
-of path~I, giving 251 brightness values.  Gaussian noise at the level
-of 4\% of the current brightness was then added to give the mock
-lightcurve.  If we assume the brightness measurements are nightly, and
-the transverse speed is 600~km/s, $r_{1/2}$ comes to 24~light-minutes,
-which is plausible for a quasar.
-
-\subsection{Model-fitting results}
-
-We now proceed to fit the mock lightcurve using MCMC, to recover the
-source parameters.
-
-Note that fitting involves three nuisance parameters, namely the
-brightness normalisation and the start- and endpoints on the path.
-The latter two are equivalent to considering the location of the
-caustic with respect to the starting point and the speed of the
-source.  The fitting method automatically marginalises over all the
-nuisance parameters.
-
-Figure \ref{fig:cc_forward} shows a fit to the mock lightcurve and
-recovery of $r_{1/2}$ and $R_n/R_p$.
-
-Figure \ref{fig:cc_backward} shows the result when a different caustic
-is used to fit, than was used to generate the mock lightcurve.  This
-mimics the unavoidable systematic error of not knowing the caustic
-exactly.  The fit is slightly degraded, but still acceptable.
-
-Figure \ref{fig:gc_forward} shows the result of attempting to a
-Gaussian source.  There is a small but statistically very significant
-misfit because a Gaussian curve cannot reproduce the part of the
-lightcurve with the slight flattening and then a second sub-peak
-(cf.~Figure~\ref{fig:char_points}) as the dark part of the crescent
-crosses the caustic.
-
-\section{Discussion}\label{sec:discussion}
-
-In the current paper, we simulate and study the resulting microlensing lightcurves of geometric crescent-shaped sources 
-and compare them with the microlensing lightcurves of other simple mathematically describable source profiles. 
-In order to mimic the behaviour of the flux of light from the source in the proximity of a fold caustic, we make use of the simple approximation described in equation (5). 
-The equation would exhaustively describe the magnification map and offer a good universal approximation for the 
-particular microlensing regime that we consider. 
-Namely, the shape of the caustic boundary in the proximity of the source in the respective plane can be approximated by
- a line due to reason that the local radius of curvature of caustic is orders of magnitude greater than the half-light
- radius of the studied source. 
-In particular cases in which the previously mentioned approximation loses its validity, the impact on the quality of the 
-lightcurves is not evenly distributed. The shape of the lightcurve will be maintained. 
-The data points corresponding to the source position before and during the overlapping of the caustic will be affected by
- smaller errors than the data points corresponding to later times. \\
-
-The first two source profiles that we consider are the uniform disk and symmetric Gaussian source. 
-Both of them can be described by a half-light radius $r_{1/2}$ and a total unlensed light flux $S_0$. 
-With the two parameters constrained the one-dimensional profiles, as well as the lightcurves of the two source, are completely determined, since no free parameter remains.
-
-The previous statement does not hold for a crescent source.
-In the case of the crescent source there are in total five parameters: the integrated flux of the source $S_0$,
- the radii of the bright/dark disk $R_p$/$R_n$ and the displacement of the centers of the two disks on the axes perpendicular and parallel to the caustic $a$ and $b$.
-
-Two of the parameters can be reduced by expressing the results in terms of $S_0$ and $r_{1/2}$. 
-The later being determinable for any set of parameters $R_p, R_n$ and $a^2+b^2$. Moreover, one of the displacement 
-parameters $b$ has no impact on the one-dimensional profile of the source that results from the projection of the source image on an axis perpendicular to the caustic.
-
-Since the one-dimensional source profile that corresponds to an axis perpendicular to the caustic contains exhaustively all the information regarding the source that can be revealed by the lightcurve,
-the value of the parameter $b$ does not have an effect on the shape of the lightcurve.
-
-Nevertheless, the $b$ parameter is relevant for the calculation of
-$r_{1/2}$.  It's qualitative effect is to decrease the value of the
-half-light radius when the absolute value of the parameter is
-increased.  With two parameters constrained and another irrelevant to
-the shape of the lightcurve, two free parameters remain $R_n$ and $a$.
-Figure 4 reveals that the lightcurve of a crescent source has more
-visible features than the other two light-curves corresponding to the
-disc and Gaussian shape. The parameters $R_n$ and $a$ have strong
-influences on the shape of the microlensed lightcurve as can be seen
-in Figures~\ref{fig:a_var}, \ref{fig:lightcurve_crescent} and
-\ref{fig:lightcurve_crescent_back}.  Moreover, the one-dimensional
-source profile corresponding to the direction perpendicular to the
-caustic reveal four characteristic points. The overlap of each of
-these points with the caustic leaves visible features on the
-lightcurve at the corresponding instances of time. In timely order,
-the instances correspond to the start of the overlap between the
-caustic and the bright disk, the start of the overlap between the
-caustic and the dark disk, the end of the overlap between the caustic
-and dark disk and finally the end of the overlap between the caustic
-and the bright disk.
-
-With the different source profiles and their corresponding lightcurves studied we can change our point of view of the system to that of an observer. 
-The observer would basically detect only the lightcurve of such a source. As described in section 4.3 the timing of the onset and offset of the previously 
-described periods can be used to estimate the values of the radii and one of the displacement parameters when assuming a geometric crescent. 
-All quantities can be estimated in terms of the relative velocity of the source in a direction perpendicular to the caustic.
-
-Furthermore, a simulated image of M87 presented in \citep{2012MNRAS.421.1517D} has been microlensed (figure 9). On the resulting lightcurve the instances 
-corresponding to the start and end of the black hole shadow and caustic overlap were distinguishable.
-
 In the case of a high-quality lightcurve with insignificant noise and measurement errors, the parameters can be obtained
@@
 Otherwise, if at one of the steps the data does not reject the model then the respective model is a good approximation.
-
-The previously mentioned abstract parameters can be related to
-physical quantities specific to the central region of a quasar.  As
-such the luminous region would correspond to the bright accretion disc
-that surrounds the black hole. The later's gravity would cause a
-shadow in the bright region limited by the extent of the event horizon
-of the black hole. Therefore, the radius of the bright disk would
-provide an estimate of the size of the accretion disk and the radius
-of the dark disc would provide and estimate of the gravitationally
-magnified Schwarzschild radius of the black hole $R_{S}^{magnified} =
-\Delta t_{dark} \cdot v_p$.  By $R_{S}^{magnified}$ we refer to the
-apparent Schwarzschild radius which is larger than the real value at
-large distances due to the black hole's own gravity.  Moreover, the
-gravitationally magnified value of the Schwarzschild radius is a
-monotonic function of the black hole's mass. Therefore, it can be used
-to estimate the mass of the black hole if it was not rotating. In the
-previous expression, the $\Delta t_{dark}$ denotes the period of
-overlap between the black hole shadow and fold caustic.  $v_p$ denotes
-the component of the relative velocity of the source and fold which is
-perpendicular to the caustic.  The respective velocity is an unknown,
-though it can be constrained on a case by case basis to an order of
-magnitude or even better. This would require the study of the dynamics
-of the stellar structure which contains the gravitational lens.  A
-better estimate of the relative velocity would facilitate a better
-estimate of the effective non-rotating black hole mass associated with
-the black hole shadow.
-
-The parameters whose values cannot be determined due to the loss of information from the directions parallel to the 
-caustic could be obtained in the eventuality in which the same source crosses multiple caustics that are not parallel. 
-Multiple crossing of caustics can reveal details of the one-dimensional flux profile corresponding to multiple distinct 
-directions which would allow the reconstruction of the two-dimensional profile analogous to the process through which an image of a CT scan is obtained.
-
-\textit{Author contributions}\\
-Prasenjit Saha provided the original idea and plan for the research project as well as multiple contributions to the analysis 
-and manuscript preparation. Mihai Tomozeiu simulated and studied the ideal behaviour of the microlensing lightcurves for the 
-different source profiles discussed and prepared the manuscript.
-Joachim Wambsganss contributed to the research planning and provided the numerical code used by Manuel Rabold to create 
-the magnification map and corresponding lightcurves used in the MCMC analysis performed by Irshad Mohammed in the last 
-part of the presented work. Both Manuel Rabold and Irshad Mohammed had large contributions in writing the "Fitting Mock Data" section.
-
-\section{Acknowledgement}
-IM is supported by Fermi Research Alliance, LLC under Contract No. De-AC02-07CH11359 with the United States Department of Energy.
-J.W. would like to acknowledge and thank the Pauli Center for Theoretical Studies of ETH Zurich and University of Zurich for 
-generous support during the Schrödinger visiting professorship in 2013.
-
-\bibliographystyle{mn2e}
-
-\def\apj{ApJ}
-\def\apjl{ApJL}
-\def\aj{AJ}
-\def\mnras{MNRAS}
-\def\aap{A\&A}
-\def\nat{nature}
-\def\araa{ARAA}
-\def\pasa{PASA}
-\bibliography{heap}
-
-\end{document}

### 2016-10-10 paper1/ms.tex
changes

@@
 The first two source profiles that we consider are the uniform disk and symmetric Gaussian source. 
-Both of them can be described by a half-light radius $r_{1/2}$ and a total unlensed light flux $S_0$. 
-With the two parameters constrained the one-dimensional profiles, as well as the lightcurves of the two source, are completely determined, since no free parameter remains.
+Both of them can be described by a half-light radius ($r_{1/2}$) and a total unlensed light flux ($S_0$). 
+With the two parameters constrained the one-dimensional profiles, as well as the lightcurves of the 
+two source, are completely determined, since no free parameter remains.
 
@@
 In the case of the crescent source there are in total five parameters: the integrated flux of the source $S_0$,
- the radii of the bright/dark disk $R_p$/$R_n$ and the displacement of the centers of the two disks on the axes perpendicular and parallel to the caustic $a$ and $b$.
+ the radii of the bright and dark disks and the displacement of the centers of the two disks on the axes perpendicular and parallel to the caustic.
 
-Two of the parameters can be reduced by expressing the results in terms of $S_0$ and $r_{1/2}$. 
-The later being determinable for any set of parameters $R_p, R_n$ and $a^2+b^2$. Moreover, one of the displacement 
-parameters $b$ has no impact on the one-dimensional profile of the source that results from the projection of the source image on an axis perpendicular to the caustic.
+%Two of the parameters can be reduced by expressing the results in 
+%terms of $S_0$ and $r_{1/2}$. 
+%The later being determinable for any set of parameters 
+%$R_p, R_n$ and $a^2+b^2$. Moreover, one of the displacement 
+%parameters $b$ has no impact on the one-dimensional profile 
+%of the source that results from the projection of the source
+% image on an axis perpendicular to the caustic.
 
-Since the one-dimensional source profile that corresponds to an axis perpendicular to the caustic contains exhaustively all the information regarding the source that can be revealed by the lightcurve,
-the value of the parameter $b$ does not have an effect on the shape of the lightcurve.
+Since the one-dimensional source profile that corresponds to an 
+axis perpendicular to the caustic contains exhaustively all the 
+information regarding the source that can be revealed by the lightcurve, 
+the value of the crescent's disks centers displacement parallel to the caustic 
+does not have an effect on the shape of the lightcurve.
 
-Nevertheless, the $b$ parameter is relevant for the calculation of
-$r_{1/2}$.  It's qualitative effect is to decrease the value of the
+Nevertheless, the respective displacement is relevant for the calculation of
+of the half-light radius.  It's qualitative effect is to 
+decrease the value of the
 half-light radius when the absolute value of the parameter is
-increased.  With two parameters constrained and another irrelevant to
-the shape of the lightcurve, two free parameters remain $R_n$ and $a$.
+increased.  In contrast to the simpler source shapes discussed before the
+crescent has two additional relevant parameters: the radius of the dark 
+disk ($R_n$) and the center displacement in the direction
+ perpendicular to the caustic $a$.
 Figure 4 reveals that the lightcurve of a crescent source has more
@@
 
-With the different source profiles and their corresponding lightcurves studied we can change our point of view of the system to that of an observer. 
-The observer would basically detect only the lightcurve of such a source. As described in section 4.3 the timing of the onset and offset of the previously 
-described periods can be used to estimate the values of the radii and one of the displacement parameters when assuming a geometric crescent. 
-All quantities can be estimated in terms of the relative velocity of the source in a direction perpendicular to the caustic.
+With the different source profiles and their corresponding lightcurves 
+studied we can change our point of view of the system to that of an observer. 
+The observer would basically detect only the lightcurve of such a source.
+ As described in section 4.3 the timing of the onset and 
+offset of the previously 
+described periods can be used to estimate the values of the radii 
+and one of the displacement parameters when assuming a geometric crescent. 
+All quantities can be estimated in terms of the relative velocity 
+of the source in a direction perpendicular to the caustic.
 
-Furthermore, a simulated image of M87 presented in \citep{2012MNRAS.421.1517D} has been microlensed (figure 9). On the resulting lightcurve the instances 
-corresponding to the start and end of the black hole shadow and caustic overlap were distinguishable.
+Furthermore, a simulated image of M87 presented in
+ \citep{2012MNRAS.421.1517D} has been microlensed (figure 9). 
+On the resulting lightcurve the instances 
+corresponding to the start and end of the black hole 
+shadow and caustic overlap were distinguishable.
 
-In the case of a high-quality lightcurve with insignificant noise and measurement errors, the parameters can be obtained 
-by simply identifying the characteristic instances of time without making use of the actual values of the magnification
- map. If the effect of the errors and the noise distorts the magnification time function enough so that the 
-characteristic epochs are not identifiable with the characteristic periods still visible, the boundaries of the periods 
-can be roughly estimated. Furthermore, if direct estimates of the parameters cannot be obtained we propose the use of a strong
-statistical tool such as Markov-Chain Monte Carlo. In Section 6  we have studied the possibility of identifying a crescent 
-source and  the possibility of recovering the respective parameters. As magnification map, we have used a complex numerical 
-one generated with the microlensing code by \cite{1999A&A...346L...5W}. In addition, we have added to the signal a Gaussian 
-noise with an SNR of 1.6.  In the experiment, we have considered all nine combinations of original source profiles and assumed fitting source models.
+In the case of a high-quality lightcurve with 
+insignificant noise and measurement errors, 
+the parameters can be obtained 
+by simply identifying the characteristic 
+instances of time without making use of 
+the actual values of the magnification
+ map. If the effect of the errors and the 
+noise distorts the magnification time function enough so that the 
+characteristic epochs are not identifiable with the characteristic 
+periods still visible, the boundaries of the periods 
+can be roughly estimated. Furthermore, if direct estimates of 
+the parameters cannot be obtained we propose the use of a strong
+statistical tool such as Markov-Chain Monte Carlo.
+ In Section 6  we have studied the possibility of identifying a crescent 
+source and the possibility of recovering the 
+respective parameters. As magnification map, we have used a complex numerical 
+one generated with the microlensing code by \cite{1999A&A...346L...5W}. 
+In addition, we have added to the signal a Gaussian 
+noise with an SNR of 1.6.  In the experiment,
+ we have considered all nine combinations of
+ original source profiles and assumed fitting source models.
 Effectively we have fitted using MCMC all three sources with all three assumed fitting models. The results of the experiment allowed us to build a 
@@
 
-The parameters whose values cannot be determined due to the loss of information from the directions parallel to the 
-caustic could be obtained in the eventuality in which the same source crosses multiple caustics that are not parallel. 
-Multiple crossing of caustics can reveal details of the one-dimensional flux profile corresponding to multiple distinct 
-directions which would allow the reconstruction of the two-dimensional profile analogous to the process through which an image of a CT scan is obtained.
+The parameters whose values cannot be determined due to the 
+loss of information from the directions parallel to the 
+caustic could be obtained in the eventuality in which 
+the same source crosses multiple caustics that are not parallel. 
+Multiple crossing of caustics can reveal details of 
+the one-dimensional flux profile corresponding to multiple distinct 
+directions which would allow the reconstruction of the 
+two-dimensional profile analogous to the process through 
+which an image of a CT scan is obtained.
 

### 2016-10-10 paper1/ms2.tex
small edits in ms2

@@
-In the case of a high-quality lightcurve with insignificant noise and measurement errors, the parameters can be obtained
-by simply identifying the characteristic instances of time without making use of the actual values of the magnification
- map. If the effect of the errors and the noise distorts the magnification time function enough so that the
-characteristic epochs are not identifiable with the characteristic periods still visible, the boundaries of the periods
-can be roughly estimated. Furthermore, if direct estimates of the parameters cannot be obtained we propose the use of a strong
-statistical tool such as Markov-Chain Monte Carlo. In Section 6  we have studied the possibility of identifying a crescent
-source and  the possibility of recovering the respective parameters. As magnification map, we have used a complex numerical
-one generated with the microlensing code by \cite{1999A&A...346L...5W}. In addition, we have added to the signal a Gaussian
-noise with an SNR of 1.6.  In the experiment, we have considered all nine combinations of original source profiles and assumed fitting source models.
-Effectively we have fitted using MCMC all three sources with all three assumed fitting models. The results of the experiment allowed us to build a
-procedure for distinguishing the shape of the source assuming that one of the three models we have considered is a good approximation.
-The procedure would be useful for observers that endeavour to gain more information about an unresolved source for which they can study
-the microlensing lightcurve.  As a first step in the procedure, one should first attempt to use MCMC with a Gaussian model assumption.
-If either the value of the $\chi^2$ or the number of rejected datapoints is large then the next step is to change the assumed source model to
-a uniform disc and redo the fitting. Finally, if the uniform disc assumption is rejected as well then the fitting should be done with a
-crescent source model assumption. If at each of the three steps the data rejects the model then the source cannot be approximated by any of the three models.
-Otherwise, if at one of the steps the data does not reject the model then the respective model is a good approximation.
+Gravitational microlensing of quasars, reviewed in
+Section~\ref{sec:microlensing} below, is a phenomenon seen in quasars
+that have been multiply imaged by galaxies.  Individual images in a
+multiple-image system can undergo sharp and seemingly random
+brightness changes.  This can occur, even for constant intrinic
+luminosity, as a consequence of two things:....
+
+-------------------------------------------------------------
+
+In Section~\ref{sec:numerics} we carry out source fitting to
+lightcurves, with both noise and systematic errors are present.  We
+generate a mock lightcurves by taking a crescent source across the
+path AB in Figure~\ref{fig:magnification_map}, and then adding noise.
+The path simulates crossing a clean but not ideal fold.  We then fit
+this lightcurve to templates from both crescent and circular-Gaussian
+sources across the path CB.  That is, the templates come from a
+similar but not identical caustic, thus deliberately generating a
+systematic error.  Despite this systematic error, a crescent source
+fits the data, while a Gaussian source is rejected from the $\chi^2$
+value.
+
+--------------------------------------------
+
+In the case of a high-quality lightcurve, the parameters of a crescent
+source can be estimated by simply identifying signatures on the
+lightcurve corresponding to transits of the particular source features
+across the caustic.  This simple idea can be supplemented with
+statistics.  We do so in Section~\ref{sec:numerics}, where we take a
+realistic magnification map generated by the microlensing code
+from \cite{1999A&A...346L...5W}, compute a caustic-crossing lightcurve
+from it, and then fit the lightcurve with other caustic-crossing
+templates.  The conclusion is that a crescent source can be
+distinguised from a circular source, on the basis of caustic-crossing
+lightcurves, if the caustic is a clean but not ideal fold.
+
+-----------------------------------------------------------------

### 2016-10-10 paper1/ms.tex

@@
 $\mu(\vec\theta)$ is more or less than unity.  If there are multiple
-mages at distinct but not observationally resolved $\vec\theta_i$ from
+images at distinct but not observationally resolved $\vec\theta_i$ from
 the same $\vec\beta$, a total brightness amplification
-of\begin{equation} \mu_{\rm total} = \sum_{i} \mu(\vec\theta_i)
+of
+
+\begin{equation} \mu_{\rm total} = \sum_{i} \mu(\vec\theta_i)
 \end{equation}
@@
 Instead we will
-use the difference which in this case is th total flux from the crescent-shaped source $S_0^C$. \\
+use the difference which in this case is the total flux from the crescent-shaped source $S_0^C$. \\
 Equation \ref{eqn:s2d} can be written as:\\
@@
 All four points mark the positions where derivative $\frac{dS_{1D}}{dp}$ is discontinuous.
-The points can be used to define three regions: $p_1 - p_2$ where $S_1D$ is convex, $p_2 - p_3$ where $S_1D$ is concave, and $p_3 - p_4$ where $S_1D$ is convex again.
+The points can be used to define three regions: $p_1 - p_2$ where $S_{1D}$ is convex, $p_2 - p_3$ where $S_{1D}$ is concave, and $p_3 - p_4$ where $S_{1D}$ is convex again.
 \\
@@
 brightness.  In real life, not only are both lens and source moving
-but the lens lens configuration, and with it the magnification
+but the lens configuration, and with it the magnification
 pattern, is also changing with time.  While the first subtlety is

### 2016-10-10 paper1/ms.tex
Prasenjit's new paragraphs

@@
 
+%Gravitational microlensing of quasars, reviewed in
+%Section~\ref{sec:microlensing} below, refers to sharp changes in the
+%observed brightness of quasars that have been lensed by an intervening
+%galaxies, without any changes in the intrinsic luminosity.
+
 Gravitational microlensing of quasars, reviewed in
-Section~\ref{sec:microlensing} below, refers to sharp changes in the
-observed brightness of quasars that have been lensed by an intervening
-galaxies, without any changes in the intrinsic luminosity.  It is a
+Section~\ref{sec:microlensing} below, is a phenomenon seen in quasars
+that have been multiply imaged by galaxies.  Individual images in a
+multiple-image system can undergo sharp and seemingly random
+brightness changes.  This can occur, even for constant intrinic
+luminosity, as a consequence of two things: It is a
 consequence of two things: the very small size of the central engine,
@@
 
-In Section~\ref{sec:source-models} we introduce the three source
-profiles used in our subsequent models and simulations: a
-constant-brightness disc, a circular Gaussian, and the crescent source
-introduced by \citep{2013MNRAS.434..765K}.  The latter is simply a
-constant-brightness disc with a smaller, non-concentric disc cut out
-of it.  We also derive the half-light radius for a crescent.  The
-half-light radius can characterise the source size for all three types
-of source.
+%In Section~\ref{sec:source-models} we introduce the three source
+%profiles used in our subsequent models and simulations: a
+%constant-brightness disc, a circular Gaussian, and the crescent source
+%introduced by \citep{2013MNRAS.434..765K}.  The latter is simply a
+%constant-brightness disc with a smaller, non-concentric disc cut out
+%of it.  We also derive the half-light radius for a crescent.  The
+%half-light radius can characterise the source size for all three types
+%of source.
+
+In Section~\ref{sec:numerics} we carry out source fitting to
+lightcurves, with both noise and systematic errors are present.  We
+generate a mock lightcurves by taking a crescent source across the
+path AB in Figure~\ref{fig:magnification_map}, and then adding noise.
+The path simulates crossing a clean but not ideal fold.  We then fit
+this lightcurve to templates from both crescent and circular-Gaussian
+sources across the path CB.  That is, the templates come from a
+similar but not identical caustic, thus deliberately generating a
+systematic error.  Despite this systematic error, a crescent source
+fits the data, while a Gaussian source is rejected from the $\chi^2$
+value.
 
@@
 
-In the case of a high-quality lightcurve with 
-insignificant noise and measurement errors, 
-the parameters can be obtained 
-by simply identifying the characteristic 
-instances of time without making use of 
-the actual values of the magnification
- map. If the effect of the errors and the 
-noise distorts the magnification time function enough so that the 
-characteristic epochs are not identifiable with the characteristic 
-periods still visible, the boundaries of the periods 
-can be roughly estimated. Furthermore, if direct estimates of 
-the parameters cannot be obtained we propose the use of a strong
-statistical tool such as Markov-Chain Monte Carlo.
- In Section 6  we have studied the possibility of identifying a crescent 
-source and the possibility of recovering the 
-respective parameters. As magnification map, we have used a complex numerical 
-one generated with the microlensing code by \cite{1999A&A...346L...5W}. 
-In addition, we have added to the signal a Gaussian 
-noise with an SNR of 1.6.  In the experiment,
- we have considered all nine combinations of
- original source profiles and assumed fitting source models.
-Effectively we have fitted using MCMC all three sources with all three assumed fitting models. The results of the experiment allowed us to build a 
-procedure for distinguishing the shape of the source assuming that one of the three models we have considered is a good approximation.  
-The procedure would be useful for observers that endeavour to gain more information about an unresolved source for which they can study 
-the microlensing lightcurve.  As a first step in the procedure, one should first attempt to use MCMC with a Gaussian model assumption. 
-If either the value of the $\chi^2$ or the number of rejected datapoints is large then the next step is to change the assumed source model to 
-a uniform disc and redo the fitting. Finally, if the uniform disc assumption is rejected as well then the fitting should be done with a 
-crescent source model assumption. If at each of the three steps the data rejects the model then the source cannot be approximated by any of the three models. 
-Otherwise, if at one of the steps the data does not reject the model then the respective model is a good approximation.
+%In the case of a high-quality lightcurve with 
+%insignificant noise and measurement errors, 
+%the parameters can be obtained 
+%by simply identifying the characteristic 
+%instances of time without making use of 
+%the actual values of the magnification
+% map. If the effect of the errors and the 
+%noise distorts the magnification time function enough so that the 
+%characteristic epochs are not identifiable with the characteristic 
+%periods still visible, the boundaries of the periods 
+%can be roughly estimated. Furthermore, if direct estimates of 
+%the parameters cannot be obtained we propose the use of a strong
+%statistical tool such as Markov-Chain Monte Carlo.
+% In Section 6  we have studied the possibility of identifying a crescent 
+%source and the possibility of recovering the 
+%respective parameters. As magnification map, we have used a complex numerical 
+%one generated with the microlensing code by \cite{1999A&A...346L...5W}. 
+%In addition, we have added to the signal a Gaussian 
+%noise with an SNR of 1.6.  In the experiment,
+% we have considered all nine combinations of
+% original source profiles and assumed fitting source models.
+%Effectively we have fitted using MCMC all three sources with all three assumed fitting models. The results of the experiment allowed us to build a 
+%procedure for distinguishing the shape of the source assuming that one of the three models we have considered is a good approximation.  
+%The procedure would be useful for observers that endeavour to gain more information about an unresolved source for which they can study 
+%the microlensing lightcurve.  As a first step in the procedure, one should first attempt to use MCMC with a Gaussian model assumption. 
+%If either the value of the $\chi^2$ or the number of rejected datapoints is large then the next step is to change the assumed source model to 
+%a uniform disc and redo the fitting. Finally, if the uniform disc assumption is rejected as well then the fitting should be done with a 
+%crescent source model assumption. If at each of the three steps the data rejects the model then the source cannot be approximated by any of the three models. 
+%Otherwise, if at one of the steps the data does not reject the model then the respective model is a good approximation.
+
+In the case of a high-quality lightcurve, the parameters of a crescent
+source can be estimated by simply identifying signatures on the
+lightcurve corresponding to transits of the particular source features
+across the caustic.  This simple idea can be supplemented with
+statistics.  We do so in Section~\ref{sec:numerics}, where we take a
+realistic magnification map generated by the microlensing code
+from \cite{1999A&A...346L...5W}, compute a caustic-crossing lightcurve
+from it, and then fit the lightcurve with other caustic-crossing
+templates.  The conclusion is that a crescent source can be
+distinguised from a circular source, on the basis of caustic-crossing
+lightcurves, if the caustic is a clean but not ideal fold.
 

### 2016-10-10 paper1/ms.tex
response

@@
 \frac{S_0^C}{2} & = \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{d_c}^{R_p} \bigg[ \sqrt{R_p^2 - p^2} \\ 
-        & - \sqrt{R_p^2 - \left(p-c\right)^2} \Theta \left(R_p^2 - \left(p-c\right)^2 \right) \bigg] dp. 
+        & - \sqrt{R_p^2 - \left(p-c\right)^2} \Theta \left(R_p^2 - \left(p-c\right)^2 \right) \bigg] \mathrm{d}p. 
 \end{aligned}
@@
 \begin{aligned}
-S_0^C &=  \frac{4S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{0}^{r_{1/2}} \left[ \Theta_2  - \Theta_1  \right] p dp; \\
+S_0^C &=  \frac{4S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{0}^{r_{1/2}} \left[ \Theta_2  - \Theta_1  \right] p \mathrm{d}p; \\
 \Theta_1 &= \bigg[ \pi - \arccos \frac{R_p^2 + p^2 -d_c^2}{2 R_p p} \\
@@
 J.W. would like to acknowledge and thank the Pauli Center for Theoretical Studies of ETH Zurich and University of Zurich for 
-generous support during the Schrödinger visiting professorship in 2013.
+generous support during the Schr\"odinger visiting professorship in 2013.
 

### 2016-10-20 paper1/ms.tex
changes
Figure 3 replaced
corrections

@@
 \cite{2013MNRAS.434..765K} to use a simple geometric crescent model to
-fit the available sub-mm interferometric data, They argued that the
+fit the available sub-mm interferometric data. They argued that the
 crescent is nothing but the silhouette of the event horizon.
@@
 brightness changes.  This can occur, even for constant intrinic
-luminosity, as a consequence of two things: It is a
-consequence of two things: the very small size of the central engine,
+luminosity, as a consequence of two things: the very small size of the central engine,
 and granularity of the mass distribution of a lensing galaxy due to
@@
 continuum \citep[e.g.,][]{2012A&A...544A..62S} but is also known in
-x-rays \citep[e.g.,][]{2016arXiv160207601N} and has been argued for at
+X-rays \citep[e.g.,][]{2016arXiv160207601N} and has been argued for at
 cm wavelengths \citep{2000A&A...358..793K}.  The effect supplies an
@@
 generate a mock lightcurves by taking a crescent source across the
-path AB in Figure~\ref{fig:magnification_map}, and then adding noise.
+path AB in figure~\ref{fig:magnification_map}, and then adding noise.
 The path simulates crossing a clean but not ideal fold.  We then fit
@@
 \begin{figure}
-\includegraphics[width = .49\textwidth]{figures/figure_rhalf.eps}
+\includegraphics[width = .49\textwidth]{figures/figure_rhalf2.eps}
 \caption{\label{fig:geom_crescent} Geometry of crescent sources: the source can be seen as a superposition of a smaller dark disk over and inside a larger brighter disk.
@@
 magnitude or even better. This would require the study of the dynamics
-of the stellar structure which contains the gravitational lens.  A
+of the stellar structure which contains the gravitational lens and the bulk
+transverse velocity of the galaxy with respect to the observer.  A
 better estimate of the relative velocity would facilitate a better

### 2016-10-20 paper1/ms.tex
corrections

@@
 applies. It is possible for the magnification to become formally
-infinite, as a result of an eigenvalues of the magnification matrix
+infinite, as a result of an eigenvalue of the magnification matrix
 (\ref{eqn:magnif-matrix}) becoming infinite.  This generically happens
@@
 so-called caustics.  Caustics can appear in the optical system, not just
-gravitational lensing.  For a point source, caustics are singularities
+during gravitational lensing.  For a point source, caustics are singularities
 of the magnification; for finite-size sources caustics correspond to
@@
 
-Due to the choice of the coordinate system the amplification factor has no dependence on the $q$ coordinate. The previous equation can be rewritten as:
+Due to the choice of the coordinate system the amplification factor 
+has no dependence on the $q$ coordinate. The previous equation can be rewritten as:
 
@@
 \end{enumerate}
-The first two sources are the typical choices used in the literature to describe the luminous parts of a quasar (Prasenjit should give some citations here). The third one is a recently proposed
-variant \citep{2013MNRAS.434..765K}.
+The first two sources are the typical choices used in the literature to describe the luminous parts of a quasar. 
+The third one is a recently proposed variant \citep{2013MNRAS.434..765K}.
 
@@
 
-The surface brightness distribution of a geometric crescent can be built by considering two disk sources of constant brightness.
-One larger disk will contribute positively to the total flux while one smaller disk
+The surface brightness distribution of a geometric crescent can be built 
+by considering two disk sources of constant brightness. One larger disk 
+will contribute positively to the total flux while one smaller disk 
 that is interior to the large one will contribute negatively. This superposition can be written for 2D as:\\
@@
 \\
-The following notations were used: $R_p, (p_{sp}, q_{sp}), R_n, (p_{sn},q_{sn})$ are the radii and coordinate of the center for the larger positive disk and smaller negative disk respectively.
+The following notations were used: $R_p, (p_{sp}, q_{sp}), R_n, (p_{sn},q_{sn})$ are the radii 
+and coordinate of the center for the larger positive disk and smaller negative disk respectively. 
 $S_0^{Dp},S_0^{Dn}$ represent the total flux of radiation received from the large and small disk. 
-From this point forward we will not use the total flux from each source.
-Instead we will
+From this point forward we will not use the total flux from each source. Instead we will 
 use the difference which in this case is the total flux from the crescent-shaped source $S_0^C$. \\
@@
 
-There are some constraints on the parameters used to define a crescent in the previously presented manner that need to be stated. First, we must impose the obvious $R_p > R_n$ relation. Secondly, 
+There are some constraints on the parameters used to define a crescent in the previously 
+presented manner that need to be stated. First, we must impose the obvious $R_p > R_n$ relation. Secondly, 
 the small disk must always be interior to the large disk:
@@
 
-For the distances between the centers of the two disks we will use the same notations as the one found in the paper \citep{2013MNRAS.434..765K}, $a \equiv p_{sn} - p_{sp}$ and $b \equiv q_{sn} - q_{sp}$.
+For the distances between the centers of the two disks we will use the same 
+notations as the one found in the paper \citep{2013MNRAS.434..765K}, $a \equiv p_{sn} - p_{sp}$ and $b \equiv q_{sn} - q_{sp}$.
 
 The half-light radius of any source is invariant to any rotational transformation. 
-In the present case of a crescent source the effective radius is dependent on the parameters $R_p$, $R_n$ and $\sqrt{a^2+b^2}\equiv c $ exclusively.
-From symmetry considerations the centroid of the source is collinear with the centers of the two disks and it is situated at a distance $d_c$ from the center of the bright disk.
-$d_c$ can be computed numerically with the use of a variation of equation \ref{eqn:s1_d}:
+In the present case of a crescent source the effective radius is dependent on the 
+parameters $R_p$, $R_n$ and $\sqrt{a^2+b^2}\equiv c $ exclusively. From symmetry
+considerations the centroid of the source is collinear with the centers of the two disks 
+and it is situated at a distance $d_c$ from the center of the bright disk. $d_c$ can be 
+computed numerically with the use of a variation of equation \ref{eqn:s1_d}:
 
@@
 \includegraphics[width = .49\textwidth]{figures/figure_rhalf2.eps}
-\caption{\label{fig:geom_crescent} Geometry of crescent sources: the source can be seen as a superposition of a smaller dark disk over and inside a larger brighter disk.
-$R_p$ and $R_n$ mark the radii of the bright and dark disk.
-$d_c$ and $c$ mark the distance from the center of the bright disk to the centroid and from the dark disk to the centroid.
-}
+\caption{\label{fig:geom_crescent} Geometry of crescent sources: the source can be seen as a superposition
+ of a smaller dark disk over and inside a larger brighter disk. $R_p$ and $R_n$ mark the radii of the 
+bright and dark disk. $d_c$ and $c$ mark the distance from the center of the bright disk to the 
+centroid and from the dark disk to the centroid.  }
 \end{figure}
@@
 
-The amount of light received by an observer from a source with a gaussian distributed brightness with $\sigma$ and total flux $S_0^G$ in the absence of any gravitational lensing is:
+The amount of light received by an observer from a source 
+with a gaussian distributed brightness with $\sigma$ and total flux $S_0^G$ in the absence of any gravitational lensing is:
 \begin{equation}
@@
 
-Analogous to the gaussian shaped source, the disk source with uniform brightness, radius $R$ and unmagnified flux $S_0^D$ has a lightcurve described by the equation:
+Analogous to the gaussian shaped source, the disk source with uniform brightness, radius $R$ and 
+unmagnified flux $S_0^D$ has a lightcurve described by the equation:
 
@@
     \includegraphics[width = 0.48\textwidth]{figures/4avar_magnification.eps}
-\caption{\label{fig:a_var} 1D profiles (top panel) and simulated lightcurves (bottom) of crescent sources with identical $R_p$ and $R_n$ ($R_n$ = 0.4 $R_p$) and different disk centers displacement (a).}
+\caption{\label{fig:a_var} 1D profiles (top panel) and simulated lightcurves 
+(bottom) of crescent sources with identical $R_p$ and $R_n$ ($R_n$ = 0.4 $R_p$) and different disk centers displacement (a).}
 \end{figure}
@@
 
-The function $p_s(t)$ can be chosen to be equal to $v_p(t-t_0) + p_{s0}$. Where $p_{s0}$ is the coordinate $p$ of the source at the initial time, and $v_p$ is the component of the velocity
-along the $p$ axis. Such a modelling of the motion of the object in the source plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
+The function $p_s(t)$ can be chosen to be equal to $v_p(t-t_0) + p_{s0}$. Where $p_{s0}$ is the coordinate $p$ 
+of the source at the initial time, and $v_p$ is the component of the velocity
+along the $p$ axis. Such a modelling of the motion of the object in the source 
+plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
 choosing the function $a$ to be constant in time.
@@
     \includegraphics[width = 0.48\textwidth]{figures/5Rn_forw_var_magnification.eps}
-\caption{\label{fig:lightcurve_crescent} 1D profiles (top panel) and simulated lightcurves (bottom) of forwards crescent sources with identical $R_p$ and $a$ ($a$ = - 0.3 $R_p$) and different dark disk radii ($R_n$).
-}
-\end{figure}
-
-In figure \ref{fig:char_points} There are four characteristic points visible on the resulting one-dimensional light profile of the crescent shaped source.
-Two of the points, p1 and p4, mark the outer boundaries of the luminous disc component.
-The other two, p2 and p3, mark the boundaries of the dark disc component.
+\caption{\label{fig:lightcurve_crescent} 1D profiles (top panel) and simulated lightcurves (bottom) 
+of forwards crescent sources with identical $R_p$ and $a$ ($a$ = - 0.3 $R_p$) and different dark disk radii ($R_n$). }
+\end{figure}
+
+In figure \ref{fig:char_points} there are four characteristic points visible on the resulting one-dimensional 
+light profile of the crescent shaped source. Two of the points, p1 and p4, mark the outer boundaries of the
+ luminous disc component. The other two, p2 and p3, mark the boundaries of the dark disc component. 
 Since $S_{1D}$ is a projection of $S_{2D}$ on a line perpendicular to the caustic, the following relation holds:
@@
 \end{equation}
-All four points mark the positions where derivative $\frac{dS_{1D}}{dp}$ is discontinuous.
-The points can be used to define three regions: $p_1 - p_2$ where $S_{1D}$ is convex, $p_2 - p_3$ where $S_{1D}$ is concave, and $p_3 - p_4$ where $S_{1D}$ is convex again.
-\\
-
-Due to the nature of the caustic and the monotonic behaviour of the magnification map on both sides of the caustic.
-The previously mentioned characteristic points are inherited by the microlensing lightcurve.
-The points on the temporal dimension $t_1, t_2, t_3$ and $t_4$ correspond to instances in time when the fold is aligned with $p_1, p_2, p_3$ and $p_4$, respectively.
+All four points mark the positions where derivative $\frac{dS_{1D}}{dp}$ is discontinuous. The points can be used 
+to define three regions: $p_1 - p_2$ where $S_{1D}$ is convex, $p_2 - p_3$ where $S_{1D}$ is concave, 
+and $p_3 - p_4$ where $S_{1D}$ is convex again. \\
+
+Due to the nature of the caustic and the monotonic behaviour of the magnification map on both sides of the caustic, the previously 
+mentioned characteristic points are inherited by the microlensing lightcurve. The points on the temporal dimension 
+$t_1, t_2, t_3$ and $t_4$ correspond to instances in time when the fold is aligned with $p_1, p_2, p_3$ and $p_4$, respectively. 
 For a constant relative velocity $v_p$ between the source and the caustic there is a simple relation between the points $p_i$ and instances $t_i$:
@@
 respective period, the magnification growth slows or even reverses
-during the first part of the period and starts to grow faster again as
-the dark disk ends its overlap with the caustic. At the point where
+and later starts to grow faster again as
+the dark disc ends its overlap with the caustic. At the point where
 the dark disc clears the caustic, the growth of the magnification is
@@
 parameter space local.  At the end of the period, the growth of the
-magnification is negative infinite. the respective point appears as a
+magnification is negative infinite. The respective point appears as a
 second saddle point on the lightcurve. Past this point the
@@
 happens when the dark region reaches the caustic and it is
-characterized by a sharp variation in magnification growth, In the
+characterized by a sharp variation in magnification growth. In the
 opposite case, the peak appears before the end of the bright disc
@@
 For the purpose of this analysis, it was desirable to mimic the
-analytical behaviour of a simple fold as much as possible, for
+analytical behaviour of a simple fold as much as possible. For
 comparing the numerical result with the analytical one, therefore, the
@@
 Note that fitting involves three nuisance parameters, namely the
-brightness normalisation and the start- and endpoints on the path.
+brightness normalisation and the start and end points on the path.
 The latter two are equivalent to considering the location of the
@@
 
-Figure \ref{fig:gc_forward} shows the result of attempting to a
+Figure \ref{fig:gc_forward} shows the result of attempting to fit a
 Gaussian source.  There is a small but statistically very significant
@@
 and compare them with the microlensing lightcurves of other simple mathematically describable source profiles. 
-In order to mimic the behaviour of the flux of light from the source in the proximity of a fold caustic, we make use of the simple approximation described in equation (5). 
+In order to mimic the behaviour of the flux of light from the source in the proximity of a fold caustic, 
+we make use of the simple approximation described in equation (5). 
 The equation would exhaustively describe the magnification map and offer a good universal approximation for the 
@@
 
-The first two source profiles that we consider are the uniform disk and symmetric Gaussian source. 
+The first two source profiles that we consider are the uniform disk and the symmetric Gaussian source. 
 Both of them can be described by a half-light radius ($r_{1/2}$) and a total unlensed light flux ($S_0$). 
@@
  perpendicular to the caustic $a$.
+
 Figure 4 reveals that the lightcurve of a crescent source has more

### 2016-10-24 paper1/ms.tex
consistent fig ref

@@
   trajectory of the center of the sources for which the lightcurves
-  presented in figure 11 are generated.  Note that this represents an
+  presented in figures \ref{fig:cc_forward}, \ref{fig:cc_backward} and \ref{fig:gc_forward} are generated.  Note that this represents an
   atypically simple region of any realistic magnification map.}
@@
 The top right-most image in figure 5 of \cite{2012MNRAS.421.1517D} has been projected to a 1D profile associated to the perpendicular 
-direction to a fold caustic approaching the image from the right. The projection is presented in the upper panel of figure 9. 
-The amplification values of the flux of light corresponding to a microlensing event are presented in the lower panel of figure 9. 
+direction to a fold caustic approaching the image from the right. The projection is presented in the upper panel of figure \ref{fig:M87_plots}. 
+The amplification values of the flux of light corresponding to a 
+microlensing event are presented in the lower panel of figure \ref{fig:M87_plots}. 
 In general, the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions 
@@
 
-Figure 4 reveals that the lightcurve of a crescent source has more
+Figure \ref{fig:lightcurve_gauss} reveals that the lightcurve of a crescent source has more
 visible features than the other two light-curves corresponding to the
@@
 Furthermore, a simulated image of M87 presented in
- \citep{2012MNRAS.421.1517D} has been microlensed (figure 9). 
+ \citep{2012MNRAS.421.1517D} has been microlensed (figure \ref{fig:M87_plots}). 
 On the resulting lightcurve the instances 

### 2016-11-23 paper1/msold1.tex
template for response2

--- 2016-10-20 paper1/ms.tex
@@
 
-\section{The General Picture }
-
-Active galactic nuclei (AGN) are thought to be powered by the accretion of
-matter from the proximal environment into a supermassive black hole.
-The radiation emitted excites the surrounding medium which becomes
-detectable as narrow line regions, broad line regions and optical
-continuum.  In the direction perpendicular to the accretion disc,
-where the medium is more transparent, the outflow tends to be
-collimated to jets.  If the orientation is such that the observer's
-view of the central region is not blocked by the accretion disc or by
-a jet, a quasar is seen \citep[e.g.,][]{1984RvMP...56..255B}.  While
-the basic mechanism \citep[originating in the work
+\section{introduction}
+
+Active galactic nuclei are thought to be powered by the
+accretion of matter from the proximal environment into a supermassive
+black hole.  The radiation emitted excites the surrounding medium
+which becomes detectable as narrow line regions, broad line regions
+and optical continuum.  Moreover, in the direction perpendicular to
+the accretion disc, where the medium is more transparent, jets will
+appear.  If a jet is oriented towards the Earth, a quasar is observed
+\citep[e.g.,][]{1984RvMP...56..255B}.  While the basic mechanism
+\citep[originating in the work
   of][]{1964ApJ...140..796S,1964SPhD....9..246Z,1969Natur.223..690L}
@@
 
-Modestly active galactic nucleii are present in the Galactic centre
-and in M87.  In these two cases, the central engines ($<0.1\rm\,mas$
-on the sky) are close to being resolved through very long baseline
-interferometry, which shows preliminary indications of the
-jet-launching structures
-\citep{2008JPhCS.131a2055D,2012Sci...338..355D,2016ApJ...820...90F}.
-Current data do not deliver images but require fitting to predefined
-models for the images.  A whole range of models have been applied,
-starting from simple geometric models to more complex physical models
+For the black holes in the Galactic centre and the centre of M87
+---the two nearest objects that barely qualify as AGN--- the central
+engines (which are $<0.1\rm\,mas$ on the sky) are close to being
+resolved through very long baseline interferometry, which shows
+preliminary indications of the jet-launching structures
+\citep{2008JPhCS.131a2055D,2012Sci...338..355D,2013MNRAS.434..765K,2016arXiv160205527F}.
+Current data do not deliver images but require fitting to
+predefined models for the images.  A whole range of models have been
+applied, starting from simple geometric models to more complex
+physical models
 \citep{2008Natur.455...78D,2011ApJ...738...38B,2009ApJ...706..497M,2010ApJ...717.1092D}.
@@
 \cite{2013MNRAS.434..765K} to use a simple geometric crescent model to
-fit the available sub-mm interferometric data. They argued that the
-crescent is nothing but the silhouette of the event horizon.
+fit the data, and argued that the crescent is nothing but the
+silhouette of the event horizon.
+
 The great majority of quasars, however, lie at redshifts beyond 2
@@
 of magnitude smaller on the sky. The direct observations of the black
-hole silhouettes of quasars are far beyond foreseeable
-instrumentation.
-
+hole silhouettes of quasars are far beyond foreseeable instrumentation.
 In the present paper, we consider a possible indirect method, related
-to \cite{1999ApJ...524...49A} and \cite{2015ApJ...814L..26M}, 
-through which gravitational
+to \cite{1999ApJ...524...49A}, through which gravitational
 microlensing could probe the black hole shadow and its proximal quasar
@@
 
-%Gravitational microlensing of quasars, reviewed in
-%Section~\ref{sec:microlensing} below, refers to sharp changes in the
-%observed brightness of quasars that have been lensed by an intervening
-%galaxies, without any changes in the intrinsic luminosity.
-
-Gravitational microlensing of quasars, reviewed in
-Section~\ref{sec:microlensing} below, is a phenomenon seen in quasars
-that have been multiply imaged by galaxies.  Individual images in a
-multiple-image system can undergo sharp and seemingly random
-brightness changes.  This can occur, even for constant intrinic
-luminosity, as a consequence of two things: the very small size of the central engine,
-and granularity of the mass distribution of a lensing galaxy due to
-stars.  The latter means that the local magnification is not a smooth
-function of source position. It contains a complicated network of
-singular curves, known as caustics.
-Figure~\ref{fig:magnification_map} shows part of a magnification map
-with a few caustics.  The lensed brightness would be given by placing
-the source on such a magnification map and integrating the surface
-brightness weighted by the magnification.  Most astrophysical sources
-straddle several caustics, and hence, their net brightness varies
-smoothly with location.  The central engine of quasars, however, is
-smaller than the typical spacing between caustics.  As a result, the
-lensed brightness undergoes sudden changes as a quasar crosses a
-caustic.  Quasar microlensing is mostly observed in the optical
-continuum \citep[e.g.,][]{2012A&A...544A..62S} but is also known in
-X-rays \citep[e.g.,][]{2016arXiv160207601N} and has been argued for at
-cm wavelengths \citep{2000A&A...358..793K}.  The effect supplies an
-upper limit on the size of the central engine
-\citep[e.g.,][]{2015ApJ...814L..26M} and can be used to study the mass
+Gravitational microlensing of quasars reviewed in
+Section~\ref{sec:microlensing} below, refers to sharp changes in the
+observed brightness of quasars that have been lensed by an intervening
+galaxies, without any changes in the intrinsic luminosity.
+Microlensing affects only the light from the innermost part of the
+quasar, such as the optical continuum
+\citep[e.g.,][]{2012A&A...544A..62S} and is a consequence of two
+things: the very small size of the central engine, and granularity of
+the mass distribution of a lensing galaxy due to stars.  The latter
+means that the local magnification is not a smooth function of source
+position. It contains a complicated network of singular curves, known
+as caustics.  Figure~\ref{fig:magnification_map} shows part of a
+magnification map with a few caustics.  The lensed brightness would be
+given by placing the source on such a magnification map and
+integrating the surface brightness weighted by the magnification.
+Most astrophysical sources straddle several caustics, and hence, their
+net brightness varies smoothly with location.  The central engine of
+quasars, however, is smaller than the typical spacing between
+caustics.  As a result, the lensed brightness undergoes sudden changes
+as a quasar crosses a caustic.  The effect supplies an upper limit on
+the size of the central engine, and can also be used to study the mass
 distribution and kinematics of stars in the lensing galaxy as well
-\citep[e.g.,][]{2012ApJ...744..111P}.  Microlensing caustics have an
-additional remarkable feature: though they can be very complicated,
-they have some universal properties well-known from catastrophe
-theory.  In particular, very close to the simplest caustics (known as
-folds), the magnification is approximately constant on one side and
+\citep[e.g.,][]{2012ApJ...744..111P}.  Caustics have an additional
+remarkable feature: though they can be very complicated, they have
+some universal properties well-known from catastrophe theory.  In
+particular, very close to the simplest caustics (known as folds), the
+magnification is approximately constant on one side and
 $\propto1/\sqrt p$ where $p$ is the transverse distance of the source
@@
 
-%In Section~\ref{sec:source-models} we introduce the three source
-%profiles used in our subsequent models and simulations: a
-%constant-brightness disc, a circular Gaussian, and the crescent source
-%introduced by \citep{2013MNRAS.434..765K}.  The latter is simply a
-%constant-brightness disc with a smaller, non-concentric disc cut out
-%of it.  We also derive the half-light radius for a crescent.  The
-%half-light radius can characterise the source size for all three types
-%of source.
-
-In Section~\ref{sec:numerics} we carry out source fitting to
-lightcurves, with both noise and systematic errors are present.  We
-generate a mock lightcurves by taking a crescent source across the
-path AB in figure~\ref{fig:magnification_map}, and then adding noise.
-The path simulates crossing a clean but not ideal fold.  We then fit
-this lightcurve to templates from both crescent and circular-Gaussian
-sources across the path CB.  That is, the templates come from a
-similar but not identical caustic, thus deliberately generating a
-systematic error.  Despite this systematic error, a crescent source
-fits the data, while a Gaussian source is rejected from the $\chi^2$
-value.
+In Section~\ref{sec:source-models} we introduce the three source
+profiles used in our subsequent models and simulations: a
+constant-brightness disc, a circular Gaussian, and the crescent source
+introduced by \citep{2013MNRAS.434..765K}.  The latter is simply a
+constant-brightness disc with a smaller, non-concentric disc cut out
+of it.  We also derive the half-light radius for a crescent.  The
+half-light radius can characterise the source size for all three types
+of source.
 
@@
 have applied the idea to observed light curves to infer properties of
-quasar accretion discs.  A similar idea appears in
-\cite{2013ApJ...769..131C}.  In this work, we take a simpler but
-arguably more robust approach: we study features in the light curves
+quasar accretion discs.  In this work, we take a simpler but arguably
+more robust approach: we study features in the light curves
 characteristic of a crescent-like source which in turn would indicate
@@
 crescent is replaced by a source based on an accretion-disc simulation
-of a black-hole environment (Figure~\ref{fig:M87_plots}).
+of a black-hole environment
+(Figures~\ref{fig:M87_image}--\ref{fig:M87_plots}).
 
@@
 suited for acquiring the necessary microlensed light-curve.
-
 \section{Microlensing}\label{sec:microlensing}
@@
 $\mu(\vec\theta)$ is more or less than unity.  If there are multiple
-images at distinct but not observationally resolved $\vec\theta_i$ from
+mages at distinct but not observationally resolved $\vec\theta_i$ from
 the same $\vec\beta$, a total brightness amplification
-of
-
-\begin{equation} \mu_{\rm total} = \sum_{i} \mu(\vec\theta_i)
+of\begin{equation} \mu_{\rm total} = \sum_{i} \mu(\vec\theta_i)
 \end{equation}
@@
 Here the magnification of a point source near a caustic is equal to
-the sum of the magnification due to other reasons $\mu_0$ 
-(such as zero point magnification and the effect of other weaker and 
-distant lenses), assumed to
+the sum of the magnification due to other reasons $\mu_0$, assumed to
 be locally constant, and a decrease with the square root of the
@@
 
-\subsection{Rotationally symmetric source with a bivariate gaussian surface brightness distribution}\label{subsec:gaussian}
+\subsection{Rotationally symmetric source with a bivariate gaussian surface brightness distribution}
 
@@
 Instead we will
-use the difference which in this case is the total flux from the crescent-shaped source $S_0^C$. \\
+use the difference which in this case is th total flux from the crescent-shaped source $S_0^C$. \\
 Equation \ref{eqn:s2d} can be written as:\\
@@
 \frac{S_0^C}{2} & = \frac{2 S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{d_c}^{R_p} \bigg[ \sqrt{R_p^2 - p^2} \\ 
-        & - \sqrt{R_p^2 - \left(p-c\right)^2} \Theta \left(R_p^2 - \left(p-c\right)^2 \right) \bigg] \mathrm{d}p. 
+        & - \sqrt{R_p^2 - \left(p-c\right)^2} \Theta \left(R_p^2 - \left(p-c\right)^2 \right) \bigg] dp. 
 \end{aligned}
@@
 \begin{aligned}
-S_0^C &=  \frac{4S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{0}^{r_{1/2}} \left[ \Theta_2  - \Theta_1  \right] p \mathrm{d}p; \\
+S_0^C &=  \frac{4S_0^C}{\pi \left(R_p^2-R_n^2 \right)} \int_{0}^{r_{1/2}} \left[ \Theta_2  - \Theta_1  \right] p dp; \\
 \Theta_1 &= \bigg[ \pi - \arccos \frac{R_p^2 + p^2 -d_c^2}{2 R_p p} \\
@@
 \begin{figure}
-\includegraphics[width = .49\textwidth]{figures/figure_rhalf2.eps}
-\caption{\label{fig:geom_crescent} Geometry of crescent sources: the source can be seen as a superposition of a smaller dark disk over and inside a larger brighter disk.
-$R_p$ and $R_n$ mark the radii of the bright and dark disk.
-$d_c$ and $c$ mark the distance from the center of the bright disk to the centroid and from the dark disk to the centroid.
-}
+\includegraphics[width = .49\textwidth]{figures/figure_rhalf.eps}
+\caption{\label{fig:geom_crescent} Geometry of crescent sources.}
 \end{figure}
@@
     \includegraphics[width = 0.48\textwidth]{figures/4source_magnification.eps}
-\caption{\label{fig:lightcurve_gauss} 1D profiles (top panel) and simulated lightcurves (bottom) of crescent, Gaussian and disk sources with identical total flux ($S_0$) and half-light radius($r_{1/2}$).
-Crescent source has $r_{1/2}$ = 0.72754 $R_p$, $R_n$ = 0.4 $R_p$, a = 0.3/-0.3 $R_p$.
-}
+\caption{\label{fig:lightcurve_gauss} Lightcurves of crescents, Gaussian and disk for sources with identical $S_0$ and $r_{1/2}$.
+Crescent source has $r_{1/2}$ = 0.72754 $R_p$, $R_n$ = 0.4 $R_p$, a = 0.3/-0.3 $R_p$ }
 \end{figure}
@@
     \includegraphics[width = 0.48\textwidth]{figures/4avar_magnification.eps}
-\caption{\label{fig:a_var} 1D profiles (top panel) and simulated lightcurves (bottom) of crescent sources with identical $R_p$ and $R_n$ ($R_n$ = 0.4 $R_p$) and different disk centers displacement (a).}
+\caption{\label{fig:lightcurve_disk} Lightcurves of crescent sources with $R_n$ = 0.4 $R_p$.}
 \end{figure}
@@
     \includegraphics[width = 0.48\textwidth]{figures/5Rn_back_var_magnification.eps}
-\caption{\label{fig:lightcurve_crescent_back} 1D profiles (top panel) and simulated lightcurves (bottom) of backwards crescent sources with identical $R_p$ and $a$ ($a$ = 0.3 $R_p$) and different dark disk radii ($R_n$).}
+\caption{\label{fig:lightcurve_crescent_back} Lightcurves of backwards crescents with  a = 0.3 $R_p$.}
 \end{figure}
@@
 along the $p$ axis. Such a modelling of the motion of the object in the source plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
-choosing the function $a$ to be constant in time.
+choosing the function $a(t)$ to be constant in time.
 
@@
     \includegraphics[width = 0.48\textwidth]{figures/5Rn_forw_var_magnification.eps}
-\caption{\label{fig:lightcurve_crescent} 1D profiles (top panel) and simulated lightcurves (bottom) of forwards crescent sources with identical $R_p$ and $a$ ($a$ = - 0.3 $R_p$) and different dark disk radii ($R_n$).
-}
-\end{figure}
-
-In figure \ref{fig:char_points} There are four characteristic points visible on the resulting one-dimensional light profile of the crescent shaped source.
+\caption{\label{fig:lightcurve_crescent} Lightcurve of forwards crescents with a = -0.3 $R_p$.}
+\end{figure}
+
+There are four characteristic points visible on the resulting one-dimensional light profile of the crescent shaped source.
 Two of the points, p1 and p4, mark the outer boundaries of the luminous disc component.
@@
 All four points mark the positions where derivative $\frac{dS_{1D}}{dp}$ is discontinuous.
-The points can be used to define three regions: $p_1 - p_2$ where $S_{1D}$ is convex, $p_2 - p_3$ where $S_{1D}$ is concave, and $p_3 - p_4$ where $S_{1D}$ is convex again.
+The points can be used to define three regions: $p_1 - p_2$ where $S_1D$ is convex, $p_2 - p_3$ where $S_1D$ is concave, and $p_3 - p_4$ where $S_1D$ is convex again.
 \\
@@
 \begin{equation}
-    R_p = \frac{v_p \times  \left( t_4 -t_1 \right)}{2}, 
-\end{equation}
-
-\begin{equation}
-        R_n = \frac{v_p \times \left( t_3 -t_2 \right)}{2}, 
-\end{equation}
-
-\begin{equation}
-        a = \frac{v_p \times \left( t_4 +t_1 - t_3 - t_2 \right)}{2}. 
+    R_p = \frac{v_p \left( t_4 -t_1 \right)}{2}, 
+\end{equation}
+
+\begin{equation}
+        R_n = \frac{v_p \left( t_3 -t_2 \right)}{2}, 
+\end{equation}
+
+\begin{equation}
+        a = \frac{v_p \left( t_4 +t_1 - t_3 - t_2 \right)}{2}. 
 \end{equation}
@@
         \includegraphics[width = 0.48\textwidth]{figures/ch_instances.eps}
-\caption{\label{fig:char_points} The characteristic points and instances 
-of a crescent source's one dimensional profile and microlensing lightcurve.
-In the top panel a typical 1D profile  of a crescent source is presented. 
-There are two light gray regions that mark the regions of the bright disk 
-that are not overlapped by the dark disk. The dark gray band marks the 
-region where the two disks overlap. $p_1$ and $p_4$ mark the start and 
-end of the bright disk in the 
-1D representation. Analogous, $p_2$ and $p_3$ mark the start and 
-end of the dark disk.
-The bottom panel represents the associated lightcurve with the points 
-$t_i$ marking the moment when point $p_i$ overlaps with the caustic.
-The gray shades are analogous to the ones in the top panel.
- }
-\end{figure}
-
-Figure \ref{fig:lightcurve_gauss} reveals that the 
-lightcurve of a crescent source has more
-visible features than the other two light-curves corresponding to the
-disc and gaussian shape. There are three regions where $S_{1D}$ and by
-inheritance, the lightcurve has distinguishable behaviour. The first
-region that would be recorded on a lightcurve plot represents the
-period of time when the bright disk begins to be overlapped by the
-caustic and stops when just before the dark disk reaches the
-caustic. During this period of time, the flux of light from the source
-is increasingly magnified. The second period starts and ends with the
-overlapping of the dark disc. As a boundary of the two regions, there
-is a distinguishable point where the slope of the magnification is
-drastically changed. This apparent discontinuity in the first
-derivative of the magnification function is caused by the caustic
-amplification of the sudden drop in the $S_{1d}$ function. During the
-respective period, the magnification growth slows or even reverses
-during the first part of the period and starts to grow faster again as
-the dark disk ends its overlap with the caustic. At the point where
-the dark disc clears the caustic, the growth of the magnification is
-infinite, which appears as a saddle point on the lightcurve. Next, the
-final period corresponds to the case when the dark disc has cleared
-the caustic and the bright disc continues to overlap with the
-caustic. During this period, the lightcurve reaches a peak that for
-most of the parameter space is global and for the rest of the
-parameter space local.  At the end of the period, the growth of the
-magnification is negative infinite. the respective point appears as a
-second saddle point on the lightcurve. Past this point the
-magnification of any finite source will decay in roughly the same
-manner.
-
-The impact of parameter $a$ on the shape of the lightcurve can be
-observed in Figure~\ref{fig:a_var}. Sources where the center of the
-dark disc reaches the caustic before the center of the bright disc are
-characterized by a smoother broad peak in contrast to the cases where
-the center of the dark disc reaches the caustic after the bright disk.
-In the case of the latter the instance when the dark disc reaches the
-caustic corresponds to larger and larger magnifications until it
-becomes a local and even global peak. The effect of the $R_n$
-parameter on the light curve is presented in
-Figures~\ref{fig:lightcurve_crescent} and
-\ref{fig:lightcurve_crescent_back}. For the particular set of
-parameters where $R_p = R_n +a$ the third period of time discussed
-previously does not exist. Particularizing further, if the value of
-the radius of the dark disc is comparable to the value of the radius
-of the bright one then the position and shape of the maximum
-magnification are strikingly different. In case the crescent reaches
-the caustic with the bright region first, the peak magnification
-happens when the dark region reaches the caustic and it is
-characterized by a sharp variation in magnification growth, In the
-opposite case, the peak appears before the end of the bright disc
-reaches the caustic and shape is smoother.
-
-\begin{figure}
-\centering
-    \includegraphics[width = 0.48\textwidth]{figures/M87_shape.eps}
-        \includegraphics[width = 0.48\textwidth]{figures/M87_lc.eps}
-\caption{\label{fig:M87_plots} The one dimensional profile obtained
-  from the numerical integration along the ordonata axis of a simulated source
-  image is presented in the upper panel. The source image was presented in 
- Dexter et al 2012 in the top right-most panel of figure 5 
-(of the respective paper) as a result of the simulations 
-presented in McKinney \& Blandford 2009. The corresponding
-  lightcurve associated to the 1D profile is presented in the lower
-  panel. The points $\rm p_2$ and $\rm p_3$ mark the start and 
-  end of the equivalent
-  dark disk and are associated with the moments of their overlap with
-  the caustic: $\rm t_2$ and $\rm t_3$. $\rm h_{w}$ denotes the maximum width of the crescent.}
-\end{figure}
+\caption{\label{fig:char_points} The characteristic points and instances of a crescent source's one dimensional profile and microlensing lightcurve. }
+\end{figure}
+
+Figure 4 reveals that the lightcurve of a crescent source has more visible features than the other two light-curves
+corresponding to the disc and gaussian shape. There are three regions where $S_{1D}$ and by inheritance, the lightcurve
+ has distinguishable behaviour. The first region that would be recorded on a lightcurve plot represents the period
+of time when the bright disk begins to be overlapped by the caustic and stops when just before the dark disk reaches
+the caustic. During this period of time, the flux of light from the source is increasingly magnified. The second
+period starts and ends with the overlapping of the dark disc. As a boundary of the two regions, there is a
+distinguishable point where the slope of the magnification is drastically changed. This apparent discontinuity in the
+first derivative of the magnification function is caused by the caustic amplification of the sudden drop in the $S_{1d}$
+function. During the respective period, the magnification growth slows or even reverses during the first part of the
+period and starts to grow faster again as the dark disk ends its overlap with the caustic. At the point where
+the dark disc clears the caustic, the growth of the magnification is infinite, which appears as a saddle point
+on the lightcurve. Next, the final period corresponds to the case when the dark disc has cleared the caustic
+and the bright disc continues to overlap with the caustic. During this period, the lightcurve reaches a peak
+that for most of the parameter space is global and for the rest of the parameter space local.
+At the end of the period, the growth of the magnification is negative infinite. the respective point appears 
+as a second saddle point on the lightcurve. Past this point the magnification of any finite source will 
+decay in roughly the same manner. \\
+
+The impact of parameter $a$ on the shape of the lightcurve can be observed in figure 5 (a-var). Sources where 
+the center of the dark disc reaches the caustic before the center of the bright disc are characterized by a smoother
+broad peak in contrast to the cases where the center of the dark disc reaches the caustic after the bright disk.
+In the case of the latter the instance when the dark disc reaches the caustic corresponds to larger and
+larger magnifications until it becomes a local and even global peak. The effect of the $R_n$  parameter on the
+light curve is presented in figures 6 and 7. For the particular set of parameters where $R_p = R_n +a$ the third
+period of time discussed previously does not exist. Particularizing further, if the value of the radius of the dark
+disc is comparable to the value of the radius of the bright one then the position and shape of the maximum magnification
+ are strikingly different. In case the crescent reaches the caustic with the bright region first, the peak magnification
+ happens when the dark region reaches the caustic and it is characterized by a sharp variation in magnification growth,
+In the opposite case, the peak appears before the end of the bright disc reaches the caustic and shape is smoother. \\
 
@@
 
+\begin{figure}
+\centering
+    \includegraphics[width = 0.48\textwidth]{figures/M87_shape.eps}
+        \includegraphics[width = 0.48\textwidth]{figures/M87_lc.eps}
+\caption{\label{fig:M87_plots} The one dimensional profile obtained
+  from the numerical integration along the ordonata axis of the source
+  image is presented in the upper panel. The corresponding
+  lightcurve associated to the 1D profile is presented in the lower
+  panel. The points p2 and p3 mark the start and end of the equivalent
+  dark disk and are associated with the moments of their overlap with
+  the caustic: t2 and t3.}
+\end{figure}
+
 \section{Fitting mock data}\label{sec:numerics}
@@
 Having used the simple model of an ideal fold to gain insight, we now
-consider the question of whether the characteristics of a crescent
-source could be discerned in real data having noise and systematic
-error as well.  Our method is to generate microlensing lightcurves
-using a widely-used microlensing code
-\citep{1990PhDT.......180W,1999A&A...346L...5W,1999JCoAM.109..353W} to
-which we have added a small extension to provide the option of
-crescent sources, and then fit these mock lightcurves to source models
-using Markov-chain Monte Carlo (MCMC).
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.5\textwidth]{figures/cc_forward.eps}
-\caption{\label{fig:cc_forward} Mock lightcurve fitted with the
-  correct source type and caustic.  The upper panel shows the mock
-  data points, and running through them is a seven-parameter best fit:
-  these being the crescent parameters $(R_p,R_n,a,b)$ and additionally
-  the start and end times and the overall normalisation.  (Ideal
-  lightcurves for the same crescent parameters appear above in
-  Figures~\ref{fig:lightcurve_gauss}, \ref{fig:a_var},
-  \ref{fig:lightcurve_crescent} and \ref{fig:char_points}.)  The curve
-  has $\chi^2=243, P=0.32$ indicating a good fit.  The lower panel
-  shows the posterior distributions for $r_{1/2}$ (solid histogram)
-  and $R_n/R_p$ (dashed histogram).}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.5\textwidth]{figures/cc_backward.eps}
-\caption{\label{fig:cc_backward} Like Figure~\ref{fig:cc_backward},
-  except that the fit uses path~II whereas the mock data still come
-  from path~I. The curve has $\chi^2=284, P=0.01$ indicating a
-  marginally acceptable fit.}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.5\textwidth]{figures/gc_forward.eps}
-\caption{\label{fig:gc_forward} Like Figure~\ref{fig:cc_backward}, but
-  showing the best fit to a Gaussian source.  The histogram shows the
-  posterior distribution of $r_{1/2}$ (there is no $R_n$).  It appears
-  to show $r_{1/2}$ as accurately recovered, but is not to be trusted,
-  because the fit has $\chi^2=344, P=3\times10^{-6}$ indicating
-  rejection of the Gaussian source.  Looking closely, the curve is a
-  little too shallow near $t=3$, and it is unable to fit the second
-  sub-peak near $t=5$.}
-\end{figure}
-
-\subsection{Numerical microlensing lightcurves}
-
-The microlensing code uses a ray-shooting technique to compute the
-gravitational lensing effect of a mass distribution consisting of
-(a)~a smooth component and (b)~a random distribution of point masses
-representing stars.  The ray shooting maps a grid of $\vec\theta$ to
-$\vec\beta$ using the lens equation (\ref{eqn:lens}).  That is, the
-rays are shot from the observer back to the source.  For the
-computation of the individual deflection angles
+move to the numerical computation of microlensing light curves with
+realistic mass distributions, and then fit source models to
+lightcurves in the presence of random and systematic errors.
+
+First we generate microlensing magnification maps using the
+microlensing code created by Joachim Wambsgans (\citep{1999A&A...346L...5W, 1990PhDT.......180W, 1999JCoAM.109..353W}, 
+\cite{1990PhDT.......180W}, \cite{1999JCoAM.109..353W}). We then generate
+mock light curves for three source models --- a crescent, a uniform
+disc and a Gaussian disc --- and then fit each mock light curve to all
+three source models. Parameter fitting and marginalising was done by
+Markov-chain Monte Carlo (MCMC).
+
+\subsection{Numerical microlensing magnification maps}
+
+The code uses a ray-shooting technique to compute the gravitational
+lensing effect of a mass distribution consisting of (a)~a smooth
+component and (b)~a random distribution of point masses representing
+stars.  The ray shooting maps a grid of $\vec\theta$ to $\vec\beta$
+using the lens equation (\ref{eqn:lens}).  That is, the rays are shot
+from the observer back to the source.
+
+For the computation of the individual deflection angles
 $\vec\alpha(\vec\theta)$ a hierarchical or tree method is used.  The
@@
 moments.  Scaled units are used, with the constant pre-factor in the
-deflection angle (\ref{eqn:alpha}) separated out.  The result of ray
-shooting is a pixel map on the $\vec\beta$ plane of the number of
-lightrays which arrive at the source plane from a particular observer.
-This intermediate result is effectively a magnification map on the
-source plane.
+deflection angle (\ref{eqn:alpha}) separated out.
+
+The result of ray shooting is a pixel map on the $\vec\beta$ plane of
+the number of lightrays which arrive at the source plane from a
+particular observer.  This intermediate result is effectively a
+magnification map on the source plane.  Once the map is created, the
+lightcurve can be obtained by specifying the transit path of the
+source across the map.  At each point on this transit line, the code
+computes equation (\ref{eqn:ft2d}) convolving the brightness
+distribution of the source with the magnification pattern of the map.
+In real life, not only are both lens and source moving but the lens
+lens configuration, and with it the magnification pattern, is also
+changing with time.  While the first subtlety is taken care of by a
+coordinate transformation in this analysis, for the second one the
+lens configuration is assumed to be constant in time.
 
@@
 below, only two point masses were included.  This was done in order to
-have clean fold caustics.  There was no smooth component.
-
-Once the map is created, the lightcurve can be obtained by specifying
-the transit path of the source across the map.  The source is moved
-along the transit line, one pixel at a time, and at each step the
-brightness distribution of the source is convolved
-(equation~\ref{eqn:ft2d}) with the magnification map to give the
-brightness.  In real life, not only are both lens and source moving
-but the lens configuration, and with it the magnification
-pattern, is also changing with time.  While the first subtlety is
-taken care of by a coordinate transformation in this analysis, for the
-second one the lens configuration is assumed to be constant in time.
+have clean fold caustics.  For the computation of the actual
+lightcurve, the code was modified to also allow for crescent shaped
+images specified through the parameter set $R_p,R_n,a,b$. Here $R_p$
+denotes the outer radius of the crescent and $R_n$ the inner one. The
+orientation of the source image with respect to the magnification map
+is specified by the parameters $a$ and $b$ as the shift of the center
+of the inner disk from the center of the outer disk in $x$- and
+$y$-direction respectively. In the original version of the code,
+gaussian and disk-shaped images were already implemented. Those are
+completely characterised by the single parameter $R_p$. The values of
+the parameters are specified in pixel units corresponding to the
+magnification map. Further, one needs to specify the start and end
+point coordinates of the path, which the center of the source image
+follows through the magnification map (see the depiction in figure
+\ref{fig:magnification_map}). Hence, the points along this path are
+specified through the number of time steps for which the computation of
+the brightness is to be carried out. Those points correspond to the
+actual measurement of the brightness of an object in the observational
+case. For each timestep the two-dimensional convolution of equation
+\ref{eqn:ft2d} is carried out numerically for the position of the
+source on the magnification map.
 
@@
 caustic perpendicularly, and on a point where the border is a fold
-caustic.  We choose the line AC in Figure~\ref{fig:magnification_map}
-and consider its two caustic-crossing segments separately:
-\begin{itemize}
-\item ``path~I'' runs from C to B;
-\item ``path II'' runs from A to B.
-\end{itemize}
-The magnification along paths I and II are very similar, but not
-identical.
-
-The mock lightcurve used a crescent source (as explained in
-\S\ref{subsec:crescent}) with parameters
-\begin{equation}
-   \left(R_p, \frac{R_n}{R_p}, \frac{a}{R_p}, \frac{b}{R_p}\right) =
-   (1.31, 0.4, 0, 0.3)
+caustic.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\hsize]{figures/data_cc.eps}
+\includegraphics[width=0.9\hsize]{figures/data_dd.eps}
+\includegraphics[width=0.9\hsize]{figures/data_gg.eps}
+\caption{\label{fig:mockdata} Three different mock datasets with
+  three different parametrizations fitted to each --- crescent (red),
+  uniform-disk (blue) and Gaussian disk (green).}
+\end{figure}
+
+\subsection{Mock light curves}
+
+Mock light curves were generated by running a source across a
+caustic, specifically along the line from point C to point B in
+Figure~\ref{fig:magnification_map}.  The model light curves, to which
+these are fitted, are made by running a source across a different
+caustic, along the line from point A to point B in the same figure.
+That is, the mock data are generated with a clean but not ideal fold
+caustic and then fitted with another such caustic.  This mimics the
+unavoidable systematic error of not knowing the caustic exactly.
+
+The crescent source had parameters (as explained in
+\S\ref{subsec:crescent}) $R_p$ and $R_n$ being the radii of the outer
+and inner circles and $(\alpha,\beta)$ being the coordinates of the
+inner circle with respect to the centre of the outer circle in the
+coordinate system associated with the image in Figure 1 and not
+associated with the caustic surface as defined in the previous
+sections.  The parameter values were
+\begin{equation}
+   (R_p, R_n, \alpha, \beta) = (50.0, 30.0, 15.0, 10.0).
 \label{eqn:cp}
 \end{equation}
-corresponding to $r_{1/2}=1$.  Unit length has 30 pixels on the
-magnification map.  The source was carried for 250 pixels along part
-of path~I, giving 251 brightness values.  Gaussian noise at the level
-of 4\% of the current brightness was then added to give the mock
-lightcurve.  If we assume the brightness measurements are nightly, and
-the transverse speed is 600~km/s, $r_{1/2}$ comes to 24~light-minutes,
-which is plausible for a quasar.
-
-\subsection{Model-fitting results}
-
-We now proceed to fit the mock lightcurve using MCMC, to recover the
-source parameters.
-
-Note that fitting involves three nuisance parameters, namely the
-brightness normalisation and the start- and endpoints on the path.
-The latter two are equivalent to considering the location of the
-caustic with respect to the starting point and the speed of the
-source.  The fitting method automatically marginalises over all the
-nuisance parameters.
-
-Figure \ref{fig:cc_forward} shows a fit to the mock lightcurve and
-recovery of $r_{1/2}$ and $R_n/R_p$.
-
-Figure \ref{fig:cc_backward} shows the result when a different caustic
-is used to fit, than was used to generate the mock lightcurve.  This
-mimics the unavoidable systematic error of not knowing the caustic
-exactly.  The fit is slightly degraded, but still acceptable.
-
-Figure \ref{fig:gc_forward} shows the result of attempting to a
-Gaussian source.  There is a small but statistically very significant
-misfit because a Gaussian curve cannot reproduce the part of the
-lightcurve with the slight flattening and then a second sub-peak
-(cf.~Figure~\ref{fig:char_points}) as the dark part of the crescent
-crosses the caustic.
+The uniform disc had the same (outer) radius as the crescent.  In the
+Gaussian source, we set $3\sigma=50$.  For convenience, below we will
+refer to $R_p$ of the Gaussian source, by which we actually mean
+$3\sigma$.
+
+Each mock light curve also has three nuisance parameters, namely the
+beginning and end of the event and the brightness normalisation.
+These are to be marginalised out by the MCMC.
+
+Figure~\ref{fig:mockdata} shows the three light curves.  Each has 250
+points regularly spaced in time, with Gaussian noise at the level of
+10\% of the current brightness.  This amounts formally to a single
+summary data point with a signal to noise of
+$10\times\sqrt{250}\simeq160$.
+
+\subsection{Model fitting and likelihood analysis}
+
+Each of the three mock light curves was fitted to all three models.
+We denote the nine possible cases with letter pairs, with the first
+letter denoting the assumed model for the fitting procedure and 
+the second letter representing
+the source: thus CG means a {\em C\/}rescent model was fitted to mock
+data from a {\em G\/}aussian source, DC stands for a uniform {\em
+  D\/}isc model fitted to mock data from a {\em C\/}rescent source,
+and so on.
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\hsize]{figures/Rp4all.eps}
+\caption{\label{fig:mcmc} Posterior probability distribution for the
+  source size $R_p$.  The vertical line is the correct value. Same
+  color represents the same dataset whereas same line style
+  corresponds to same model fit.  The legend gives the reduced
+  $\chi^2$ of the best fit in each case.}
+\end{figure}
+
+Figure~\ref{fig:mcmc} shows the posterior probability distribution of
+$R_p$ for all nine cases, along with the minimum reduced $\chi^2$ for
+each case.  The area under each curve is unity. Note that the height
+of the curves are not the likelihood, they are probability densities
+in parameter space.
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\hsize]{figures/Rhalf_RnRp.eps}
+  \includegraphics[width=0.9\hsize]{figures/aRp_bRp.eps}
+\caption{\label{fig:crescentfit} The 2$\sigma$ contours (or error ellipses)
+  of the crescent model when fitting with the three different
+  datasets.}
+\end{figure}
+
+Let us first consider the three cases where a crescent model was
+fitted.  These are the solid curves in Figure~\ref{fig:mcmc}, with the
+colours of the curves indicating the source.  Meanwhile,
+Figure~\ref{fig:crescentfit} shows $2\sigma$ of the inferred parameter
+values.
+
+\begin{enumerate}
+
+\item[1) {\bf CC}:] solid red curve in Figure~\ref{fig:mcmc} and red
+  ellipses in Figure~\ref{fig:crescentfit}.  In this case, a light
+  curve from a crescent source was being fitted to a crescent model.
+  The fit gives reduced $\chi^2$ close to unity, as expected.  The
+  recovered parameter values are near or slightly outside the
+  $2\sigma$ ellipses.  This is expected since we added a small
+  systematic error, by using different caustics (though both clean
+  folds) for the mock data and the model fit.  The apparent degeneracy
+  between $a$ and $b$, seen in Figure~\ref{fig:crescentfit}, is also
+  expected, since only the distance of the crescent's small circle
+  from the caustic influences the magnification.
+
+\item[2) {\bf CD}:] solid blue curve and blue ellipses.  Here a
+  crescent model was fitted to a light curve from a uniform disc.  The
+  situation is formally a crescent with $R_n=0$ and $a,b$ arbitrary,
+  and this shows in the blue ellipses in the recovered parameter
+  values.  The redundant parameters $\alpha$ and $\beta$, in effect, allow the
+  model to partly fit the noise, and hence, the $\chi^2$ is somewhat
+  lower than in the previous case.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\hsize]{figures/burnin_cg.eps}
+\caption{\label{fig:burnin} The $\chi^2$ of each step of the full MCMC
+  for one case.}
+\end{figure}
+
+\item[3) {\bf CG}:] Solid green curve and green ellipses.
+  Additionally, Figure~\ref{fig:burnin} shows the progress of
+  reduction of $\chi^2$ in this case, where a crescent-source model is
+  used to fit data from a Gaussian source.  The parameter values are
+  seemingly tightly constrained, but the recovered $R_p$ is completely
+  wrong since the correct value is far from the green curve and green
+  ellipses.  The reduced $\chi^2$ is only 0.85, indicative of
+  over-fitting.  We can see what has happened from the red curve in
+  the bottom panel of Figure~\ref{fig:mockdata}.  The best-fit model
+  covers only a small part of the light curve.  That is, the fitting
+  procedure has exploited the nuisance parameters to find a good fit
+  to mainly noise.
+
+\end{enumerate}
+
+Next we consider the three cases where a disc model is fitted.  These
+correspond to the dashed curves in Figure~\ref{fig:mcmc}.  There is
+only one interesting parameter to fit, the disc radius $R_p$, and no
+equivalent of Figure~\ref{fig:crescentfit} is needed.
+
+\begin{enumerate}
+
+\item[4) {\bf DC}:] red dashed curve.  On fitting a disc model to a
+  crescent-source source, the recovered $R_p$ is incorrect, but the
+  reduced $\chi^2$ of 1.29 signals that the data reject the model.
+
+\item[5) {\bf DD}:] blue dashed curve.  When the correct model is
+  fitted to a disc-source, the best fit $\chi^2=0.91$ is good and
+  $R_p$ is recovered within the uncertainty estimate.
+
+\item[6) {\bf DG}:] green dashed curve.  On fitting a disc model to a
+  Gaussian-source model, the recovered $R_p$ is incorrect, but the
+  reduced $\chi^2$ gives no signal that something is amiss.  It
+  appears that a Gaussian source could mimic a uniform disc.
+
+\end{enumerate}
+
+Finally, we consider the three cases where a Gaussian model is fitted.
+These correspond to the dotted curves in Figure~\ref{fig:mcmc}.  Again,
+there is only one interesting parameter to fit, the Gaussian
+$3\sigma$-radius, which we have called $R_p$.
+
+\begin{enumerate}
+
+\item[7) {\bf GC}:] red dotted curve.  When a Gaussian-source model is
+  fitted to data from a crescent source, the recovered $R_p$ is wrong
+  but the reduced $\chi^2=1.29$ shows the data rejecting the model.
+
+\item[8) {\bf GD}:] blue dotted curve.  When a Gaussian-source model
+  is fitted to data from a uniform disc, the recovered $R_p$ is wrong
+  but the model is rejected anyway.
+
+\item[9) {\bf GG}:] green dotted curve.  When data from a Gaussian is
+  fitted to the correct model, $R_p$ is recovered within it's estimated
+  uncertainty and the best fit reduced $\chi^2$ is close to unity.
+
+\end{enumerate}
+
+The above results suggest the following strategy for fitting a
+lightcurve from an unknown source profile: first try a Gaussian-source
+model; if the data reject that model, try a uniform disc; if the
+uniform disc is also rejected by the data, try a crescent model.  Our
+numerical experiments indicate --- assuming one of the three sources
+models are correct --- that the reduced $\chi^2$ would unmask the
+correct model, and its parameters would be correctly recovered.
 
@@
  radius of the studied source. 
-%In particular cases in which the previously mentioned approximation loses its validity, the impact on the quality of the 
-%lightcurves is not evenly distributed. The shape of the lightcurve will be maintained. 
-%The data points corresponding to the source position before and during the overlapping of the caustic will be affected by
-% smaller errors than the data points corresponding to later times. \\
+In particular cases in which the previously mentioned approximation loses its validity, the impact on the quality of the 
+lightcurves is not evenly distributed. The shape of the lightcurve will be maintained. 
+The data points corresponding to the source position before and during the overlapping of the caustic will be affected by
+ smaller errors than the data points corresponding to later times. \\
 
 The first two source profiles that we consider are the uniform disk and symmetric Gaussian source. 
-Both of them can be described by a half-light radius ($r_{1/2}$) and a total unlensed light flux ($S_0$). 
-With the two parameters constrained the one-dimensional profiles, as well as the lightcurves of the 
-two source, are completely determined, since no free parameter remains.
+Both of them can be described by a half-light radius $r_{1/2}$ and a total unlensed light flux $S_0$. 
+With the two parameters constrained the one-dimensional profiles, as well as the lightcurves of the two source, are completely determined, since no free parameter remains.
 
@@
 In the case of the crescent source there are in total five parameters: the integrated flux of the source $S_0$,
- the radii of the bright and dark disks and the displacement of the centers of the two disks on the axes perpendicular and parallel to the caustic.
-
-%Two of the parameters can be reduced by expressing the results in 
-%terms of $S_0$ and $r_{1/2}$. 
-%The later being determinable for any set of parameters 
-%$R_p, R_n$ and $a^2+b^2$. Moreover, one of the displacement 
-%parameters $b$ has no impact on the one-dimensional profile 
-%of the source that results from the projection of the source
-% image on an axis perpendicular to the caustic.
-
-Since the one-dimensional source profile that corresponds to an 
-axis perpendicular to the caustic contains exhaustively all the 
-information regarding the source that can be revealed by the lightcurve, 
-the value of the crescent's disks centers displacement parallel to the caustic 
-does not have an effect on the shape of the lightcurve.
-
-Nevertheless, the respective displacement is relevant for the calculation of
-of the half-light radius.  It's qualitative effect is to 
-decrease the value of the
-half-light radius when the absolute value of the parameter is
-increased.  In contrast to the simpler source shapes discussed before the
-crescent has two additional relevant parameters: the radius of the dark 
-disk ($R_n$) and the center displacement in the direction
- perpendicular to the caustic $a$.
-Figure 4 reveals that the lightcurve of a crescent source has more
-visible features than the other two light-curves corresponding to the
-disc and Gaussian shape. The parameters $R_n$ and $a$ have strong
-influences on the shape of the microlensed lightcurve as can be seen
-in Figures~\ref{fig:a_var}, \ref{fig:lightcurve_crescent} and
-\ref{fig:lightcurve_crescent_back}.  Moreover, the one-dimensional
-source profile corresponding to the direction perpendicular to the
-caustic reveal four characteristic points. The overlap of each of
-these points with the caustic leaves visible features on the
-lightcurve at the corresponding instances of time. In timely order,
-the instances correspond to the start of the overlap between the
-caustic and the bright disk, the start of the overlap between the
-caustic and the dark disk, the end of the overlap between the caustic
-and dark disk and finally the end of the overlap between the caustic
-and the bright disk.
-
-With the different source profiles and their corresponding lightcurves 
-studied we can change our point of view of the system to that of an observer. 
-The observer would basically detect only the lightcurve of such a source.
- As described in section 4.3 the timing of the onset and 
-offset of the previously 
-described periods can be used to estimate the values of the radii 
-and one of the displacement parameters when assuming a geometric crescent. 
-All quantities can be estimated in terms of the relative velocity 
-of the source in a direction perpendicular to the caustic.
-
-Furthermore, a simulated image of M87 presented in
- \citep{2012MNRAS.421.1517D} has been microlensed (figure 9). 
-On the resulting lightcurve the instances 
-corresponding to the start and end of the black hole 
-shadow and caustic overlap were distinguishable.
-
-%In the case of a high-quality lightcurve with 
-%insignificant noise and measurement errors, 
-%the parameters can be obtained 
-%by simply identifying the characteristic 
-%instances of time without making use of 
-%the actual values of the magnification
-% map. If the effect of the errors and the 
-%noise distorts the magnification time function enough so that the 
-%characteristic epochs are not identifiable with the characteristic 
-%periods still visible, the boundaries of the periods 
-%can be roughly estimated. Furthermore, if direct estimates of 
-%the parameters cannot be obtained we propose the use of a strong
-%statistical tool such as Markov-Chain Monte Carlo.
-% In Section 6  we have studied the possibility of identifying a crescent 
-%source and the possibility of recovering the 
-%respective parameters. As magnification map, we have used a complex numerical 
-%one generated with the microlensing code by \cite{1999A&A...346L...5W}. 
-%In addition, we have added to the signal a Gaussian 
-%noise with an SNR of 1.6.  In the experiment,
-% we have considered all nine combinations of
-% original source profiles and assumed fitting source models.
-%Effectively we have fitted using MCMC all three sources with all three assumed fitting models. The results of the experiment allowed us to build a 
-%procedure for distinguishing the shape of the source assuming that one of the three models we have considered is a good approximation.  
-%The procedure would be useful for observers that endeavour to gain more information about an unresolved source for which they can study 
-%the microlensing lightcurve.  As a first step in the procedure, one should first attempt to use MCMC with a Gaussian model assumption. 
-%If either the value of the $\chi^2$ or the number of rejected datapoints is large then the next step is to change the assumed source model to 
-%a uniform disc and redo the fitting. Finally, if the uniform disc assumption is rejected as well then the fitting should be done with a 
-%crescent source model assumption. If at each of the three steps the data rejects the model then the source cannot be approximated by any of the three models. 
-%Otherwise, if at one of the steps the data does not reject the model then the respective model is a good approximation.
-
-In the case of a high-quality lightcurve, the parameters of a crescent
-source can be estimated by simply identifying signatures on the
-lightcurve corresponding to transits of the particular source features
-across the caustic.  This simple idea can be supplemented with
-statistics.  We do so in Section~\ref{sec:numerics}, where we take a
-realistic magnification map generated by the microlensing code
-from \cite{1999A&A...346L...5W}, compute a caustic-crossing lightcurve
-from it, and then fit the lightcurve with other caustic-crossing
-templates.  The conclusion is that a crescent source can be
-distinguised from a circular source, on the basis of caustic-crossing
-lightcurves, if the caustic is a clean but not ideal fold.
+ the radii of the bright/dark disk $R_p$/$R_n$ and the displacement of the centers of the two disks on the axes perpendicular and parallel to the caustic $a$ and $b$.
+
+Two of the parameters can be reduced by expressing the results in terms of $S_0$ and $r_{1/2}$. 
+The later being determinable for any set of parameters $R_p, R_n$ and $a^2+b^2$. Moreover, one of the displacement 
+parameters $b$ has no impact on the one-dimensional profile of the source that results from the projection of the source image on an axis perpendicular to the caustic.
+
+Since the one-dimensional source profile that corresponds to an axis perpendicular to the caustic contains exhaustively all the information regarding the source that can be revealed by the lightcurve,
+the value of the parameter $b$ does not have an effect on the shape of the lightcurve.
+
+Nevertheless, the $b$ parameter is relevant for the calculation of $r_{1/2}$. 
+It's qualitative effect is to decrease the value of the half-light radius when the absolute value of the parameter is increased.  
+With two parameters constrained and another irrelevant to the shape of the lightcurve, two free parameters remain $R_n$ and $a$. 
+Figure 4 reveals that the lightcurve of a crescent source has more visible features than the other two light-curves 
+corresponding to the disc and Gaussian shape. The parameters $R_n$ and $a$ have strong influences on the shape of 
+the microlensed lightcurve as can be seen in figures 5, 6 and 7.
+Moreover, the one-dimensional source profile corresponding to the direction perpendicular to the caustic reveal four characteristic points.
+The overlap of each of these points with the caustic leaves visible features on the lightcurve at the corresponding instances of time.
+In timely order, the instances correspond to the start of the overlap between the caustic and the bright disk, the start of the overlap between the caustic and the dark disk, the end of the overlap between the caustic and dark disk and finally the end of the overlap between the caustic and the bright disk.
+
+With the different source profiles and their corresponding lightcurves studied we can change our point of view of the system to that of an observer. 
+The observer would basically detect only the lightcurve of such a source. As described in section 4.3 the timing of the onset and offset of the previously 
+described periods can be used to estimate the values of the radii and one of the displacement parameters when assuming a geometric crescent. 
+All quantities can be estimated in terms of the relative velocity of the source in a direction perpendicular to the caustic.
+
+Furthermore, a simulated image of M87 presented in \citep{2012MNRAS.421.1517D} has been microlensed (figure 9). On the resulting lightcurve the instances 
+corresponding to the start and end of the black hole shadow and caustic overlap were distinguishable.
+
+In the case of a high-quality lightcurve with insignificant noise and measurement errors, the parameters can be obtained 
+by simply identifying the characteristic instances of time without making use of the actual values of the magnification
+ map. If the effect of the errors and the noise distorts the magnification time function enough so that the 
+characteristic epochs are not identifiable with the characteristic periods still visible, the boundaries of the periods 
+can be roughly estimated. Furthermore, if direct estimates of the parameters cannot be obtained we propose the use of a strong
+statistical tool such as Markov-Chain Monte Carlo. In Section 6  we have studied the possibility of identifying a crescent 
+source and  the possibility of recovering the respective parameters. As magnification map, we have used a complex numerical 
+one generated with the microlensing code by \cite{1999A&A...346L...5W}. In addition, we have added to the signal a Gaussian 
+noise with an SNR of 1.6.  In the experiment, we have considered all nine combinations of original source profiles and assumed fitting source models.
+Effectively we have fitted using MCMC all three sources with all three assumed fitting models. The results of the experiment allowed us to build a 
+procedure for distinguishing the shape of the source assuming that one of the three models we have considered is a good approximation.  
+The procedure would be useful for observers that endeavour to gain more information about an unresolved source for which they can study 
+the microlensing lightcurve.  As a first step in the procedure, one should first attempt to use MCMC with a Gaussian model assumption. 
+If either the value of the $\chi^2$ or the number of rejected datapoints is large then the next step is to change the assumed source model to 
+a uniform disc and redo the fitting. Finally, if the uniform disc assumption is rejected as well then the fitting should be done with a 
+crescent source model assumption. If at each of the three steps the data rejects the model then the source cannot be approximated by any of the three models. 
+Otherwise, if at one of the steps the data does not reject the model then the respective model is a good approximation.
 
@@
 magnitude or even better. This would require the study of the dynamics
-of the stellar structure which contains the gravitational lens and the bulk
-transverse velocity of the galaxy with respect to the observer.  A
+of the stellar structure which contains the gravitational lens.  A
 better estimate of the relative velocity would facilitate a better
@@
 
-The parameters whose values cannot be determined due to the 
-loss of information from the directions parallel to the 
-caustic could be obtained in the eventuality in which 
-the same source crosses multiple caustics that are not parallel. 
-Multiple crossing of caustics can reveal details of 
-the one-dimensional flux profile corresponding to multiple distinct 
-directions which would allow the reconstruction of the 
-two-dimensional profile analogous to the process through 
-which an image of a CT scan is obtained.
+The parameters whose values cannot be determined due to the loss of information from the directions parallel to the 
+caustic could be obtained in the eventuality in which the same source crosses multiple caustics that are not parallel. 
+Multiple crossing of caustics can reveal details of the one-dimensional flux profile corresponding to multiple distinct 
+directions which would allow the reconstruction of the two-dimensional profile analogous to the process through which an image of a CT scan is obtained.
 
@@
 J.W. would like to acknowledge and thank the Pauli Center for Theoretical Studies of ETH Zurich and University of Zurich for 
-generous support during the Schr\"odinger visiting professorship in 2013.
+generous support during the Schrödinger visiting professorship in 2013.
 

### 2016-11-23 paper1/msold2.tex
template for response2

--- 2016-10-24 paper1/ms.tex
@@
 brightness changes.  This can occur, even for constant intrinic
-luminosity, as a consequence of two things: the very small size of the central engine,
-and granularity of the mass distribution of a lensing galaxy due to
-stars.  The latter means that the local magnification is not a smooth
-function of source position. It contains a complicated network of
-singular curves, known as caustics.
+luminosity, as a consequence of two things: the very small size of the
+central engine, and granularity of the mass distribution of a lensing
+galaxy due to stars.  The latter means that the local magnification is
+not a smooth function of source position. It contains a complicated
+network of singular curves, known as caustics.
 Figure~\ref{fig:magnification_map} shows part of a magnification map
@@
 continuum \citep[e.g.,][]{2012A&A...544A..62S} but is also known in
-X-rays \citep[e.g.,][]{2016arXiv160207601N} and has been argued for at
-cm wavelengths \citep{2000A&A...358..793K}.  The effect supplies an
-upper limit on the size of the central engine
+X-rays \citep[e.g.,][]{2016arXiv160207601N}.  In principle, it exists
+at all wavelengths \citep[and has been argued for at cm wavelengths
+  by][]{2000A&A...358..793K} but will be washed out if the light is
+from larger regions than the central engine of the quasar.  The effect
+thus supplies an upper limit on the size of the central engine
 \citep[e.g.,][]{2015ApJ...814L..26M} and can be used to study the mass
@@
 \centering
-    \includegraphics[width = 0.48\textwidth]{figures/S1D_all.eps}
-    \includegraphics[width = 0.48\textwidth]{figures/4source_magnification.eps}
-\caption{\label{fig:lightcurve_gauss} 1D profiles (top panel) and simulated lightcurves (bottom) of crescent, Gaussian and disk sources with identical total flux ($S_0$) and half-light radius($r_{1/2}$).
-Crescent source has $r_{1/2}$ = 0.72754 $R_p$, $R_n$ = 0.4 $R_p$, a = 0.3/-0.3 $R_p$.
-}
-\end{figure}
-
-\subsection{Lightcurve of the disk shaped source}
-
-Analogous to the gaussian shaped source, the disk source with uniform brightness, radius $R$ and 
-unmagnified flux $S_0^D$ has a lightcurve described by the equation:
-
-\begin{equation}
-\begin{aligned}
- F^D(t) &= \int_{-\infty}^\infty  \left( \mu_0 + \frac{C_0}{\sqrt{p}} \Theta \left( p \right) \right) \\
-    & \bigg[ \frac{2 S_0^D}{ \pi R} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}}  \Theta \left(R^2 - \left(p-p_s(t) \right)^2 \right) \bigg] \mathrm{d}p.
-\end{aligned}
-\end{equation}
-
-which is equivalent to:
-\begin{equation}
- F^D(t) = \mu_0 S_0^D + \frac{2 C_0 S_0^D}{\pi R} \int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \frac{1}{\sqrt{p}} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}} \mathrm{d}p.
-\end{equation}
-
-\begin{figure}
-\centering
-    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_a.eps}
-    \includegraphics[width = 0.48\textwidth]{figures/4avar_magnification.eps}
-\caption{\label{fig:a_var} 1D profiles (top panel) and simulated lightcurves 
-(bottom) of crescent sources with identical $R_p$ and $R_n$ ($R_n$ = 0.4 $R_p$) and different disk centers displacement (a).}
-\end{figure}
-
-\subsection{Lightcurve of the crescent shaped source}
-
-The lightcurve of a crescent shaped source with unamplified flux $S_0^C$, radii $R_p$, $R_n$ and center displacement $a(t)$ is:
-
-\begin{equation}
-\begin{aligned}
- F^c(t) &= \mu_0 S_0^C + C_0 \frac{2 S_0^C}{\pi \left( R_p^2 -R_n^2 \right) } \\
-    &\bigg[ \int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) \right)^2 }{p}} \mathrm{d}p \\
-    &  -  \int_{max(0, p_s(t) - a(t) - R)}^{max(0, p_s(t) -a(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) +a(t) \right)^2 }{p}} \mathrm{d}p  \bigg] .
-\end{aligned}
-\end{equation}
-
-\begin{figure}
-\centering
-    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_poz.eps}
-    \includegraphics[width = 0.48\textwidth]{figures/5Rn_back_var_magnification.eps}
-\caption{\label{fig:lightcurve_crescent_back} 1D profiles (top panel) and simulated lightcurves (bottom) of backwards crescent sources with identical $R_p$ and $a$ ($a$ = 0.3 $R_p$) and different dark disk radii ($R_n$).}
-\end{figure}
-
-The function $p_s(t)$ can be chosen to be equal to $v_p(t-t_0) + p_{s0}$. Where $p_{s0}$ is the coordinate $p$ 
-of the source at the initial time, and $v_p$ is the component of the velocity
-along the $p$ axis. Such a modelling of the motion of the object in the source 
-plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
-choosing the function $a$ to be constant in time.
-
-\begin{figure}
-\centering
-    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_neg.eps}
-    \includegraphics[width = 0.48\textwidth]{figures/5Rn_forw_var_magnification.eps}
-\caption{\label{fig:lightcurve_crescent} 1D profiles (top panel) and simulated lightcurves (bottom) 
-of forwards crescent sources with identical $R_p$ and $a$ ($a$ = - 0.3 $R_p$) and different dark disk radii ($R_n$). }
-\end{figure}
-
-In figure \ref{fig:char_points} there are four characteristic points visible on the resulting one-dimensional 
-light profile of the crescent shaped source. Two of the points, p1 and p4, mark the outer boundaries of the
- luminous disc component. The other two, p2 and p3, mark the boundaries of the dark disc component. 
-Since $S_{1D}$ is a projection of $S_{2D}$ on a line perpendicular to the caustic, the following relation holds:
-\begin{equation}
-    p_2-p_1 = R_p -R_n - a.
-\end{equation}
-In addition there are two other obvious relations the characteristic points which are independent of the projection:
-\begin{equation}
-    p_4 -p_1 = 2 R_p,
-\end{equation}
-
-\begin{equation}
-        p_3 -p_2 = 2 R_n.
-\end{equation}
-All four points mark the positions where derivative $\frac{dS_{1D}}{dp}$ is discontinuous. The points can be used 
-to define three regions: $p_1 - p_2$ where $S_{1D}$ is convex, $p_2 - p_3$ where $S_{1D}$ is concave, 
-and $p_3 - p_4$ where $S_{1D}$ is convex again. \\
-
-Due to the nature of the caustic and the monotonic behaviour of the magnification map on both sides of the caustic, the previously 
-mentioned characteristic points are inherited by the microlensing lightcurve. The points on the temporal dimension 
-$t_1, t_2, t_3$ and $t_4$ correspond to instances in time when the fold is aligned with $p_1, p_2, p_3$ and $p_4$, respectively. 
-For a constant relative velocity $v_p$ between the source and the caustic there is a simple relation between the points $p_i$ and instances $t_i$:
-
-\begin{equation}
-    t_j - t_i = \frac{p_j - p_i}{v_p}.
-\end{equation}
-
-With the use of the previous four equations the following identities can be written:
-
-\begin{equation}
-    R_p = \frac{v_p \times  \left( t_4 -t_1 \right)}{2}, 
-\end{equation}
-
-\begin{equation}
-        R_n = \frac{v_p \times \left( t_3 -t_2 \right)}{2}, 
-\end{equation}
-
-\begin{equation}
-        a = \frac{v_p \times \left( t_4 +t_1 - t_3 - t_2 \right)}{2}. 
-\end{equation}
-
-\begin{figure}
-\centering
         \includegraphics[width = 0.48\textwidth]{figures/ch_points.eps}
@@
 \end{figure}
+
+\begin{figure}
+\centering
+    \includegraphics[width = 0.48\textwidth]{figures/S1D_all.eps}
+    \includegraphics[width = 0.48\textwidth]{figures/4source_magnification.eps}
+\caption{\label{fig:lightcurve_gauss} 1D profiles (top panel) and simulated lightcurves (bottom) of crescent, Gaussian and disk sources with identical total flux ($S_0$) and half-light radius($r_{1/2}$).
+Crescent source has $r_{1/2}$ = 0.72754 $R_p$, $R_n$ = 0.4 $R_p$, a = 0.3/-0.3 $R_p$.
+}
+\end{figure}
+
+\subsection{Lightcurve of the disk shaped source}
+
+Analogous to the gaussian shaped source, the disk source with uniform brightness, radius $R$ and 
+unmagnified flux $S_0^D$ has a lightcurve described by the equation:
+
+\begin{equation}
+\begin{aligned}
+ F^D(t) &= \int_{-\infty}^\infty  \left( \mu_0 + \frac{C_0}{\sqrt{p}} \Theta \left( p \right) \right) \\
+    & \bigg[ \frac{2 S_0^D}{ \pi R} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}}  \Theta \left(R^2 - \left(p-p_s(t) \right)^2 \right) \bigg] \mathrm{d}p.
+\end{aligned}
+\end{equation}
+
+which is equivalent to:
+\begin{equation}
+ F^D(t) = \mu_0 S_0^D + \frac{2 C_0 S_0^D}{\pi R} \int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \frac{1}{\sqrt{p}} \sqrt{1 - \frac{\left( p-p_s(t) \right)^2}{R^2}} \mathrm{d}p.
+\end{equation}
+
+\begin{figure}
+\centering
+    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_a.eps}
+    \includegraphics[width = 0.48\textwidth]{figures/4avar_magnification.eps}
+\caption{\label{fig:a_var} 1D profiles (top panel) and simulated lightcurves 
+(bottom) of crescent sources with identical $R_p$ and $R_n$ ($R_n$ = 0.4 $R_p$) and different disk centers displacement (a).}
+\end{figure}
+
+\subsection{Lightcurve of the crescent shaped source}
+
+The lightcurve of a crescent shaped source with unamplified flux $S_0^C$, radii $R_p$, $R_n$ and center displacement $a(t)$ is:
+
+\begin{equation}
+\begin{aligned}
+ F^c(t) &= \mu_0 S_0^C + C_0 \frac{2 S_0^C}{\pi \left( R_p^2 -R_n^2 \right) } \\
+    &\bigg[ \int_{max(0, p_s(t) - R)}^{max(0, p_s(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) \right)^2 }{p}} \mathrm{d}p \\
+    &  -  \int_{max(0, p_s(t) - a(t) - R)}^{max(0, p_s(t) -a(t) + R)} \sqrt{\frac{R_p^2 - \left( p-p_s(t) +a(t) \right)^2 }{p}} \mathrm{d}p  \bigg] .
+\end{aligned}
+\end{equation}
+
+\begin{figure}
+\centering
+    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_poz.eps}
+    \includegraphics[width = 0.48\textwidth]{figures/5Rn_back_var_magnification.eps}
+\caption{\label{fig:lightcurve_crescent_back} 1D profiles (top panel) and simulated lightcurves (bottom) of backwards crescent sources with identical $R_p$ and $a$ ($a$ = 0.3 $R_p$) and different dark disk radii ($R_n$).}
+\end{figure}
+
+The function $p_s(t)$ can be chosen to be equal to $v_p(t-t_0) + p_{s0}$. Where $p_{s0}$ is the coordinate $p$ 
+of the source at the initial time, and $v_p$ is the component of the velocity
+along the $p$ axis. Such a modelling of the motion of the object in the source 
+plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
+choosing the function $a$ to be constant in time.
+
+\begin{figure}
+\centering
+    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_neg.eps}
+    \includegraphics[width = 0.48\textwidth]{figures/5Rn_forw_var_magnification.eps}
+\caption{\label{fig:lightcurve_crescent} 1D profiles (top panel) and simulated lightcurves (bottom) 
+of forwards crescent sources with identical $R_p$ and $a$ ($a$ = - 0.3 $R_p$) and different dark disk radii ($R_n$). }
+\end{figure}
+
+In figure \ref{fig:char_points} there are four characteristic points visible on the resulting one-dimensional 
+light profile of the crescent shaped source. Two of the points, p1 and p4, mark the outer boundaries of the
+ luminous disc component. The other two, p2 and p3, mark the boundaries of the dark disc component. 
+Since $S_{1D}$ is a projection of $S_{2D}$ on a line perpendicular to the caustic, the following relation holds:
+\begin{equation}
+    p_2-p_1 = R_p -R_n - a.
+\end{equation}
+In addition there are two other obvious relations the characteristic points which are independent of the projection:
+\begin{equation}
+    p_4 -p_1 = 2 R_p,
+\end{equation}
+
+\begin{equation}
+        p_3 -p_2 = 2 R_n.
+\end{equation}
+All four points mark the positions where derivative $\frac{dS_{1D}}{dp}$ is discontinuous. The points can be used 
+to define three regions: $p_1 - p_2$ where $S_{1D}$ is convex, $p_2 - p_3$ where $S_{1D}$ is concave, 
+and $p_3 - p_4$ where $S_{1D}$ is convex again. \\
+
+Due to the nature of the caustic and the monotonic behaviour of the magnification map on both sides of the caustic, the previously 
+mentioned characteristic points are inherited by the microlensing lightcurve. The points on the temporal dimension 
+$t_1, t_2, t_3$ and $t_4$ correspond to instances in time when the fold is aligned with $p_1, p_2, p_3$ and $p_4$, respectively. 
+For a constant relative velocity $v_p$ between the source and the caustic there is a simple relation between the points $p_i$ and instances $t_i$:
+
+\begin{equation}
+    t_j - t_i = \frac{p_j - p_i}{v_p}.
+\end{equation}
+
+With the use of the previous four equations the following identities can be written:
+
+\begin{equation}
+    R_p = \frac{v_p \times  \left( t_4 -t_1 \right)}{2}, 
+\end{equation}
+
+\begin{equation}
+        R_n = \frac{v_p \times \left( t_3 -t_2 \right)}{2}, 
+\end{equation}
+
+\begin{equation}
+        a = \frac{v_p \times \left( t_4 +t_1 - t_3 - t_2 \right)}{2}. 
+\end{equation}
 
@@
 lightcurve with the slight flattening and then a second sub-peak
-(cf.~Figure~\ref{fig:char_points}) as the dark part of the crescent
-crosses the caustic.
+(cf.~Figure~\ref{fig:char_points}) as the dark part of the crescent crosses the caustic.  
+A hint of
+such a feature is present in recent caustic-crossing observations of
+the well-known system Q2237+0305 \citep[see Figure~3
+in][]{2015ApJ...814L..26M} though interpretation as a crescent source
+is speculation at present.
 

### 2017-05-10 paper1/ms.tex
figure movement
Prasenjit's add to Section 5
responding to comments 2

@@
   \ref{fig:lightcurve_crescent} and \ref{fig:char_points}.)  The curve
-  has $\chi^2=243, P=0.32$ indicating a good fit.  The lower panel
-  shows the posterior distributions for $r_{1/2}$ (solid histogram)
-  and $R_n/R_p$ (dashed histogram).}
+  has $\chi^2_{(244)}=243, P=0.32$ indicating a good fit.  (The
+  subscript 244 refers to degrees of freedom: 251~points less
+  7~parameters.)  The lower panel shows the posterior distributions
+  for $r_{1/2}$ (solid histogram) and $R_n/R_p$ (dashed histogram).
+  Vertical dashes mark the correct values.}
 \end{figure}
@@
   except that the fit uses path~II whereas the mock data still come
-  from path~I. The curve has $\chi^2=284, P=0.01$ indicating a
+  from path~I. The curve has $\chi^2_{(244)}=284, P=0.01$ indicating a
   marginally acceptable fit.}
@@
   to show $r_{1/2}$ as accurately recovered, but is not to be trusted,
-  because the fit has $\chi^2=344, P=3\times10^{-6}$ indicating
+  because the fit has $\chi^2_{(247)}=344, P=3\times10^{-6}$ indicating
   rejection of the Gaussian source.  Looking closely, the curve is a
@@
 the magnification map and corresponding lightcurves used in the MCMC analysis performed by Irshad Mohammed in the last 
-part of the presented work. Both Manuel Rabold and Irshad Mohammed had large contributions in writing the "Fitting Mock Data" section.
+part of the presented work. Both Manuel Rabold and Irshad Mohammed had large contributions in writing the ``Fitting Mock Data'' section.
 

### 2017-05-10 fitting/summaryfig.py

@@ -9,8 +9,9 @@ from rhalf import Rhalf
 def mockdata():
     global pars
     lcurv.readmap()
-    Rp = 39.4
+    Rp = 39.4              # scaled by 30
     Rn,a,b = 0.4, 0, 0.3
+    print('r_half = ',Rhalf(1,Rn,a,b)*Rp/30)
     pars = [ 51, 290, 1.0, Rp,  Rn, a, b ]
     pars = np.array(pars)
     print(pars)
@@ -78,10 +79,12 @@ def plothist(p):
     rmin,rmax,nr = 0,1.2,61
     p.hist(rh,np.linspace(rmin,rmax,nr),weights=wt,
            histtype='step',lw=2,color='black')
+    p.plot([1,1],[.43,.48],lw=2,color='black')
     fl = max(rns)
     if fl > 0:
         p.hist(rns,np.linspace(rmin,rmax,nr),weights=wt,
                histtype='step',lw=2,color='black',linestyle='dashed')
+        p.plot([.4,.4],[.43,.48],lw=2,color='black',linestyle='dashed')
         p.set_xlabel('inferred $R_n/R_p$ and $r_{1/2}$',labelpad=10)
     else:
         p.set_xlabel('inferred $r_{1/2}$',labelpad=10, fontsize=16)

### 2017-06-12 paper1/ms.tex
edits by Mihai

@@
 continuum.  In the direction perpendicular to the accretion disc,
-where the medium is more transparent, the outflow tends to be
+where the medium is more transparent, the outflowing material tends to be
 collimated to jets.  If the orientation is such that the observer's
@@
 lensed brightness undergoes sudden changes as a quasar crosses a
-caustic.  Quasar microlensing is mostly observed in the optical
-continuum \citep[e.g.,][]{2012A&A...544A..62S} but is also known in
+caustic.  Quasar microlensing has been first observed in the optical
+range \citep[e.g.,][]{2012A&A...544A..62S} but is also known in
 X-rays \citep[e.g.,][]{2016arXiv160207601N}.  In principle, it exists
@@
 \citep[e.g.,][]{2015ApJ...814L..26M} and can be used to study the mass
-distribution and kinematics of stars in the lensing galaxy as well
+distribution of stars in the lensing galaxy as well
 \citep[e.g.,][]{2012ApJ...744..111P}.  Microlensing caustics have an
@@
 opposite case, the peak appears before the end of the bright disc
-reaches the caustic and shape is smoother.
+reaches the caustic and the shape of the lightcurve is smoother.
 

### 2017-06-26 paper1/ms.tex
editing

@@
 lightcurves, with both noise and systematic errors are present.  We
-generate three lightcurves by taking a uniform disc, a circular
-Gaussian, and a crescent across the path AB in
-Figure~\ref{fig:magnification_map}, and then adding noise.  The path
-simulates crossing a clean but not ideal fold.  In addition, we
-generate templates lightcurves by running the three source types, with
-various parameter values, across the path CB.  That is, the templates
-come from a similar but not identical caustic, thus deliberately
-generating a systematic error.  We then fit the noisy lightcurves to
-the templates using Markov chain Monte-Carlo.  We find that the
-correct source type can be inferred from the $\chi^2$ values.  The
-parameter values can also be inferred. The fitting errors are larger
-than the formal uncertainties, which is expected in the presence of
-systematic errors, but still appear acceptable.
+generate a mock lightcurves by taking a crescent source across the
+path AB in figure~\ref{fig:magnification_map}, and then adding noise.
+The path simulates crossing a clean but not ideal fold.  We then fit
+this lightcurve to templates from both crescent and circular-Gaussian
+sources across the path CB.  That is, the templates come from a
+similar but not identical caustics, thus deliberately generating a
+systematic error.  Despite this systematic error, a crescent source
+fits the data, while a Gaussian source is rejected from the $\chi^2$
+value.
 
@@
 (equation~\ref{eqn:ft2d}) with the magnification map to give the
-brightness.  In real life, not only are both lens and source moving
-but the lens configuration, and with it the magnification
-pattern, is also changing with time.  While the first subtlety is
-taken care of by a coordinate transformation in this analysis, for the
-second one the lens configuration is assumed to be constant in time.
-
-For the purpose of this analysis, it was desirable to mimic the
-analytical behaviour of a simple fold as much as possible. For
-comparing the numerical result with the analytical one, therefore, the
-path of the source was chosen, so that it intersects the border of the
-caustic perpendicularly, and on a point where the border is a fold
-caustic.  We choose the line AC in Figure~\ref{fig:magnification_map}
-and consider its two caustic-crossing segments separately:
-\begin{itemize}
-\item ``path~I'' runs from C to B;
-\item ``path II'' runs from A to B.
-\end{itemize}
-The magnification along paths I and II are very similar, but not
-identical.
-
-The mock lightcurve used a crescent source (as explained in
-\S\ref{subsec:crescent}) with parameters
+observable total brightness.  In real life, not only are both lens and
+source moving but the lens configuration, and with it the
+magnification pattern, is also changing with time.  The first subtlety
+is taken care of by a coordinate transformation in this analysis; the
+second one is not considered, and the lens configuration is assumed to
+be constant in time.
+
+For the purpose of this analysis, it is desirable to mimic the
+analytical behaviour of a simple fold as much as possible.
+Accordingly, we generated a mock lightcurve by taking a crescent
+source along the path CB in Figure~\ref{fig:magnification_map}.  The
+crescent parameters (as defined in \S\ref{subsec:crescent}) were
 \begin{equation}
@@
 corresponding to $r_{1/2}=1$.  Unit length has 30 pixels on the
-magnification map.  The source was carried for 250 pixels along part
-of path~I, giving 251 brightness values.  Gaussian noise at the level
-of 4\% of the current brightness was then added to give the mock
-lightcurve.  If we assume the brightness measurements are nightly, and
-the transverse speed is 600~km/s, $r_{1/2}$ comes to 24~light-minutes,
-which is plausible for a quasar.
+magnification map.  The source was carried for 250 pixels along the
+path CB, giving 251 brightness values.  Gaussian noise at 4\% of the
+current brightness was then added to give the mock data.
+
+The time unit for the light curve is taken to be $r_{1/2}/v_{\!\perp}$
+where $v_{\!\perp}$ is the velocity transverse to the caustic.  If we
+assume the brightness measurements are nightly, and
+$v_{\!\perp}=\rm200\,km/s$, the length of the microlensing event (250
+days) is comparable to typical observed events \cite[see
+  e.g.,][]{2015ApJ...814L..26M}. The implied $r_{1/2}$ is
+$\sim10^{14}\rm\,cm$, which is at the lower end of values inferred for
+microlensed X-ray sources, whereas for optical sources, the inferred
+$r_{1/2}$ is one to two order of magnitude larger \citep[see, e.g.,
+  Figure~8 of][]{0004-637X-769-1-53}.
+
+The mock data appear in Figures
+\ref{fig:cc_forward}--\ref{fig:gc_forward} below.  The example values
+abovr for $r_{1/2}$, $v_{\!\perp}$ and the observing frequency are not
+important.  What is important is that the observations have enough
+sampling and signal-to-noise that the characteristic features of a
+crescent source (see e.g., Figure~\ref{fig:char_points}) are not
+washed out.
 
@@
 nuisance parameters.
+
+Widely used \citep[e.g.,][]{2010ApJ...712..658P}.
 
@@
 shadow and caustic overlap were distinguishable.
-
-%In the case of a high-quality lightcurve with 
-%insignificant noise and measurement errors, 
-%the parameters can be obtained 
-%by simply identifying the characteristic 
-%instances of time without making use of 
-%the actual values of the magnification
-% map. If the effect of the errors and the 
-%noise distorts the magnification time function enough so that the 
-%characteristic epochs are not identifiable with the characteristic 
-%periods still visible, the boundaries of the periods 
-%can be roughly estimated. Furthermore, if direct estimates of 
-%the parameters cannot be obtained we propose the use of a strong
-%statistical tool such as Markov-Chain Monte Carlo.
-% In Section 6  we have studied the possibility of identifying a crescent 
-%source and the possibility of recovering the 
-%respective parameters. As magnification map, we have used a complex numerical 
-%one generated with the microlensing code by \cite{1999A&A...346L...5W}. 
-%In addition, we have added to the signal a Gaussian 
-%noise with an SNR of 1.6.  In the experiment,
-% we have considered all nine combinations of
-% original source profiles and assumed fitting source models.
-%Effectively we have fitted using MCMC all three sources with all three assumed fitting models. The results of the experiment allowed us to build a 
-%procedure for distinguishing the shape of the source assuming that one of the three models we have considered is a good approximation.  
-%The procedure would be useful for observers that endeavour to gain more information about an unresolved source for which they can study 
-%the microlensing lightcurve.  As a first step in the procedure, one should first attempt to use MCMC with a Gaussian model assumption. 
-%If either the value of the $\chi^2$ or the number of rejected datapoints is large then the next step is to change the assumed source model to 
-%a uniform disc and redo the fitting. Finally, if the uniform disc assumption is rejected as well then the fitting should be done with a 
-%crescent source model assumption. If at each of the three steps the data rejects the model then the source cannot be approximated by any of the three models. 
-%Otherwise, if at one of the steps the data does not reject the model then the respective model is a good approximation.
 

### 2017-06-26 fitting/summaryfig.py

@@ -72,7 +72,7 @@ def plotcurves(p):
     p.errorbar(t,b,yerr=db,linestyle='none',color='gray')
     p.plot(t,m,color='black')
     p.set_yticks(np.arange(0,7,2))
-    p.set_xlabel('$t [r_{1/2}/v]$',labelpad=10, fontsize=16)
+    p.set_xlabel('$t [r_{1/2}/v{\!_\perp}]$',labelpad=10, fontsize=16)
     p.set_ylabel('magnification', fontsize=16)
 
 def plothist(p):

### 2017-06-27 paper1/ms.tex
editing

@@
 constant-brightness disc, a circular Gaussian, and the crescent source
-introduced by \citep{2013MNRAS.434..765K}.  The latter is simply a
+introduced by \cite{2013MNRAS.434..765K}.  The latter is simply a
 constant-brightness disc with a smaller, non-concentric disc cut out
@@
 
-We now proceed to fit the mock lightcurve using MCMC, to recover the
-source parameters.
-
-Note that fitting involves three nuisance parameters, namely the
-brightness normalisation and the start and end points on the path.
-The latter two are equivalent to considering the location of the
-caustic with respect to the starting point and the speed of the
-source.  The fitting method automatically marginalises over all the
-nuisance parameters.
-
-Widely used \citep[e.g.,][]{2010ApJ...712..658P}.
+We now proceed to fit the mock lightcurve.  The basic fitting
+technique is the same as used in several previous studies
+\citep[e.g.,][]{2010ApJ...712..658P}.  Markov-Chain Monte-Carlo is
+used to fit source parameters and additional nuisance parameters (see
+below) jointly, and then marginalise out all uninteresting parameters,
+leaving posterior distributions for any chosen parameter.
+
+In addition to the source parameters, the fitting procedure includes
+three nuisance parameters, namely the brightness normalisation and the
+start and end points on the path.  The latter two are equivalent to
+considering the location of the caustic with respect to the starting
+point and the velocity component of the source transverse to the
+caustic (that is, $v_{\!\perp}$).  The orientation of the source track
+is another nuisance parameter, which we do not include it separately.
+This is because, for a pure fold, changing the orientation of the
+source track is equivalent to changing $v_{\!\perp}$ and the
+orientation of the source (see Figure~\ref{fig:infinite_fold}), all
+parameters that are included.  In the present case, the caustic is a
+clean but non-ideal fold, hence disregarding the orientation of the
+source track is expected to introduce some systematic error.
 

### 2017-06-27 paper1/ms.tex
editing

@@
 lensed brightness undergoes sudden changes as a quasar crosses a
-caustic.  Quasar microlensing has been first observed in the optical
-range \citep[e.g.,][]{2012A&A...544A..62S} but is also known in
+caustic.  Quasar microlensing is observable at optical
+wavelengths \citep[e.g.,][]{2012A&A...544A..62S} and in
 X-rays \citep[e.g.,][]{2016arXiv160207601N}.  In principle, it exists
@@
 crescent sources, and then fit these mock lightcurves to source models
-using Markov-chain Monte Carlo (MCMC).
+using Markov-chain Monte Carlo (MCMC).  The general strategy is the same
+as in previous studies \citep[e.g.,][]{2010ApJ...712..658P} though
+the details are different.
 
@@
 We now proceed to fit the mock lightcurve.  The basic fitting
-technique is the same as used in several previous studies
-\citep[e.g.,][]{2010ApJ...712..658P}.  Markov-Chain Monte-Carlo is
+technique is the same as used in   Markov-Chain Monte-Carlo is
 used to fit source parameters and additional nuisance parameters (see
@@
 caustic (that is, $v_{\!\perp}$).  The orientation of the source track
-is another nuisance parameter, which we do not include it separately.
-This is because, for a pure fold, changing the orientation of the
-source track is equivalent to changing $v_{\!\perp}$ and the
-orientation of the source (see Figure~\ref{fig:infinite_fold}), all
-parameters that are included.  In the present case, the caustic is a
-clean but non-ideal fold, hence disregarding the orientation of the
-source track is expected to introduce some systematic error.
-
-Figure \ref{fig:cc_forward} shows a fit to the mock lightcurve and
-recovery of $r_{1/2}$ and $R_n/R_p$.
-
-Figure \ref{fig:cc_backward} shows the result when a different caustic
-is used to fit, than was used to generate the mock lightcurve.  This
-mimics the unavoidable systematic error of not knowing the caustic
-exactly.  The fit is slightly degraded, but still acceptable.
-
-Figure \ref{fig:gc_forward} shows the result of attempting to fit a
-Gaussian source.  There is a small but statistically very significant
-misfit because a Gaussian curve cannot reproduce the part of the
-lightcurve with the slight flattening and then a second sub-peak
-(cf.~Figure~\ref{fig:char_points}) as the dark part of the crescent crosses the caustic.  
-A hint of
-such a feature is present in recent caustic-crossing observations of
-the well-known system Q2237+0305 \citep[see Figure~3
-in][]{2015ApJ...814L..26M} though interpretation as a crescent source
-is speculation at present.
+is another nuisance parameter, which we do not include it, because for
+a pure fold, changing the orientation of the source track is
+equivalent to changing $v_{\!\perp}$ and the orientation of the source
+(see Figure~\ref{fig:infinite_fold}), all parameters that are
+included.  Since the caustic is not a perfect fold, disregarding the
+orientation of the source track can be expected to introduce some
+systematic error.
+
+Figures \ref{fig:cc_forward}--\ref{fig:gc_forward} show the numerical
+results.  Each of these figures shows the mock data (the same in all
+three cases) along with best fits and the fitted values and
+uncertainties on $r_{1/2}$ and (if applicable) $R_n/R_p$.
+
+Figure \ref{fig:cc_forward} shows a test of the fitting procedure.
+The computer compares the mock data to different templates
+lightcurves.  All the templates come from taking a crescent source
+along the path CB in Figure~\ref{fig:magnification_map}, but with
+varied crescent parameters and varied start and finish points.  As the
+figure shows, the lightcurve can be fitted to within the noise, and
+the parameters $r_{1/2}$ and $R_n/R_p$ can be recovered.
+
+Figure \ref{fig:cc_backward} attempts a somewhat more realistic
+situation.  The mock lightcurve is the same as before, but the
+templates come from taking a crescent source along the path AB in
+Figure~\ref{fig:magnification_map}.  That is, the mock data and the
+templates come from crossings of different caustics.  This mimics the
+unavoidable systematic error of not knowing the caustic exactly.  The
+fit is slightly degraded, but still acceptable.  The parameters
+$r_{1/2}$ and $R_n/R_p$ are still recovered.  The parameters $a/R_p$
+and $b/R_p$ were not correctly recovered, which is understandable from
+the track change.
+
+Figure \ref{fig:gc_forward} shows the result of attempting to fit
+templates from a Gaussian source to the same mock data.  There is a
+small but statistically very significant misfit because a Gaussian
+source cannot reproduce the part of the lightcurve with the slight
+flattening and then a second sub-peak
+(cf.~Figure~\ref{fig:char_points}) as the dark part of the crescent
+crosses the caustic.  A hint of such a feature is present in recent
+caustic-crossing observations of the well-known system Q2237+0305
+\citep[see Figure~3 in][]{2015ApJ...814L..26M} though interpretation
+as a crescent source is speculation at present.
 

### 2017-06-28 paper1/ms.tex
edits

@@
 
-In Section~\ref{sec:source-models} we introduce the three source
-profiles used in our subsequent models and simulations: a
+In Section~\ref{sec:source-models} we introduce three source profiles
+which are used in our subsequent models and simulations: a
 constant-brightness disc, a circular Gaussian, and the crescent source
@@
 lightcurves, with both noise and systematic errors are present.  We
-generate a mock lightcurves by taking a crescent source across the
-path AB in figure~\ref{fig:magnification_map}, and then adding noise.
-The path simulates crossing a clean but not ideal fold.  We then fit
-this lightcurve to templates from both crescent and circular-Gaussian
-sources across the path CB.  That is, the templates come from a
-similar but not identical caustics, thus deliberately generating a
-systematic error.  Despite this systematic error, a crescent source
-fits the data, while a Gaussian source is rejected from the $\chi^2$
-value.
+generate a mock lightcurve by taking a crescent source across the path
+CB in Figure~\ref{fig:magnification_map}, and then adding noise.  The
+path simulates crossing a clean but not ideal fold.  We then fit this
+lightcurve to templates from both crescent and circular-Gaussian
+sources along the paths AB as well as CB.  That is, the templates from
+a similar but not identical caustic are also used, thus deliberately
+generating a systematic error.  Despite this systematic error, a
+crescent source fits the data, while a Gaussian source is rejected
+from the $\chi^2$ value.
 
@@
 implications of the results presented in the previously mentioned
-sections.  One of the most interesting implication is the possibility
-to estimate the black hole mass from the reconstructed
+sections.  One of the most interesting implications is the possibility
+of estimating the black hole mass from the reconstructed
 parameters. This further requires good approximations of the relative
@@
 \citep{2001PASA...18..207W}.  Extended sources wash out this
-micro-caustic structure, but the optical-continuum source of quasars
-is even smaller.
+micro-caustic structure, but the optical-continuum and X-ray sources
+of quasars are even smaller.
 
@@
 For a microlensing event the lightcurve can be written for an
-undefined source shape as:
-\begin{equation}
- F(t) = \int_{-\infty}^\infty \int_{-\infty}^\infty S_{2D}(p-p_s(t), q-q_s(t)) \mu_t(p) \mathrm{d}q \mathrm{d}p
+arbitrary source shape as:
+\begin{equation}
+ F(t) = \int_{-\infty}^\infty \int_{-\infty}^\infty S_{2D}(p-p_s(t), q-q_s(t)) \mu_t(p) \mathrm{d}q \mathrm{d}p.
  \label{eqn:ft2d}
@@
 dependent quantities in the right hand side of the equation are the
-coordinates of the center of the source $p_s,q_s$ and by construct
-$S_{2D}$.
-
-Due to the choice of the coordinate system the amplification factor 
+coordinates of the center of the source $p_s(t),q_s(t)$.
+
+Due to the choice of coordinate system the amplification factor 
 has no dependence on the $q$ coordinate. The previous equation can be rewritten as:
@@
 
-In the present study we are analysing three types of sources with different surface brightness: 
+In the present study we analyse three types of sources with different surface brightness: 
 \begin{enumerate}
@@
 \begin{equation}
-r_{1/2} = \sqrt{ln(4)} \sigma.
+r_{1/2} = \sqrt{\ln4}\,\sigma.
 \end{equation}
@@
 
-One can construct mathematically a disk source with constant surface brightness and radius $R$ using a stepfunction:
+One can construct mathematically a disk source with constant surface-brightness and radius $R$ using a stepfunction:
 \begin{equation}
@@
 \end{equation}
-\\
 The half-light radius of a uniform disk source is $R/\sqrt{2}$.
@@
 \end{equation}
-\\
-The following notations were used: $R_p, (p_{sp}, q_{sp}), R_n, (p_{sn},q_{sn})$ are the radii 
-and coordinate of the center for the larger positive disk and smaller negative disk respectively. 
+The following notations were used: $R_p$ and $(p_{sp}, q_{sp})$ are
+the radii and coordinate of the center for the larger positive disk, 
+while $R_n$ $(p_{sn},q_{sn})$ correspond to the smaller negative disk. 
 $S_0^{Dp},S_0^{Dn}$ represent the total flux of radiation received from the large and small disk. 
@@
 \\
-Analogous for the linear brightness function:
+Analogously for the linear brightness function:
 
@@
 
-There are some constraints on the parameters used to define a crescent in the previously 
-presented manner that need to be stated. First, we must impose the obvious $R_p > R_n$ relation. Secondly, 
-the small disk must always be interior to the large disk:
+There are some constraints on the parameters used to define a crescent
+in the previously presented manner that need to be stated. First, we
+must impose the obvious $R_p \geq R_n$ relation. Secondly, the small disk
+must always be interior to the large disk:
 \begin{equation}
@@
 \end{equation}
-
 For the distances between the centers of the two disks we will use the same 
@@
 
-The half-light radius of any source is invariant to any rotational transformation. 
-In the present case of a crescent source the effective radius is dependent on the 
-parameters $R_p$, $R_n$ and $\sqrt{a^2+b^2}\equiv c $ exclusively. From symmetry
-considerations the centroid of the source is collinear with the centers of the two disks 
-and it is situated at a distance $d_c$ from the center of the bright disk. $d_c$ can be 
-computed numerically with the use of a variation of equation \ref{eqn:s1_d}:
+The half-light radius of any source is invariant under any rotational
+transformation.  In the present case of a crescent source the
+effective radius is dependent on the parameters $R_p$, $R_n$ and
+$\sqrt{a^2+b^2}\equiv c $ exclusively. From symmetry considerations
+the centroid of the source is collinear with the centers of the two
+disks and it is situated at a distance $d_c$ from the center of the
+bright disk. $d_c$ can be computed numerically with the use of a
+variation of equation \ref{eqn:s1_d}:
 
@@
 
-With the position of the centroid determined, the half-light radius can be also be computed numerically:
+With the position of the centroid determined, the half-light radius
+can be also be computed through a numerical integral:
 
@@
 Using equation \ref{eqn:ft} and the one-dimensional flux function
-presented in the previous section one can compute numerically the
+presented in the previous section one can numerically compute the
 lightcurves of the three extended sources for the simplified
@@
 \end{equation}
-
 which can be simplified to:
@@
 \end{equation}
-
 which is equivalent to:
@@
 
-The function $p_s(t)$ can be chosen to be equal to $v_p(t-t_0) + p_{s0}$. Where $p_{s0}$ is the coordinate $p$ 
+The function $p_s(t)$ can be chosen to be equal to $v_p(t-t_0) + p_{s0}$ where $p_{s0}$ is the coordinate $p$ 
 of the source at the initial time, and $v_p$ is the component of the velocity
-along the $p$ axis. Such a modelling of the motion of the object in the source 
+along the $p$ axis. Such a model of the motion of the object in the source 
 plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
@@
 
-In figure \ref{fig:char_points} there are four characteristic points visible on the resulting one-dimensional 
+In Figure \ref{fig:char_points} there are four characteristic points visible on the resulting one-dimensional 
 light profile of the crescent shaped source. Two of the points, p1 and p4, mark the outer boundaries of the
@@
 \end{equation}
-All four points mark the positions where derivative $\frac{dS_{1D}}{dp}$ is discontinuous. The points can be used 
+All four points mark the positions where the derivative $\frac{dS_{1D}}{dp}$ is discontinuous. The points can be used 
 to define three regions: $p_1 - p_2$ where $S_{1D}$ is convex, $p_2 - p_3$ where $S_{1D}$ is concave, 
@@
 parameter on the light curve is presented in
-Figures~\ref{fig:lightcurve_crescent} and
-\ref{fig:lightcurve_crescent_back}. For the particular set of
+Figures~\ref{fig:lightcurve_crescent_back} and
+\ref{fig:lightcurve_crescent}. For the particular set of
 parameters where $R_p = R_n +a$ the third period of time discussed
@@
 The top right-most image in figure 5 of \cite{2012MNRAS.421.1517D} has been projected to a 1D profile associated to the perpendicular 
-direction to a fold caustic approaching the image from the right. The projection is presented in the upper panel of figure \ref{fig:M87_plots}. 
+direction to a fold caustic approaching the image from the right. The projection is presented in the upper panel of Figure \ref{fig:M87_plots}. 
 The amplification values of the flux of light corresponding to a 
-microlensing event are presented in the lower panel of figure \ref{fig:M87_plots}. 
+microlensing event are presented in the lower panel of Figure \ref{fig:M87_plots}. 
 In general, the behaviour of the lightcurve is similar to the geometric crescent source with the caveat that the outer regions 
@@
 crescent sources, and then fit these mock lightcurves to source models
-using Markov-chain Monte Carlo (MCMC).  The general strategy is the same
+using Markov-chain Monte Carlo (MCMC).  The basic technique is the same
 as in previous studies \citep[e.g.,][]{2010ApJ...712..658P} though
@@
 
-The time unit for the light curve is taken to be $r_{1/2}/v_{\!\perp}$
-where $v_{\!\perp}$ is the velocity transverse to the caustic.  If we
+The time unit for the light curve is taken to be $r_{1/2}/v_p$
+where $v_p$ is the velocity transverse to the caustic.  If we
 assume the brightness measurements are nightly, and
-$v_{\!\perp}=\rm200\,km/s$, the length of the microlensing event (250
+$v_p=\rm200\,km/s$, the length of the microlensing event (250
 days) is comparable to typical observed events \cite[see
@@
 The mock data appear in Figures
-\ref{fig:cc_forward}--\ref{fig:gc_forward} below.  The example values
-abovr for $r_{1/2}$, $v_{\!\perp}$ and the observing frequency are not
+\ref{fig:cc_forward}--\ref{fig:gc_forward}.  The example values
+above for $r_{1/2}$, $v_p$ and the observing frequency are not
 important.  What is important is that the observations have enough
@@
 
-We now proceed to fit the mock lightcurve.  The basic fitting
-technique is the same as used in   Markov-Chain Monte-Carlo is
-used to fit source parameters and additional nuisance parameters (see
-below) jointly, and then marginalise out all uninteresting parameters,
-leaving posterior distributions for any chosen parameter.
+We now proceed to fit the mock lightcurve.  MCMC is used to fit source
+parameters and additional nuisance parameters (see below) jointly, and
+then marginalise out all uninteresting parameters, leaving posterior
+distributions for any chosen parameter.
 
@@
 point and the velocity component of the source transverse to the
-caustic (that is, $v_{\!\perp}$).  The orientation of the source track
-is another nuisance parameter, which we do not include it, because for
-a pure fold, changing the orientation of the source track is
-equivalent to changing $v_{\!\perp}$ and the orientation of the source
-(see Figure~\ref{fig:infinite_fold}), all parameters that are
-included.  Since the caustic is not a perfect fold, disregarding the
-orientation of the source track can be expected to introduce some
-systematic error.
+caustic (that is, $v_p$).  The orientation of the source track is
+another nuisance parameter, but we do not include it, because for a
+pure fold, changing the orientation of the source track is equivalent
+to changing $v_p$ and the orientation of the source (see
+Figure~\ref{fig:infinite_fold}), all parameters that are included.
+Since the caustic is not a perfect fold, disregarding the orientation
+of the source track can be expected to introduce some systematic
+error.
 
@@
 Figure \ref{fig:cc_forward} shows a test of the fitting procedure.
-The computer compares the mock data to different templates
-lightcurves.  All the templates come from taking a crescent source
-along the path CB in Figure~\ref{fig:magnification_map}, but with
-varied crescent parameters and varied start and finish points.  As the
-figure shows, the lightcurve can be fitted to within the noise, and
-the parameters $r_{1/2}$ and $R_n/R_p$ can be recovered.
+The computer compares the mock data to different template
+lightcurves. The templates come from taking a crescent source along
+the path CB in Figure~\ref{fig:magnification_map}, but with varied
+crescent parameters and varied start and finish points.  As the figure
+shows, the lightcurve can be fitted to within the noise, and the
+parameters $r_{1/2}$ and $R_n/R_p$ can be recovered.
 

### 2017-06-28 fitting/summaryfig.py

@@ -72,7 +72,7 @@ def plotcurves(p):
     p.errorbar(t,b,yerr=db,linestyle='none',color='gray')
     p.plot(t,m,color='black')
     p.set_yticks(np.arange(0,7,2))
-    p.set_xlabel('$t [r_{1/2}/v{\!_\perp}]$',labelpad=10, fontsize=16)
+    p.set_xlabel('$t [r_{1/2}/v_p]$',labelpad=10, fontsize=16)
     p.set_ylabel('magnification', fontsize=16)
 
 def plothist(p):

### 2017-06-29 paper1/ms.tex
submitted

@@
  \renewcommand{\theenumi}{(\arabic{enumi})}
-  \item a rotationally symmetric source with a bivariate gaussian surface brightness distribution,
+  \item a rotationally symmetric source with a bivariate Gaussian surface brightness distribution,
   \item a disk source with constant surface brightness distribution,
@@
 
-\subsection{Rotationally symmetric source with a bivariate gaussian surface brightness distribution}\label{subsec:gaussian}
-
-A symmetric 2D gaussian can be described mathematically as:
+\subsection{Rotationally symmetric source with a bivariate Gaussian surface brightness distribution}\label{subsec:gaussian}
+
+A symmetric 2D Gaussian can be described mathematically as:
 
@@
 Although such a definition for a source would have non-zero surface/linear brightness for any coordinate $p,q$, the amount of light received by a detector from outside a $3 \sigma$ disk centered at $p_s, q_s$
-would be insignificant. For a gaussian distributed surface brightness source the half-light radius is directly proportional to the parameter $\sigma$ according to the equation:
+would be insignificant. For a Gaussian distributed surface brightness source the half-light radius is directly proportional to the parameter $\sigma$ according to the equation:
 \begin{equation}
@@
 
-One can construct mathematically a disk source with constant surface-brightness and radius $R$ using a stepfunction:
+One can construct mathematically a disk source with constant surface brightness and radius $R$ using a stepfunction:
 \begin{equation}
@@
 
-\subsection{Lightcurve of the gaussian source}
+\subsection{Lightcurve of the Gaussian source}
 
 The amount of light received by an observer from a source 
-with a gaussian distributed brightness with $\sigma$ and total flux $S_0^G$ in the absence of any gravitational lensing is:
+with a Gaussian distributed brightness with $\sigma$ and total flux $S_0^G$ in the absence of any gravitational lensing is:
 \begin{equation}
@@
         \includegraphics[width = 0.48\textwidth]{figures/ch_instances.eps}
-\caption{\label{fig:char_points} The characteristic points and instances
-of a crescent source's one dimensional profile and microlensing lightcurve.
-In the top panel a typical 1D profile  of a crescent source is presented.
-There are two light gray regions that mark the regions of the bright disk
-that are not overlapped by the dark disk. The dark gray band marks the
-region where the two disks overlap. $p_1$ and $p_4$ mark the start and
-end of the bright disk in the
-1D representation. Analogous, $p_2$ and $p_3$ mark the start and
-end of the dark disk.
-The bottom panel represents the associated lightcurve with the points
-$t_i$ marking the moment when point $p_i$ overlaps with the caustic.
-The gray shades are analogous to the ones in the top panel.
- }
+\caption{\label{fig:char_points} The characteristic points and
+  instances of a crescent source's one dimensional profile and
+  microlensing lightcurve.  In the top panel a typical 1D profile of a
+  crescent source is presented.  There are two light gray regions that
+  mark the regions of the bright disk that are not overlapped by the
+  dark disk. The dark gray band marks the region where the two disks
+  overlap. The points $p_1$ and $p_4$ mark the start and end of the
+  bright disk in the 1D representation. Analogously, $p_2$ and $p_3$
+  mark the start and end of the dark disk.  The bottom panel
+  represents the associated lightcurve with the points $t_i$ marking
+  the moment when point $p_i$ overlaps with the caustic.  The gray
+  shades are analogous to the ones in the top panel.  }
 \end{figure}
@@
 
-Analogous to the gaussian shaped source, the disk source with uniform brightness, radius $R$ and 
+Analogous to the Gaussian shaped source, the disk source with uniform brightness, radius $R$ and 
 unmagnified flux $S_0^D$ has a lightcurve described by the equation:
@@
 visible features than the other two light-curves corresponding to the
-disc and gaussian shape. There are three regions where $S_{1D}$ and by
+disc and Gaussian shape. There are three regions where $S_{1D}$ and by
 inheritance, the lightcurve has distinguishable behaviour. The first
@@
 \caption{\label{fig:M87_plots} The one dimensional profile obtained
-  from the numerical integration along the ordonata axis of a simulated source
-  image is presented in the upper panel. The source image was presented in 
- Dexter et al 2012 in the top right-most panel of figure 5 
-(of the respective paper) as a result of the simulations 
-presented in McKinney \& Blandford 2009. The corresponding
-  lightcurve associated to the 1D profile is presented in the lower
-  panel. The points $\rm p_2$ and $\rm p_3$ mark the start and 
-  end of the equivalent
-  dark disk and are associated with the moments of their overlap with
-  the caustic: $\rm t_2$ and $\rm t_3$. $\rm h_{w}$ denotes the maximum width of the crescent.}
+  from the numerical integration along the ordonata axis of a
+  simulated source image is presented in the upper panel. The source
+  image was presented in Dexter et al (2012) in the top right-most
+  panel of Figure~5 (of the respective paper) as a result of the
+  simulations presented in McKinney \& Blandford (2009). The
+  corresponding lightcurve associated to the 1D profile is presented
+  in the lower panel. The points $\rm p_2$ and $\rm p_3$ mark the
+  start and end of the equivalent dark disk and are associated with
+  the moments of their overlap with the caustic: $t_2$ and $t_3$;
+  $h_{w}$ denotes the maximum width of the crescent.}
 \end{figure}
@@
 source could be discerned in real data having noise and systematic
-error as well.  Our method is to generate microlensing lightcurves
+error as well.  Do do so, we first generate a microlensing lightcurve
 using a widely-used microlensing code
@@
 which we have added a small extension to provide the option of
-crescent sources, and then fit these mock lightcurves to source models
-using Markov-chain Monte Carlo (MCMC).  The basic technique is the same
-as in previous studies \citep[e.g.,][]{2010ApJ...712..658P} though
-the details are different.
+crescent sources.  We then fit model or template lightcurves ---which
+are generated with varied source and other parameters--- to the mock
+lightcurve.  The basic technique is the same as in previous studies
+\citep[e.g.,][]{2010ApJ...712..658P} though the details are different.
 
@@
   \includegraphics[width=0.5\textwidth]{figures/cc_backward.eps}
-\caption{\label{fig:cc_backward} Like Figure~\ref{fig:cc_backward},
-  except that the fit uses path~II whereas the mock data still come
-  from path~I. The curve has $\chi^2_{(244)}=284, P=0.01$ indicating a
-  marginally acceptable fit.}
+\caption{\label{fig:cc_backward} Like Figure~\ref{fig:cc_forward},
+  except that the fit is derived from a slightly different caustic
+  than the one used to generate the mock data.  The curve has
+  $\chi^2_{(244)}=284, P=0.01$ indicating a marginally acceptable
+  fit.}
 \end{figure}
@@
   \includegraphics[width=0.5\textwidth]{figures/gc_forward.eps}
-\caption{\label{fig:gc_forward} Like Figure~\ref{fig:cc_backward}, but
-  showing the best fit to a Gaussian source.  The histogram shows the
-  posterior distribution of $r_{1/2}$ (there is no $R_n$).  It appears
-  to show $r_{1/2}$ as accurately recovered, but is not to be trusted,
-  because the fit has $\chi^2_{(247)}=344, P=3\times10^{-6}$ indicating
-  rejection of the Gaussian source.  Looking closely, the curve is a
-  little too shallow near $t=3$, and it is unable to fit the second
-  sub-peak near $t=5$.}
+\caption{\label{fig:gc_forward} Like Figures~\ref{fig:cc_forward} and
+  \ref{fig:cc_backward}, but showing the best fit to a Gaussian
+  source.  The histogram shows the posterior distribution of $r_{1/2}$
+  (there is no $R_n$).  It appears to show $r_{1/2}$ as accurately
+  recovered, but is not to be trusted, because the fit has
+  $\chi^2_{(247)}=344, P=3\times10^{-6}$ indicating rejection of the
+  Gaussian source.  Looking closely, the curve is a little too shallow
+  near $t=3$, and it is unable to fit the second sub-peak near $t=5$.}
 \end{figure}
@@
 Nevertheless, the respective displacement is relevant for the calculation of
-of the half-light radius.  It's qualitative effect is to 
+of the half-light radius.  Its qualitative effect is to 
 decrease the value of the
@@
 influences on the shape of the microlensed lightcurve as can be seen
-in Figures~\ref{fig:a_var}, \ref{fig:lightcurve_crescent} and
-\ref{fig:lightcurve_crescent_back}.  Moreover, the one-dimensional
+in Figures~\ref{fig:a_var}, \ref{fig:lightcurve_crescent_back} and
+\ref{fig:lightcurve_crescent}.  Moreover, the one-dimensional
 source profile corresponding to the direction perpendicular to the
@@
 these points with the caustic leaves visible features on the
-lightcurve at the corresponding instances of time. In timely order,
+lightcurve at the corresponding instances of time. In time order,
 the instances correspond to the start of the overlap between the
@@
 Furthermore, a simulated image of M87 presented in
- \citep{2012MNRAS.421.1517D} has been microlensed (figure \ref{fig:M87_plots}). 
-On the resulting lightcurve the instances 
-corresponding to the start and end of the black hole 
-shadow and caustic overlap were distinguishable.
+\cite{2012MNRAS.421.1517D} has been microlensed
+(Figure~\ref{fig:M87_plots}).  On the resulting lightcurve the
+instances corresponding to the start and end of the black hole shadow
+and caustic overlap were distinguishable.
 
@@
 previous expression, the $\Delta t_{dark}$ denotes the period of
-overlap between the black hole shadow and fold caustic.  $v_p$ denotes
-the component of the relative velocity of the source and fold which is
-perpendicular to the caustic.  The respective velocity is an unknown,
-though it can be constrained on a case by case basis to an order of
-magnitude or even better. This would require the study of the dynamics
-of the stellar structure which contains the gravitational lens and the bulk
-transverse velocity of the galaxy with respect to the observer.  A
-better estimate of the relative velocity would facilitate a better
-estimate of the effective non-rotating black hole mass associated with
-the black hole shadow.
+overlap between the black hole shadow and fold caustic, while $v_p$
+denotes the component of the relative velocity of the source and fold
+which is perpendicular to the caustic.  The respective velocity is an
+unknown, though it can be constrained on a case by case basis to an
+order of magnitude or even better. This would require the study of the
+dynamics of the stellar structure which contains the gravitational
+lens and the bulk transverse velocity of the galaxy with respect to
+the observer.  A better estimate of the relative velocity would
+facilitate a better estimate of the effective non-rotating black hole
+mass associated with the black hole shadow.
 

### 2017-09-11 fitting/summaryfig2.py
edits

@@ -55,7 +55,7 @@ def gencurves(fname):
     lcurv.writecurves(q)
 
 
-def plotcurves(p):
+def plotcurves(p,lbl):
     fil = open('curves.txt')
     all = fil.readlines()
     t = []
@@ -70,43 +70,18 @@ def plotcurves(p):
         db.append(float(s[3]))
     p.axis([0,6,0,7])
     p.errorbar(t,b,yerr=db,linestyle='none',color='gray')
-    p.plot(t,m,color='black')
+    p.plot(t,m,label=lbl) #,color='black')
     p.set_yticks(np.arange(0,7,2))
     p.set_xlabel('$t [r_{1/2}/v_p]$',labelpad=10, fontsize=16)
     p.set_ylabel('magnification', fontsize=16)
 
-def plothist(p):
-    rmin,rmax,nr = 0,1.2,61
-    p.hist(rh,np.linspace(rmin,rmax,nr),weights=wt,
-           histtype='step',lw=2,color='black')
-    p.plot([1,1],[.43,.48],lw=2,color='black')
-    fl = max(rns)
-    if fl > 0:
-        p.hist(rns,np.linspace(rmin,rmax,nr),weights=wt,
-               histtype='step',lw=2,color='black',linestyle='dashed')
-        p.plot([.4,.4],[.43,.48],lw=2,color='black',linestyle='dashed')
-        p.set_xlabel('inferred $R_n/R_p$ and $r_{1/2}/v_p$',labelpad=10)
-    else:
-        p.set_xlabel('inferred $r_{1/2}/v_p$',labelpad=10, fontsize=16)
-    p.axis([rmin,rmax,0,0.5])
-
-
-
 mockdata()
 
-for fname in ['gc_forward','cc_forward','cc_backward']:
+fig = pl.figure()
+p = fig.add_subplot(111)
+for fname in ['cc_forward','cc_backward','gc_forward']:
     gencurves(fname)
+    plotcurves(p,fname[:7])
+pl.legend()
 
-    fig = pl.figure(figsize=(5,10))
-    fig.subplots_adjust(hspace=0.2)
-
-    p = fig.add_subplot(211)
-    plotcurves(p)
-
-    p = fig.add_subplot(212)
-    plothist(p)
-    p.get_yaxis().set_visible(False)
-
-    fig.savefig('../paper1/figures/%s.eps'%fname)
-
-    pl.show()
+pl.show()

### 2017-09-11 morefigs/m_general.py

@@ -1,0 +1,479 @@
+def delimitate (tip):
+    print str(tip)*70
+    
+def print_format (script_name, input_set):
+    delimitate ("=")
+    print "Format:"
+    print "> python ", script_name, input_set
+    delimitate ("-")
+
+def in_sphere (point, center, radius):
+    x, y, z = point[0]-center[0], point[1]-center[1], point[2]-center[2]
+    if (x**2+y**2+z**2) > radius**2:
+        ok = False
+    else:
+        ok = True
+    return ok
+
+def load_npz(name):
+    file_input = numpy.load(name)
+    print file_input.files()
+
+    
+
+def status (previous, counting, total_number):
+    if counting > previous * total_number/100.:
+        print previous,"%"
+        previous = previous+1
+    return previous
+
+def simple_inquiry():
+    answer = str(raw_input(">"))
+    valid_answer = False
+    while not valid_answer:
+        if answer.lower() in ["yes", "y", "ye", "da","d", "ja", "si"]:
+            yes = True
+            valid_answer = True    
+        if answer.lower() is "no":
+            yes = False
+            valid_answer = True
+        if answer.lower() not in ["no","yes","y", "ye", "da","d", "ja", "si"]:
+            valid_answer = False 
+    return yes
+
+import pylab as pt
+def plot2d_scatter(x,y, title, x_label, y_label, file_save, non_zero, ratio, mark_zero_point):
+    pt.plot(x,y,'r.')
+    pt.xlabel(str(x_label))
+    pt.ylabel(str(y_label))
+    pt.title(str(title))
+    if str(non_zero) == "non_zero" :
+        pt.ylim(min(y)*ratio, max(y)*ratio)
+        pt.xlim(min(x)*ratio, max(y)*ratio)
+    else:
+        pt.xlim(min(x), max(x)*ratio)
+        pt.ylim(0, max(y)*ratio)
[…]

### 2017-09-11 morefigs/sf_ML_4source_lightcurves.py

@@ -1,0 +1,266 @@
+import numpy
+import scipy
+from scipy import integrate
+import m_general
+from matplotlib import pyplot
+
+
+def magnification_function(p, mu_0, C_0):
+    if p < 0:
+        return mu_0
+    elif p == 0:
+        return mu_0
+    else:
+
+        return mu_0 + C_0/p**0.5
+        
+
+def crescent_1D(S_0, R_p, R_n, p_sp, p_sn,  p):
+    factor = 2* S_0/numpy.pi/(R_p**2-R_n**2)
+    if (R_p**2 > (p-p_sp)**2) and (R_n**2 > (p-p_sn)**2):
+        value = ((R_p**2-(p-p_sp)**2))**0.5 -  ((R_n**2-(p-p_sn)**2))**0.5
+    elif (R_p**2 > (p-p_sp)**2) and (R_n**2 < (p-p_sn)**2):
+        value = (R_p**2-(p-p_sp)**2)**0.5 
+    else:
+        value = 0
+    return value*factor 
+
+def gaussian_1D(S_0G, sig, p_s, p):
+    return S_0G/((2*numpy.pi)**0.5*sig)*numpy.exp(-(p-p_s)**2/2/sig**2)
+
+def disk_1D(S_0D, R, p_s, p):
+    factor = 2* S_0D/numpy.pi/R * (1-(p-p_s)**2/R**2)**0.5
+    if R**2 > (p-p_s)**2:
+        return factor
+    else:
+        return 0
+
+
+def function(p, mu_0, C_0, S_0, parameters, type = "crescent"):  #gaussian parameters= [sig, p_s], disk parameters= [ R, p_s] 
+    if type not in ["crescent", "disk", "gaussian"]:
+	print "Shape of source unrecognized"
+
+    if type == "crescent":
+    	return magnification_function(p, mu_0, C_0)* crescent_1D(S_0, parameters[0], parameters[1], parameters[2], parameters[3], p)
+    if type == "disk":
+        return magnification_function(p, mu_0, C_0)* disk_1D(S_0, parameters[0], parameters[1], p)
+    if type == "gaussian":
+        return magnification_function(p, mu_0, C_0)* gaussian_1D(S_0, parameters[0], parameters[1], p)
+
+
[…]

### 2017-09-11 morefigs/sf_ML_source_shapes.py

@@ -1,0 +1,228 @@
+import numpy
+
+def theta(value):
+	lista_poz = numpy.where(value > 0)[0]
+	lista_neg = numpy.where(value < 0)[0]
+	theta = value.copy()
+	theta[lista_poz] = numpy.ones(len(lista_poz))
+	theta[lista_neg] = numpy.zeros(len(lista_neg))
+	return theta
+
+
+def zerofy(value):
+	lista = numpy.where(value**2 > 100000)[0]
+	value[lista] = numpy.zeros(len(lista))
+
+
+def S1D(time, shift, r, S_0):
+	lista_non = numpy.where(1-((time-shift)/r)**2 < 0)[0]
+	S1D = time.copy()
+	S1D = 2*S_0/numpy.pi/r*(1-((time-shift)/r)**2)**0.5*theta(r**2 - (time-shift)**2)
+	S1D[lista_non] = numpy.zeros(len(lista_non))
+	return S1D
+
+
+S_0 = 1.
+r_half = 1.
+
+time = numpy.linspace(-2., 2., 1000)
+
+#gaussian
+sigma = r_half/ (numpy.log(4)**0.5)
+S1D_g = S_0/(2*numpy.pi)**0.5/sigma*numpy.exp(-time**2/2/sigma**2)
+print "sigma", sigma
+
+#disk
+r = r_half*2**0.5
+shift = 0
+S1D_d = S1D(time, 0, r, S_0)# 2*S_0/numpy.pi/r*(1-((time-shift)/r)**2)**0.5*theta(r**2 - (time-shift)**2)
+
+print "radius", r 
+
+#crescent
+R_p = r_half/0.72754
+R_n = 0.4*R_p
+a1 = 0.3*R_p
+a2 = -0.3*R_p
+
+
+S_0dp = S_0*R_p**2/(R_p**2-R_n**2)
+S_0dn = S_0*R_n**2/(R_p**2-R_n**2)
+
+
+
+S1D_dp = S1D(time, 0, R_p, S_0dp)
+
+S1D_dn1 = S1D(time, a1, R_n, S_0dn)
+
+
+S1D_dn2 = S1D(time, a2, R_n, S_0dn)
+
+zerofy(S1D_dp)
+zerofy(S1D_dn1)
+zerofy(S1D_dn2)
+
[…]

### 2017-09-11 paper1/ms.tex

@@
 engine of quasars. We present a toy model with a crescent-shaped source crossing a fold caustic. The silhouette 
-of a black hole over an accretion disk tends to produce roughly crescent sources. When a crescent-shaped source 
+of a black hole over an accretion disc tends to produce roughly crescent sources. When a crescent-shaped source 
 crosses a fold caustic, the resulting light curve is noticeably different from the case of a circular luminosity 
@@
 In Section~\ref{sec:numerics} we carry out source fitting to
-lightcurves, with both noise and systematic errors are present.  We
+lightcurves, where both noise and systematic errors are present.  We
 generate a mock lightcurve by taking a crescent source across the path
@@
 \includegraphics[width = .49\textwidth]{figures/infinite_fold_ar.eps}
-\caption{\label{fig:infinite_fold} Source profiles for Gaussian, disk,
+\caption{\label{fig:infinite_fold} Source profiles for Gaussian, disc,
   crescents (right) and magnification map for an infinite fold
@@
 
-\subsection{Disk source with constant surface brightness distribution}
+\subsection{Disc source with constant surface brightness distribution}
 
@@
 \end{equation}
-The half-light radius of a uniform disk source is $R/\sqrt{2}$.
+The half-light radius of a uniform disc source is $R/\sqrt{2}$.
 
@@
 The surface brightness distribution of a geometric crescent can be built 
-by considering two disk sources of constant brightness. One larger disk 
-will contribute positively to the total flux while one smaller disk 
+by considering two disc sources of constant brightness. One larger disc 
+will contribute positively to the total flux while one smaller disc 
 that is interior to the large one will contribute negatively. This superposition can be written for 2D as:\\
@@
 The following notations were used: $R_p$ and $(p_{sp}, q_{sp})$ are
-the radii and coordinate of the center for the larger positive disk, 
-while $R_n$ $(p_{sn},q_{sn})$ correspond to the smaller negative disk. 
-$S_0^{Dp},S_0^{Dn}$ represent the total flux of radiation received from the large and small disk. 
+the radii and coordinate of the center for the larger positive disc, 
+while $R_n$ $(p_{sn},q_{sn})$ correspond to the smaller negative disc. 
+$S_0^{Dp},S_0^{Dn}$ represent the total flux of radiation received from the large and small disc. 
 From this point forward we will not use the total flux from each source. Instead we will 
@@
 in the previously presented manner that need to be stated. First, we
-must impose the obvious $R_p \geq R_n$ relation. Secondly, the small disk
-must always be interior to the large disk:
+must impose the obvious $R_p \geq R_n$ relation. Secondly, the small disc
+must always be interior to the large disc:
 \begin{equation}
@@
 \end{equation}
-For the distances between the centers of the two disks we will use the same 
-notations as the one found in the paper \citep{2013MNRAS.434..765K}, $a \equiv p_{sn} - p_{sp}$ and $b \equiv q_{sn} - q_{sp}$.
+For the distances between the centers of the two discs we will use the same 
+notations as the one found in \cite{2013MNRAS.434..765K}, $a \equiv p_{sn} - p_{sp}$ and $b \equiv q_{sn} - q_{sp}$.
 
@@
 the centroid of the source is collinear with the centers of the two
-disks and it is situated at a distance $d_c$ from the center of the
-bright disk. $d_c$ can be computed numerically with the use of a
+discs and it is situated at a distance $d_c$ from the center of the
+bright disc. $d_c$ can be computed numerically with the use of a
 variation of equation \ref{eqn:s1_d}:
@@
 \caption{\label{fig:geom_crescent} Geometry of crescent sources: the source can be seen as a superposition
- of a smaller dark disk over and inside a larger brighter disk. $R_p$ and $R_n$ mark the radii of the 
-bright and dark disk. $d_c$ and $c$ mark the distance from the center of the bright disk to the 
-centroid and from the dark disk to the centroid.  }
+ of a smaller dark disc over and inside a larger brighter disc. $R_p$ and $R_n$ mark the radii of the 
+bright and dark disc. $d_c$ and $c$ mark the distance from the center of the bright disc to the 
+centroid and from the dark disc to the centroid.  }
 \end{figure}
@@
     \includegraphics[width = 0.48\textwidth]{figures/4source_magnification.eps}
-\caption{\label{fig:lightcurve_gauss} 1D profiles (top panel) and simulated lightcurves (bottom) of crescent, Gaussian and disk sources with identical total flux ($S_0$) and half-light radius($r_{1/2}$).
+\caption{\label{fig:lightcurve_gauss} 1D profiles (top panel) and simulated lightcurves (bottom) of crescent, Gaussian and disc sources with identical total flux ($S_0$) and half-light radius($r_{1/2}$).
 Crescent source has $r_{1/2}$ = 0.72754 $R_p$, $R_n$ = 0.4 $R_p$, a = 0.3/-0.3 $R_p$.
@@
 
-\subsection{Lightcurve of the disk shaped source}
+\subsection{Lightcurve of the disc shaped source}
 
@@
 \caption{\label{fig:a_var} 1D profiles (top panel) and simulated lightcurves 
-(bottom) of crescent sources with identical $R_p$ and $R_n$ ($R_n$ = 0.4 $R_p$) and different disk centers displacement (a).}
+(bottom) of crescent sources with identical $R_p$ and $R_n$ ($R_n$ = 0.4 $R_p$) and different disc centers displacement (a).}
 \end{figure}
@@
     \includegraphics[width = 0.48\textwidth]{figures/5Rn_back_var_magnification.eps}
-\caption{\label{fig:lightcurve_crescent_back} 1D profiles (top panel) and simulated lightcurves (bottom) of backwards crescent sources with identical $R_p$ and $a$ ($a$ = 0.3 $R_p$) and different dark disk radii ($R_n$).}
+\caption{\label{fig:lightcurve_crescent_back} 1D profiles (top panel) and simulated lightcurves (bottom) of backwards crescent sources with identical $R_p$ and $a$ ($a$ = 0.3 $R_p$) and different dark disc radii ($R_n$).}
 \end{figure}
@@
 \caption{\label{fig:lightcurve_crescent} 1D profiles (top panel) and simulated lightcurves (bottom) 
-of forwards crescent sources with identical $R_p$ and $a$ ($a$ = - 0.3 $R_p$) and different dark disk radii ($R_n$). }
+of forwards crescent sources with identical $R_p$ and $a$ ($a$ = - 0.3 $R_p$) and different dark disc radii ($R_n$). }
 \end{figure}
@@
 region that would be recorded on a lightcurve plot represents the
-period of time when the bright disk begins to be overlapped by the
-caustic and stops when just before the dark disk reaches the
+period of time when the bright disc begins to be overlapped by the
+caustic and stops when just before the dark disc reaches the
 caustic. During this period of time, the flux of light from the source
@@
 characterized by a smoother broad peak in contrast to the cases where
-the center of the dark disc reaches the caustic after the bright disk.
+the center of the dark disc reaches the caustic after the bright disc.
 In the case of the latter the instance when the dark disc reaches the
@@
   near $t=3$, and it is unable to fit the second sub-peak near $t=5$.}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.5\textwidth]{compare-mag.eps}
+\caption{``Visiting Figure''.  Only in red/blue version, not for the paper.}
 \end{figure}
@@
 The mock data appear in Figures
-\ref{fig:cc_forward}--\ref{fig:gc_forward}.  The example values
-above for $r_{1/2}$, $v_p$ and the observing frequency are not
-important.  What is important is that the observations have enough
-sampling and signal-to-noise that the characteristic features of a
-crescent source (see e.g., Figure~\ref{fig:char_points}) are not
+\ref{fig:cc_forward}--\ref{fig:gc_forward}.  Illustrative values have
+been adopted for the frequency of observations and for the noise
+level.  It is, of course, important is that the observations have
+enough sampling and signal-to-noise that the characteristic features
+of a crescent source (see e.g., Figure~\ref{fig:char_points}) are not
 washed out.
@@
 
-The first two source profiles that we consider are the uniform disk and the symmetric Gaussian source. 
+The first two source profiles that we consider are the uniform disc and the symmetric Gaussian source. 
 Both of them can be described by a half-light radius ($r_{1/2}$) and a total unlensed light flux ($S_0$). 
@@
 In the case of the crescent source there are in total five parameters: the integrated flux of the source $S_0$,
- the radii of the bright and dark disks and the displacement of the centers of the two disks on the axes perpendicular and parallel to the caustic.
+ the radii of the bright and dark discs and the displacement of the centers of the two discs on the axes perpendicular and parallel to the caustic.
 
@@
 information regarding the source that can be revealed by the lightcurve, 
-the value of the crescent's disks centers displacement parallel to the caustic 
+the value of the crescent's discs centers displacement parallel to the caustic 
 does not have an effect on the shape of the lightcurve.
@@
 crescent has two additional relevant parameters: the radius of the dark 
-disk ($R_n$) and the center displacement in the direction
+disc ($R_n$) and the center displacement in the direction
  perpendicular to the caustic $a$.
@@
 the instances correspond to the start of the overlap between the
-caustic and the bright disk, the start of the overlap between the
-caustic and the dark disk, the end of the overlap between the caustic
-and dark disk and finally the end of the overlap between the caustic
-and the bright disk.
+caustic and the bright disc, the start of the overlap between the
+caustic and the dark disc, the end of the overlap between the caustic
+and dark disc and finally the end of the overlap between the caustic
+and the bright disc.
 
@@
 shadow in the bright region limited by the extent of the event horizon
-of the black hole. Therefore, the radius of the bright disk would
-provide an estimate of the size of the accretion disk and the radius
+of the black hole. Therefore, the radius of the bright disc would
+provide an estimate of the size of the accretion disc and the radius
 of the dark disc would provide and estimate of the gravitationally
@@
 \def\aap{A\&A}
-\def\nat{nature}
+\def\nat{Nature}
 \def\araa{ARAA}

### 2017-09-14 paper1/ms.tex
editing remaining points

@@
   \includegraphics[width=0.5\textwidth]{compare-mag.eps}
-\caption{``Visiting Figure''.  Only in red/blue version, not for the paper.}
+\caption{[{\bf Temporary.} This figure is included only the in
+    red/blue version, and is not for the paper.] Overlay from the
+  three preceding figures.}
 \end{figure}
@@
 caustic (that is, $v_p$).  The orientation of the source track is
-another nuisance parameter, but we do not include it, because for a
-pure fold, changing the orientation of the source track is equivalent
-to changing $v_p$ and the orientation of the source (see
-Figure~\ref{fig:infinite_fold}), all parameters that are included.
-Since the caustic is not a perfect fold, disregarding the orientation
-of the source track can be expected to introduce some systematic
-error.
+another nuisance parameter.  For a pure fold, changing the orientation
+of the source track is equivalent to changing $v_p$ and the
+orientation of the source (see Figure~\ref{fig:infinite_fold}).  Since
+the latter two parameters are already accounted for, we do not include
+the orientation as a separate parameter.  Since the caustic is not a
+perfect fold, disregarding the orientation of the source track will
+introduce some systematic error.
 

### 2017-09-15 paper1/ms.tex
editing

@@
   7~parameters.)  The lower panel shows the posterior distributions
-  for $r_{1/2}$ (solid histogram) and $R_n/R_p$ (dashed histogram).
+  for $r_{1/2}/v_p$ (solid histogram) and $R_n/R_p$ (dashed histogram).
   Vertical dashes mark the correct values.}
@@
   \ref{fig:cc_backward}, but showing the best fit to a Gaussian
-  source.  The histogram shows the posterior distribution of $r_{1/2}$
-  (there is no $R_n$).  It appears to show $r_{1/2}$ as accurately
+  source.  The histogram shows the posterior distribution of $r_{1/2}/v_p$
+  (there is no $R_n$).  It appears to show $r_{1/2}/v_p$ as accurately
   recovered, but is not to be trusted, because the fit has
@@
 caustic (that is, $v_p$).  The orientation of the source track is
-another nuisance parameter.  For a pure fold, changing the orientation
-of the source track is equivalent to changing $v_p$ and the
-orientation of the source (see Figure~\ref{fig:infinite_fold}).  Since
-the latter two parameters are already accounted for, we do not include
-the orientation as a separate parameter.  Since the caustic is not a
-perfect fold, disregarding the orientation of the source track will
-introduce some systematic error.
+another nuisance parameter.  For an ideal fold, changing the
+orientation of the source track is equivalent to changing $v_p$ and
+the orientation of the source (see Figure~\ref{fig:infinite_fold}).
+Since the latter two parameters are accounted for, we do not include
+the orientation of the track as a separate parameter.  Since the
+caustic is not an ideal fold, disregarding the orientation of the
+source track can be expected to introduce some systematic error.
 
@@
 three cases) along with best fits and the fitted values and
-uncertainties on $r_{1/2}$ and (if applicable) $R_n/R_p$.
+uncertainties on $r_{1/2}/v_p$ and (if applicable) $R_n/R_p$.
 
@@
 shows, the lightcurve can be fitted to within the noise, and the
-parameters $r_{1/2}$ and $R_n/R_p$ can be recovered.
+parameters $r_{1/2}/v_p$ and $R_n/R_p$ can be recovered.
 
@@
 Figure~\ref{fig:magnification_map}.  That is, the mock data and the
-templates come from crossings of different caustics.  This mimics the
-unavoidable systematic error of not knowing the caustic exactly.  The
-fit is slightly degraded, but still acceptable.  The parameters
-$r_{1/2}$ and $R_n/R_p$ are still recovered.  The parameters $a/R_p$
-and $b/R_p$ were not correctly recovered, which is understandable from
-the track change.
+templates come from crossings of different caustics.  This is intended
+to mimic the systematic errors arising from not knowing the caustic
+exactly.  As Figure~\ref{fig:cc_backward} shows, the fit is slightly
+degraded, but still acceptable.  The parameters $r_{1/2}/v_p$ and
+$R_n/R_p$ are still recovered.  The parameters $a/R_p$ and $b/R_p$
+(not shown in the Figure) giving the orientation of the crescent were
+not correctly recovered; but this is understandable, because the AB
+and CB tracks have different orientations with respect to the caustics
+crossed.  The slight degradation of the fit can be understood as
+follows.  Although the caustics are clean folds, they have some
+curvature, and moreover the curvature is not identical in the two
+caustics used for the mock data and the template.
 

### 2017-10-10 paper1/ms.tex
ready to resubmit

@@
 (not shown in the Figure) giving the orientation of the crescent were
-not correctly recovered; but this is understandable, because the AB
-and CB tracks have different orientations with respect to the caustics
+not correctly recovered; but this expected, because the AB and CB
+tracks have different orientations with respect to the caustics
 crossed.  The slight degradation of the fit can be understood as
-follows.  Although the caustics are clean folds, they have some
-curvature, and moreover the curvature is not identical in the two
-caustics used for the mock data and the template.
+follows.  The source track will not, in general, be normal to the
+caustic, but will instead have a non-normal velocity component.  If
+the caustic is perfectly straight, the velocity component along the
+caustic makes no difference.  (Hence our model only considers only the
+normal velocity component $v_p$).  But in fact the caustics, although
+clean folds do have some curvature.  The combination of a curving
+caustic and non-normal velocity of the source makes the orientation of
+the source twist with respect to the caustic.  This twist would be
+different for crossings along the AB and CB tracks.
 

### 2017-11-17 paper1/msold4.tex
further edits and response

--- 2017-10-10 paper1/ms.tex
@@
 \title{Microlensing as a possible probe of event-horizon structure in quasars}
-\author[Tomozeiu et al]{Mihai Tomozeiu\thanks{mihai@physik.uzh.ch},$^{1,2}$ 
+\author[Tomozeiu et al]{Mihai Tomozeiu$^{1,2}$
 Irshad Mohammed,$^{1,2,3}$ 
@@
 
-\begin{figure}
-\centering
-  \includegraphics[width=0.5\textwidth]{compare-mag.eps}
-\caption{[{\bf Temporary.} This figure is included only the in
-    red/blue version, and is not for the paper.] Overlay from the
-  three preceding figures.}
-\end{figure}
+%\begin{figure}
+%\centering
+%  \includegraphics[width=0.5\textwidth]{compare-mag.eps}
+%\caption{[{\bf Temporary.} This figure is included only the in
+%    red/blue version, and is not for the paper.] Overlay from the
+%  three preceding figures.}
+%\end{figure}
 
@@
 \def\pasa{PASA}
-\bibliography{heap}
+\bibliography{ms}
 

### 2017-11-17 fitting/summaryfig.py

@@ -77,20 +77,31 @@ def plotcurves(p):
 
 def plothist(p):
     rmin,rmax,nr = 0,1.2,61
-    p.hist(rh,np.linspace(rmin,rmax,nr),weights=wt,
+    print estim(rh)
+    p.hist(rh,np.linspace(rmin,rmax,nr),weights=wt,normed=True,
            histtype='step',lw=2,color='black')
-    p.plot([1,1],[.43,.48],lw=2,color='black')
+    p.plot([1,1],[20,23],lw=2,color='black')
     fl = max(rns)
     if fl > 0:
-        p.hist(rns,np.linspace(rmin,rmax,nr),weights=wt,
+        print estim(rns)
+        p.hist(rns,np.linspace(rmin,rmax,nr),weights=wt,normed=True,
                histtype='step',lw=2,color='black',linestyle='dashed')
-        p.plot([.4,.4],[.43,.48],lw=2,color='black',linestyle='dashed')
+        p.plot([.4,.4],[20,23],lw=2,color='black',linestyle='dashed')
         p.set_xlabel('inferred $R_n/R_p$ and $r_{1/2}/v_p$',labelpad=10)
     else:
         p.set_xlabel('inferred $r_{1/2}/v_p$',labelpad=10, fontsize=16)
-    p.axis([rmin,rmax,0,0.5])
+    p.set_ylabel('posterior distribution',labelpad=10, fontsize=16)
+    p.axis([rmin,rmax,0,24])
 
-
+def estim(x):
+    N = len(x)
+    lo = int(.05*N)
+    mid = int(0.5*N)
+    hi = int(.95*N)  # 90% confidence
+    y = sorted(x)
+    res = y[mid],y[hi]-y[mid],y[lo]-y[mid],y[0],y[-1]
+    res = [int(100*x+0.5)*0.01 for x in res]
+    return res
 
 mockdata()
 
@@ -105,7 +116,7 @@ for fname in ['gc_forward','cc_forward','cc_backward']:
 
     p = fig.add_subplot(212)
     plothist(p)
-    p.get_yaxis().set_visible(False)
+#    p.get_yaxis().set_visible(False)
 
     fig.savefig('../paper1/figures/%s.eps'%fname)
 

### 2017-11-26 paper1/ms.tex
checking...

@@
 
-\subsection{Lightcurve of the crescent shaped source}
+\subsection{Lightcurve of the crescent shaped source}\label{subsec:lcurvcres}
 
@@
 \caption{\label{fig:M87_plots} The one dimensional profile obtained
-  from the numerical integration along the ordonata axis of a
+  from the numerical integration along the ordinate axis of a
   simulated source image is presented in the upper panel. The source
@@
 source could be discerned in real data having noise and systematic
-error as well.  Do do so, we first generate a microlensing lightcurve
+error as well.  To do so, we first generate a microlensing lightcurve
 using a widely-used microlensing code
@@
   7~parameters.)  The lower panel shows the posterior distributions
-  for $r_{1/2}/v_p$ (solid histogram) and $R_n/R_p$ (dashed histogram).
-  Vertical dashes mark the correct values.}
+  for $r_{1/2}/v_p$ (solid histogram) and $R_n/R_p$ (dashed
+  histogram).  Vertical dashes mark the correct values.  Inferred are
+  $r_{1/2}/v_p=0.90^{+0.15}_{-.07}$ and $R_n/R_p=0.32^{+0.19}_{-0.08}$
+  (90\% confidence).}
 \end{figure}
@@
   $\chi^2_{(244)}=284, P=0.01$ indicating a marginally acceptable
-  fit.}
+  fit. Inferred are $r_{1/2}/v_p=0.92^{+0.06}_{-0.04}$ and
+  $R_n/R_p=0.32^{+0.12}_{-0.07}$ (90\% confidence).}
 \end{figure}
@@
   Gaussian source.  Looking closely, the curve is a little too shallow
-  near $t=3$, and it is unable to fit the second sub-peak near $t=5$.}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.5\textwidth]{compare-mag.eps}
-\caption{[{\bf Temporary.} This figure is included only the in
-    red/blue version, and is not for the paper.] Overlay from the
-  three preceding figures.}
+  near $t=3$, and it is unable to fit the second sub-peak near $t=5$.
+  Inferred is $r_{1/2}/v_p=0.91^{+0.05}_{-0.01}$ (90\% confidence).}
 \end{figure}
@@
 been adopted for the frequency of observations and for the noise
-level.  It is, of course, important is that the observations have
+level.  It is, of course, important that the observations have
 enough sampling and signal-to-noise that the characteristic features
@@
 degraded, but still acceptable.  The parameters $r_{1/2}/v_p$ and
-$R_n/R_p$ are still recovered.  The parameters $a/R_p$ and $b/R_p$
+$R_n/R_p$ are still recovered, though less well; the correct
+value of $r_{1/2}/v_p$ lies outside 90\% of the MCMC distribution,
+but within 99\%.   The parameters $a/R_p$ and $b/R_p$
 (not shown in the Figure) giving the orientation of the crescent were
@@
 Namely, the shape of the caustic boundary in the proximity of the source in the respective plane can be approximated by
- a line due to reason that the local radius of curvature of caustic is orders of magnitude greater than the half-light
+ a line due to reason that the local radius of curvature of the caustic is orders of magnitude greater than the half-light
  radius of the studied source. 
@@
 
-Nevertheless, the respective displacement is relevant for the calculation of
+Nevertheless, the respective displacement is relevant for the calculation
 of the half-light radius.  Its qualitative effect is to 
@@
 The observer would basically detect only the lightcurve of such a source.
- As described in section 4.3 the timing of the onset and 
+ As described in section~\ref{subsec:lcurvcres} the timing of the onset and 
 offset of the previously 
@@
 
-In the case of a high-quality lightcurve, the parameters of a crescent
-source can be estimated by simply identifying signatures on the
+In the case of a high-quality lightcurve, provided the caustic is a
+fairly clean fold, or equivalently, the source is much smaller than
+the local radius of curvature of the caustic, the crossing times of 
+the bright and dark areas (cf.~Figure~\ref{fig:char_points}) could
+be estimated by simply identifying signatures on the
 lightcurve corresponding to transits of the particular source features
 across the caustic.  This simple idea can be supplemented with
-statistics.  We do so in Section~\ref{sec:numerics}, where we take a
+statistics.  In Section~\ref{sec:numerics}, we take a
 realistic magnification map generated by the microlensing code
@@
 denotes the component of the relative velocity of the source and fold
-which is perpendicular to the caustic.  The respective velocity is an
-unknown, though it can be constrained on a case by case basis to an
-order of magnitude or even better. This would require the study of the
-dynamics of the stellar structure which contains the gravitational
-lens and the bulk transverse velocity of the galaxy with respect to
-the observer.  A better estimate of the relative velocity would
-facilitate a better estimate of the effective non-rotating black hole
-mass associated with the black hole shadow.
+which is perpendicular to the caustic.
+
+% The respective velocity is an
+% unknown, though it can be constrained on a case by case basis to an
+% order of magnitude or even better. This would require the study of the
+% dynamics of the stellar structure which contains the gravitational
+% lens and the bulk transverse velocity of the galaxy with respect to
+% the observer.  A better estimate of the relative velocity would
+% facilitate a better estimate of the effective non-rotating black hole
+% mass associated with the black hole shadow.
 
